### 2017-02-01 introduction.tex
Add files via upload

\chapter{Introduction}

\section{Defining the urban surface}
 	Sometimes I have no idea what I am doing.

 \section{The urban radiation budget}

 \section{The structure of the following sections}

### 2017-02-01 paper1.tex
Add files via upload

\chapter{A method to correct upwelling longwave radiation to estimate hemispherical urban surface temperature}

\section{Introduction}
Thermal infrared (TIR) remote sensing of land surface temperature (T\textsubscript{surf}) has emerged as a primary research focus in climatology, as researchers seek to better describe spatiotemporal patterns of T\textsubscript{surf} globally and better understand how anthropogenic modification of earth's surface influences land T\textsubscript{surf} with links to the climate at-large. Over the last two decades, use of thermal remote sensing of surface climates has expanded significantly --- both in terms of the volume and breadth of remote sensed study and its explicative importance in climatology as a discipline. Thermal remote sensing of earth's surface has applications over a wide range of disciplines: from informing micro-, urban-, and global-scale climate models, to aiding decision making and mitigation praxis with respect to climate change and the \gls{uhi}.

Within urban climatology, a combination of satellite, aerial, and ground-based thermal remote sensors have been integral in elucidating the spatial \cite{Roth1989}, temporal \cite{Peng2012}, and geometric \cite{Voogt1997} effects of the built environment on land T\textsubscript{surf}; in evaluating and partitioning urban surface energy balances \cite{BastiaanssenW.G.M.1998, Yamaguchi2005} and; in characterizations of the relationship between surface and \gls{blayer} air temperatures (T\textsubscript{air}) \cite{Stoll1992}. These advances have been aided by substantial improvements in sensor ground, spectral, and radiometric resolutions, and by the proliferation of both large-scale public satellite remote sensing campaigns and low-cost aerial and near-ground thermography. However, in spite of it's widespread usage, several questions concerning the use and validity of urban remote thermal remote sensing, first posed in Roth et al. 1989 \cite{Roth1989}, have yet to be sufficiently answered, viz,

\begin{enumerate}
	\item What is the nature of the surface 'seen' by a thermal remote sensor?
	\item How does T\textsubscript{surf} observed by a remote sensor relate to the 'true' temperature governing the surface-atmosphere interface?
\end{enumerate}

\noindent In this paper, we seek to examine question two by introducing and evaluating a method for atmospheric and emissivity correction of near-ground hemispherical TIR --- measured via \gls{pyrg} --- for hemispherical radiometric temperature (T\textsubscript{hem}) retrieval. These measures are common to most urban energy balance assessments and thus constitute a hitherto untapped method for urban T\textsubscript{surf} analysis. A companion paper responds to question one through an analysis of a climatology of T\textsubscript{hem} and derived surface UHI (sUHI), to quantify geometric and temporal biases across multiple methods for remote sensing of urban T\textsubscript{surf}.

\section{Atmospheric effects on TIR}

Although most thermal remote sensors operate within one of the \gls{atmwind}[s] --- where atmospheric effects are greatly reduced --- virtually any remote sensed TIR signal is subject to radiative effects from the intervening atmosphere between the surface and the sensor. Over much of the thermal infrared \gls{waveband} the atmosphere emits radiation and absorbs a fraction of radiation emitted by the surface. In addition, the earth (and indeed the urban surface) reflects a portion of incoming solar radiation. Thus, a TIR signal measured from a remote sensing platform is almost certainly not equal to the ground emitted signal. These effects can lead to differences between the 'true' radiometric T\textsubscript{surf} and the remote sensed T\textsubscript{surf} of over 10 k for satellite platforms \cite{Cooper1989} and over 6 K for near-ground sensors \cite{Meier2011}. Moreover, because atmospheric and emissivity effects are a function of non-uniform and spatiotemporally variant surface and atmospheric properties, their associated errors change depending on instrument type, study location, and ambient conditions. As intersite and time-sequential analysis is a significant goal of most thermal remote sensed studies (urban or otherwise) these effects cannot be ignored.

Spectral transmission of longwave radiation through a given layer of atmosphere varies significantly based on wavelength, path geometry, and column absorber content (primarily H\textsubscript{2}0, C0\textsubscript{2}, and to a lesser extent 0\textsubscript{3}, N\textsubscript{2}0, CO, CH\textsubscript{4}, and 0\textsubscript{2} \cite{Miskolczi1993}). Variation in band-by-band longwave transmittance with path length is greatest at urban scales (approximately 1 to 50 meters), where transparent spectral bands can quickly become opaque with small changes in path length or absorber content - illustrated in figure \ref{spectransheight}. Thus, transmittance of TIR near the surface is highly dependent on surface-sensor-sun geometry, instrument properties, and synoptic conditions. Indeed, accurate assessment of atmospheric influence on TIR may be most complex when measured near the surface.

\begin{figure}[H]
	\includegraphics[width=\textwidth]{spectransheight}
	\caption{Spectral transmission of water vapor as a function of height. Black and grey shading indicate opaque (0\% transmittance) and transparent (100\% transmittance) bands respectively. Redo this figure with wavelength and colorbar?}
	\label{spectransheight}
\end{figure}

The radiation passing through a layer of atmosphere from an emitting surface can be described in a number of ways. \textit{Spectral directional radiance} $ R (\lambda)$, at height $ z $ and from a direction defined by zenith angle \(\theta\), and azimuth angle \(\phi\) is commonly written as

\begin{equation}
R^\uparrow_z (\lambda, \theta, \phi) = \tau_\lambda \epsilon R^\uparrow_0(\lambda) + (1-\epsilon) R^\downarrow_{sky} (\lambda) + (1-\tau_\lambda) R^\uparrow_{atm}(\lambda)
\end{equation}

\noindent where \(\epsilon\) is surface emissivity, \(\tau\)\textsubscript{\( \lambda \)} is spectral "slab" transmittance through the layer between the emitting surface $( z  = 0) $ and $ z $. The same spectral TIR signal, as measured by a narrow-FOV sensor mounted at height $ z $, passes through an instrument shield (or dome) with spectral transmittance \(\tau_d\), and is integrated over the sensor waveband (bounded by \(\lambda_1\) and \(\lambda_2\)) to yield a \textit{directional radiance} $L'$ as 'seen' by the sensor

\begin{equation}
L'_z (\theta, \phi) = \int_{\lambda_1}^{\lambda_2} \tau_d(\lambda) R^\uparrow_z(\lambda) ~ d\lambda
\end{equation}

\noindent which, integrated over the hemisphere with respect to zenith \(\theta\) angle and azimuth \(\phi\) angle, yields an \textit{irradiance} $ L $ at height $ z $,

\begin{equation}
L_z = \int_{0}^{2\pi} \int_{0}^{\pi/2} L_z'(\theta, \phi) ~ \cos\theta \sin\theta ~ d\theta d\phi
\end{equation}

\subsection{Relating TIR and surface temperature}

TIR received by a remote sensor can be related to the emitting body's temperature in a number of ways --- each producing different conceptions of T\textsubscript{surf} from different instrument and sensor types. As such, the term "surface temperature" with respect to a remote sensed TIR is vague and can refer to several definitions of "surface" and "temperature" Thus, proper terminology must be attached to land T\textsubscript{surf} inferred from TIR. Definitions and nomenclature conventions for multiple methods for T\textsubscript{surf} retrieval are discussed at length in Norman \& Becker, (1995)\cite{Norman1995}.

Directional radiance $ L'_z $ detected from a narrow-FOV sensor operating over waveband (\(\lambda_1\) --- \(\lambda_2\)) and viewing the surface from some orientation described by \(\theta\), \(\phi\) can be used to infer a \textit{directional brightness temperature} T'\textsubscript{bright}(\(\theta\), \(\phi\)) via an inversion of the Planck function multiplied by normalized sensor response integrated over the sensor waveband,

\begin{equation}
L'_z(T'_{bright} (\theta, \phi)) = \bigints_{\lambda_1}^{\lambda_2} \cfrac{f(\lambda) C_1}{\pi \lambda^5 \left( \exp\left(\cfrac{C_2}{\lambda T'_{bright} (\theta, \phi)}\right)\right)}
\end{equation}

\noindent where C\textsubscript{1} = $ 3.7404 \cdot 10^8 $ $ W\mu^4 m^{-2} $, C\textsubscript{2} = $ 14387 \mu K $, and relative sensor response $ r(\lambda) $ normalized via,

\begin{equation}
1 ~ =  \int_{\lambda_1}^{\lambda_2} r(\lambda) ~ d\lambda
\end{equation}

Similarly irradiance $L_z$ received by a broadband hemispherical sensor (such as a pyrgeometer), can be used to infer a \textit{hemispherical brightness temperature} T\textsubscript{bright} through an inversion of the Stefan-Boltzmann law,

\begin{equation}
\label{stefb1}
L_z = \overline{r} (\sigma T_{bright}^4)
\end{equation}

\noindent where $ \sigma $ is the Stefan-Boltzmann constant and $ \overline{r} $ is Planck weighted average broadband sensor response computed as,

\begin{equation}
\overline{r} = \cfrac{\int R(\lambda) ~ r(\lambda)}{\int R(\lambda)}
\end{equation}

\noindent with $ R(\lambda) $ computed from a Planck function at an approximated $T_{bright}$.

In addition, a simple approximation of $T_{bright}$ can be inferred from directional radiance using equation \ref{stefb1} by replacing $ L_z $ with $L'_z$ multiplied by a constant. This method is commonly used to infer T\textsubscript{bright} from infrared thermometers (IRT) operating over the atmospheric window (where atmospheric correction magnitudes are small and T\textsubscript{bright} is a reasonably accurate approximation of T\textsubscript{surf}). However, constants must be calibrated for the range of expected T\textsubscript{surf} as the relationship between $ L_z $ and $L'_z$  is not perfectly linear.

Inversions of uncorrected directional radiances or irradiances using the Planck function or the Stefan-Boltzmann law yield a temperature equal to that of a blackbody emitting the same amount of radiation as detected by the sensor. Since $ L_z $ is unlikely to be equal to $ L_0 $, T\textsubscript{bright, 0} and T\textsubscript{bright, z} often show significant variation based on sensor characteristics, ambient conditions, and surface-sensor geometry. Hence, remote sensed T\textsubscript{bright} is generally considered only a rough approximation representation of kinetic or thermodynamic T\textsubscript{surf}.

To retrieve a more accurate estimation of the 'true' T\textsubscript{surf}, the same inversions can be applied to TIR measurements after correction for atmospheric and surface emissivity effects (e.g. transformations of the TIR signal to represent that at $ z = 0 $ if the surface was emitting as a perfect blackbody) to yield a direction radiometric surface temperature T\textsubscript{rad} for atmospheric and emissivity corrected directional radiances, a hemispherical radiometric surface temperature T\textsubscript{hem} for atmospheric and emissivity corrected irradiances. $ T\textsubscript{rad} (\theta, \phi) $ and T\textsubscript{hem} provide a better approximation of the ‘true’ T\textsubscript{surf} by representing the temperature at which the surface is radiating integrated over the sensor FOV.

\subsection{Atmospheric correction of near-ground TIR} \label{Atmospheric correction of near-ground TIR}

A large number of correction routines have been developed to remove atmospheric and emissivity effects from aerial and satellite TIR signals and derive accurate T\textsubscript{rad}. Methods range from simple mono- \cite{Qin2001} and split-window \cite{Wan1996} routines for single- and multi-channel remote sensors to schemes that integrate a radiative transfer code to isolate the surface emitted signal from interfering signals. Boundary conditions are standard across most correction methods: often requiring vertical profiles of T\textsubscript{air}, humidity, pressure, and aerosol content to remove atmospheric effects, and assessments of surface radiative properties to correct for emissivity effects. However, correction methods are generally instrument (or at least platform) specific and difficult to generalize across sensor and platform types. Few methods exist for correction of ground based remote sensed TIR --- none of which are robust enough to correct irradiances upwelling from rough terrain measured  wide-\gls{fov}(FOV) sensor. In part, this is due to the fact that near ground thermal sensors have only recently become sensitive enough to require correction. However, a new breed of more accurate radiometers and thermal imagers should prompt a critical reevaluation of this assumption.

Atmospheric correction of near-ground remote sensed TIR is subject to a unique set of challenges compared to satellite and aerial platforms. Near-ground remote sensors have complex, multiple line-of-sight (LOS) geometry - illustrated in figure 2.1 for a wide-FOV instrument. Surface to sensor path lengths vary significantly over the sensor FOV, even when measured close (less than 5m) above the surface. This creates the potential for non-uniform atmospheric effects over the sensor FOV and necessitates a multi-LOS correction to retrieve accurate T\textsubscript{rad}. The addition of 3-dimensional surface geometry further amplifies this effect, as some paths may intersect with raised vertical, sloped, and horizontal features. In effect, with near-ground wide-FOV sensors, surface geometry is non-trivial and must be represented in atmospheric correction routines. In contrast, over a scene retrieved via satellite, spatial viability in surface geometry and LOS angle have a negligible effect on path length. Atmospheric correction routines for satellite retrieved TIR, therefore, assume uniform or single-LOS geometry because the TIR signal passes through a relatively constant volume of atmosphere over the projected sensor FOV, regardless of surface geometry. This greatly reduces the complexity of correction routines for satellite or aerial thermography.

\begin{figure}[!ht]
	\includegraphics[width=\textwidth]{lineofsight}
	\label{lineofsight}
	\caption{Variable path geometry inherent with wide-FOV near-ground sensor visualized over an idealized 2-dimensional urban area.}
\end{figure}

Several multi-LOS atmospheric correction routines have been developed for near-ground TIR: Meier et al., 2011 describes a correction method for oblique angled urban thermography \cite{Meier2011}. Path lengths were calculated for each pixel over the FOV of a tower-mounted thermal imager angled obliquely toward the urban surface. A correction factor was then computed and applied using a radiative transfer code initialized using isothermal, isohumal atmospheric profiles for each time step. Applied over a time-series of images, the result is a continuous atmospherically corrected series of thermal images with each pixel representing a different T\textsubscript{rad}. However, the method uses visual-band images and a GIS database to represent urban geometry for each pixel's LOS - a technique not possible with a pyrgeometer, which returns a single integrated irradiance at each time-step. Moreover, the target instrument operates over a narrow waveband, reducing the magnitude and variance in atmospheric transmission over the sensor response curve. Thus, the method is not directly generalizable to correction TIR as measured via pyrgeometer.  Kotani \& Sugita, 2009 describes a method for correction of wide-FOV (pyrgeometer) TIR Irradiances \cite{Kotani2009a} over planar terrain. However, these methods are limited to narrow-FOV thermal imagers and sensors over planar surfaces respectively. A method which combines Meier et al., 2009's representation of complex surface geometry and Kotani \& Sugita, 2009's broadband hemispherical integration is needed for atmospheric correction of urban TIR measured from a downward facing pyrgeometer. A method to correct hemispherical broadband TIR as measured from a downward facing pyrgeometer needs to combine the two for accurate T\textsubscript{hem} retrieval over rough terrain.

\section{A "rolling lookup table" method for hemispherical atmospheric correction}

%alt title:method to retrieve atmospherically corrected hemispherical surface temperatures from near ground TIR

The "rolling lookup-table" method described in this study uses a sensor view model in conjunction with a radiative transfer code to model hemispherical irradiances upwelling from a simplified isothermal 3-dimensional representation of the target study area. In summary, the method (depicted in figure \ref{flow}) uses vertical profiles of measured T\textsubscript{air} and humidity to model at-sensor spectral radiances at 5$^{\circ}$ increments over the sensor FOV for a predefined range of possible T\textsubscript{hem} at each time-step. Spectral directional radiances are convolved by a dome transmittance curve, integrated over the sensor waveband, and weighted for their respective angular view factor. Weighted directional radiances are then integrated over the hemisphere and aggregated into a lookup table (LUT) of modeled irradiance---T\textsubscript{hem} pairings for each timestep, unique to the vertical profile of measured T\textsubscript{air} and humidity. Finally, for each time-step, measured irradiances are matched with the closest modeled irradiances in the associated LUT to return an atmospherically corrected hemispherical surface temperature (T\textsubscript{hem}). This process is repeated at 30 minute intervals to yield a continuous climatology of urban T\textsubscript{hem} for surface urban heat island (sUHI) analysis. The following sections introduce the study area and describe the sensor view modeling, radiative transfer, and post-processing steps of the method.

\begin{figure}[!ht]
	\includegraphics[width=\textwidth]{flow}
	\label{flow}
	\caption{A workflow schematic depicting the input, model, and output-processing steps of a "rolling lookup table method" for hemispherical radiometric surface temperature retrial.}
\end{figure}

\subsection{Study area}

As discussed in Section \ref{Atmospheric correction of near-ground TIR}, atmospheric correction of longwave irradiances measured from downward-facing, near-ground, wide-FOV sensors must account for surface geometry - particularly when mounted to view complex surface geometry. Hence, routines to retrieve atmospherically corrected urban T\textsubscript{hem} from upwelling longwave irradiances are inherently site specific. However, it is important to note that although correction magnitudes described in this paper are not generalizable, the correction method described in this paper can readily be adapted to different study sites, pyrgeometer types, and unique surface geometries.

With methodological generalizability in mind, a "rolling lookup table" atmospheric correction method was developed to retrieve radiometric T\textsubscript{hem} for a climatology of upwelling longwave irradiances measured from above the Sperrstrasse urban street canyon in Basel, Switzerland. The site, instrumented as a part of the Basel Urban Boundary Layer Experiment (BUBBLE) \cite{Rotach2005}, has an approximately northeast-southwest orientation and site morphology representative of local climate zone (LCZ) 2\footnote{Site surroundings can be described by the following morphological parameters: mean building height: 14.6m, plan aspect ratio: 0.54, complete aspect ratio: 1.92, local canyon aspect ratio: 1.0, and average shortwave albedo: 0.11 \cite{Rotach2005}.} \cite{Stewart2012}. LCZ classification was based on an assessment of surface characteristics in a 250m circular area extending from the Sperrstrasse tower using a 1m raster digital building model (DBM). Thus, the morphological parameters identified in \cite{Rotach2005} are representative of the vast majority of the pyrgeometer footprint. However, it should be noted that vegetation was not included in the DBM, and thus is not represented in morphological assessment or the sensor view model.

For the nine month period between November 2001 and July 2002, a triangular lattice tower was installed skewed towards the southeast facing wall near the center of the Sperrstrasse canyon to observe a full suite of meteorological, radiation, and flux parameters. Profiles of T\textsubscript{air} and humidity were measured from seven heights extending from 2.5m to 31.5m above the canyon floor (with the highest observation level at approximately 2.17 times mean building height). Upwelling and downwelling short/longwave fluxes were measured at the lowest and highest tower heights, with an additional downward facing pyrgeometer mounted at roof level in the center of the canyon. In addition, during a summertime intense operation period (IOP) an array of narrow-FOV IRTs was installed to view individual facet surface temperatures (T\textsubscript{facet}) of approximately the same urban patch viewed by the pyrgeometer.

The BUBBLE Sperrstrasse site was chosen for two primary reasons: 1. The study site provided a long-term climatology of radiation and meteorological variables for a representative mid-latitude city over a diverse range of synoptic conditions. This allowed for examination of urban T\textsubscript{surf}, atmospheric correction magnitudes, and sUHI magnitudes over a wide range of mid-latitude conditions. 2. T\textsubscript{facet} measured during the summertime IOP allowed for direct climatological comparison of T\textsubscript{hem} to wall (T\textsubscript{wall}), roof (T\textsubscript{roof}), and road (T\textsubscript{road}) surface temperatures of both the northwest- and southeast-facing sides of the canyon. In addition, several hypothetical sensor views of the canyon were simulated by weighting to represent different geometric representations of the canyon, including a nadir/plan view, an oblique south-facing, and an oblique north-fac from narrow-FOV sensors and a complete, 3-dimensional view of the canyon. This facilitated comparison of T\textsubscript{hem} against different instrument types and platforms to identify and quantify biases in each method.  In each representation, different weightings were applied to T\textsubscript{wall}, T\textsubscript{roof}, and T\textsubscript{road} on both sides of the canyon to represent different sensor orientations and sampling regimes. Weightings for each representation are described in table \ref{weightings}. 
\begin{table}[!ht]
	\centering
	\caption{Weightings to produce T'\textsubscript{rad} for different geometric representations of the Basel street canyon.}
	\label{weightings}
	\begin{tabular*}{\textwidth}{l@{\extracolsep{\fill}} p{1cm}p{1cm}p{1cm}p{1cm}p{1.6cm}}
		\toprule 
		& Road & Northwest & Southeast & Northwest & Southeast \\ 
		&  & Roof & Roof & Wall & Wall \\ 	\midrule

		Complete & 0.33 & 0.16 & 0.16 & 0.16 & 0.16 \\

		Nadir & 0.46 & 0.27 & 0.27 & 0.00 & 0.00 \\

		Oblique (south-facing) & 0.20 & 0.25 & 0.25 & 0.00 & 0.30 \\

		Oblique (north-facing) & 0.20 & 0.25 & 0.25 & 0.30 & 0.00 \\ 
		\bottomrule
	\end{tabular*} 
\end{table}

\subsection{Modeling path lengths of 3-dimensional terrain}

Radiative transfer of broadband longwave radiation in highly dependent on sensor-surface geometry. To calculate surface-sensor path lengths, the Surface-Sensor-Sun Urban Model (S\textsuperscript{3}UM) \cite{Soux2004} is initialized with a simplified, orthogonal 3-dimensional DBM representing the area immediately surrounding the sensor. In S\textsuperscript{3}UM, urban features (roads, roofs, buildings, etc.) are constructed from an array of horizontal and stacked vertical patches (representing rooftops/roads and walls respectively). After indicating sensor FOV and position over the DBM, the model determines which patches have an unobstructed line of sight to the sensor and calculates the path distance from "seen" patches to the sensor. Path lengths are binned at 5$^{\circ}$ increments from the nadir and averaged to return an azimuthally averaged path length for each 5$^{\circ}$ bin. Using azimuthally averaged path lengths in lieu of a full 3-dimensional radiative transfer model significantly reduces the computational time required to model each irradiance-T\textsubscript{hem} pairing, as at-sensor radiance can be computed independent of azimuth angle. Radiative transfer is thus reduced to a two dimensional problem (shown in panel (b) of figure \ref{23radtran}) in which an azimuthally averaged radiance is simulated for each zenith angle. The view factors required to weight directional radiances in the hemispherical integrated are computed While generating azimuthally averaged path lengths, S\textsuperscript{3}UM calculates view factors for each solid angle bin used in weighting 2-dimensional angular radiances during the hemispherical integration.  By using azimuthally averaged path lengths in lieu of  However, because hemispherical TIR sensor returns a bulk irradiance, the sensor signal does not need to be computed for each point of the

\begin{figure}[!ht]
	\includegraphics[width=\textwidth]{23radtran}
	\label{23radtran}
	\caption{A typical 2-dimensional (b) radiative transfer schematic adapted for an idealized 3-dimensional urban area (a). In 3-dimensions, path length for a given zenith angle can change significantly with azimuth angle. Dotted lines indicate the potential for absorption by the intervening atmospheric layer}
\end{figure}

\subsection{Modeling hemispherical irradiances}
With path length geometries calculates inS\textsuperscript{3}UM, MODTRAN 4.1 is initialized with 30-minute averaged profiles of T\textsubscript{air} and humidity to model at-sensor spectral directional radiances over a waveband of 0---2500cm\textsuperscript{-1}. Aerosol and above canyon meterological conditions are informed by the Mid-Latitude Summer standard atmosphere when dayttime maximum T\textsubscript{air} is greater than 10$^{\circ}$. For days with dayttime T\textsubscript{air} maximum below 10$^{\circ}$, the Mid-Latitude Winter standard atmosphere is substituted. In spite of significant computational efficiency gains from reducing 3-dimensional radiative transfer to azimuthally averaged 2-dimensions, the method still requires significant computational time to write input cards and model spectral radiances - in this case, with a path length increment of 5 and a LUT of 31 modeled irradiance---T\textsubscript{hem} pairings, each time-step requires 496 independent MODTRAN runs

\section{Evaluation of the method using profiles of upwelling longwave radiation over a homogenous planar surface}

MODTRAN has been shown to effectively model near-ground radiative transfer at urban-scale path lengths in 2-dimensions. However, the method detailed in this paper includes significant post-processing to ‘collate’ point-to-point radiances into 3-dimensional hemispherical irradiances. In addition, the method accounts for site-specific surface geometry, which has a significant influence on boundary-layer transmittance of broadband longwave radiation. Thus, prior to deriving a climatology of radiometric urban Them, the method was evaluated over a simple surface with known surface characteristics. Thus, divergences are solely the result of differential atmospheric effects with increasing height above ground, which should be accounted for by the model. It is assumed that each pyrgeometer in this validation views a patch with approximately the same temperature.

Profiles of upwelling longwave radiation, downwelling shortwave, air temperature, and humidity at 2m, 10m, and 30m were obtained from the Payerne station, located 50 miles southwest of Basel, Switzerland in a cultivated field. To evaluate the method, we used a modified version of the workflow described in XXXX. Brightness temperature calculated from the lowest upwelling longwave measurement was used to model irradiances at 10m and 30m at each time-step. Downwelling shortwave radiation was used to categorize test days based on cloud cover. The daytime air temperature profile was modified to replicate typical lapse rates in the Sperrstrasse canyon. Unlike, Kotani \& Sugita (2009), an isothermal atmospheric profile was not sufficient to accurately model upwelling longwave divergences (nor fluxes) of planar terrain. Although thermal stratification is likely to be relatively small by day in an urban canyon \cite{Nakamura1988} --- where strong microscale contrasts in Tsurf foster canyon mixing and neutral stability --- the large daytime Tsurf-T\textsubscript{air}  differential and the path-length/transmittance gradient can create large positive divergences (5 – 15 Wm\textsuperscript{-2},CHECK THIS). As such, we suggest using a full canyon T\textsubscript{air} /humidity profile to

Modeled upwelling longwave at 10m and 30m and divergences are compared to their measured counterparts in figure a. Both fluxes and divergence show strong correlations, thus we conclude: 1. Irradiances measured at 2m are free from significant atmospheric influence – as an uncorrected irradiance provided an accurate Tsurf for modeling of 10m and 30m Irradiances As such, brightness Them and radiometric Them are approximately equal when z < 2m. Although, it should be noted that a 2m sensor is not representative of canyon geometry and should not be used to derive urban Them. 2. The method is sensitive enough to model divergences in a large layer above a flat surface. Urban divergences are likely to be smaller with less frequent and less severe stable stratifications. Thus, we can safely make the assumption that the method is effective over complex terrain – provided path length geometries are accurately replicated in SUM.

### 2017-02-01 paper2.tex
Add files via upload

\chapter{A climatology of urban surface heat islands derived from hemispherical radiometric surface temperatures}

\section{Introduction}

The temperature of the surface is integral in understanding, predicting, and modeling boundary-layer air temperature patterns, surface energy balances, and, in urban areas, has important implications for human thermal comfort and building energy usage. Urban modification of surface geometry and thermal, radiative, moisture, and aerodynamic properties results in differential surface heating and cooling patterns and strong microscale spatiotemporal variations in urban surface temperature. Integrated up to larger scales, urban areas tend to store and generate more heat relative to non-built surroundings and, as a result, are generally warmer --- a phenomenon termed the urban heat island effect (UHI). As such, accurate, spatiotemporally continuous and geometrically representative characterization of surface temperatures has long been a goal in urban climatology. The proliferation of satellite and aerial thermal infrared (TIR) remote sensing has enabled spatially-extensive, large-scale characterizations of surface climates at ever improving spatial and spectral resolutions. Such campaigns have elucidated urban surface temperature (T\textsubscript{surf}) and surface urban heat island (sUHI) patterns globally at large spatial scales. However, technological improvements in TIR remote sensing have yet to address a three potential sources of error when applied in urban areas:

\begin{enumerate}
	\item Geometric undersampling of 3-dimensional terrain
	\item Temporal discontinuity in overpass cycles and sensor sampling regimes
	\item Clear-sky bias
\end{enumerate}

These biases present a potentially significant source of error by failing to capture micro-scale temporal and geometric variations in urban T\textsubscript{surf}.

Inter-site comparison is the crux of UHI analysis, thus it is imperative that urban surface temperature measurements are accurate and representative of coherent urban patches. Meta-analysis of air temperature UHI literature shows that these goals are not often not satisfied \cite{Stewart2011}. Given the relative difficulty in retrieving accurate, representative urban surface temperatures, similar conclusions are likely for sUHI analysis. In spite of this fact, and the short period over which large-scale generalizable methods for urban surface temperature acquisition have been available, study of sUHI has expanded significantly in the last twenty years \cite{Peng2012,Voogt2003}. This paper presents a method to derive hemispherical radiometric urban T\textsubscript{surf} (Them) from hemispherical upwelling TIR as measured by inverted pyrgeometers. This method was developed to addresses biases inherent in traditional methods for urban T\textsubscript{surf} retrieval by providing temporally continuous, geometrically representative\footnote{Them is not perfectly representative of urban geometry. This is best illustrated by visualizing an urban area from the perspective of a downward facing ‘fish-eye’ camera. However, it is undoubtedly more representative of urban geometry than most 2-dimensional views of the surface. This fact is discussed in more depth in sectionXXXX.}  urban T\textsubscript{surf} under all-sky conditions for sUHI analysis. These measurements are often made as a part of the net radiation determination for urban energy balance studies.

\section{Bias in Thermal remote sensing}
Geometric biases in remote sensing of the urban surface are a result of its 3-dimensional, convoluted structure. Compared to flat terrain, complex urban surface geometry modifies receipt of incoming solar radiation and traps a portion of reflected solar and outgoing terrestrial radiation. Differential heating and shading patterns of vertical, sloped, and horizontal urban facets manifests in significant micro-scale contrasts in surface temperature. As such, measured radiometric surface temperature varies based on sensor field-of-view, viewing angle and direction, and sun-surface geometry – this directional dependence of urban surface temperature is termed ‘effective thermal anisotropy’ \cite{Voogt1998a}. Traditional satellite or airborne remote sensing platforms, by viewing the surface in the nadir, sample only a fraction of the complete urban surface and fail to capture this effect – leading to directional biases in measured urban surface temperature. Geometric undersampling by a remote sensor in the nadir manifests in an overestimation of daytime temperatures and underestimation of nighttime temperatures \cite{Adderley2015}. However, the magnitude and diurnal pattern of this bias is dependent on urban patch characteristics (canyon height-to-width ratio, canyon orientation and materials, vegetation coverage, etc.) and sensor viewing angle. Thus, parameterization schemes to account for urban effective anisotropy are difficult to generalize across urban sites, sensor types, and sensor orientations.

In addition to undersampling the urban surface, most thermal remote sensing platforms yield an instantaneous ‘snap-shot’ and cannot characterize temporally continuous T\textsubscript{surf} patterns. Temporal discontinuities in thermal remote sensing result in a number of sources of bias operating over a wide range of time scales. Aerial and satellite thermal remote sensing require clear sky conditions (clouds are opaque with respect to thermal infrared radiation). Hence, long term satellite characterizations of sUHI are biased towards conditions that maximize macro-scale urban-rural contrasts in surface temperature. This results in an overestimation of “all-sky” sUHI. Although

Over a day, sUHI is generally largest in the late afternoon, near solar noon, and just after sunset. Continuous sUHI analysis indicates that it exhibits a large diurnal amplitude. Satellite overpass cycles rarely coincide with sUHI maximums and are not standard across cities or platforms. Thus, analysis of sUHI patterns and magnitudes across cities and instrument platforms is difficult.  High-frequency fluctuations in surface temperature add an additional – and understudied – bias to thermal remote sensing of urban surface temperature. Time series analysis of urban surface temperature shows significant microscale (second to minute) fluctuations in temperature \cite{Christen2012}. Most thermal remote sensors observe surface temperatures instantaneously (rather than temporally averaged) and are potentially contaminated by microscale fluctuations. This is particularly salient in urban environments, where a large variety of fabric materials can produce significant directional contrasts in thermal admittance – and thus spatial variations in the magnitude of microscale fluctuations depending on the facet material types viewed by the sensor. The effect of this phenomenon on thermal remote sensing has not been extensively studied, however, the magnitude of microscale fluctuations in surface temperature is significant relative to a typical sUHI signal and thus constitutes a potentially large source of bias.

Both geometric and temporal shortcomings limit the representativity of traditional remote sensed evaluations of urban T\textsubscript{surf} and sUHI. The magnitude of these biases has not been extensively studied in climatological form. This paper presents a method to both assess the magnitude of and overcome these biases.

### 2017-02-01 uw-ethesis-frontpgs.tex
Add files via upload

% T I T L E   P A G E
% -------------------
% Last updated Nov 1, 2016, by Stephen Carr, IST-Client Services
% The title page is counted as page `i' but we need to suppress the
% page number.  We also don't want any headers or footers.
\pagestyle{empty}
\pagenumbering{roman}

% The contents of the title page are specified in the "titlepage"
% environment.
\begin{titlepage}
        \begin{center}
        \vspace*{1.0cm}

        \Huge
        {\bf A Method to Assess the Surface Urban Heat Island Effect Using Hemispherical Radiometric Surface Temperatures}

        \vspace*{1.0cm}

        \normalsize
        by \\

        \vspace*{1.0cm}

        \Large
        Michael A. Allen \\

        \vspace*{3.0cm}

        \normalsize
        A thesis \\
        presented to the University of Western Ontario\\ 
        in fulfillment of the \\
        thesis requirement for the degree of \\
        Master of Science \\

        \vspace*{2.0cm}

        London, Ontario, Canada, 2017 \\

        \vspace*{1.0cm}

        \copyright\ Michael A. Allen 2017 \\
        \end{center}
\end{titlepage}

% The rest of the front pages should contain no headers and be numbered using Roman numerals starting with `ii'
\pagestyle{plain}
\setcounter{page}{2}

\cleardoublepage % Ends the current page and causes all figures and tables that have so far appeared in the input to be printed.
% In a two-sided printing style, it also makes the next page a right-hand (odd-numbered) page, producing a blank page if necessary.

% D E C L A R A T I O N   P A G E
% -------------------------------
  % The following is a sample Delaration Page as provided by the GSO
  % December 13th, 2006.  It is designed for an electronic thesis.
  \noindent
I hereby declare that I am the sole author of this thesis. This is a true copy of the thesis, including any required final revisions, as accepted by my examiners.

  \bigskip

  \noindent
I understand that my thesis may be made electronically available to the public.

\cleardoublepage

% A B S T R A C T
% ---------------

\begin{center}\textbf{Abstract}\end{center}

This is the abstract.

\cleardoublepage

% A C K N O W L E D G E M E N T S
% -------------------------------

\begin{center}\textbf{Acknowledgements}\end{center}

These are my Acknowledgments
\cleardoublepage

% T A B L E   O F   C O N T E N T S
% ---------------------------------
\renewcommand\contentsname{Table of Contents}
\tableofcontents
\cleardoublepage
\phantomsection    % allows hyperref to link to the correct page

% L I S T   O F   T A B L E S
% ---------------------------
\addcontentsline{toc}{chapter}{List of Tables}
\listoftables
\cleardoublepage
\phantomsection		% allows hyperref to link to the correct page

% L I S T   O F   F I G U R E S
% -----------------------------
\addcontentsline{toc}{chapter}{List of Figures}
\listoffigures
\cleardoublepage
\phantomsection		% allows hyperref to link to the correct page

% GLOSSARIES (Lists of definitions, abbreviations, symbols, etc. provided by the glossaries-extra package)
% -----------------------------
\printglossaries
\cleardoublepage
\phantomsection		% allows hyperref to link to the correct page

% Change page numbering back to Arabic numerals
\pagenumbering{arabic}

### 2017-02-01 paper1.tex
Add files via upload

@@
 
-Although most thermal remote sensors operate within one of the \gls{atmwind}[s] --- where atmospheric effects are greatly reduced --- virtually any remote sensed TIR signal is subject to radiative effects from the intervening atmosphere between the surface and the sensor.
+Although most thermal remote sensors operate within one of the \gls{atmwind}[s] --- where atmospheric effects are greatly reduced --- virtually any remote sensed TIR signal is subject to radiative effects from the layer of atmosphere between the surface and the sensor.
 Over much of the thermal infrared \gls{waveband} the atmosphere emits radiation and absorbs a fraction of radiation emitted by the surface.
-In addition, the earth (and indeed the urban surface) reflects a portion of incoming solar radiation.
-Thus, a TIR signal measured from a remote sensing platform is almost certainly not equal to the ground emitted signal.
-These effects can lead to differences between the 'true' radiometric T\textsubscript{surf} and the remote sensed T\textsubscript{surf} of over 10 k for satellite platforms \cite{Cooper1989} and over 6 K for near-ground sensors \cite{Meier2011}.
-Moreover, because atmospheric and emissivity effects are a function of non-uniform and spatiotemporally variant surface and atmospheric properties, their associated errors change depending on instrument type, study location, and ambient conditions.
+Thus, a remote sensed TIR signal almost certainly not equal to the ground emitted signal.
+For an isothermal, homogeneous surface-atmosphere system, at-sensor spectral radiance at height ($z$) can be described by a function deviating from a Planck curve at T\textsubscript{surf} based on the spectral transmittance of the intervening atmosphere, with the magnitude of that deviation governed by the difference between Planck curves at T\textsubscript{surf} and the ambient T\textsubscript{air}.
+As shown for two path geometries in \ref{spectairtsurf} for an atmosphere with water vapor content of  12 g/m\textsuperscript{3} and aerosol and trace gas profiles from the mid-latitude summer standard atmosphere, an at-sensor spectral radiance signal deviates significantly from the surface emitted spectral radiance curve.
+A less transmissive atmosphere increases the potential for deviation in the at-sensor signal from the ground emitted signal, while the difference between T\textsubscript{air} and T\textsubscript{surf} determines the resulting magnitude of atmospheric influence on the spectral TIR signal.
+
+\begin{figure}[!ht]
+	\includegraphics[width=\textwidth]{spectairtsurf}
+	\caption{At-sensor spectral directional radiances computed in MODTRAN 4.1 \cite{Berk1987} for short (30m) and long (3000m) path lengths (z) with Planck curves indicating spectral radiances at T\textsubscript{air} = 305 K and T\textsubscript{surf} = 295 K.}
+	\label{spectairtsurf}
+\end{figure}
+
+Atmospheric effects can lead to differences between the 'true' radiometric T\textsubscript{surf} and the remote sensed T\textsubscript{surf} of over 10 k for satellite platforms \cite{Cooper1989} and over 6 K for near-ground sensors \cite{Meier2011}.
+Moreover, because atmospheric and emissivity effects are a function of non-uniform and spatiotemporally variant surface and atmospheric properties, their associated errors change depending on instrument type, surface-sensor geometry, study location, and ambient conditions.
 As intersite and time-sequential analysis is a significant goal of most thermal remote sensed studies (urban or otherwise) these effects cannot be ignored.
 
-Spectral transmission of longwave radiation through a given layer of atmosphere varies significantly based on wavelength, path geometry, and column absorber content (primarily H\textsubscript{2}0, C0\textsubscript{2}, and to a lesser extent 0\textsubscript{3}, N\textsubscript{2}0, CO, CH\textsubscript{4}, and 0\textsubscript{2} \cite{Miskolczi1993}).
-Variation in band-by-band longwave transmittance with path length is greatest at urban scales (approximately 1 to 50 meters), where transparent spectral bands can quickly become opaque with small changes in path length or absorber content - illustrated in figure \ref{spectransheight}.
-Thus, transmittance of TIR near the surface is highly dependent on surface-sensor-sun geometry, instrument properties, and synoptic conditions.
+Spectral transmission of longwave radiation through a given layer of atmosphere is dependent on total column absorber content (the principal broadband TIR absorbers are H\textsubscript{2}0, C0\textsubscript{2}, and to a lesser extent 0\textsubscript{3}, N\textsubscript{2}0, CO, CH\textsubscript{4}, and 0\textsubscript{2} \cite{Miskolczi1993}).
+Holding vertical absorber content constant, variation in band-by-band TIR transmittance with path length is greatest at urban scales (approximately 1 to 50 meters), where transparent spectral bands can quickly become opaque with small changes in path length or absorber content - illustrated in figure \ref{spectransheight}.
+Thus, transmittance of TIR near the surface is highly dependent on surface-sensor geometry, instrument spectral response, and atmospheric absorber content.
 Indeed, accurate assessment of atmospheric influence on TIR may be most complex when measured near the surface.
@@
 Black and grey shading indicate opaque (0\% transmittance) and transparent (100\% transmittance) bands respectively.
-Redo this figure with wavelength and colorbar?}
+Redo this figure with wavelength and colorbar no totle and the tick marks in back on the outside?}
 	\label{spectransheight}
@@
 
-The radiation passing through a layer of atmosphere from an emitting surface can be described in a number of ways.
+Radiation passing through a layer of atmosphere from an emitting surface can be described in a number of ways.
 \textit{Spectral directional radiance} $ R (\lambda)$, at height $ z $ and from a direction defined by zenith angle \(\theta\), and azimuth angle \(\phi\) is commonly written as
@@
 Methods range from simple mono- \cite{Qin2001} and split-window \cite{Wan1996} routines for single- and multi-channel remote sensors to schemes that integrate a radiative transfer code to isolate the surface emitted signal from interfering signals.
-Boundary conditions are standard across most correction methods: often requiring vertical profiles of T\textsubscript{air}, humidity, pressure, and aerosol content to remove atmospheric effects, and assessments of surface radiative properties to correct for emissivity effects.
-However, correction methods are generally instrument (or at least platform) specific and difficult to generalize across sensor and platform types.
-Few methods exist for correction of ground based remote sensed TIR --- none of which are robust enough to correct irradiances upwelling from rough terrain measured  wide-\gls{fov}(FOV) sensor.
-In part, this is due to the fact that near ground thermal sensors have only recently become sensitive enough to require correction.
+Boundary conditions are standard across most correction methods: generally requiring vertical profiles of T\textsubscript{air}, humidity, pressure, and aerosol content to remove atmospheric effects, and assessments of surface radiative properties to correct for emissivity effects.
+However, correction methods are often instrument (or at least platform) specific and difficult to generalize across sensor and platform types.
+Few methods exist for correction of ground based remote sensed TIR --- none of which are robust enough to correct irradiances upwelling from rough terrain measured via wide-\gls{fov} (FOV) radiometers.
+In part, this is due to the fact that until recently, errors inherent in radiometer measurements were large relative to atmospheric effects.
 However, a new breed of more accurate radiometers and thermal imagers should prompt a critical reevaluation of this assumption.
 
-Atmospheric correction of near-ground remote sensed TIR is subject to a unique set of challenges compared to satellite and aerial platforms.
-Near-ground remote sensors have complex, multiple line-of-sight (LOS) geometry - illustrated in figure 2.1 for a wide-FOV instrument.
-Surface to sensor path lengths vary significantly over the sensor FOV, even when measured close (less than 5m) above the surface.
+Atmospheric correction of near-ground remote sensed TIR is subject to a unique set of challenges compared to traditional satellite and aerial platforms.
+Near-ground and wide-FOV remote sensors have complex, multiple line-of-sight (LOS) geometry - illustrated in figure 2.1 for a downward facing hemispherical radiometer.
+Surface-sensor geometry varies significantly over the sensor FOV, even when measured close (less than 5m) above the surface.
+The addition of 3-dimensional surface geometry further amplifies this effect, as some paths may intersect with raised vertical, sloped, and horizontal features.
 This creates the potential for non-uniform atmospheric effects over the sensor FOV and necessitates a multi-LOS correction to retrieve accurate T\textsubscript{rad}.
-The addition of 3-dimensional surface geometry further amplifies this effect, as some paths may intersect with raised vertical, sloped, and horizontal features.
 In effect, with near-ground wide-FOV sensors, surface geometry is non-trivial and must be represented in atmospheric correction routines.
@@
 Atmospheric correction routines for satellite retrieved TIR, therefore, assume uniform or single-LOS geometry because the TIR signal passes through a relatively constant volume of atmosphere over the projected sensor FOV, regardless of surface geometry.
-This greatly reduces the complexity of correction routines for satellite or aerial thermography.
+This greatly increases the complexity of correction routines for near-ground wide-FOV radiometry.
 
@@
 
-Radiative transfer of broadband longwave radiation in highly dependent on sensor-surface geometry.
-To calculate surface-sensor path lengths, the Surface-Sensor-Sun Urban Model (S\textsuperscript{3}UM) \cite{Soux2004} is initialized with a simplified, orthogonal 3-dimensional DBM representing the area immediately surrounding the sensor.
-In S\textsuperscript{3}UM, urban features (roads, roofs, buildings, etc.) are constructed from an array of horizontal and stacked vertical patches (representing rooftops/roads and walls respectively).
-After indicating sensor FOV and position over the DBM, the model determines which patches have an unobstructed line of sight to the sensor and calculates the path distance from "seen" patches to the sensor.
-Path lengths are binned at 5$^{\circ}$ increments from the nadir and averaged to return an azimuthally averaged path length for each 5$^{\circ}$ bin.
-Using azimuthally averaged path lengths in lieu of a full 3-dimensional radiative transfer model significantly reduces the computational time required to model each irradiance-T\textsubscript{hem} pairing, as at-sensor radiance can be computed independent of azimuth angle.
-Radiative transfer is thus reduced to a two dimensional problem (shown in panel (b) of figure \ref{23radtran}) in which an azimuthally averaged radiance is simulated for each zenith angle.
-The view factors required to weight directional radiances in the hemispherical integrated are computed While generating azimuthally averaged path lengths, S\textsuperscript{3}UM calculates view factors for each solid angle bin used in weighting 2-dimensional angular radiances during the hemispherical integration.
-By using azimuthally averaged path lengths in lieu of  However, because hemispherical TIR sensor returns a bulk irradiance, the sensor signal does not need to be computed for each point of the
+The sheer number of unique path length geometries inherent with wide-FOV radiometry of rough terrain makes full 3-dimensional radiative transfer simulation difficult and computationally intensive --- particularly when correcting a climatology of T\textsubscript{hem}.
+In this method, to improve model efficiency, radiances are calculated for azimuthally averaged path lengths that represent average surface terrain for each solid angle "slice" of the sensor FOV.
+Thus, hemispherical radiative transfer is reduced to a 2-dimensional problem (shown in panel (b) of figure \ref{23radtran}).
+This greatly reduces the computational time required to model each irradiance-T\textsubscript{hem} pairing, as angular radiances can be computed as a function of zenith angle alone and subsequently weighted and integrated 3-dimensionally over the hemisphere.
+
+To calculate surface-sensor geometries, the Surface-Sensor-Sun Urban Model (SUM) \cite{Soux2004} is initialized with a simplified, orthogonal 3-dimensional DBM representing the street canyons and courtyards immediately surrounding the sensor.
+SUM uses a four-dimensional array to represent surface morphology   with three spatial dimensions ($x$, $y$, and $z$), with $z$ representing height above the $x$, $y$ plane.
+SUM also includes an additional fourth dimension, where properties describing each cell are stored (in this case, distance from the point to the sensor).
+After specifying sensor FOV and position relative to the DBM, the model determines which patches have an unobstructed line of sight to the sensor and calculates the distance from "seen" patches to the sensor.
+Path lengths are binned at 5$^{\circ}$ increments of zenith angle and averaged to return an azimuthally-independent mean path length for each bin.
+In addition, while simulating path length geometries, SUM calculates view factors for each solid angle "slice", which are later used to weight 2-dimensional angular radiances in the hemispherical integration post-processing steps.
 
@@
 \subsection{Modeling hemispherical irradiances}
-With path length geometries calculates inS\textsuperscript{3}UM, MODTRAN 4.1 is initialized with 30-minute averaged profiles of T\textsubscript{air} and humidity to model at-sensor spectral directional radiances over a waveband of 0---2500cm\textsuperscript{-1}.
-Aerosol and above canyon meterological conditions are informed by the Mid-Latitude Summer standard atmosphere when dayttime maximum T\textsubscript{air} is greater than 10$^{\circ}$.
-For days with dayttime T\textsubscript{air} maximum below 10$^{\circ}$, the Mid-Latitude Winter standard atmosphere is substituted.
-In spite of significant computational efficiency gains from reducing 3-dimensional radiative transfer to azimuthally averaged 2-dimensions, the method still requires significant computational time to write input cards and model spectral radiances - in this case, with a path length increment of 5 and a LUT of 31 modeled irradiance---T\textsubscript{hem} pairings, each time-step requires 496 independent MODTRAN runs
+With path length geometries calculated in SUM, irradiances are modeled time-sequentially using the MODerate resolution atmospheric TRANsmission 4.1 radiative transfer code (MODTRAN) \cite{Berk1987}.
+At each time-step, a range of potential T\textsubscript{hem} is selected using T\textsubscript{bright} subtracted by some constant (6 K during nighttime runs and 4 K for daytime runs).
+In MODTRAN, at-sensor spectral radiances for each path length are modeled over a waveband of 0---2500cm\textsuperscript{-1} at 0.5 K intervals over the range of potential T\textsubscript{hem}.
+Atmospheric profiles are constructed from 30-minute averages of T\textsubscript{air} and humidity with aerosol and above-sensor conditions are informed by the mid-latitude summer standard atmosphere when daytime maximum T\textsubscript{air} is greater than 10 $^{\circ}$C (the mid-latitude winter profile is substituted on days with T\textsubscript{air, max} of less than 10 $^{\circ}$C).
+Spectral radiances are convolved by a dome transmittance.
 

### 2017-02-02 paper1.tex
Add files via upload

@@
 For an isothermal, homogeneous surface-atmosphere system, at-sensor spectral radiance at height ($z$) can be described by a function deviating from a Planck curve at T\textsubscript{surf} based on the spectral transmittance of the intervening atmosphere, with the magnitude of that deviation governed by the difference between Planck curves at T\textsubscript{surf} and the ambient T\textsubscript{air}.
-As shown for two path geometries in \ref{spectairtsurf} for an atmosphere with water vapor content of  12 g/m\textsuperscript{3} and aerosol and trace gas profiles from the mid-latitude summer standard atmosphere, an at-sensor spectral radiance signal deviates significantly from the surface emitted spectral radiance curve.
+As shown for two path geometries in figure \ref{spectairtsurf} for an isothermal atmosphere with water vapor content of 12 \si{\gram\per\meter\cubed} and aerosol and trace gas profiles from the mid-latitude summer standard atmosphere, an at-sensor spectral radiance signal deviates significantly from the surface emitted spectral radiance curve.
 A less transmissive atmosphere increases the potential for deviation in the at-sensor signal from the ground emitted signal, while the difference between T\textsubscript{air} and T\textsubscript{surf} determines the resulting magnitude of atmospheric influence on the spectral TIR signal.
@@
 	\includegraphics[width=\textwidth]{spectairtsurf}
-	\caption{At-sensor spectral directional radiances computed in MODTRAN 4.1 \cite{Berk1987} for short (30m) and long (3000m) path lengths (z) with Planck curves indicating spectral radiances at T\textsubscript{air} = 305 K and T\textsubscript{surf} = 295 K.}
+	\caption{At-sensor spectral directional radiances computed in MODTRAN 4.1 \cite{Berk1987} for short (30 \si{m}) and long (3000 \si{m}) path lengths (z) with Planck curves indicating spectral radiances at T\textsubscript{air} = 305 \si{K} and T\textsubscript{surf} = 295 \si{K}.}
 	\label{spectairtsurf}
@@
 
-Atmospheric effects can lead to differences between the 'true' radiometric T\textsubscript{surf} and the remote sensed T\textsubscript{surf} of over 10 k for satellite platforms \cite{Cooper1989} and over 6 K for near-ground sensors \cite{Meier2011}.
+Atmospheric effects can lead to differences between the 'true' radiometric T\textsubscript{surf} and the remote sensed T\textsubscript{surf} of over 10 \si{K} for satellite platforms \cite{Cooper1989} and over 6 \si{K} for near-ground sensors \cite{Meier2011}.
 Moreover, because atmospheric and emissivity effects are a function of non-uniform and spatiotemporally variant surface and atmospheric properties, their associated errors change depending on instrument type, surface-sensor geometry, study location, and ambient conditions.
@@
 Spectral transmission of longwave radiation through a given layer of atmosphere is dependent on total column absorber content (the principal broadband TIR absorbers are H\textsubscript{2}0, C0\textsubscript{2}, and to a lesser extent 0\textsubscript{3}, N\textsubscript{2}0, CO, CH\textsubscript{4}, and 0\textsubscript{2} \cite{Miskolczi1993}).
-Holding vertical absorber content constant, variation in band-by-band TIR transmittance with path length is greatest at urban scales (approximately 1 to 50 meters), where transparent spectral bands can quickly become opaque with small changes in path length or absorber content - illustrated in figure \ref{spectransheight}.
+Holding vertical absorber content constant, variation in band-by-band TIR transmittance with path length is greatest at urban scales (approximately 1 to 50 \si{m}), where transparent spectral bands can quickly become opaque with small changes in path length or absorber content - illustrated in figure \ref{spectransheight}.
 Thus, transmittance of TIR near the surface is highly dependent on surface-sensor geometry, instrument spectral response, and atmospheric absorber content.
@@
 
-\noindent where $ \sigma $ is the Stefan-Boltzmann constant and $ \overline{r} $ is Planck weighted average broadband sensor response computed as,
+\noindent where $ \sigma $ is the Stefan-Boltzmann constant and $ \overline{r} $ is Planck weighted mean broadband sensor response computed as,
 
@@
 In addition, a simple approximation of $T_{bright}$ can be inferred from directional radiance using equation \ref{stefb1} by replacing $ L_z $ with $L'_z$ multiplied by a constant.
-This method is commonly used to infer T\textsubscript{bright} from infrared thermometers (IRT) operating over the atmospheric window (where atmospheric correction magnitudes are small and T\textsubscript{bright} is a reasonably accurate approximation of T\textsubscript{surf}).
-However, constants must be calibrated for the range of expected T\textsubscript{surf} as the relationship between $ L_z $ and $L'_z$  is not perfectly linear.
+This method is commonly used to infer T\textsubscript{bright} from infrared thermometers (IRT) operating over the atmospheric window, where atmospheric effects are minimal and T\textsubscript{bright} is a reasonably accurate approximation of T\textsubscript{surf}.
+However, constants must be calibrated for the range of expected T\textsubscript{surf} as the relationship between $ L_z $ and $L'_z$  is not perfectly linear with respect to emitter temperature.
 
-Inversions of uncorrected directional radiances or irradiances using the Planck function or the Stefan-Boltzmann law yield a temperature equal to that of a blackbody emitting the same amount of radiation as detected by the sensor.
-Since $ L_z $ is unlikely to be equal to $ L_0 $, T\textsubscript{bright, 0} and T\textsubscript{bright, z} often show significant variation based on sensor characteristics, ambient conditions, and surface-sensor geometry.
-Hence, remote sensed T\textsubscript{bright} is generally considered only a rough approximation representation of kinetic or thermodynamic T\textsubscript{surf}.
+Inversions of uncorrected $ L_z $ or $L'_z$ yield a temperature equal to that of a blackbody emitting the same amount of radiation as detected by the sensor.
+Since $ L_z $ ($L'_z$ ) is unlikely to be equal to $ L_0 $ ($L'_0$), T\textsubscript{bright} at $ z = 0 $ and T\textsubscript{bright} at $ z $ often show significant deviation based on sensor characteristics, ambient conditions, and surface-sensor geometry.
+Hence, remote sensed T\textsubscript{bright} is generally considered only a rough approximation representation of radiometric, kinetic, or thermodynamic T\textsubscript{surf}.
 
 To retrieve a more accurate estimation of the 'true' T\textsubscript{surf}, the same inversions can be applied to TIR measurements after correction for atmospheric and surface emissivity effects (e.g.
-transformations of the TIR signal to represent that at $ z = 0 $ if the surface was emitting as a perfect blackbody) to yield a direction radiometric surface temperature T\textsubscript{rad} for atmospheric and emissivity corrected directional radiances, a hemispherical radiometric surface temperature T\textsubscript{hem} for atmospheric and emissivity corrected irradiances.
-$ T\textsubscript{rad} (\theta, \phi) $ and T\textsubscript{hem} provide a better approximation of the ‘true’ T\textsubscript{surf} by representing the temperature at which the surface is radiating integrated over the sensor FOV.
+modification of the remote sensed TIR signal to represent that at $ z = 0 $ of a homogeneous blackbody emitter) to yield a direction radiometric surface temperature  T\textsubscript{rad} for atmospheric and emissivity corrected directional radiances, a hemispherical radiometric surface temperature T\textsubscript{hem} for atmospheric and emissivity corrected irradiances.
+T\textsubscript{rad} and T\textsubscript{hem} provide a better approximation of the ‘true’ T\textsubscript{surf} by representing the temperature at which the viewed surface is radiating, integrated over the sensor FOV.
 
@@
 
-To calculate surface-sensor geometries, the Surface-Sensor-Sun Urban Model (SUM) \cite{Soux2004} is initialized with a simplified, orthogonal 3-dimensional DBM representing the street canyons and courtyards immediately surrounding the sensor.
+To calculate surface-sensor geometries, the Surface-Sensor-Sun Urban Model (SUM) \cite{Soux2004} is initialized with a simplified, orthogonal 3-dimensional DBM representing geometry of the area surrounding the sensor.
 SUM uses a four-dimensional array to represent surface morphology   with three spatial dimensions ($x$, $y$, and $z$), with $z$ representing height above the $x$, $y$ plane.
-SUM also includes an additional fourth dimension, where properties describing each cell are stored (in this case, distance from the point to the sensor).
-After specifying sensor FOV and position relative to the DBM, the model determines which patches have an unobstructed line of sight to the sensor and calculates the distance from "seen" patches to the sensor.
+An additional fourth dimension is used to store information describing each cell (in this case, distance from the point to the sensor).
+After specifying sensor position and FOV, the model determines which patches have an unobstructed line of sight to the sensor and calculates distance from "seen" patches to the sensor.
 Path lengths are binned at 5$^{\circ}$ increments of zenith angle and averaged to return an azimuthally-independent mean path length for each bin.
@@
 In 3-dimensions, path length for a given zenith angle can change significantly with azimuth angle.
-Dotted lines indicate the potential for absorption by the intervening atmospheric layer}
+Dotted lines indicate absorption by the intervening atmospheric layer.}
 \end{figure}
@@
 \subsection{Modeling hemispherical irradiances}
-With path length geometries calculated in SUM, irradiances are modeled time-sequentially using the MODerate resolution atmospheric TRANsmission 4.1 radiative transfer code (MODTRAN) \cite{Berk1987}.
-At each time-step, a range of potential T\textsubscript{hem} is selected using T\textsubscript{bright} subtracted by some constant (6 K during nighttime runs and 4 K for daytime runs).
-In MODTRAN, at-sensor spectral radiances for each path length are modeled over a waveband of 0---2500cm\textsuperscript{-1} at 0.5 K intervals over the range of potential T\textsubscript{hem}.
-Atmospheric profiles are constructed from 30-minute averages of T\textsubscript{air} and humidity with aerosol and above-sensor conditions are informed by the mid-latitude summer standard atmosphere when daytime maximum T\textsubscript{air} is greater than 10 $^{\circ}$C (the mid-latitude winter profile is substituted on days with T\textsubscript{air, max} of less than 10 $^{\circ}$C).
-Spectral radiances are convolved by a dome transmittance.
+With path length geometries calculated in SUM, irradiances are modeled time-sequentially using version 4.1 of the MODerate resolution atmospheric TRANsmission radiative transfer code (MODTRAN) \cite{Berk1987}.
+At each time-step, T\textsubscript{bright} inferred from the measured irradiance and subtracted by a constant (6 \si{K} during nighttime runs and 4 \si{K} for daytime runs) is used to select a range of potential T\textsubscript{hem} which forms the bounds of the LUT for the target time-step.
+At-sensor spectral radiances for each path length are then modeled over a waveband of 0---2500 \si{cm^{-1}} at 0.5 \si{K} intervals over the LUT range.
+Atmospheric profiles are retrieved from 30-minute averages of T\textsubscript{air} and humidity, with aerosol and above-sensor conditions are informed by the mid-latitude summer standard atmosphere when daytime maximum T\textsubscript{air} is greater than 10 \si{\degreeCelsius} (the mid-latitude winter profile is substituted on days with maximum T\textsubscript{air} of less than 10 \si{\degreeCelsius}) \cite{Kantor1962}.
 
-\section{Evaluation of the method using profiles of upwelling longwave radiation over a homogenous planar surface}
+In this method, a "typical" longwave bandpass (approximately 4---42 \si{\micro\meter}) was extended to include much longer wavelengths.
+This was done for two reasons: 1.
+To accurately represent a longwave curve for typical urban T\textsubscript{surf}, which have significant emittance into wavelengths much longer than 42 \si{\micro\meter}.
+2.
+To replicate the signal from a silicone-domed pyrgeometer, which is transmissive of radiation in wavelengths.
+A "typical" longwave bandpass underestimates a pyrgeometer signal by approximately 5---9 \si{\watt\per\square\meter}, depending on emitter temperature.
+Figure \ref{dometrans} comparison of ‘typical’ and ‘extended’ longwave bandpasses.
+
+\begin{figure}[!ht]
+	\includegraphics[width=\textwidth]{dometrans}
+	\label{dometrans}
+	\caption{ddddd}
+\end{figure}
+
+\section{Evaluation of the method using profiles of upwelling longwave radiation over a homogeneous planar surface}
 

### 2017-02-03 paper1.tex
Add files via upload

@@
 
-In this method, a "typical" longwave bandpass (approximately 4---42 \si{\micro\meter}) was extended to include much longer wavelengths.
+In this method, a "typical" longwave bandpass, approximately 250 - 2215 \si{cm^{-1}} (4 - 42 \si{\micro\meter}), was extended to include much shorter wavenumber (longer wavelengths).
 This was done for two reasons: 1.
-To accurately represent a longwave curve for typical urban T\textsubscript{surf}, which have significant emittance into wavelengths much longer than 42 \si{\micro\meter}.
+To accurately represent a longwave curve for typical urban T\textsubscript{surf}, which have significant emittance into wavenumber shorter than 250 \si{cm^{-1}} (wavelengths much longer than 42 \si{\micro\meter}).
 2.

### 2017-04-11 paper1.tex
Add files via upload

--- 2017-02-02 paper1.tex
@@
-\chapter{A method to correct upwelling longwave radiation to estimate hemispherical urban surface temperature}
+\chapter{A method to correct upwelling longwave radiation to estimate hemispherical urban surface temperature}\label{paper1}
 
 \section{Introduction}
-Thermal infrared (TIR) remote sensing of land surface temperature (T\textsubscript{surf}) has emerged as a primary research focus in climatology, as researchers seek to better describe spatiotemporal patterns of T\textsubscript{surf} globally and better understand how anthropogenic modification of earth's surface influences land T\textsubscript{surf} with links to the climate at-large.
-Over the last two decades, use of thermal remote sensing of surface climates has expanded significantly --- both in terms of the volume and breadth of remote sensed study and its explicative importance in climatology as a discipline.
+Thermal infrared (TIR) remote sensing of land surface temperature (T\textsubscript{surf}) has emerged as a primary research focus in climatology, as researchers seek to better describe spatiotemporal patterns of T\textsubscript{surf} globally and better understand how anthropogenic modification of earth's surface influences land T\textsubscript{surf} and impacts climate at various scales.
+Over the last two decades, the application of thermal remote sensing to study surface climates has expanded significantly.
 Thermal remote sensing of earth's surface has applications over a wide range of disciplines: from informing micro-, urban-, and global-scale climate models, to aiding decision making and mitigation praxis with respect to climate change and the \gls{uhi}.
@@
 Within urban climatology, a combination of satellite, aerial, and ground-based thermal remote sensors have been integral in elucidating the spatial \cite{Roth1989}, temporal \cite{Peng2012}, and geometric \cite{Voogt1997} effects of the built environment on land T\textsubscript{surf}; in evaluating and partitioning urban surface energy balances \cite{BastiaanssenW.G.M.1998, Yamaguchi2005} and; in characterizations of the relationship between surface and \gls{blayer} air temperatures (T\textsubscript{air}) \cite{Stoll1992}.
-These advances have been aided by substantial improvements in sensor ground, spectral, and radiometric resolutions, and by the proliferation of both large-scale public satellite remote sensing campaigns and low-cost aerial and near-ground thermography.
+These advances have been aided by substantial improvements in sensor spatial, spectral, and radiometric resolutions, and by the proliferation of both large-scale public satellite remote sensing campaigns and low-cost aerial and near-ground thermography.
 However, in spite of it's widespread usage, several questions concerning the use and validity of urban remote thermal remote sensing, first posed in Roth et al.
@@
 
-\noindent In this paper, we seek to examine question two by introducing and evaluating a method for atmospheric and emissivity correction of near-ground hemispherical TIR --- measured via \gls{pyrg} --- for hemispherical radiometric temperature (T\textsubscript{hem}) retrieval.
+\noindent In this paper, we seek to examine question two by introducing and evaluating a method for atmospheric correction of near-ground hemispherical TIR --- measured via \gls{pyrg} --- for hemispherical radiometric temperature (T\textsubscript{hem, r}) retrieval.
 These measures are common to most urban energy balance assessments and thus constitute a hitherto untapped method for urban T\textsubscript{surf} analysis.
-A companion paper responds to question one through an analysis of a climatology of T\textsubscript{hem} and derived surface UHI (sUHI), to quantify geometric and temporal biases across multiple methods for remote sensing of urban T\textsubscript{surf}.
+A companion paper responds to question one through an analysis of an 8-month climatology of T\textsubscript{hem, r} and derived surface UHI magnitudes (sUHI), to quantify geometric and temporal biases across multiple methods for remote sensing of urban T\textsubscript{surf}.
 
@@
 
-Although most thermal remote sensors operate within one of the \gls{atmwind}[s] --- where atmospheric effects are greatly reduced --- virtually any remote sensed TIR signal is subject to radiative effects from the layer of atmosphere between the surface and the sensor.
+Although most thermal remote sensors operate within one of the \gls{atmwind}[s] - where atmospheric effects are minimal - virtually any remote sensed TIR signal is subject to radiative effects from the layer of atmosphere between the surface and the sensor.
 Over much of the thermal infrared \gls{waveband} the atmosphere emits radiation and absorbs a fraction of radiation emitted by the surface.
 Thus, a remote sensed TIR signal almost certainly not equal to the ground emitted signal.
-For an isothermal, homogeneous surface-atmosphere system, at-sensor spectral radiance at height ($z$) can be described by a function deviating from a Planck curve at T\textsubscript{surf} based on the spectral transmittance of the intervening atmosphere, with the magnitude of that deviation governed by the difference between Planck curves at T\textsubscript{surf} and the ambient T\textsubscript{air}.
-As shown for two path geometries in figure \ref{spectairtsurf} for an isothermal atmosphere with water vapor content of 12 \si{\gram\per\meter\cubed} and aerosol and trace gas profiles from the mid-latitude summer standard atmosphere, an at-sensor spectral radiance signal deviates significantly from the surface emitted spectral radiance curve.
-A less transmissive atmosphere increases the potential for deviation in the at-sensor signal from the ground emitted signal, while the difference between T\textsubscript{air} and T\textsubscript{surf} determines the resulting magnitude of atmospheric influence on the spectral TIR signal.
+Spectral radiance received by a sensor at height ($z$) can be described by a function deviating from a Planck curve defined by T\textsubscript{surf} based on the spectral transmittance of the intervening atmosphere, with the magnitude of that deviation determined by the difference between Planck curves calculated from T\textsubscript{surf} and ambient T\textsubscript{air}.
+A radiative transfer code, such as MODTRAN \cite{Berk1987}, can be used to model the effect of the atmosphere on a remote sensed TIR signal.
+Radiative transfer simulations run in MODTRAN version 4.1 for two path lengths are shown in figure \ref{spectairtsurf}, for spectral radiance emitted from a planar surface at 300 \si{kelvin} through an isothermal atmosphere at 290 \si{\kelvin}, water vapor content of 12 \si{\gram\per\meter\cubed}, and aerosol and trace gas profiles from the mid-latitude summer standard atmosphere \cite{Kantor1962}.
+At both short and long path lengths, radiance in each band shows some deviation from the ground emitted signal.
+Larger deviations are observed at longer path lengths, where a significant portion of the resulting signal is emitted from that atmosphere, even in bands with high spectral transmittance.
+Long path lengths and a less transmissive atmosphere increase the potential for deviation in a remote sensed TIR signal from the ground emitted signal, while the difference between T\textsubscript{air} and T\textsubscript{surf} determines the resulting magnitude of atmospheric influence.
 
-\begin{figure}[!ht]
+\begin{figure}[H]
 	\includegraphics[width=\textwidth]{spectairtsurf}
-	\caption{At-sensor spectral directional radiances computed in MODTRAN 4.1 \cite{Berk1987} for short (30 \si{m}) and long (3000 \si{m}) path lengths (z) with Planck curves indicating spectral radiances at T\textsubscript{air} = 305 \si{K} and T\textsubscript{surf} = 295 \si{K}.}
+	\caption{At-sensor spectral directional radiances computed in MODTRAN 4.1 \cite{Berk1987} for short (30 \si{m}) and long (3000 \si{m}) path lengths (z) with Planck curves indicating spectral radiances at T\textsubscript{surf} = 300 \si{K} and T\textsubscript{air} = 290 \si{K}.}
 	\label{spectairtsurf}
@@
 Atmospheric effects can lead to differences between the 'true' radiometric T\textsubscript{surf} and the remote sensed T\textsubscript{surf} of over 10 \si{K} for satellite platforms \cite{Cooper1989} and over 6 \si{K} for near-ground sensors \cite{Meier2011}.
-Moreover, because atmospheric and emissivity effects are a function of non-uniform and spatiotemporally variant surface and atmospheric properties, their associated errors change depending on instrument type, surface-sensor geometry, study location, and ambient conditions.
-As intersite and time-sequential analysis is a significant goal of most thermal remote sensed studies (urban or otherwise) these effects cannot be ignored.
+Moreover, because atmospheric effects are a function of non-uniform and spatiotemporally variant surface and atmospheric properties, their associated errors change depending on instrument type, surface-sensor geometry, study location, and ambient conditions.
+As intersite and time-continuous analysis is a significant goal of most thermal remote sensed studies (urban or otherwise) these effects cannot be ignored.
 
-Spectral transmission of longwave radiation through a given layer of atmosphere is dependent on total column absorber content (the principal broadband TIR absorbers are H\textsubscript{2}0, C0\textsubscript{2}, and to a lesser extent 0\textsubscript{3}, N\textsubscript{2}0, CO, CH\textsubscript{4}, and 0\textsubscript{2} \cite{Miskolczi1993}).
-Holding vertical absorber content constant, variation in band-by-band TIR transmittance with path length is greatest at urban scales (approximately 1 to 50 \si{m}), where transparent spectral bands can quickly become opaque with small changes in path length or absorber content - illustrated in figure \ref{spectransheight}.
+Spectral transmission of longwave radiation through a given layer of atmosphere is dependent on total column absorber content (the principal broadband TIR absorbers are H\textsubscript{2}O, CO\textsubscript{2}, and to a lesser extent O\textsubscript{3}, N\textsubscript{2}O, CO, CH\textsubscript{4}, and O\textsubscript{2} \cite{Miskolczi1993}).
+Holding vertical absorber content constant through a layer of atmosphere, variation in band-by-band TIR transmittance with path length is greatest at urban scales (approximately 1 to 50 \si{\meter}), where transparent spectral bands can quickly become opaque with small changes in path length or absorber content - illustrated in figure \ref{spectransheight}.
 Thus, transmittance of TIR near the surface is highly dependent on surface-sensor geometry, instrument spectral response, and atmospheric absorber content.
@@
 	\caption{Spectral transmission of water vapor as a function of height.
-Black and grey shading indicate opaque (0\% transmittance) and transparent (100\% transmittance) bands respectively.
-Redo this figure with wavelength and colorbar no totle and the tick marks in back on the outside?}
+Run in MODTRAN 4.1 for radiance emitted from a planar surface at 300 \si{\kelvin} through an atmosphere with water vapor content of 8 \si{\gram\per\meter\cubed}.
+Sampled at heights of 1, 5, 10, 15, 20, 25, and 30 \si{\meter}.}
 	\label{spectransheight}
@@
 
-Radiation passing through a layer of atmosphere from an emitting surface can be described in a number of ways.
-\textit{Spectral directional radiance} $ R (\lambda)$, at height $ z $ and from a direction defined by zenith angle \(\theta\), and azimuth angle \(\phi\) is commonly written as
+\subsection{Describing radiation as received by a remote sensor}
+
+Radiation passing through a layer of atmosphere from an emitting surface towards a sensor can be described in a number of ways: \textit{Spectral directional radiance} $R^\uparrow_z (\lambda, \theta, \phi)$, at height $ z $, wavelength $\lambda$ and from a direction defined by viewingzenith angle \(\theta\), and azimuth angle \(\phi\) is commonly written as
 
 \begin{equation}
-R^\uparrow_z (\lambda, \theta, \phi) = \tau_\lambda \epsilon R^\uparrow_0(\lambda) + (1-\epsilon) R^\downarrow_{sky} (\lambda) + (1-\tau_\lambda) R^\uparrow_{atm}(\lambda)
+R^\uparrow_z (\lambda, \theta, \phi) = \tau_\lambda \epsilon_\lambda R^\uparrow_0(\lambda) + (1-\epsilon_\lambda) R^\downarrow_{sky} (\lambda) + (1-\tau_\lambda) R^\uparrow_{atm}(\lambda)
 \end{equation}
 
-\noindent where \(\epsilon\) is surface emissivity, \(\tau\)\textsubscript{\( \lambda \)} is spectral "slab" transmittance through the layer between the emitting surface $( z  = 0) $ and $ z $.
-The same spectral TIR signal, as measured by a narrow-FOV sensor mounted at height $ z $, passes through an instrument shield (or dome) with spectral transmittance \(\tau_d\), and is integrated over the sensor waveband (bounded by \(\lambda_1\) and \(\lambda_2\)) to yield a \textit{directional radiance} $L'$ as 'seen' by the sensor
+\noindent where \(\epsilon\)\textsubscript{\( \lambda \)} is spectral surface emissivity, \(\tau\)\textsubscript{\( \lambda \)} is spectral "slab" transmittance through the layer between the emitting surface $( z  = 0) $ and $ z $.
+Directional radiances upwelling from the atmosphere $R^\uparrow_{atm}$, and the surface $R^\uparrow_0$, and downwelling from the sky $R^\downarrow_{sky}$, can be described spectrally or integrated over a waveband bounded by \(\lambda_1\) and \(\lambda_2\) as
 
 \begin{equation}
-L'_z (\theta, \phi) = \int_{\lambda_1}^{\lambda_2} \tau_d(\lambda) R^\uparrow_z(\lambda) ~ d\lambda
+R(\theta, \phi) = \bigints_{\lambda_1}^{\lambda_2} R(\lambda,\theta, \phi) ~ d\lambda =\displaystyle\frac{\epsilon_\lambda C_1}{\pi \lambda^5 \left( \exp\left(\cfrac{C_2}{\lambda T}\right)\right)}
+\end{equation}
+
+\noindent where C\textsubscript{1} = $ 3.7404 \cdot 10^8 $ $ W\mu^4 m^{-2} $, C\textsubscript{2} = $ 14387 \mu K $, and $T$ is emitter temperature.
+
+Measured by a narrow-FOV sensor mounted at height $ z $, $R^\uparrow_z (\lambda, \theta, \phi)$ passes through an instrument filter (or dome) with spectral transmittance \(\tau_d\), and is integrated over the sensor waveband to yield a \textit{directional radiance} $L_z'$ as 'seen' by the sensor
+
+\begin{equation}
+L_z' (\theta, \phi) = \int_{\lambda_1}^{\lambda_2} \tau_d(\lambda) R^\uparrow_z(\lambda) ~ d\lambda
 \end{equation}
@@
 
-TIR received by a remote sensor can be related to the emitting body's temperature in a number of ways --- each producing different conceptions of T\textsubscript{surf} from different instrument and sensor types.
-As such, the term "surface temperature" with respect to a remote sensed TIR is vague and can refer to several definitions of "surface" and "temperature" Thus, proper terminology must be attached to land T\textsubscript{surf} inferred from TIR.
-Definitions and nomenclature conventions for multiple methods for T\textsubscript{surf} retrieval are discussed at length in Norman \& Becker, (1995)\cite{Norman1995}.
+TIR received by a remote sensor can be related to a surface temperature in a number of ways --- each producing different conceptions of T\textsubscript{surf} from different instrument and sensor types.
+As such, the term "surface temperature" with respect to a remote sensed TIR is vague and can refer to several definitions of "surface" and "temperature".
+Thus, proper terminology must be attached to land T\textsubscript{surf} inferred from TIR.
+Definitions and nomenclature conventions for multiple methods for T\textsubscript{surf} retrieval are discussed at length in Norman \& Becker, (1995) \cite{Norman1995}.
 
-Directional radiance $ L'_z $ detected from a narrow-FOV sensor operating over waveband (\(\lambda_1\) --- \(\lambda_2\)) and viewing the surface from some orientation described by \(\theta\), \(\phi\) can be used to infer a \textit{directional brightness temperature} T'\textsubscript{bright}(\(\theta\), \(\phi\)) via an inversion of the Planck function multiplied by normalized sensor response integrated over the sensor waveband,
-
-\begin{equation}
-L'_z(T'_{bright} (\theta, \phi)) = \bigints_{\lambda_1}^{\lambda_2} \cfrac{f(\lambda) C_1}{\pi \lambda^5 \left( \exp\left(\cfrac{C_2}{\lambda T'_{bright} (\theta, \phi)}\right)\right)}
-\end{equation}
-
-\noindent where C\textsubscript{1} = $ 3.7404 \cdot 10^8 $ $ W\mu^4 m^{-2} $, C\textsubscript{2} = $ 14387 \mu K $, and relative sensor response $ r(\lambda) $ normalized via,
-
-\begin{equation}
-1 ~ =  \int_{\lambda_1}^{\lambda_2} r(\lambda) ~ d\lambda
-\end{equation}
-
-Similarly irradiance $L_z$ received by a broadband hemispherical sensor (such as a pyrgeometer), can be used to infer a \textit{hemispherical brightness temperature} T\textsubscript{bright} through an inversion of the Stefan-Boltzmann law,
+ Irradiance $L_z$ received by a broadband hemispherical sensor (such as a pyrgeometer), can be used to infer a \textit{hemispherical brightness temperature} T\textsubscript{hem, b} through an inversion of the Stefan-Boltzmann law,
 
@@
 \label{stefb1}
-L_z = \overline{r} (\sigma T_{bright}^4)
+T_{hem,~ b} = \sqrt[4]{\frac{L_z}{\overline{r}\sigma}}
 \end{equation}
@@
 \begin{equation}
-\overline{r} = \cfrac{\int R(\lambda) ~ r(\lambda)}{\int R(\lambda)}
+\label{meansens}
+\overline{r} = \cfrac{\int R(\lambda) ~ r(\lambda)~d\lambda}{\int R(\lambda) ~ d\lambda}
 \end{equation}
 
-\noindent with $ R(\lambda) $ computed from a Planck function at an approximated $T_{bright}$.
+\noindent with $ R(\lambda) $ computed from a Planck function at an approximated emitter temperature.
 
-In addition, a simple approximation of $T_{bright}$ can be inferred from directional radiance using equation \ref{stefb1} by replacing $ L_z $ with $L'_z$ multiplied by a constant.
-This method is commonly used to infer T\textsubscript{bright} from infrared thermometers (IRT) operating over the atmospheric window, where atmospheric effects are minimal and T\textsubscript{bright} is a reasonably accurate approximation of T\textsubscript{surf}.
-However, constants must be calibrated for the range of expected T\textsubscript{surf} as the relationship between $ L_z $ and $L'_z$  is not perfectly linear with respect to emitter temperature.
+A \textit{directional brightness temperature} T\textsubscript{bright} $(\theta, \phi)$ from some viewing able described $\theta$ and $\phi$ by can be inferred from directional radiance via equation \ref{stefb1} by replacing $ L_z $ with $L_z'$ multiplied by a constant.
+This method is commonly used to infer T\textsubscript{bright} $(\theta, \phi)$ from infrared thermometers (IRT) operating over the atmospheric window - where atmospheric effects are minimal and T\textsubscript{bright} is a reasonably accurate approximation of T\textsubscript{surf}.
+However, constants must be calibrated for the range of expected T\textsubscript{surf} as the relationship between $ L_z $ and $L_z'$  is not perfectly linear with respect to emitter temperature.
 
-Inversions of uncorrected $ L_z $ or $L'_z$ yield a temperature equal to that of a blackbody emitting the same amount of radiation as detected by the sensor.
-Since $ L_z $ ($L'_z$ ) is unlikely to be equal to $ L_0 $ ($L'_0$), T\textsubscript{bright} at $ z = 0 $ and T\textsubscript{bright} at $ z $ often show significant deviation based on sensor characteristics, ambient conditions, and surface-sensor geometry.
-Hence, remote sensed T\textsubscript{bright} is generally considered only a rough approximation representation of radiometric, kinetic, or thermodynamic T\textsubscript{surf}.
+Inversions of uncorrected $ L_z $ or \textbf{$L_z'$} yield a temperature equal to that of a blackbody emitting the same amount of radiation as detected by the sensor.
+Since $ L_z $ is unlikely to be equal to $ L_0 $ and, by extension,  $L_z'$ is unlikely to be equal to $L_0'$, T\textsubscript{hem, b} at $ z = 0 $ and T\textsubscript{hem, b} at $ z $ often show significant deviation based on sensor characteristics, ambient conditions, and surface-sensor geometry.
+Hence, T\textsubscript{bright} and T\textsubscript{hem, b} are generally considered only a rough approximation representation of radiometric, kinetic, or thermodynamic T\textsubscript{surf}.
 
-To retrieve a more accurate estimation of the 'true' T\textsubscript{surf}, the same inversions can be applied to TIR measurements after correction for atmospheric and surface emissivity effects (e.g.
-modification of the remote sensed TIR signal to represent that at $ z = 0 $ of a homogeneous blackbody emitter) to yield a direction radiometric surface temperature  T\textsubscript{rad} for atmospheric and emissivity corrected directional radiances, a hemispherical radiometric surface temperature T\textsubscript{hem} for atmospheric and emissivity corrected irradiances.
-T\textsubscript{rad} and T\textsubscript{hem} provide a better approximation of the ‘true’ T\textsubscript{surf} by representing the temperature at which the viewed surface is radiating, integrated over the sensor FOV.
+To retrieve a more accurate estimation of the 'true' T\textsubscript{surf}, the same inversions can be applied to TIR measurements after correction for atmospheric effects (e.g.
+modification of the remote sensed TIR signal to represent the same signal at $ z = 0 $ emitted from a homogeneous, isothermal, blackbody emitter) to yield a \textit{directional radiometric surface temperature} T\textsubscript{rad} from atmospheric corrected directional radiances and a \textit{hemispherical radiometric surface temperature} T\textsubscript{hem, r} from atmospherically corrected irradiances.
+T\textsubscript{rad} and T\textsubscript{hem, r} provide a better approximation of the ‘true’ T\textsubscript{surf} by representing the temperature at which emitting surfaces are radiating, integrated over the sensor FOV.
 
@@
 A large number of correction routines have been developed to remove atmospheric and emissivity effects from aerial and satellite TIR signals and derive accurate T\textsubscript{rad}.
-Methods range from simple mono- \cite{Qin2001} and split-window \cite{Wan1996} routines for single- and multi-channel remote sensors to schemes that integrate a radiative transfer code to isolate the surface emitted signal from interfering signals.
-Boundary conditions are standard across most correction methods: generally requiring vertical profiles of T\textsubscript{air}, humidity, pressure, and aerosol content to remove atmospheric effects, and assessments of surface radiative properties to correct for emissivity effects.
-However, correction methods are often instrument (or at least platform) specific and difficult to generalize across sensor and platform types.
-Few methods exist for correction of ground based remote sensed TIR --- none of which are robust enough to correct irradiances upwelling from rough terrain measured via wide-\gls{fov} (FOV) radiometers.
-In part, this is due to the fact that until recently, errors inherent in radiometer measurements were large relative to atmospheric effects.
+Methods range from simple mono- \cite{Qin2001} and split-window \cite{Wan1996} routines for single- and multi-channel remote sensors, to schemes that integrate a radiative transfer code to isolate the surface emitted signal from interfering signals.
+Boundary conditions are standard across most correction methods: generally requiring vertical profiles of T\textsubscript{air}, humidity, pressure, and aerosol content to remove atmospheric effects, and surface radiative properties to correct for emissivity effects.
+However, correction methods are often instrument (or, at the very least, least platform) specific and difficult to generalize across sensor and platform types.
+Few methods exist for correction of ground based remote sensed TIR --- none of which are robust enough to correct irradiances measured from a near-ground wide-\gls{fov} (FOV) radiometer mounted to view rough terrain.
+
+%In part, this is due to the fact that until recently, errors inherent in radiometer measurements were large relative to atmospheric effects.
 However, a new breed of more accurate radiometers and thermal imagers should prompt a critical reevaluation of this assumption.
@@
 Atmospheric correction of near-ground remote sensed TIR is subject to a unique set of challenges compared to traditional satellite and aerial platforms.
-Near-ground and wide-FOV remote sensors have complex, multiple line-of-sight (LOS) geometry - illustrated in figure 2.1 for a downward facing hemispherical radiometer.
-Surface-sensor geometry varies significantly over the sensor FOV, even when measured close (less than 5m) above the surface.
-The addition of 3-dimensional surface geometry further amplifies this effect, as some paths may intersect with raised vertical, sloped, and horizontal features.
-This creates the potential for non-uniform atmospheric effects over the sensor FOV and necessitates a multi-LOS correction to retrieve accurate T\textsubscript{rad}.
+Wide-FOV remote sensors have complex, multiple line-of-sight (LOS) path length geometries - illustrated in Figure \ref{lineofsight} for a downward facing pyrgeometer.
+Surface-sensor geometry varies significantly over the sensor FOV as some path lengths intersect with raised vertical, sloped, and horizontal features.
+This creates the potential for non-uniform atmospheric effects over the sensor FOV and necessitates a multi-LOS correction to retrieve accurate T\textsubscript{hem, r}.
 In effect, with near-ground wide-FOV sensors, surface geometry is non-trivial and must be represented in atmospheric correction routines.
@@
 Atmospheric correction routines for satellite retrieved TIR, therefore, assume uniform or single-LOS geometry because the TIR signal passes through a relatively constant volume of atmosphere over the projected sensor FOV, regardless of surface geometry.
-This greatly increases the complexity of correction routines for near-ground wide-FOV radiometry.
 
-\begin{figure}[!ht]
+\begin{figure}[H]
 	\includegraphics[width=\textwidth]{lineofsight}
+	\caption{Variable path geometry inherent with wide-FOV near-ground sensor visualized over an idealized 2-dimensional urban area.}
 	\label{lineofsight}
-	\caption{Variable path geometry inherent with wide-FOV near-ground sensor visualized over an idealized 2-dimensional urban area.}
 \end{figure}
 
-Several multi-LOS atmospheric correction routines have been developed for near-ground TIR: Meier et al., 2011 describes a correction method for oblique angled urban thermography \cite{Meier2011}.
-Path lengths were calculated for each pixel over the FOV of a tower-mounted thermal imager angled obliquely toward the urban surface.
-A correction factor was then computed and applied using a radiative transfer code initialized using isothermal, isohumal atmospheric profiles for each time step.
-Applied over a time-series of images, the result is a continuous atmospherically corrected series of thermal images with each pixel representing a different T\textsubscript{rad}.
-However, the method uses visual-band images and a GIS database to represent urban geometry for each pixel's LOS - a technique not possible with a pyrgeometer, which returns a single integrated irradiance at each time-step.
-Moreover, the target instrument operates over a narrow waveband, reducing the magnitude and variance in atmospheric transmission over the sensor response curve.
-Thus, the method is not directly generalizable to correction TIR as measured via pyrgeometer.
-Kotani \& Sugita, 2009 describes a method for correction of wide-FOV (pyrgeometer) TIR Irradiances \cite{Kotani2009a} over planar terrain.
-However, these methods are limited to narrow-FOV thermal imagers and sensors over planar surfaces respectively.
-A method which combines Meier et al., 2009's representation of complex surface geometry and Kotani \& Sugita, 2009's broadband hemispherical integration is needed for atmospheric correction of urban TIR measured from a downward facing pyrgeometer.
-A method to correct hemispherical broadband TIR as measured from a downward facing pyrgeometer needs to combine the two for accurate T\textsubscript{hem} retrieval over rough terrain.
+Several multi-LOS correction routines have been developed for to remove atmospheric effects on near-ground remote sensed TIR: Meier et al., 2011 describes a correction method for oblique angled thermal imagery of urban terrain \cite{Meier2011}.
+In the method, spatially distributed LOS geometries were calculated by linking each image pixel to the corresponding 3-d coordinates of its viewpoint on a digital building model (DBM).
+Path lengths were then calculated as the distance between each pixel's corresponding location on the DBM and the sensor represented as the vanishing point of a 3-dimensional pyramidic projection from the sensor's location in the DBM.
+A pixel-by-pixel correction was then applied to remove atmospheric effects and retrieve T\textsubscript{rad} for each pixel at 30 minute intervals, resulting in a time-continuous climatology of urban T\textsubscript{rad}.
+However, the method uses thermal images in conjunction with a DBM to calculate path length geometries for each pixel's LOS - a technique not possible with a pyrgeometer, which returns a single integrated value over the sensor FOV.
+Moreover, the target instrument operates over a narrow waveband with relatively uniform spectral sensor response, reducing the magnitude and variance in atmospheric transmission over the sensor response curve.
+Thus, the method is not directly generalizable to correct TIR measured via pyrgeometer.
+
+Kotani \& Sugita, 2009 describes a method for correction of wide-FOV (pyrgeometer) TIR irradiances \cite{Kotani2009a} over a homogeneous planar surface.
+In this method, path lengths were calculated for six sensor heights.
+Radiances were then modeled using LOWTRAN initialized at $5^{\circ}$ intervals and integrated over the hemisphere to retrieve irradiances for a suite of T\textsubscript{surf}, and ambient T\textsubscript{air} and humidities.
+The resulting lookup table (LUT) of values is then used to correct $L_z$ and quantify atmospheric effects on remote sensed $L_z$ measured from several sensor heights.
+Methods outlined in Meier et al., 2011 and Kotani \& Sugita, 2009 are limited to narrow-FOV thermal imagers and planar terrain respectively.
+A method which combines Meier et al., 2011's representation of complex surface geometry and Kotani \& Sugita, 2009's broadband hemispherical integration is needed for atmospheric correction of urban TIR measured via downward facing pyrgeometer.
 
@@
 
-The "rolling lookup-table" method described in this study uses a sensor view model in conjunction with a radiative transfer code to model hemispherical irradiances upwelling from a simplified isothermal 3-dimensional representation of the target study area.
-In summary, the method (depicted in figure \ref{flow}) uses vertical profiles of measured T\textsubscript{air} and humidity to model at-sensor spectral radiances at 5$^{\circ}$ increments over the sensor FOV for a predefined range of possible T\textsubscript{hem} at each time-step.
+The "rolling lookup-table" method described in this study uses a sensor view model in conjunction with a radiative transfer code to model hemispherical irradiances upwelling from a simplified isothermal 3-dimensional representation of the urban surface.
+In summary, the method (depicted in Figure \ref{flow}) uses vertical profiles of measured T\textsubscript{air} and humidity to model at-sensor spectral radiances at 5$^{\circ}$ increments over the sensor FOV for a predetermined range of possible T\textsubscript{hem, r} at each time-step.
 Spectral directional radiances are convolved by a dome transmittance curve, integrated over the sensor waveband, and weighted for their respective angular view factor.
-Weighted directional radiances are then integrated over the hemisphere and aggregated into a lookup table (LUT) of modeled irradiance---T\textsubscript{hem} pairings for each timestep, unique to the vertical profile of measured T\textsubscript{air} and humidity.
-Finally, for each time-step, measured irradiances are matched with the closest modeled irradiances in the associated LUT to return an atmospherically corrected hemispherical surface temperature (T\textsubscript{hem}).
-This process is repeated at 30 minute intervals to yield a continuous climatology of urban T\textsubscript{hem} for surface urban heat island (sUHI) analysis.
-The following sections introduce the study area and describe the sensor view modeling, radiative transfer, and post-processing steps of the method.
+Weighted directional radiances are then integrated over the hemisphere and aggregated into a LUT of modeled irradiance - T\textsubscript{hem, r} pairings for each time step, unique to the vertical profile of measured T\textsubscript{air} and humidity.
+Finally, for each time step, measured irradiances are matched with the closest modeled irradiances in the LUT to return an atmospherically corrected radiometric hemispherical surface temperature.
+This process is repeated at 30 minute intervals to yield a continuous climatology of urban T\textsubscript{hem, r} for surface urban heat island (sUHI) analysis.
+The following sections introduce the study area for which the method was developed and describe the sensor view model, radiative transfer, and post-processing steps of the correction method.
 
@@
 	\includegraphics[width=\textwidth]{flow}
+	\caption{A workflow schematic depicting the input, model, and output-processing steps of a "rolling lookup table method" for hemispherical radiometric surface temperature retrial.}
 	\label{flow}
-	\caption{A workflow schematic depicting the input, model, and output-processing steps of a "rolling lookup table method" for hemispherical radiometric surface temperature retrial.}
 \end{figure}
@@
 
-As discussed in Section \ref{Atmospheric correction of near-ground TIR}, atmospheric correction of longwave irradiances measured from downward-facing, near-ground, wide-FOV sensors must account for surface geometry - particularly when mounted to view complex surface geometry.
-Hence, routines to retrieve atmospherically corrected urban T\textsubscript{hem} from upwelling longwave irradiances are inherently site specific.
-However, it is important to note that although correction magnitudes described in this paper are not generalizable, the correction method described in this paper can readily be adapted to different study sites, pyrgeometer types, and unique surface geometries.
+As discussed in section \ref{Atmospheric correction of near-ground TIR}, atmospheric correction of longwave irradiances measured from downward-facing, near-ground, wide-FOV sensors must account for complex surface geometry.
+Thus, routines to retrieve atmospherically corrected urban T\textsubscript{hem, r} from upwelling longwave irradiances are inherently site specific.
+However, it is important to note that although correction \textit{magnitudes} described in this paper are not necessarily generalizable, the correction \textit{method} described in this paper can readily be adapted to different study sites, sensor types, and unique surface geometries.
 
-With methodological generalizability in mind, a "rolling lookup table" atmospheric correction method was developed to retrieve radiometric T\textsubscript{hem} for a climatology of upwelling longwave irradiances measured from above the Sperrstrasse urban street canyon in Basel, Switzerland.
+With methodological generalizability in mind, a "rolling lookup table" atmospheric correction method was developed to retrieve radiometric T\textsubscript{hem, r} for a climatology of upwelling longwave irradiances measured from above the Sperrstrasse street canyon in Basel, Switzerland.
 The site, instrumented as a part of the Basel Urban Boundary Layer Experiment (BUBBLE) \cite{Rotach2005}, has an approximately northeast-southwest orientation and site morphology representative of local climate zone (LCZ) 2\footnote{Site surroundings can be described by the following morphological parameters: mean building height: 14.6m, plan aspect ratio: 0.54, complete aspect ratio: 1.92, local canyon aspect ratio: 1.0, and average shortwave albedo: 0.11 \cite{Rotach2005}.} \cite{Stewart2012}.
-LCZ classification was based on an assessment of surface characteristics in a 250m circular area extending from the Sperrstrasse tower using a 1m raster digital building model (DBM).
-Thus, the morphological parameters identified in \cite{Rotach2005} are representative of the vast majority of the pyrgeometer footprint.
-However, it should be noted that vegetation was not included in the DBM, and thus is not represented in morphological assessment or the sensor view model.
+LCZ classification was based on an assessment of surface characteristics in a 250m circular area extending from the Sperrstrasse tower using a 1 \si{m} raster digital building model (DBM).
+Thus, the morphological parameters identified in \cite{Rotach2005} are representative of a majority of the pyrgeometer footprint.
+Vegetation was not included in the DBM, and thus is not represented in morphological assessment or the sensor view model, however, the street canyon and surrounding area has little vegetation.
 
-For the nine month period between November 2001 and July 2002, a triangular lattice tower was installed skewed towards the southeast facing wall near the center of the Sperrstrasse canyon to observe a full suite of meteorological, radiation, and flux parameters.
-Profiles of T\textsubscript{air} and humidity were measured from seven heights extending from 2.5m to 31.5m above the canyon floor (with the highest observation level at approximately 2.17 times mean building height).
-Upwelling and downwelling short/longwave fluxes were measured at the lowest and highest tower heights, with an additional downward facing pyrgeometer mounted at roof level in the center of the canyon.
-In addition, during a summertime intense operation period (IOP) an array of narrow-FOV IRTs was installed to view individual facet surface temperatures (T\textsubscript{facet}) of approximately the same urban patch viewed by the pyrgeometer.
+For the eight month period between December 2001 and July 2002, a triangular lattice tower was installed within the Sperrstrasse street canyon.
+It's location was skewed towards the southeast facing wall near the along-canyon center of the canyon.
+Instruments to observe a full suite of meteorological variables and fluxes of heat, mass, and momentum are mounted at various levels on the tower.
+Profiles of T\textsubscript{air} and humidity were measured at seven heights extending from 2.5 \si{m} to 31.5 \si{m} above the canyon floor (with the highest observation level at approximately 2.17 times mean roof level).
+Upwelling and downwelling short/longwave fluxes were obtained from radiometers mounted at the lowest and highest measurement levels, with an additional downward facing pyrgeometer mounted at roof level near the center of the street canyon.
+In addition, during a summertime intense operation period (IOP) an array of narrow-FOV IRTs was installed to sample representative individual facet surface temperatures (T\textsubscript{facet}).
 
 The BUBBLE Sperrstrasse site was chosen for two primary reasons: 1.
-The study site provided a long-term climatology of radiation and meteorological variables for a representative mid-latitude city over a diverse range of synoptic conditions.
-This allowed for examination of urban T\textsubscript{surf}, atmospheric correction magnitudes, and sUHI magnitudes over a wide range of mid-latitude conditions.
+The site provided a long-term climatology of radiation and meteorological variables for a representative mid-latitude city.
+This allowed for examination of urban T\textsubscript{surf}, sUHI, and atmospheric correction magnitudes over a wide range of representative mid-latitude conditions.
 2.
-T\textsubscript{facet} measured during the summertime IOP allowed for direct climatological comparison of T\textsubscript{hem} to wall (T\textsubscript{wall}), roof (T\textsubscript{roof}), and road (T\textsubscript{road}) surface temperatures of both the northwest- and southeast-facing sides of the canyon.
-In addition, several hypothetical sensor views of the canyon were simulated by weighting to represent different geometric representations of the canyon, including a nadir/plan view, an oblique south-facing, and an oblique north-fac from narrow-FOV sensors and a complete, 3-dimensional view of the canyon.
-This facilitated comparison of T\textsubscript{hem} against different instrument types and platforms to identify and quantify biases in each method.
-In each representation, different weightings were applied to T\textsubscript{wall}, T\textsubscript{roof}, and T\textsubscript{road} on both sides of the canyon to represent different sensor orientations and sampling regimes.
-Weightings for each representation are described in table \ref{weightings}.
-\begin{table}[!ht]
+In inclusion of T\textsubscript{facet} over the IOP allows for investigation of the effect of sensor-surface geometry on remote sensed urban T\textsubscript{surf} for common urban T\textsubscript{surf} measurement strategies.
+This directional dependence of urban T\textsubscript{surf} - termed "effective anisotropy" - refers to the notion that a TIR signal upwelling from rough terrain viewed from a remote sensor changes based on the combined effects of surface-sensor-sun geometry and the convoluted, 3-dimensional structure of the urban surface \cite{Voogt1998a}.
+Facet surface temperatures measured during the summertime IOP allow for direct climatological comparison of T\textsubscript{hem, r} to common remote sensed representations of the urban surface constructed from weighted averages of wall (T\textsubscript{wall}), road (T\textsubscript{road}), and roof (T\textsubscript{roof}) temperatures.
+Table \ref{weightings} describes T\textsubscript{facet} weightings for nadir remote sensed (T\textsubscript{plan}) and complete (T\textsubscript{comp}) representations of urban surface temperature.
+At each time step, both T\textsubscript{hem, r} and T\textsubscript{plan} were divided by T\textsubscript{comp} and averaged over the IOP to yield normalized mean T\textsubscript{hem, r} and T\textsubscript{plan} at 30-minute intervals.
+Through comparison of T\textsubscript{hem, r} to T\textsubscript{plan} and T\textsubscript{comp} we investigate the effect of sensor-surface geometry on remote sensed urban T\textsubscript{surf} and quantify directional biases in common urban T\textsubscript{surf} measurements.
+
+\begin{table}[H]
 	\centering
-	\caption{Weightings to produce T'\textsubscript{rad} for different geometric representations of the Basel street canyon.}
+	\caption{Weights applied to individual facet temperature components in order to calculate complete and nadir temperatures for the Basel Sperrstrasse street canyon.}
 	\label{weightings}
@@
 		&  & Roof & Roof & Wall & Wall \\ 	\midrule
-
-		Complete & 0.33 & 0.16 & 0.16 & 0.16 & 0.16 \\
-
-		Nadir & 0.46 & 0.27 & 0.27 & 0.00 & 0.00 \\
-
-		Oblique (south-facing) & 0.20 & 0.25 & 0.25 & 0.00 & 0.30 \\
-
-		Oblique (north-facing) & 0.20 & 0.25 & 0.25 & 0.30 & 0.00 \\ 
+		Complete (T\textsubscript{comp}) & 0.33 & 0.16 & 0.16 & 0.16 & 0.16 \\ 
+		Nadir (T\textsubscript{plan}) & 0.46 & 0.27 & 0.27 & 0.00 & 0.00 \\ 
 		\bottomrule
@@
 
-The sheer number of unique path length geometries inherent with wide-FOV radiometry of rough terrain makes full 3-dimensional radiative transfer simulation difficult and computationally intensive --- particularly when correcting a climatology of T\textsubscript{hem}.
-In this method, to improve model efficiency, radiances are calculated for azimuthally averaged path lengths that represent average surface terrain for each solid angle "slice" of the sensor FOV.
-Thus, hemispherical radiative transfer is reduced to a 2-dimensional problem (shown in panel (b) of figure \ref{23radtran}).
-This greatly reduces the computational time required to model each irradiance-T\textsubscript{hem} pairing, as angular radiances can be computed as a function of zenith angle alone and subsequently weighted and integrated 3-dimensionally over the hemisphere.
+The sheer number of unique path length geometries inherent with wide-FOV radiometry of urban areas makes full 3-dimensional radiative transfer simulation difficult and computationally intensive - particularly when correcting a long-term climatology of irradiances or T\textsubscript{hem, r}.
+In this method, to improve efficiency, radiances are calculated for azimuthally averaged path lengths that represent average surface-sensor geometry for each solid angle "slice" of the sensor FOV.
+Thus, hemispherical radiative transfer is reduced to a 2-dimensional problem as shown in Figure \ref{23radtran}.
+This greatly reduces the computational time required to model each irradiance - T\textsubscript{hem, r} pairing, as angular radiances can be computed as a function of zenith angle alone and subsequently weighted and integrated 3-dimensionally over the hemisphere.
 
-To calculate surface-sensor geometries, the Surface-Sensor-Sun Urban Model (SUM) \cite{Soux2004} is initialized with a simplified, orthogonal 3-dimensional DBM representing geometry of the area surrounding the sensor.
+To calculate surface-sensor geometries, the Surface-Sensor-Sun Urban Model (SUM) \cite{Soux2004} is initialized with a simplified, orthogonal 3-dimensional DBM representing surface geometry of the area surrounding the sensor.
 SUM uses a four-dimensional array to represent surface morphology with three spatial dimensions ($x$, $y$, and $z$), with $z$ representing height above the $x$, $y$ plane.
@@
 
-\begin{figure}[!ht]
+\begin{figure}[H]
 	\includegraphics[width=\textwidth]{23radtran}
+	\caption{Path lengths over an urban surface.
+a) A downward radiometer mounted above a simplified three-dimensional urban surface showing b) the component radiances and path lengths from a vertical slice.
+Dotted lines indicate radiation streams subject to absorption by the intervening atmospheric layer.}
 	\label{23radtran}
-	\caption{A typical 2-dimensional (b) radiative transfer schematic adapted for an idealized 3-dimensional urban area (a).
-In 3-dimensions, path length for a given zenith angle can change significantly with azimuth angle.
-Dotted lines indicate absorption by the intervening atmospheric layer.}
 \end{figure}
@@
 \subsection{Modeling hemispherical irradiances}
-With path length geometries calculated in SUM, irradiances are modeled time-sequentially using version 4.1 of the MODerate resolution atmospheric TRANsmission radiative transfer code (MODTRAN) \cite{Berk1987}.
-At each time-step, T\textsubscript{bright} inferred from the measured irradiance and subtracted by a constant (6 \si{K} during nighttime runs and 4 \si{K} for daytime runs) is used to select a range of potential T\textsubscript{hem} which forms the bounds of the LUT for the target time-step.
-At-sensor spectral radiances for each path length are then modeled over a waveband of 0---2500 \si{cm^{-1}} at 0.5 \si{K} intervals over the LUT range.
-Atmospheric profiles are retrieved from 30-minute averages of T\textsubscript{air} and humidity, with aerosol and above-sensor conditions are informed by the mid-latitude summer standard atmosphere when daytime maximum T\textsubscript{air} is greater than 10 \si{\degreeCelsius} (the mid-latitude winter profile is substituted on days with maximum T\textsubscript{air} of less than 10 \si{\degreeCelsius}) \cite{Kantor1962}.
+With path length geometries calculated in SUM, irradiances are modeled for each time step using version 4.1 of the MODerate resolution atmospheric TRANsmission radiative transfer code (MODTRAN) \cite{Berk1987}.
+At each time-step, a range of potential temperatures is specified (T\textsubscript{spec}) based on T\textsubscript{hem, b} calculated from the measured irradiance with
 
-In this method, a "typical" longwave bandpass (approximately 4---42 \si{\micro\meter}) was extended to include much longer wavelengths.
-This was done for two reasons: 1.
-To accurately represent a longwave curve for typical urban T\textsubscript{surf}, which have significant emittance into wavelengths much longer than 42 \si{\micro\meter}.
-2.
-To replicate the signal from a silicone-domed pyrgeometer, which is transmissive of radiation in wavelengths.
-A "typical" longwave bandpass underestimates a pyrgeometer signal by approximately 5---9 \si{\watt\per\square\meter}, depending on emitter temperature.
-Figure \ref{dometrans} comparison of ‘typical’ and ‘extended’ longwave bandpasses.
+\begin{equation}
+	T_{spec, ~min} = T_{hem, ~b} - 6 ~K 
+\end{equation}
 
-\begin{figure}[!ht]
+\noindent defining a minimum T\textsubscript{spec} before iterating over
+
+\begin{equation}
+T_{spec, ~i} = T_{spec, ~min} + \displaystyle\sum_{i = 1}^{n} 0.5n, ~1 \leq n \leq 32
+\end{equation}
+
+\noindent to retrieve a LUT of T\textsubscript{spec}.
+
+After the LUT is defined, at-sensor spectral radiances for each path length/zenith angle are modeled at an emissivity of 0.95 over a waveband of 0 - 2300 \si{cm^{-1}} for each T\textsubscript{spec}.
+Profiles of T\textsubscript{air} and humidity are retrieved from 30-minute averages of conditions observed at the Sperrstrasse site.
+Aerosol, trace gas absorber, and above-sensor T\textsubscript{air} and humidity conditions are defined by the mid-latitude summer standard atmosphere when daytime T\textsubscript{air, max} \textgreater 10 \si{\degreeCelsius} (the mid-latitude winter profile is substituted on days where T\textsubscript{air, max} \textless 10 \si{\degreeCelsius}) \cite{Kantor1962}.
+
+In this method, a "typical" longwave bandpass, approximately 250 - 2300 \si{\centi\meter^{-1}} (4 - 42 \si{\micro\meter}), was extended to include much smaller wavenumber (longer wavelengths).
+This was done for two reasons: 1) To accurately represent spectral longwave curves for typical urban T\textsubscript{surf}, which show significant emittance in wavenumber smaller than 250 \si{cm^{-1}} (wavelengths much longer than 42 \si{\micro\meter}).
+2) To replicate the spectral signal "seen" by a silicone-domed pyrgeometer, which continues to transmit radiation at wavenumber \textless 250 \si{cm^{-1}} (wavelengths \textgreater 42 \si{\micro\meter}).
+A "typical" longwave bandpass underestimates a pyrgeometer signal by approximately 7 - 10 \si{\watt\per\square\meter}, depending on emitter temperature.
+A comparison of ‘typical’ and ‘extended’ longwave bandpasses is shown in Figure \ref{dometrans}.
+
+\begin{figure}[H]
 	\includegraphics[width=\textwidth]{dometrans}
+	\caption{A comparison of dome transmittance for a Kipp \& Zonen silicone domed pyrgeometer and at-sensor radiance (T\textsubscript{surf} = 300 \si{K}, T\textsubscript{air} = 300 \si{K}, $z$ = 30 \si{m})  for "typical" and extended bandpasses.}
 	\label{dometrans}
-	\caption{ddddd}
 \end{figure}
 
-\section{Evaluation of the method using profiles of upwelling longwave radiation over a homogeneous planar surface}
+At-sensor spectral radiances computed by MODTRAN are convolved by a dome transmittance curve and integrated over the bandpass via
 
-MODTRAN has been shown to effectively model near-ground radiative transfer at urban-scale path lengths in 2-dimensions.
-However, the method detailed in this paper includes significant post-processing to ‘collate’ point-to-point radiances into 3-dimensional hemispherical irradiances.
-In addition, the method accounts for site-specific surface geometry, which has a significant influence on boundary-layer transmittance of broadband longwave radiation.
-Thus, prior to deriving a climatology of radiometric urban Them, the method was evaluated over a simple surface with known surface characteristics.
-Thus, divergences are solely the result of differential atmospheric effects with increasing height above ground, which should be accounted for by the model.
-It is assumed that each pyrgeometer in this validation views a patch with approximately the same temperature.
+\begin{equation}
+\label{work1}
+	L'_z (\theta, \phi) = \frac{\bigints_{\nu_1}^{\nu_2} R(\theta, \phi, \lambda) ~ r(\lambda) ~ d\lambda}{\overline{r}}
+\end{equation}
 
-Profiles of upwelling longwave radiation, downwelling shortwave, air temperature, and humidity at 2m, 10m, and 30m were obtained from the Payerne station, located 50 miles southwest of Basel, Switzerland in a cultivated field.
-To evaluate the method, we used a modified version of the workflow described in XXXX.
-Brightness temperature calculated from the lowest upwelling longwave measurement was used to model irradiances at 10m and 30m at each time-step.
-Downwelling shortwave radiation was used to categorize test days based on cloud cover.
-The daytime air temperature profile was modified to replicate typical lapse rates in the Sperrstrasse canyon.
-Unlike, Kotani \& Sugita (2009), an isothermal atmospheric profile was not sufficient to accurately model upwelling longwave divergences (nor fluxes) of planar terrain.
-Although thermal stratification is likely to be relatively small by day in an urban canyon \cite{Nakamura1988} --- where strong microscale contrasts in Tsurf foster canyon mixing and neutral stability --- the large daytime Tsurf-T\textsubscript{air}  differential and the path-length/transmittance gradient can create large positive divergences (5 – 15 Wm\textsuperscript{-2},CHECK THIS).
-As such, we suggest using a full canyon T\textsubscript{air} /humidity profile to
+\noindent to yield a directional radiance $ L'_z $ for each $\theta$, with $\overline{r}$ as mean dome transmittance over a waveband of $\nu_1 $= 0 \si{cm^{-1}} to $\nu_2$ = 2300 \si{cm^{-1}}.
+$ L'_z (\theta) $ are then multiplied by their associated angular view factor weighting ($\Phi$) and integrated over the hemisphere to yield an irradiance $ L_z $ for the target T\textsubscript{hem, r}
 
-Modeled upwelling longwave at 10m and 30m and divergences are compared to their measured counterparts in figure a.
-Both fluxes and divergence show strong correlations, thus we conclude: 1.
-Irradiances measured at 2m are free from significant atmospheric influence – as an uncorrected irradiance provided an accurate Tsurf for modeling of 10m and 30m Irradiances As such, brightness Them and radiometric Them are approximately equal when z < 2m.
-Although, it should be noted that a 2m sensor is not representative of canyon geometry and should not be used to derive urban Them.
+\begin{equation}
+\label{work2}
+	L_z (T\textsubscript{spec}) = \int_{0}^{2\pi}\int_{0}^{\pi/2} L'_z (\phi, \theta) ~ \Phi (\phi) ~ d\theta d\phi
+\end{equation}
+
+$L_z (T\textsubscript{spec}) $ is representative of at-sensor irradiance upwelling from the urban surface described in SUM at the target specified temperature T\textsubscript{hem, r} for the measured T\textsubscript{air}, humidity, aerosol, and trace gas profile.
+The process is repeated to retrieve $L_z$ for the range of potential T\textsubscript{hem, r} at the given time-step.
+Irradiance - T\textsubscript{hem, r} pairings are then aggregated into a LUT.
+Finally, the measured irradiance from BUBBLE is matched with its closest modeled irradiance to yield an atmospherically corrected, radiometric hemispherical surface temperature for the given time step.
+The workflow is repeated at 30-minute intervals to yield a continuous time series of T\textsubscript{hem, r}.
+Atmospheric correction magnitudes can then be calculated as the difference between T\textsubscript{hem, r} and T\textsubscript{hem, b}, with T\textsubscript{hem, b} calculated via:
+
+\begin{equation}
+	\label{stefboltz2}
+	T_{hem,~ b} = \sqrt[4]{\frac{L_z}{\sigma}}
+\end{equation}.
+
+\section{Evaluation of the method using profiles of upwelling longwave radiation over a homogeneous planar surface} \label{sensitivity}
+
+MODTRAN has been shown to effectively model near-ground radiative transfer at urban-scale path lengths in 2-dimensions \cite{Hoch2005} \cite{Hoch2007}.
+However, the method described in this paper accounts for complex 3-dimensional surface geometry, which can exert significant influence on atmospheric transmittance.
+In addition, the method includes significant post-processing to weight and integrate point-to-point spectral directional radiances to derive 3-dimensional, hemispherical irradiances.
+Thus, prior to deriving a climatology of urban T\textsubscript{hem, r}, the method was evaluated using profiles of upwelling longwave irradiances measured over a simple planar surface.
+
+%Although downward facing pyrgeometers mounted at different heights do not have the same radiative source area.
+Over a simple relatively homogeneous surface, the signal measured by each sensor should be radiating from an approximately isothermal surface area.
+As such, differences in the upwelling longwave signal between sensor heights (a phenomenon termed longwave divergence) are solely the result of differential atmospheric effects, which should be accounted for by the model.
+
+For a continuous 14-day period, profiles of up/downwelling short/longwave radiation, T\textsubscript{air}, and humidity measured from 2 \si{m}, 10 \si{m}, and 30 \si{m} were obtained from an instrumented tower in Payerne, Switzerland.
+The tower, installed over a cultivated field as a part of the Baseline Surface Radiation Network (BSRN), is located approximately 100 \si{\kilo\meter} southwest of the BUBBLE Sperrstrasse tower and is subject to similar summertime conditions as the study site.
+The study period was chosen to include a comprehensive range of T\textsubscript{surf}, T\textsubscript{air}, humidities, and cloud coverages to represent typical summertime conditions over which atmospheric corrections show the greatest variation.
+To evaluate the method, we used a modified version of the workflow described in Figure \ref{flow}.
+First, azimuthally averaged path lengths over a planar surface were calculated in SUM at 5\si{\degree} increments of zenith angle over the sensor FOV for the 10 \si{\meter} and 30 \si{\meter} sensors.
+For each time step, T\textsubscript{hem, b}, calculated from the lowest $L_z$ measurement, was used to model irradiances at 10 \si{m} and 30 \si{m} at 30-minute intervals using concurrent profiles of T\textsubscript{air} and humidity.
+Daytime T\textsubscript{air} observed at the highest measurement level was anomalously high compared to typical summertime lapse rates above planar vegetated terrain \cite{Oke1987}.
+Thus, the T\textsubscript{air} profile was modified to be more representative of the neutral or weakly unstable daytime conditions and weakly stable nighttime conditions.
+It is likely that these modifications better represent the T\textsubscript{air} profile in the layer between 10 and 30 \si{\meter} than a simple interpolation between the 10 \si{\meter} T\textsubscript{air} and 30 \si{\meter} T\textsubscript{air} measurements.
+Contrary to the findings in Kotani \& Sugita (2009), an isothermal atmospheric profile was not sufficient to accurately model upwelling fluxes over planar terrain.
+Although thermal stratification is relatively small by day in an urban canyon \cite{Nakamura1988} - where strong microscale contrasts in T\textsubscript{surf} foster strong canyon mixing and neutral stability - the large daytime T\textsubscript{surf} - T\textsubscript{air} differential and the path-length/transmittance gradient can create large differences in upwelling longwave forced by inter-canyon thermal contrasts - a phenomenon termed radiative divergence.
+By underestimating near-surface air temperatures and overestimating roughness-layer T\textsubscript{air}, positive divergences by day are underestimated by an isothermal profile.
+As such, we suggest using a full canyon T\textsubscript{air} and humidity profile to most accurately model near ground fluxes.
+
+Modeled and measured upwelling longwave at 10 \si{m} and 30 \si{m} fluxes and divergences - calculated as the difference between 30 \si{\meter} and 10 \si{\meter} fluxes - are compared in Figure \ref{fluxes}.
+Both fluxes and divergences show strong correlation, thus we conclude: 1.
+Irradiances measured near screen level (approximately 2 \si{\meter} above ground level) do not include significant atmospheric influence - as T\textsubscript{bright} inferred from $L_{z=2~\si{\meter}}$ was sufficient for modeling irradiances at 10 \si{m} and 30 \si{m}.
+As such, T\textsubscript{bright} and T\textsubscript{hem, r} are reasonably equal when $z$ = 2 \si{m}.
+However, it should be noted that in urban areas a 2 \si{m} sensor is not sufficiently representative of canyon geometry and should not be used to derive urban T\textsubscript{hem, r}.
 2.
-The method is sensitive enough to model divergences in a large layer above a flat surface.
-Urban divergences are likely to be smaller with less frequent and less severe stable stratifications.
-Thus, we can safely make the assumption that the method is effective over complex terrain – provided path length geometries are accurately replicated in SUM.
+MODTRAN can accurately model longwave fluxes and divergences above a flat, homogeneous surface over a wide range of temperatures, humidities, and cloud coverages and large resulting contrasts in near-ground stability.
+At 30 \si{\meter} under typical summertime profiles of humidity, CO\textsubscript{2}, and O\textsubscript{3}, view factor weighted hemispherical transmittance over a planar surface is approximately 55\% - meaning 45\% of a broadband hemispherical TIR signal at 30 \si{\meter} is emitted by the atmosphere, rather than the target surface.
+As such, it is imperative that T\textsubscript{air} profiles are well represented in MODTRAN to accurately model the atmospheric component of remote sensed TIR signal.
+As profiles of T\textsubscript{air} in the layer between the surface and the sensor at the Sperrstrasse site are generally subject to neutral or weakly unstable conditions, with only rare stable stratifications, urban longwave divergences are likely to be smaller than those observed at Payerne - and well approximated by MODTRAN.
+Thus, provided path length geometries are accurately replicated in by the sensor view model, we can reasonably assume that the method has similar effectiveness at the urban site.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=\textwidth]{eval}
+	\caption{A comparison of 10 \si{m} and 30 \si{m} measured and modeled longwave fluxes and divergences - calculated as the difference between 30 \si{\meter} and 10 \si{\meter} fluxes - over the 14 day evaluation period at Payerne, Switzerland.}
+	\label{fluxes}
+\end{figure}
+
+\begin{table}[H]
+	\centering
+	\caption{Statistical performance of the Payerne evaluation. RMSE\textit{\textsubscript{s}} and RMSE\textit{\textsubscript{u}} represent the systemic and unsystematic RMSE respectively. n = 378}
+	\label{evalstats}
+	\begin{tabular*}{\textwidth}{l@{\extracolsep{\fill}} cccr}
+		\toprule 
+		Statistic & 10 \si{\meter} flux & 30 \si{\meter} flux & Divergence & Citation \\  \midrule
+		Slope & 1.002 & 1.011 & 0.792 &  \\ 
+		Intercept, (\si{\watt\per\meter\squared}) & 0.422 & $-$3.560 & $-$0.583 &   \\ 
+		\textit{R\textsuperscript{2}} & 0.995 & 0.996 & 0.863 & \\ 
+		MAE,  (\si{\watt\per\meter\squared}) & 1.781 & 1.325 & 1.191 & \cite{Willmott1985a} \\ 
+		RMSE,  (\si{\watt\per\meter\squared}) & 2.204 & 1.690 & 1.423 & \cite{Willmott1985a} \\ 
+		RMSE\textit{\textsubscript{s}},  (\si{\watt\per\meter\squared}) & 1.444 & 0.851 & 1.028 & \cite{Willmott1985a} \\ 
+		RMSE\textit{\textsubscript{u}},  (\si{\watt\per\meter\squared}) & 1.666 & 1.460 & 0.985 & \cite{Willmott1985a} \\ 
+		\textit{d} & 0.999 & 0.999 & 0.959 & \cite{Willmott2012} \\
+		\bottomrule
+	\end{tabular*} 
+\end{table}
+
+\section{Results}
+
+\subsection{Atmospheric correction magnitudes}
+
+Atmospheric correction magnitudes binned for season and time of day are show in Figure \ref{bx_ac}.
+Summertime correction magnitudes show significant day-to-day and diurnal variation.
+Variance is largest in summer near solar noon, where different synoptic conditions can force correction magnitudes of up to $8$ \si{\kelvin} in clear-sky hot conditions and below $0$ \si{\kelvin} under cooler overcast conditions.
+In winter, correction magnitudes display less variance and are smaller - approximately $2$ \si{\kelvin} over a range of representative wintertime synoptic conditions.
+
+Over both summer and winter seasons, correction magnitudes are strongly correlated with $\Delta$T\textsubscript{hem, r - air}.
+Figure \ref{season_fit} shows the effect of seasonality and cloud cover on the relationship between correction magnitudes and $\Delta$T\textsubscript{hem, r - air}.
+The relationship is strongest and most clear in summer when variations in cloud cover can lead to large day-to-day contrasts in solar input and $\Delta$T\textsubscript{surf - air}, which translates into a wide range of correction magnitudes.
+
+Figure \ref{nofit} illustrates the the relationship between correction magnitude and meteorological variables commonly measured as a part of climatological assessments of urban environments.
+A strong positive relationship is observed between correction magnitudes and incoming solar radiation.
+T\textsubscript{hem, r} and T\textsubscript{air} display more complex relationships with correction magnitude, with a large cluster of observed correction magnitudes situated at a local minimum between $10$ and $20$ \si{\degreeCelsius}, weakly negative correlation at temperatures lower than $10$ \si{\degreeCelsius}, and weakly positive correlation at temperatures greater than $20$ \si{\degreeCelsius}.
+Atmospheric water vapor content did not appear to exert significant control over correction magnitudes.
+
+Figure \ref{humtest} shows hemispherical atmospheric transmittance as a function of vapor pressure for pyrgeometers at 10 \si{\meter} and 30 \si{\meter} above a planar surface.
+Hemispherical transmittances are modeled in MODTRAN 4.1 as the fraction of surface emission reaching the sensor for the path length and view factor geometries calculated in the method evaluation in Section \ref{sensitivity}.
+Hemispherical transmittance falls sharply with increasing vapor pressure between 0 and 5 \si{\hecto\pascal}, but is relatively constant with vapor pressure \textgreater 5 \si{\hecto\pascal}.
+This results in the weak correlation observed in Figure \ref{nofit} between humidity and correction magnitude.
+
+%Correction magnitudes are most strongly dependent on incoming solar radiation - which exerts significant control over seasonal and diurnal variance in the T\textsubscript{hem, r} - T\textsubscript{air} differential.
+These observations are largely made up of nighttime and cloudy points during the summer months with small or weakly negative T\textsubscript{surf} - T\textsubscript{air} differentials.
+During the winter months, a negative relationship between T\textsubscript{hem, r}/T\textsubscript{air} and correction magnitudes is observed at lower overall T\textsubscript{hem, r} and T\textsubscript{air}.
+In summer months, a positive relationship between T\textsubscript{hem, r}/T\textsubscript{air} and correction magnitudes is observed at higher overall temperatures.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=15cm,
+	height=7in,
+	keepaspectratio]{bxsac_season_clima_pres}
+	\caption{Atmospheric correction magnitudes calculated as the difference between T\textsubscript{hem, r} and T\textsubscript{bright} for the duration of the study period.
+Grey shading indicates nighttime.}
+	\label{bx_ac}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=15cm,
+	height=7in,
+	keepaspectratio]{season_fit}
+	\caption{Correction magnitude versus $\Delta$T\textsubscript{plan - hem, r} binned first for season and the then for cloud cover in the summer months.}
+	\label{season_fit}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=15cm,
+	height=7in,
+	keepaspectratio]{nofit}
+	\caption{Correction magnitude versus meteorological variables for the daytime hours of the study period. }
+	\label{nofit}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=15cm,
+	height=7in,
+	keepaspectratio]{hum_test}
+	\caption{Hemispherical at-sensor atmospheric transmittance as a function of vapor pressure for two pyrgeometer heights (10 \si{\meter} and 30 \si{\meter}) over a planar surface.
+Calculated as the fraction of surface emission reaching the sensor height using MODTRAN 4.1 with T\textsubscript{surf} = T\textsubscript{air} = 300 \si{\kelvin} with water vapor as the sole atmospheric absorber.}
+	\label{humtest}
+\end{figure}
+
+%\subsection{Relating correction magnitudes to climatological parameters}
+
+%Correction magnitudes show strong diurnal and seasonal relationships with the T\textsubscript{surf} - T\textsubscript{air} differential - shown in figure \ref{season_fit}.
+As such, parameters that determine the T\textsubscript{surf} - T\textsubscript{air} differential exert similar control over correction magnitudes, most notably incoming shortwave radiation, T\textsubscript{surf}, and T\textsubscript{air}, shown in figures \ref{nofit}.
+Correction magnitudes are greatest with large shortwave radiation input and with high T\textsubscript{air} and T\textsubscript{hem, r}, during which contrasts between T\textsubscript{air} and T\textsubscript{surf} are largest.
+Conversely, correction magnitudes are smallest under conditions that suppress contrasts between T\textsubscript{surf} and T\textsubscript{air} - when solar input is small, and with low T\textsubscript{air} and T\textsubscript{hem, r}.
+In summer, variations in solar input based on time of day and cloud cover produce large fluctuations in the T\textsubscript{surf} - T\textsubscript{air} differential and significant variation in correction magnitude with time of day and synoptic conditions.
+Thus, maximum correction magnitudes occur on hot, clear-sky days, when solar input, T\textsubscript{air} and T\textsubscript{hem, r} are also maximized.
+Conversely, overcast and nighttime conditions with zero or low incoming solar radiation produce smaller or negative T\textsubscript{surf} - T\textsubscript{air} differentials and suppress correction magnitudes.
+In winter, a lower solar angle and shorter day length reduce overall solar input and the diurnal amplitude of T\textsubscript{hem, r} and T\textsubscript{air}.
+This suppresses both the overall size and diurnal amplitude of the T\textsubscript{surf} - T\textsubscript{air} differential and has a similar effect on winter correction magnitudes.
+
+%Atmospheric transmittance of longwave radiation is primarily determined by column water vapor content.
+However, correction magnitudes are only weakly correlated with humidity (figure \ref{corr_abshum}).
+This supports the notion that atmospheric transmittance determines the \textit{potential} for atmospheric effects on a remote sensed TIR signal, but not necessarily the \textit{magnitude} of atmospheric influence.
+In the absence of a strong contrast between T\textsubscript{surf} and T\textsubscript{air}, absorption and reemission of TIR by water vapor or other absorbers has little effect on a remote sensed signal.
+Contra, with a strong contrast between T\textsubscript{surf} and T\textsubscript{air}, but little absorber content, the atmosphere is highly transmissive and does not yield strong atmospheric effects.
+
+%\begin{figure}[H]
+%	\centering
+%	\includegraphics[width=15cm,
+%	height=7in,
+%	keepaspectratio]{corr_sd}
+%	\caption{Correction magnitude versus the T\textsubscript{hem, r}-T\textsubscript{air} differential for summer months of the study period (April 15\textsuperscript{th} - July 7\textsuperscript{th}).
+Contour shading indicates frequency (higher frequencies in red, lower in yellow).}
+%	\label{corr_sd}
+%\end{figure}
+%
+%\begin{figure}[H]
+%	\centering
+%	\includegraphics[width=15cm,
+%	height=7in,
+%	keepaspectratio]{corr_wd}
+%	\caption{Correction magnitude versus the T\textsubscript{hem, r}-T\textsubscript{air} differential for winter and spring months of the study period (Dec 1\textsuperscript{st} - April 14\textsuperscript{th}).
+Contour shading indicates frequency (higher frequencies in red, lower in yellow).}
+%	\label{corr_wd}
+%\end{figure}
+%
+%\begin{figure}[H]
+%	\centering
+%	\includegraphics[width=15cm,
+%	height=7in,
+%	keepaspectratio]{corr_sds}
+%	\caption{Correction magnitude versus the T\textsubscript{hem, r}-T\textsubscript{air} differential for clear sky summertime days over the study period.
+Contour shading indicates frequency (higher frequencies in red, lower in yellow).}
+%	\label{corr_sds}
+%\end{figure}
+%
+%\begin{figure}[H]
+%	\centering
+%	\includegraphics[width=15cm,
+%	height=7in,
+%	keepaspectratio]{corr_sdc}
+%	\caption{Correction magnitude versus the T\textsubscript{hem, r}-T\textsubscript{air} differential for partly cloudy and overcast summertime days over the study period.
+Contour shading indicates frequency (higher frequencies in red, lower in yellow).}
+%	\label{corr_sdc}
+%\end{figure}
+%
+%\begin{figure}[H]
+%	\centering
+%	\includegraphics[width=15cm,
+%	height=7in,
+%	keepaspectratio]{corr_kdown}
+%	\caption{Correction magnitude versus incoming shortwave radiation for the daytime hours of the study period. Contour shading indicates frequency (higher frequencies in red, lower in yellow).}
+%	\label{corr_kdown}
+%\end{figure}
+%
+%\begin{figure}[H]
+%	\centering
+%	\includegraphics[width=15cm,
+%	height=7in,
+%	keepaspectratio]{corr_them}
+%	\caption{Correction magnitude versus T\textsubscript{hem, r} over the duration of the study period. Contour shading indicates frequency (higher frequencies in red, lower in yellow).}
+%	\label{corr_them}
+%\end{figure}
+%
+%\begin{figure}[H]
+%	\centering
+%	\includegraphics[width=15cm,
+%	height=7in,
+%	keepaspectratio]{corr_tair}
+%	\caption{Correction magnitude versus T\textsubscript{air} over the duration of the study period. Contour shading indicates frequency (higher frequencies in red, lower in yellow).}
+%	\label{corr_tair}
+%\end{figure}
+%
+%\begin{figure}[H]
+%	\centering
+%	\includegraphics[width=15cm,
+%	height=7in,
+%	keepaspectratio]{corr_abshum}
+%	\caption{Correction magnitude versus atmospheric water vapor content for the daytime hours of the study period. Contour shading indicates frequency (higher frequencies in red, lower in yellow).}
+%	\label{corr_abshum}
+%\end{figure}
+
+\section{Comparing T\textsubscript{surf} from different sensor geometries}
+
+%\begin{figure}[H]
+%	\centering
+%	\includegraphics[width=15cm,
+%	height=7in,
+%	keepaspectratio]{tc_quarts2}
+%	\caption{A comparison of T\textsubscript{surf} from sensor geometries described in Table \ref{weightings}. T\textsubscript{surf} is normalized against complete %T\textsubscript{surf}.}
+%	\label{tc_quarts}
+%\end{figure}
+
+Figure \ref{tc_nadhem} shows a comparison of mean normalized T\textsubscript{hem, r} and T\textsubscript{plan} calculated as T\textsubscript{x} divided by T\textsubscript{comp}, where T\textsubscript{x} is T\textsubscript{hem, r} or T\textsubscript{plan}, and averaged over the IOP.
+Both T\textsubscript{hem, r} and T\textsubscript{plan} overestimate T\textsubscript{comp} by day and underestimate T\textsubscript{comp} by night.
+Mean daytime overestimation of T\textsubscript{comp} by T\textsubscript{hem, r} is smaller and less variable than that by T\textsubscript{plan}.
+In the figure, overestimation of T\textsubscript{comp} is observed at values above $1$ and underestimation by values below $1$.
+
+Mean normalized T\textsubscript{hem, r}, T\textsubscript{plan}, and $\Delta$T\textsubscript{plan - hem, r} are shown in figure \ref{tc_mean_314} along with clear sky and overcast sky case days.
+While mean T\textsubscript{hem, r} and T\textsubscript{plan} display similar overestimations of T\textsubscript{comp}, under clear sky conditions, overestimation by T\textsubscript{plan} is much larger than that by T\textsubscript{hem, r}.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=15cm,
+	height=7in,
+	keepaspectratio]{nad_hem_314}
+	\caption{A comparison of normalized T\textsubscript{hem, r} and T\textsubscript{plan} over the IOP. Shading indicates the area bounded by the first and third quartiles. n = 14}
+	\label{tc_nadhem}
+\end{figure}
+
+%As shown in figure \ref{tc_mean},
+
+%\begin{figure}[H]
+%	\centering
+%	\includegraphics[width=15cm,
+%	height=7in,
+%	keepaspectratio]{tc_mean3}
+%	\caption{Normalized T\textsubscript{hem, r}, T\textsubscript{plan}, and T\textsubscript{plan} minus T\textsubscript{hem, r} averaged over the IOP and for clear sky and overcast case days.}
+%	\label{tc_mean}
+%\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=15cm,
+	height=7in,
+	keepaspectratio]{tc_mean_314}
+	\caption{Normalized T\textsubscript{hem, r}, T\textsubscript{plan}, and $\Delta$T\textsubscript{plan - hem, r} averaged over the IOP and for clear sky and overcast case days. n = 14}
+	\label{tc_mean2}
+\end{figure}
+
+\section{Discussion}
+
+\subsection{Controls on atmospheric correction magnitude}
+
+Atmospheric correction magnitudes are highly dependent on factors that control $\Delta$T\textsubscript{surf - air}.
+With a nonzero $\Delta$T\textsubscript{surf - air}, TIR emitted from the surface and absorbed by the atmosphere is re-emitted at a different temperature, modifying the at-sensor TIR signal.
+Thus, the strength of the T\textsubscript{surf} to T\textsubscript{air} differential is directly related to the magnitude of atmospheric effects on a remote sensed TIR signal.
+This results in a strong positive relationship between atmospheric correction magnitude and $\Delta$T\textsubscript{hem, r - air}.
+Results show that conditions that maximize microscale surface to atmosphere thermal contrasts entail large correction magnitudes and conditions that suppress differences between T\textsubscript{surf} and T\textsubscript{air} show much smaller or negative correction magnitudes.
+
+Of the meteorological variables tested, results show that correction magnitudes are most strongly dependent on solar input.
+Solar radiation incident on the surface forces microscale thermal contrasts between the surface and the air, directly controlling the T\textsubscript{surf} - T\textsubscript{air} differential.
+The relationships between correction magnitude and T\textsubscript{hem, r} and T\textsubscript{air} individually are weaker and more complex.
+Correction magnitudes are smallest at temperatures between approximately $10$ and $20$ \si{\degreeCelsius} - largely made up of morning and cloudy observations during the summer months where thermal contrasts between T\textsubscript{surf} and T\textsubscript{air} are small or weakly negative, forcing small correction magnitudes.
+During the winter months, a negative relationship between T\textsubscript{hem, r}/T\textsubscript{air} and correction magnitudes is observed, likely a result of the effect of shifting spectral emittance curves combined with non-uniform sensor response - discussed in section \ref{Persistence in correction magnitudes}.
+On clear sky summer days, positive relationships between surface and air temperatures and correction magnitudes is observed as an increased solar input foster higher overall temperatures and large contrasts between T\textsubscript{surf} and T\textsubscript{air}.
+
+Results show significant seasonal, day to day, and diurnal variation in correction magnitudes.
+The range of observed correction magnitudes is largest in summer, when a wide range of mid latitude synoptic conditions result in significant contrasts in daytime solar input and similarly large variance in the T\textsubscript{surf} - T\textsubscript{air} differential.
+Maximum correction magnitudes over the study period (often approaching $7$ to $8$ \si{\kelvin}) occur frequently near solar noon on clear, hot, humid days, during which intense solar heating of the surface produces a large T\textsubscript{surf} - T\textsubscript{air} differential.
+Minimum correction magnitudes near $-1$ \si{\kelvin} occurs consistently on clear, calm nights following hot days, when T\textsubscript{surf} can dip below T\textsubscript{air}.
+In winter, decreased variance in overall solar input and lower solar angles result in less intense solar heating of the surface and smaller T\textsubscript{surf} - T\textsubscript{air} contrasts.
+Thus, winter correction magnitudes are smaller and more consistent with a mean of approximately $2$ \si{\kelvin}.
+
+\subsection{The effect of non-uniform pyrgeometer spectral dome transmittance on correction magnitudes}\label{Persistence in correction magnitudes}
+
+A pyrgeometer's dome transmittance is spectrally non-uniform.
+Thus, the "dome effect" a pyrgeometer's exerts on a given incoming TIR signal changes as a function of emission temperature (T\textsubscript{emiss}), as the shape of a given spectral radiance curve is temperature dependent.
+This is illustrated in Figure \ref{nonuniform} by comparing the relative peaks and shape of a pyrgeometer dome transmittance curve to a Planckian spectral radiance curve (R\textsubscript{Planck}) at T = 300 \si{\kelvin} and the same curve shifted to smaller wavenumber to represent the spectral radiance peak of a curve at T = 270 \si{\kelvin} (R\textsubscript{Planck, s}).
+When integrated, the two curves have the same radiance, but different spectral peaks and different resulting dome effects.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=15cm,
+	height=7in,
+	keepaspectratio]{nonuniform}
+	\caption{Spectral dome transmittance (\textit{r}) for a Kipp \& Zonen pyrgeometer overlaid with a Planckian spectral radiance curve at T = 300 \si{\kelvin} and the same curve shifted to represent the shape and peak of a spectral radiance curve at T = 270 \si{\kelvin}.}
+	\label{nonuniform}
+\end{figure}
+
+Spectrally non-uniform pyrgeometer dome transmittance results in differential dome effects on the two spectral radiance curves in Figure \ref{nonuniform} and two different radiance signals "seen" by the pyrgeometer.
+Thus, calibration routines designed to remove dome effects and retrieve accurate irradiance values are temperature specific and subject to errors based on the difference between T\textsubscript{emiss} and the temperature at which the instrument was calibrated (T\textsubscript{calib}).
+The magnitude of this error can be expressed as differences Planck weighted mean sensor response ($\overline{r} $) between T\textsubscript{emiss}, computed from equation \ref{meansens} - where lower $\overline{r} $ indicates a larger dome effect on the signal received by the sensor.
+In addition, this error can be expressed as a temperature bias (T\textsubscript{bias}) computed as the difference between a reference T\textsubscript{calib} and a dome adjusted temperature (T\textsubscript{adj}).
+T\textsubscript{adj} is computed by applying a spectral shift to R\textsubscript{Planck} calculated for T = 300 \si{\kelvin} so that its peak spectral radiance matches the wavenumber of peak spectral emission for the target T\textsubscript{emiss}\footnote{In this sensitivity test, the waveband was truncated from 1 - 2500 \si{cm^{-1}} to 80 - 2500 \si{cm^{-1}} to accommodate shorter wavenumber when shifting R\textsubscript{Planck} calculated at the reference T\textsubscript{calib} to lower temperatures where peak spectral emission occurs at smaller wavenumber.
+This results in a slight underestimation of the total radiance at a given T\textsubscript{emiss}, and thus, slightly negative T\textsubscript{bias} at T\textsubscript{calib}, where no spectral shift has been applied.
+When computed from total in band radiance, T\textsubscript{bias} at the reference T\textsubscript{calib} is zero.} - a similar shift is illustrated for a reference T\textsubscript{calib} = 300 \si{\kelvin} and T\textsubscript{emiss} = 270 \si{\kelvin} in Figure \ref{nonuniform}.
+Following the spectral shift, spectral radiances were convolved by the dome transmittance function, integrated over the waveband and hemisphere, and used to derive T\textsubscript{adj} via equation \ref{stefb1} with $\overline{r}$ at T\textsubscript{calib}.
+Table \ref{nonuniformsen} shows $\overline{r}$ and T\textsubscript{bias} and for a range of T\textsubscript{emiss}.
+
+\begin{table}[H]
+	\centering
+	\caption{Planck weighted mean sensor response and T\textsubscript{bias} for a suite of common urban T\textsubscript{surf}.}
+	\label{nonuniformsen}
+	\begin{tabular}{lp{2cm}c}
+		\toprule 
+		Emission & & T\textsubscript{bias}, \si{\kelvin} \\
+		temperature, \si{\kelvin}& $\overline{r}$, \% & computed for T\textsubscript{calib} = 300 \si{\kelvin} \\ \midrule
+		%Emission temperature, \si{\kelvin}& $\overline{r}$, \% & T\textsubscript{adj}, \si{\kelvin} \\ \midrule
+		260 & 48.3 & -1.67 \\ 
+		265 & 48.5 & -1.49 \\ 
+		270 & 48.7 & -1.32 \\ 
+		275 & 48.9 & -1.13 \\ 
+		280 & 49.1 & -0.95 \\ 
+		285 & 49.3 & -0.75 \\ 
+		290 & 49.4 & -0.56 \\ 
+		295 & 49.6 & -0.36 \\ 
+		300 & 49.7 & -0.15 \\ 
+		305 & 50.0 & 0.05 \\ 
+		310 & 50.1 & 0.26  \\ 
+		315 & 50.3 & 0.47 \\ 
+		320 & 50.4 & 0.69  \\
+		\bottomrule
+	\end{tabular} 
+\end{table}
+
+The magnitude of T\textsubscript{bias} increases as the difference between the reference T\textsubscript{calib} and T\textsubscript{emiss} increases, with approximately equal rates of change with increasing and decreasing T\textsubscript{emiss}.
+Thus, dome effects have a larger effect on the pyrgeometer received signal when comparing irradiances or T\textsubscript{hem} over a wide range of temperatures, or when T\textsubscript{emiss} is significantly different from T\textsubscript{calib}.
+As pyrgeometers used in the BUBBLE campaign were calibrated outdoors under clear-sky, mid-latitude springtime conditions \cite{Christen2005}, the dome effect present in uncorrected BUBBLE $L_z$ measurements is likely smaller in summer than in winter, and is well represented by Table \ref{nonuniformsen}.
+
+By convolving modeled spectral radiances by a dome transmittance curve, the correction method presented in this study accounts for dome effects.
+Thus, correction magnitudes include both atmospheric and dome effects in aggregate (and, crudely, emissivity effects, as modeled radiances are calculated with $\epsilon$ = 0.95 and include a reflected L\textsubscript{sky} component).
+While the individual effects are difficult to separate, temperature dependent differential dome effects are visible in correction magnitudes as a persistent minor positive correction in winter.
+Relative to T\textsubscript{calib}, colder T\textsubscript{emiss} in winter results in a larger dome effects.
+This manifests in artificially low T\textsubscript{hem, b} and increased wintertime correction magnitudes.
+Additionally, modeled irradiances include a reflected downwelling component, which, although small, is sensitive to T\textsubscript{air} and humidity profiles above the sensor height.
+Because this data was not consistently measured, mid-latitude summer and winter standard atmospheric profiles were supplemented.
+The climate of Basel is characterized by mild, humid winters (K\"oppen classification: Cfb).
+As a result, above-sensor T\textsubscript{air} and humidity values are likely underestimated by the standard wintertime atmosphere, reducing the modeled at-sensor reflected component of downwelling longwave in winter and the overall modeled at-sensor signal.
+This results in an increase in both derived T\textsubscript{hem, r} and correction magnitudes as measured irradiances are matched with irradiances modeled at slightly higher T\textsubscript{surf} to make up for the underestimated at-sensor L\textsubscript{sky} signal.
+
+%Differential dome effects are visible in atmospheric corrections as a persistent minor positive correction in winter.
+During winter T\textsubscript{arg} This has the effect of positively weighting spectral bands with less energy and negatively weighting bands with more energy, resulting in a small systemic underestimation of the broadband signal which is removed during instrument calibration.
+However, calibration routines are specific to a range of T\textsubscript{surf} and the effect of the sensor's non-uniform spectral response changes based on emitter temperature.
+Lower emitter temperatures shift the peak of the Planckian curve closer to bands with low relative sensor response.
+Thus, T\textsubscript{bright} inverted from pyrgeometer measured irradiance via \ref{stefb1} is systemically lower than those derived using a routine that accounts for nonuniform sensor response.
+In other words, because a pyrgeometer does not respond uniformly over a spectral longwave curve and operates over a slightly restricted longwave bandpass, T\textsubscript{bright} will always include some small underestimation of T\textsubscript{rad}, even when atmospheric influence is negligible.
+The size of this underestimation is inversely related to T\textsubscript{surf} - increasing the base magnitude of wintertime corrections compared to summertime corrections.
+Additionally, modeled irradiances include a reflected downwelling component, which, although small, is sensitive to T\textsubscript{air} and humidity profiles above the sensor height.
+Because this data was not consistently measured, mid-latitude summer and winter standard atmospheric profiles were supplemented.
+Basel's climate is characterized by mild, humid winters (K\"oppen classification: Cfb).
+As a result, above-sensor profiles of T\textsubscript{air} and humidity at study site are likely underestimated by the standard wintertime atmosphere, reducing the modeled at-sensor reflected component of downwelling longwave in winter.
+This results in an increase in both derived T\textsubscript{hem, r} and correction magnitudes.
+
+\subsection{The effect of sensor sampling geometry on remote sensed T\textsubscript{surf}}
+
+Results show that sensor sampling regimes have a significant effect on remotely sensed T\textsubscript{surf}.
+Both T\textsubscript{hem, r} and T\textsubscript{plan} overestimate T\textsubscript{comp} by day and underestimate T\textsubscript{comp} by night.
+However, over/underestimations from geometric effects are highly dependent on synoptic conditions, particularly for a nadir view of the urban surface.
+Overestimation of T\textsubscript{comp} under clear sky conditions is much larger for T\textsubscript{plan} than T\textsubscript{hem, r}.
+Representing the surface from the nadir, over half of the T\textsubscript{plan} signal from the Sperrstrasse canyon is generated by rooftops with both a large diurnal T\textsubscript{surf} amplitude and the highest daytime facet T\textsubscript{max}.
+Undersampling of wall facets by T\textsubscript{plan} also has a significant effect on the daytime clear-sky warming bias, as walls show much cooler daytime T\textsubscript{max} than T\textsubscript{roof} or T\textsubscript{road} and have a moderating effect on T\textsubscript{comp} and T\textsubscript{hem, r} by day, particularly on hot clear sky days.
+Mean daytime overestimations by T\textsubscript{plan} just after solar noon are approximately $4$ \si{K} over the IOP, approaching $8$ \si{K} on the clear sky case day.
+T\textsubscript{hem, r} shows smaller overestimations of approximately $3$ \si{K} over the IOP and $5$ \si{K} on the clear sky case day.
+Urban T\textsubscript{surf} and the sUHI are often measured from a satellite TIR remote sensors that retrieve T\textsubscript{plan} exclusively under clear-sky conditions and often sample in the nadir.
+Thus, overestimation of daytime urban T\textsubscript{comp} inherent in the satellite T\textsubscript{surf} record is best represented by T\textsubscript{plan} on the clear sky case day.
+
+Nighttime underestimations of T\textsubscript{comp} by nadir and hemispherical views are a result of an undersampling of wall and road facets respectively.
+T\textsubscript{hem, r}, derived from a pyrgeometer skewed towards the south-facing canyon wall oversamples the nearest roof and underestimates the canyon floor compared to T\textsubscript{comp}.
+This results in a large underestimation of nighttime T\textsubscript{comp}.
+However, nighttime underestimation - and to some extent, daytime overestimation - by T\textsubscript{hem, r} is the combined result of a sampling bias inherent in a hemispherical view of the surface and improper sensor placement, discussed in greater detail in section \ref{Methodological limitations and considerations}.
+Thus, both of these bias can likely be reduced significantly by adjusting sensor placement to best represent surface geometry.
+
+\section{Sensor placement sensitivity testing} \label{Methodological limitations and considerations}
+
+ When measured from a near ground downward facing remote sensor, upwelling TIR is sensitive to sensor height and position relative to surface geometry \cite{Adderley2015, Roberts2010}.
+Below 3 times mean roof level, remote sensed TIR varies from a 'true', geometrically representative, TIR signal based on a comparison between view factor proportions of and those of the urban surface.
+For an idealized, orthogonal, block-like urban array, Roberts 2010 found that facet view factor proportions - and thus measured TIR - varied significantly with sensor position and height.
+To investigate the effect of sensor placement on T\textsubscript{hem, r} for a simplified representation of the Sperrstrasse canyon site, we used SUM to vary sensor position over a simplified DBM and calculated normalized view factor proportions for a 160\si{\degree} FOV sensor.
+For nine sensor positions, show in in Figure \ref{placement}, facet view factor proportions normalized to unity are shown in Figure \ref{bars1} for a sensor height of 2.17 times mean building height and in Figure \ref{bars45} for a sensor height of 3 times mean building height.
+Approximated actual facet surface area proportions for the Sperrstrasse canyon are included for reference.
+A fully geometrically representative sensor viewing the urban surface would have facet view factor proportions equal to actual facet surface area proportions.
+
+ Normalized facet view factors show significant variation based on sensor position at both test heights.
+At the lower test height, in the horizontal direction (moving perpendicular to the canyon axis), sensor positions at the canyon center (positions 4, 5, and 6) are most representative of actual facet view factors, with decreasing representivity at the edge of the canyon (positions 1, 2, and 3), and over the rooftops (positions 7, 8, and 9).
+In the vertical direction (moving parallel to the canyon axis), positions slightly offset from the roof edge (positions 2, 5, and 8) agree most closely with actual canyon surface area proportions, with decreasing representivity in positions further from the roof edge.
+Roof and road view factor proportions are most strongly affected by sensor placement.
+Positions directly over buildings show large biases towards roof view factors and positions in the center of the canyon showing slight biases towards road facets.
+In all positions roof view factor was overestimated (save position 4), and wall view factor was underestimated.
+
+ \begin{figure}[H]
+	\centering
+	\includegraphics[width=15cm,
+	height=7in,
+	keepaspectratio]{placement1}
+	\caption{A plan view of the simplified DBM showing the nine test sensor placements. Actual sensor placement at the Sperrstrasse canyon is approximated by number 3.}
+	\label{placement}
+\end{figure}
+
+ \begin{figure}[H]
+	\centering
+	\includegraphics[width=15cm,
+	height=7in,
+	keepaspectratio]{bars2_ann}
+	\caption{Normalized wall, road, and roof view factors for nine sensor positions viewing the simplified street canyon array from 2.17 times mean building height.
+Actual normalized view factors refer to surface area proportions for the three facet types in the Sperrstrasse street canyon.
+}
+	\label{bars1}
+ \end{figure}
+
+Results in Roberts 2010 and Adderley et al.
+2015 show that sensitivity to sensor placement decreases with increasing sensor height  Above a height of approximately 3 to 5 times mean building height, depending on surface geometry, a TIR signal becomes spatially invariant and view factor proportions are constant across all sensor placements.
+Results from the higher test height support findings in Roberts 2010 and Adderley et al.
+2015, with decreased spatial variance in facet view factor proportions across the nine sensor positions.
+However increased spatial stability in view factor proportions does not necessarily entail increased geometric representativity.
+For all sensor placements, road view factor proportion is increased and wall view factor is reduced.
+This results in large overestimations of road view factor for nearly all positions except positions 3 and 9, which include a large overestimation of roof view factor.
+Thus, although increasing sensor height does increase spatial invariance, it can lead to decreased representivity for some positions - in this case, positions 1 and 4.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=15cm,
+	height=7in,
+	keepaspectratio]{bars45_ann}
+	\caption{Normalized wall, road, and roof view factors for nine sensor positions viewing the simplified street canyon array from 3 times mean building height.
+Actual normalized view factors refer to surface area proportions for the three facet types in the Sperrstrasse street canyon.
+}
+	\label{bars45}
+\end{figure}
+
+To visualize how sensor placement affects inferred surface temperatures, T\textsubscript{hem, r} is calculated using view factors as weights in an average of T\textsubscript{wall}, T\textsubscript{road}, and T\textsubscript{roof}.
+Temperatures are normalized by T\textsubscript{comp} and averaged at 30-minute intervals over the IOP, the results of which are shown in figures \ref{vf_all} and \ref{vf_all45} for sensor heights of 2.17 and 3 times mean building height respectively.
+As with view factor proportions, calculated T\textsubscript{hem, r} is highly dependent on relative to T\textsubscript{comp}.
+For the lower sensor height, as predicted by view factor proportions, positions at the center of the canyon (positions 4, 5, and 6) and edge of the building (positions 1, 4, and 7) follow most closely to T\textsubscript{comp}.
+By day, all sensor positions overestimate T\textsubscript{comp}, from a systemic bias towards rooftop facets.
+The degree of overestimation is largely determined by rooftop view factor, with larger overestimations coming from positions most biased towards hot rooftop surfaces.
+At night, most sensor positions (save position 4) underestimate T\textsubscript{comp}, again from a bias towards cool rooftop facets.
+Position 4 weakly overestimates nighttime T\textsubscript{comp}, from a slight view factor bias towards road facets with suppressed nocturnal cooling from canyon radiation trapping.
+
+At the higher test height, T\textsubscript{hem, r} is more representative of T\textsubscript{comp} for positions 2, 3, and 5 - 9.
+T\textsubscript{hem, r} from positions 1 and 4 displays a slight systemic overestimation of T\textsubscript{comp}.
+For these positions, although roof view factor proportion is well represented, road view factor is significantly overestimated, oversampling warmer nighttime road temperatures.
+For all other positions daytime overestimation and nighttime underestimation of T\textsubscript{comp} is reduced.
+As is the case with view factor proportions, T\textsubscript{hem, r} is more spatially stable based on sensor placement.
+
+   \begin{figure}[H]
+ 	\centering
+ 	\includegraphics[width=15cm,
+ 	height=7in,
+ 	keepaspectratio]{vf_all}
+ 	\caption{Mean T\textsubscript{hem, r} normalized against T\textsubscript{comp} for each sensor position the 14-day IOP. Shaded area indicates quartiles one through three.}
+ 	\label{vf_all}
+ \end{figure}
+
+  \begin{figure}[H]
+	\centering
+	\includegraphics[width=15cm,
+	height=7in,
+	keepaspectratio]{vf_all45}
+	\caption{Mean T\textsubscript{hem, r} normalized against T\textsubscript{comp} for each sensor position the 14-day IOP. Shaded area indicates quartiles one through three.}
+	\label{vf_all45}
+\end{figure}
+
+Both Roberts 2010 and Adderley et al.
+2015 did not find a perfectly representative sensor position for typical low- and mid-rise urban street canyons, this is supported by sensitivity testing performed in this study.
+As such, under conditions with strong microscale spatial contrasts in T\textsubscript{surf}, T\textsubscript{hem, r} is unlikely to be equal T\textsubscript{comp} regardless of sensor placement.
+However, sensor view models, such as SUM, can be used to optimize sensor placement with information about surface geometry to retrieve the most accurate measured irradiances and derived T\textsubscript{hem, r} for a given surface geometry.
+
+ \section{Conclusions}
+
+Accurate, geometrically representative, spatially extensive, and temporally continuous thermal remote sensing of complex terrain (urban or otherwise) is difficult.
+It is unlikely that a single sensor will satisfy spatial, geometric, and temporal goals.
+Thus, a combination of satellite, aerial, and in-situ observational methods has been integral in isolating and elucidating the urban effect on surface temperature and its spatial, temporal, and geometric characteristics.
+However, neither technological advancement nor critical analysis of these methods has provided a satisfactory answer to the questions posed in Roth et al.
+1989 viz urban thermal remote sensing.
+As such, the true spatial, temporal, and geometric nature of the urban effect on surface temperature remains unknown.
+
+This paper details a  method for urban T\textsubscript{surf} retrieval using atmospherically corrected hemispherical radiometric surface temperatures derived from continuous near-ground measurements of thermal infrared radiation.
+Atmospherically corrected, hemispherical urban T\textsubscript{surf} is derived using a method which combines a sensor view model (SUM) to represent urban surface geometry and a point-to-point radiative transfer model (MODTRAN 4.1) to model irradiances upwelling from complex surface terrain in 3-dimensions.
+At each time step, irradiances are modeled using profiles of T\textsubscript{air} and water vapor content for a range of possible T\textsubscript{hem, r}.
+Irradiance - T\textsubscript{hem, r} pairings are aggregated into a LUT and the measured irradiance is matched with the closest modeled irradiance to retrieve T\textsubscript{hem, r} for a given time step.
+Repeated at 30-minute intervals, this method is used to derive an 8-month climatology of urban T\textsubscript{hem, r} for the Basel Sperrstrasse canyon, instrument as a part of the BUBBLE campaign.
+T\textsubscript{hem, r} at each time step is a view factor weighted temperature representative of the  at which the surfaces contained within the sensor FOV are emitting longwave radiation.
+
+Atmospheric corrections are large compared to typical sUHI magnitudes \cite{Peng2012} and vary significantly based on season, time of day, and synoptic conditions.
+Correction magnitudes are largest on hot, clear sky summertime days - with large solar input and high T\textsubscript{hem, r} and T\textsubscript{air} - and smallest by night and under overcast conditions, during which solar input is suppressed and T\textsubscript{hem, r} and T\textsubscript{air} are low.
+Correction magnitudes are most strongly correlated with the T\textsubscript{surf} - T\textsubscript{air} differential and incoming solar radiation and weakly correlated with T\textsubscript{hem, r} and T\textsubscript{air}.
+Water vapor content did not appear to exert strong control over correction magnitudes.
+
+Comparison of T\textsubscript{surf} from nadir, hemispherical, and complete representations of the Basel Sperrstrasse street canyon show that a hemispherical view is more geometrically representative of the complete surface temperature than a sensor viewing in the nadir.
+However, sensor placement sensitivity tests show that T\textsubscript{hem, r} is highly dependent on sensor placement and height.
+Overestimations by T\textsubscript{plan} and T\textsubscript{hem, r} are greatest under clear sky conditions, but is particularly large for T\textsubscript{plan}.
+This is particularly important when one considers that satellite observations of urban T\textsubscript{surf} are only possible under clear sky conditions (when T\textsubscript{plan} is least representative of T\textsubscript{comp}).
+Significant variability in over/underestimation of T\textsubscript{comp} by T\textsubscript{hem, r} and T\textsubscript{plan} based on time of day and synoptic conditions make these biases difficult to generalize to other urban study sites with different surface geometries and characteristics, canyon orientations, and macro climates.
+However, the measures used to derive T\textsubscript{hem, r} are common to most energy balance assessments and constitute an untapped, but promising, resource to improve understanding of how the built environment affects land T\textsubscript{surf} and to quantify the geometric and temporal biases inherent in satellite TIR remote sensing of complex terrain.

### 2017-04-11 paper2.tex
Add files via upload

@@
 The temperature of the surface is integral in understanding, predicting, and modeling boundary-layer air temperature patterns, surface energy balances, and, in urban areas, has important implications for human thermal comfort and building energy usage.
-Urban modification of surface geometry and thermal, radiative, moisture, and aerodynamic properties results in differential surface heating and cooling patterns and strong microscale spatiotemporal variations in urban surface temperature.
-Integrated up to larger scales, urban areas tend to store and generate more heat relative to non-built surroundings and, as a result, are generally warmer --- a phenomenon termed the urban heat island effect (UHI).
-As such, accurate, spatiotemporally continuous and geometrically representative characterization of surface temperatures has long been a goal in urban climatology.
-The proliferation of satellite and aerial thermal infrared (TIR) remote sensing has enabled spatially-extensive, large-scale characterizations of surface climates at ever improving spatial and spectral resolutions.
-Such campaigns have elucidated urban surface temperature (T\textsubscript{surf}) and surface urban heat island (sUHI) patterns globally at large spatial scales.
+Urban modification of surface geometry and thermal, radiative, moisture, and aerodynamic properties results in differential surface heating and cooling patterns and strong microscale spatiotemporal variations in urban surface temperature (T\textsubscript{surf}).
+Integrated up to larger scales, urban areas tend to store and generate more heat relative to non-built surroundings which manifests in elevated T\textsubscript{surf} and T\textsubscript{air} --- a phenomenon termed the urban heat island effect (UHI).
+To foster a more complete understanding of the effect of urban areas climates across scales, accurate, spatiotemporally continuous and geometrically representative characterization of T\textsubscript{surf} in cities has long been a goal in urban climatology.
+The proliferation of satellite and aerial thermal infrared (TIR) remote sensing has enabled spatially-extensive characterizations of surface climates at ever improving spatial and spectral resolutions.
+Such campaigns have elucidated urban T\textsubscript{surf} and surface urban heat island (sUHI) patterns globally at large spatial scales.
 However, technological improvements in TIR remote sensing have yet to address a three potential sources of error when applied in urban areas:
@@
 
-These biases present a potentially significant source of error by failing to capture micro-scale temporal and geometric variations in urban T\textsubscript{surf}.
+\noindent These biases present a potentially significant source of error by failing to capture micro-scale temporal and geometric variations in urban T\textsubscript{surf} and sUHI.
 
-Inter-site comparison is the crux of UHI analysis, thus it is imperative that urban surface temperature measurements are accurate and representative of coherent urban patches.
-Meta-analysis of air temperature UHI literature shows that these goals are not often not satisfied \cite{Stewart2011}.
-Given the relative difficulty in retrieving accurate, representative urban surface temperatures, similar conclusions are likely for sUHI analysis.
-In spite of this fact, and the short period over which large-scale generalizable methods for urban surface temperature acquisition have been available, study of sUHI has expanded significantly in the last twenty years \cite{Peng2012,Voogt2003}.
-This paper presents a method to derive hemispherical radiometric urban T\textsubscript{surf} (Them) from hemispherical upwelling TIR as measured by inverted pyrgeometers.
-This method was developed to addresses biases inherent in traditional methods for urban T\textsubscript{surf} retrieval by providing temporally continuous, geometrically representative\footnote{Them is not perfectly representative of urban geometry.
+Inter-site comparison is the crux of UHI analysis, thus it is imperative that urban T\textsubscript{surf} measurements are representative of coherent urban patches and free from confounding influences (i.e.
+atmospheric and emissivity effects).
+Meta-analysis of air temperature (T\textsubscript{air}) UHI literature shows that these goals are rarely satisfied \cite{Stewart2011}.
+Given the relative difficulty in retrieving accurate, representative urban T\textsubscript{surf}, similar conclusions are likely for sUHI analysis.
+In spite of this fact, and the short period over which large-scale, generalizable methods for measuring urban T\textsubscript{surf} have been available, study of sUHI via TIR remote sensing has expanded significantly in the last twenty years \cite{Peng2012,Voogt2003}.
+Using the method described in \ref{paper1} we derive an 8-month climatology of hemispherical radiometric urban T\textsubscript{surf} (T\textsubscript{hem, r}) from time continuous near-ground upwelling TIR as measured from an inverted pyrgeometer.
+The method was developed to address and overcome biases inherent in traditional methods for urban T\textsubscript{surf} retrieval by providing temporally continuous, geometrically representative\footnote{A hemispherical view is not perfectly geometrically representative of urban surface geometry.
 This is best illustrated by visualizing an urban area from the perspective of a downward facing ‘fish-eye’ camera.
-However, it is undoubtedly more representative of urban geometry than most 2-dimensional views of the surface.
-This fact is discussed in more depth in sectionXXXX.}  urban T\textsubscript{surf} under all-sky conditions for sUHI analysis.
+Geometric sampling biases are a result of lens distortions and the sensor cosine response.
+However, by sampling the surface 3-dimensionally, T\textsubscript{hem, r} is more representative of urban geometry than conventional 2-dimensional views of the surface.
+The geometric representivity of urban T\textsubscript{hem, r} is discussed in section \ref{Methodological limitations and considerations}.} urban T\textsubscript{surf} under all-sky conditions for sUHI analysis.
 These measurements are often made as a part of the net radiation determination for urban energy balance studies.
 
-\section{Bias in Thermal remote sensing}
+\subsection{Bias in thermal remote sensing}
 Geometric biases in remote sensing of the urban surface are a result of its 3-dimensional, convoluted structure.
 Compared to flat terrain, complex urban surface geometry modifies receipt of incoming solar radiation and traps a portion of reflected solar and outgoing terrestrial radiation.
-Differential heating and shading patterns of vertical, sloped, and horizontal urban facets manifests in significant micro-scale contrasts in surface temperature.
-As such, measured radiometric surface temperature varies based on sensor field-of-view, viewing angle and direction, and sun-surface geometry – this directional dependence of urban surface temperature is termed ‘effective thermal anisotropy’ \cite{Voogt1998a}.
-Traditional satellite or airborne remote sensing platforms, by viewing the surface in the nadir, sample only a fraction of the complete urban surface and fail to capture this effect – leading to directional biases in measured urban surface temperature.
-Geometric undersampling by a remote sensor in the nadir manifests in an overestimation of daytime temperatures and underestimation of nighttime temperatures \cite{Adderley2015}.
-However, the magnitude and diurnal pattern of this bias is dependent on urban patch characteristics (canyon height-to-width ratio, canyon orientation and materials, vegetation coverage, etc.) and sensor viewing angle.
-Thus, parameterization schemes to account for urban effective anisotropy are difficult to generalize across urban sites, sensor types, and sensor orientations.
+The resulting microscale spatiotemporal contrasts in urban T\textsubscript{surf} create a directional dependence in observed urban T\textsubscript{surf} when measured from conventional narrow-field-of-view (FOV) TIR remote sensing platforms.
+Thus, remote sensed urban T\textsubscript{surf} varies based on sensor FOV, viewing angle and direction, and sun-surface geometry - this directional dependence of urban surface temperature is termed "effective thermal anisotropy" \cite{Voogt1998a}.
+Traditional satellite or airborne remote sensing platforms, by viewing the surface in the nadir, sample only a fraction of the complete urban surface and fail to capture this effect - leading to spatiotemporally variant directional biases of up to 10 \si{\kelvin} in observed urban T\textsubscript{surf} \cite{Voogt1995}.
+In general, geometric undersampling by a remote sensor in the nadir manifests as an overestimation of daytime T\textsubscript{surf} and an underestimation of nighttime T\textsubscript{surf} \cite{Adderley2015}.
+However, the magnitude and diurnal and seasonal behaviors of this bias are dependent on sensor viewing geometry and unique site characteristics (canyon height-to-width ratio, canyon orientation and materials, vegetation coverage, etc.).
+Thus, parameterization schemes to account for urban effective anisotropy are difficult to generalize across urban sites, sensor types, and sensor-surface geometries.
 
-In addition to undersampling the urban surface, most thermal remote sensing platforms yield an instantaneous ‘snap-shot’ and cannot characterize temporally continuous T\textsubscript{surf} patterns.
-Temporal discontinuities in thermal remote sensing result in a number of sources of bias operating over a wide range of time scales.
+In addition to undersampling the urban surface, most thermal remote sensing platforms yield an instantaneous ‘snap-shot’ and cannot characterize time-continuous T\textsubscript{surf} patterns without sacrificing ground resolution.
+Temporal discontinuities in thermal remote sensing result in myriad of potential sources of bias over a wide range of time scales.
 Aerial and satellite thermal remote sensing require clear sky conditions (clouds are opaque with respect to thermal infrared radiation).
-Hence, long term satellite characterizations of sUHI are biased towards conditions that maximize macro-scale urban-rural contrasts in surface temperature.
+Hence, long term satellite characterizations of sUHI are biased towards conditions that maximize urban-rural contrasts in T\textsubscript{surf}.
 This results in an overestimation of “all-sky” sUHI.
-Although
-
-Over a day, sUHI is generally largest in the late afternoon, near solar noon, and just after sunset.
-Continuous sUHI analysis indicates that it exhibits a large diurnal amplitude.
-Satellite overpass cycles rarely coincide with sUHI maximums and are not standard across cities or platforms.
+At diurnal scales, satellite overpass cycles rarely coincide with sUHI maximums and are not standard across cities or platforms.
 Thus, analysis of sUHI patterns and magnitudes across cities and instrument platforms is difficult.
-High-frequency fluctuations in surface temperature add an additional – and understudied – bias to thermal remote sensing of urban surface temperature.
-Time series analysis of urban surface temperature shows significant microscale (second to minute) fluctuations in temperature \cite{Christen2012}.
-Most thermal remote sensors observe surface temperatures instantaneously (rather than temporally averaged) and are potentially contaminated by microscale fluctuations.
-This is particularly salient in urban environments, where a large variety of fabric materials can produce significant directional contrasts in thermal admittance – and thus spatial variations in the magnitude of microscale fluctuations depending on the facet material types viewed by the sensor.
-The effect of this phenomenon on thermal remote sensing has not been extensively studied, however, the magnitude of microscale fluctuations in surface temperature is significant relative to a typical sUHI signal and thus constitutes a potentially large source of bias.
+At smaller time scales still, time-continuous analysis of urban T\textsubscript{surf} shows significant microscale (second to minute) fluctuations in temperature \cite{Christen2012}.
+Most thermal remote sensors provide instantaneous T\textsubscript{surf} (rather than temporally averaged) and are potentially contaminated by high-frequency microscale fluctuations in urban T\textsubscript{surf}.
+This is particularly salient in urban environments, where a large variety of fabric materials can produce significant directional contrasts in thermal admittance - and thus spatial variations in the magnitude of microscale fluctuations depending on the facet material types viewed by the sensor.
+The effect of this phenomenon on thermal remote sensing has not been extensively studied, however, the magnitude of microscale fluctuations in T\textsubscript{surf} is significant relative to a typical sUHI signal and thus constitutes a potentially large source of bias.
 
 Both geometric and temporal shortcomings limit the representativity of traditional remote sensed evaluations of urban T\textsubscript{surf} and sUHI.
-The magnitude of these biases has not been extensively studied in climatological form.
-This paper presents a method to both assess the magnitude of and overcome these biases.
+The magnitude of these biases has not been extensively studied, particularly from a long term, climatological perspective.
+This study presents the first time-continuous, climatological analysis of sUHI retrieved from radiometric hemispherical urban T\textsubscript{surf} retrieved via the correction method described in Chapter \ref{paper1}.
+These measures are used to assess the magnitude of and overcome geometric and temporal biases inherent in urban TIR remote sensing.
+
+\section{Methods}
+
+This section describes the study area and provides a brief overview of the multiple line-of-sight atmospheric correction method used to derive a time-continuous eight month climatology urban T\textsubscript{hem, r} for sUHI analysis.
+A thorough discussion of the method is included in Chapter \ref{paper1}.
+
+\subsection{A method to retrieve hemispherical radiometric urban T\textsubscript{surf}}
+
+Remote sensing of TIR is subject to atmospheric influence from gaseous and aerosol absorbers between the surface and the sensor.
+A pyrgeometer's broad waveband and wide FOV increase the potential for significant atmospheric influence on an observed TIR signal.
+These effects can result in differences of up to 8 \si{\kelvin} between the 'true' radiometric T\textsubscript{surf} and brightness T\textsubscript{surf} derived from remote sensed observations of upwelling TIR.
+Atmospheric effects are particularly important when comparing irradiances or remote sensed T\textsubscript{surf} across different study sites and times, as differences in surface geometry, instrument height, and ambient conditions can introduce large spatiotemporal contrasts in the magnitude of atmospheric influence.
+This is particularly salient for near-ground, broadband radiometers, as band by band spectral atmospheric transmittance can change significantly with path length, shown in Figure \ref{spectransheight}.
+As inter-site comparison is paramount in understanding the urban effect on T\textsubscript{surf} and is the basis of sUHI analysis, accurate atmospheric correction of remote sensed T\textsubscript{surf} is of prime importance.
+Thus, urban T\textsubscript{hem, r} in this study are derived using a multiple line-of-sight correction routine which accounts for the 3-dimensionality of the urban surface, changing atmospheric conditions, and spectrally non-uniform sensor response.
+For each time step, the method generates of modeled at-sensor irradiances for a range of potential T\textsubscript{hem, r} using profiles of measured T\textsubscript{air} and humidity.
+A corrected T\textsubscript{hem, r} for the target time step is then retrieved by matching the measured irradiance value to the closed modeled irradiance - T\textsubscript{hem, r} pairing.
+The work flow, described in Figure \ref{flow}, is repeated at 30-minute intervals over the eight month study period to retrieve a climatology of urban T\textsubscript{hem, r} for sUHI analysis.
+
+First, surface-to-sensor path length and view factor geometries are calculated with the surface-sensor-sun urban model (SUM) \cite{Soux2004}.
+Using a digital building model, SUM calculates path lengths from the surface to the sensor and angular view factors by projecting the pyrgeometer FOV onto a simplified (orthogonal) representation of the surface, determining which points in the DBM are 'seen' by the sensor, and calculating the distance from each 'seen' point to the sensor.
+Path lengths are binned at 5\si{\degree} intervals and averaged over the azimuth angle.
+Finally, view factors are calculated for each 5\si{\degree} angular as each bin's proportion of the total view factor and normalized to unity.
+
+Second, version 4.1 of the MODTRAN radiative transfer code \cite{Berk1987} is used to model spectral at-sensor radiances for each angular path length.
+Runs are initialized using profiles of T\textsubscript{air} and water vapor content collected concurrently at each time step.
+Spectral radiances are modeled over a bandpass of 1 to 2500 \si{cm^{-1}} at an emissivity of 0.95 and convolved by an extended dome transmittance function to ensure that modeled radiances replicate the actual remote sensed signal.
+Consultation with radiometer manufacturers made clear the need to extend the modeled bandpass beyond a typical longwave bandpass (approximately 250 - 2500 \si{cm^{-1}}), as silicone domed pyrgeometers are transmissive of radiation in much shorter wavenumber (longer wavelengths) than 250 \si{cm^{-1}} (40 \si{\micro\meter}).
+
+Third, radiances for each angular path length are integrated over the waveband and multiplied their respective normalized view factors.
+Weighted and corrected angular radiances are integrated over the hemisphere to yield a modeled at-sensor irradiance as ‘seen’ by the pyrgeometer for the target T\textsubscript{hem, r}.
+
+For each time step, the above steps are repeated at an interval of 0.5 \si{\kelvin} for a predefined range of potential T\textsubscript{hem, r}, with results aggregated into a lookup table relating modeled irradiances to corrected T\textsubscript{hem, r} for the observed ambient T\textsubscript{air} and humidities.
+The measured irradiance is then matched with the closest modeled irradiance to return the associated corrected T\textsubscript{hem, r} for the target time step.
+sUHI magnitudes for the study area are calculated as the difference between urban T\textsubscript{hem, r} and rural T\textsubscript{hem, b} with rural T\textsubscript{hem, b} calculated via
+
+\begin{equation}
+\label{ruralt}
+T_{hem,~ b} = \sqrt[4]{\frac{\epsilon L_z^\uparrow + (1 - \epsilon)L_{sky}^\downarrow}{\sigma}}
+\end{equation}.
+
+\noindent where $L_z^\uparrow$ is at-sensor upwelling longwave radiation, $L_{sky}^\downarrow$ is at-sensor reflected downwelling longwave radiation, and emissivity ($\epsilon$) = 0.98.
+
+Sensitivity tests included in Section \ref{sensitivity} of the companion paper show that screen level irradiances ($ z $ = ~2 \si{\meter}) are not subject to significant effects from the intervening atmosphere and do not require atmospheric correction.
+Rural irradiances in this study were measured from approximately 2 \si{\meter} above ground, thus rural T\textsubscript{hem, b} is approximately equal to T\textsubscript{hem, r}.
+
+\subsection{Study area}
+
+T\textsubscript{hem, r} are retrieved from radiation and meteorological data collected from December 2001 through July 2002 as a part of the Basel Urban Boundary Layer Experiment (BUBBLE) campaign conducted in Basel, Switzerland \cite{Rotach2005}.
+Urban data were observed from a tower located near city center at the "Basel Sperrstrasse" site.
+Rural reference data were observed at the "Lange Erlen" site approximately 6 \si{\kilo \meter} to the north east in the outskirts of the city of Basel.
+Urban and rural site morphologies and observed variables are described in Table \ref{morphbspr}.
+
+\begin{table}[H]
+	\centering
+	\caption{A description of morphological and measured variables for urban and rural sites. Modified from Rotach et al., 2005 to include only relevant parameters \cite{Rotach2005}.}
+	\label{morphbspr}
+	\begin{tabular*}{\textwidth}{p{3.75cm} p{2.25cm}p{3.5cm}p{2.75cm}p{2.75cm}}
+		\toprule 
+		Site & Location & Morphological & Meteorological & Radiation \\ 
+		& Height & Characteristics\footnote{} & Variables\footnote{} & [No of levels]\footnote{} \\ 	\midrule
+
+		Basel Sperrstrasse \newline \textit{Urban} \textit{street canyon} \newline LCZ: 2 & 47.57\si{\degree} N \newline 7.60\si{\degree} E \newline 255 \si{\meter} a.s.l.
+& $z_H$ = 14.6  \si{\meter} \newline $\sigma_H $ = 6.9 \si{\meter} \newline H/W = 0.54 \newline $\lambda_C $ = 0.37\newline $\lambda_S$ = 1.0 \newline $\alpha$ = 11.0\% & T\textsubscript{air} [7] \newline H [7] \newline WV [12] \newline WD [1] \newline P [1] & L\textsubscript{up} [3] \newline L\textsubscript{down} [5] \newline S\textsubscript{up} [2] \newline S\textsubscript{down} [3] \\
+		& & & & \\
+		Lange Erlen \newline \textit{Rural parkland} \newline LCZ: B & 47.59\si{\degree} N \newline 7.65\si{\degree} E \newline 240 \si{\meter} a.s.l.
+& $\alpha$ = 21.4\%  & T\textsubscript{air} [4] \newline H [4] \newline WV [3] \newline WD [1] &  L\textsubscript{up} [1] \newline L\textsubscript{down} [1] \newline S\textsubscript{up} [1] \newline S\textsubscript{down} [1]  \\
+		\bottomrule
+	\end{tabular*} 
+		\raggedright
+		\textsuperscript{2} Morphological parameters for the Sperrstrasse site were calculated for a 250 \si{\meter} circular area surrounding the study sites using the method described in \cite{Grimmond1999}.
+$z_H$: average building height, $\sigma_H $: standard deviation of building height, $\lambda_P $: plan aspect ratio, $\lambda_C $: complete aspect ratio, H/W: local canyon height to width ratio, $\alpha$: surface albedo.
+\\
+		\textsuperscript{3} H: humidity, WV: wind velocity, WD: wind direction, P: pressure. \\
+		\textsuperscript{4} L\textsubscript{up}: upwelling longwave radiation, L\textsubscript{down}: downwelling longwave radiation, S\textsubscript{up}: upwelling shortwave radiation, S\textsubscript{down}: downwelling shortwave radiation.
+\end{table}
+
+During an intensive observation period (IOP) in late June/early July 2002, the Sperrstrasse canyon was instrumented with an array of infrared thermometers (IRT) to view individual canyon facet T\textsubscript{surf} to sample representative road, roof, and wall surface temperatures (T\textsubscript{road}, T\textsubscript{roof}, and T\textsubscript{wall} respectively).
+Plan and complete aspect ratios for the Sperrstrasse canyon are used to compute weighting schemes to retrieve complete and plan surface temperatures (T\textsubscript{comp} and T\textsubscript{plan} respectively) at 30 minute intervals over the IOP.
+T\textsubscript{comp} represents an area weighted average complete urban T\textsubscript{surf}, while T\textsubscript{plan} represents the Sperrstrasse site as viewed by a satellite or aerial remote sensor in the nadir.
+Weighting schemes for T\textsubscript{plan} and T\textsubscript{comp} are included in Table \ref{weightings}.
+Urban T\textsubscript{surf} derived from complete, plan, and hemispherical representations of the Sperrstrasse site are used to investigate the effect of sensor-surface geometry on remote sensed T\textsubscript{surf} and sUHI, and to quantify the magnitude of geometric biases time-continuously over a range of synoptic conditions.
+For comparison of sUHI to air temperature UHI, canopy layer air temperature UHI (clUHI) magnitudes are calculated as the difference between urban and rural T\textsubscript{air} measured from 2 \si{\meter} above ground level.
+
+\section{Results}
+
+\subsection{Seasonality in sUHI magnitudes}
+
+Figure \ref{heatsuhi} shows seasonal variance in diurnal patterns of mean monthly hemispherical sUHI magnitudes at 30 minute intervals for the eight month study period.
+sUHI displays significant seasonal variance, particularly by day, and is stronly controlled by day length and solar angle.
+Nighttime sUHI does not vary significantly across seasons.
+
+Figure \ref{heatcluhi} shows seasonal variance in diurnal patterns of mean monthly clUHI calculated at 30 minute intervals for the eight month study period.
+clUHI development is strongly controlled by the timing of sunset/sunrise cycles and displays the largest seasonal variance in the hours immediately after sunset.
+Daytime clUHI does not vary significantly across seasons.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=15cm,height=7in,keepaspectratio]{heatsuhi}
+	\caption{A heatmap of mean half hourly hemispherical sUHI for each month calculated at 30-minute intervals over the eight month study period.}
+	\label{heatsuhi}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=15cm,height=7in,keepaspectratio]{heatatuhi}
+	\caption{A heatmap of mean half hourly canopy layer UHI for each month calculated at 30-minute intervals from T\textsubscript{air} measured at approximately 2 \si{\meter} above ground level over the eight month study period.}
+	\label{heatcluhi}
+\end{figure}
+
+\subsection{The effect of sensor-surface geometry on sUHI}
+
+Figure \ref{bx_suhi_compare} shows sUHI magnitudes calculated from urban T\textsubscript{comp}, T\textsubscript{plan}, and T\textsubscript{hem}.
+Compared to complete sUHI; nadir and hemispherical views of the surface overestimate sUHI by day and underestimate sUHI by night.
+sUHI from a nadir view shows the greatest diurnal variance in sUHI, particularly under clear sky 'satellite friendly' conditions.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=15cm,height=7in,keepaspectratio]{bx_suhi_compare}
+	\caption{A comparison of sUHI magnitudes from complete, hemispherical, and nadir remote sensed representations of the Sperrstrasse canyon over the IOP.
+Each plot includes mean sUHI, as well case days representing sUHI under clear sky and clouded sky conditions.}
+	\label{bx_suhi_compare}
+\end{figure}
+
+\subsection{The effect of meteorological conditions on sUHI}
+
+Figures \ref{meteo_kdown}, \ref{meteo_hum}, \ref{meteo_wv}, and \ref{meteo_atuhi} show sUHI magnitudes as a function of incoming solar radiation, water vapor content, wind velocity, and clUHI.
+Winter hours are omitted in Figures \ref{meteo_hum}, and \ref{meteo_wv}, and \ref{meteo_atuhi} as sUHI is small and displays minimal variation in winter.
+Both winter and nighttime hours are omitted in Figure \ref{meteo_kdown} to examine daytime sUHI under clear sky and clouded sky conditions.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=15cm,height=7in,keepaspectratio]{meteo_kdown}
+	\caption{sUHI magnitude versus incoming solar radiation binned for clear sky (when the sum of daytime K\textsubscript{down} \textgreater 10000 \si{\watt \per \square \meter})and cloudy sky days (when the sum of daytime K\textsubscript{down} \textless 10000 \si{\watt \per \square \meter})}
+	\label{meteo_kdown}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=15cm,height=7in,keepaspectratio]{meteo_hum}
+	\caption{sUHI magnitude versus atmospheric water vapor content measured at 2 \si{\meter}. Similar patterns are observed when rural water vapor content is substituted.}
+	\label{meteo_hum}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=15cm,height=7in,keepaspectratio]{meteo_wv}
+	\caption{Bees nest plot}
+	\label{meteo_wv}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=15cm,height=7in,keepaspectratio]{meteo_atuhi}
+	\caption{Bees nest plot}
+	\label{meteo_atuhi}
+\end{figure}
+
+\section{Discussion}
+
+\subsection{Diurnal patterns of sUHI}
+
+sUHI in Basel from complete, nadir, and hemispherical representations of T\textsubscript{surf} displays two diurnal peaks, the first in the late afternoon and a second just after sunset.
+Clear sky conditions amplify this phenomena.
+A late afternoon peak in sUHI is likely the result of the urban-rural contrasts in thermal admittance (and by relation, urban-rural soil moisture contrasts).
+The thermal admittance of many urban surfaces, particularly rooftops, tends to be lower than rural and vegetated environments \cite{Spronken-Smith1998} – particularly in temperate climates where rural soil moisture is largely maintained through summer.
+Lower urban thermal admittance manifests in a larger diurnal thermal amplitude and thus, greater daytime surface heating and nighttime surface cooling.
+Nadir and hemispherical views are biased towards low thermal admittance rooftop facets, and thus display large daytime peaks in sUHI.
+
+A second peak in sUHI after sunset is the result of differences in relative cooling rates of urban and rural surfaces.
+In this case, nighttime sUHI is likely the result of canyon radiation trapping from obstructed canyon sky-view-factors.
+The second peak sUHI is largest on following clear sky conditions when the potential for significant differentials in urban-rural nocturnal cooling rates is maximized.
+A second peak is observed across all representations of urban T\textsubscript{surf} but is most apparent in complete and nadir representations of sUHI, which, for the Sperrstrasse site, sample T\textsubscript{road} to a greater extent than the hemispherical view.
+A weak sUHI is maintained through the nighttime as urban and rural cooling rates reach parity in the hours after sunset \cite{Oke2017}.
+Minimum sUHI is observed just after sunrise, where sUHI magnitudes are occasionally negative.
+At low solar angles, shading by 3-dimensional urban geometry suppresses heating rates for a large portion of urban surface area, reducing overall urban heating rates in early morning.
+As solar angles increase through the morning and more facets experience direct isolation, sUHI develops, with nadir and hemispherical representations displaying more rapid intensification than complete sUHI from a bias towards low thermal admittance and directly sunlit rooftop surfaces.
+
+\subsection{Seasonal patterns of sUHI}
+
+Hemispherical sUHI displays significant seasonality, both in terms of its diurnal pattern and overall magnitude.
+Seasonal variance in sUHI is greatest by day, where large seasonal contrasts in cloud cover frequency and solar angle exert strong control on daytime heating rates and sUHI development.
+Peak sUHI occurs near the summer solstice, when days are longest and isolation on clear sky days is most intense.
+In winter, daytime sUHI is suppressed from a lower solar input and frequent cloud cover.
+This shifts and weakens peak sUHI in winter to shortly after sunset.
+Nighttime sUHI is relatively consistent across seasons, largely a product of contrasts in urban-rural cooling rates just after sunset resulting from radiation trapping by canyon geometry.
+Similar patterns are observed in seasonal variance of canopy layer UHI, albeit concentrated at different times of day.
+
+In contrast to sUHI, clUHI is greatest just after sunset and is strongly controlled by sunset/sunrise timing.
+Peak clUHI occurs just after sunset across seasons, differing only in terms of magnitude, with a larger peak clUHI in summer.
+However, in this study, it should be note that summertime clUHI magnitudes in the hours just after sunset are slightly higher than expected \cite{Runnalls2000} – possibly the result of cold air draining from hilly terrain west of the rural site.
+Seasonality in clUHI results from seasonal contrasts in cloud cover frequency.
+Increased wintertime cloud cover reduces radiative losses by the canyon air volume after sunset, suppressing mean wintertime clUHI magnitudes.
+Daytime clUHI is consistently small across all seasons.
+Thus, although sUHI and clUHI are controlled by similar climatic controls and have a similar micro meteorological genesis, they do not display the same seasonal patterns.
+Thus, \textit{a surface urban heat island is not simply a stronger, more pronounced, canopy layer urban heat island.}
+
+\subsection{Sensor-surface-sun geometries and sUHI}
+
+Sensor-surface geometry exerts significant influence on observed sUHI magnitudes.
+By undersampling vertical and sloped urban surfaces and oversampling low thermal admittance fabric materials, nadir sUHI overestimates mean afternoon complete sUHI by 3.41 \si{\kelvin} and up to 5.96 \si{\kelvin} under clear sky “satellite friendly” conditions.
+Thus, peak nadir sUHI overestimates the peak 'true' complete sUHI by nearly double under clear sky conditions.
+As clear sky conditions are the only conditions under which satellite TIR remote sensors can operate, and make up a large fraction of the urban T\textsubscript{surf} literature, these biases cannot be ignored.
+By sampling the urban surface in 3-dimensions, hemispherical sUHI yields a much smaller mean daytime overestimation (0.54 \si{\kelvin}), particularly under clear sky conditions (0.93 \si{\kelvin}), providing a more geometrically representative depiction of complete urban T\textsubscript{surf} and sUHI, in addition to allowing for time-continuous analysis.
+Cloud cover suppresses the influence of sensor-surface geometry on remote sensed sUHI by reducing microscale geometric and spatiotemporal contrasts in urban T\textsubscript{surf}, resulting in less facet-scale heterogeneity in urban T\textsubscript{surf} and reducing the overall magnitude of sUHI.
+
+\subsection{Meteorological controls on sUHI}
+
+sUHI is highly dependent on meteorological conditions, which account for much of the variation in daytime sUHI magnitudes.
+Conditions that foster the largest microscale contrasts in urban T\textsubscript{surf} also entail the largest sUHI magnitudes.
+Contra, conditions that suppress microscale contrasts in T\textsubscript{surf} show similarly reduced sUHI magnitudes.
+As such, incoming shortwave radiation and wind velocity display positive and negative relationships with sUHI magnitudes respectively.
+Under clear sky conditions, strong solar heating of the surfaec combined with differences in urban and rural thermal properties fosters the development of large daytime urban-rural thermal contrasts.
+Thus, sUHI\textsubscript{max} coincides with insolation\textsubscript{max}.
+Under cloudy conditions, decreased direct solar heating of the surface suppresses the development of urban-rural thermal contrasts, resulting in a reduced sUHI.
+$\Delta$T\textsubscript{hem, r - air} has a similar positive effect on sUHI, owing to its dependence on solar insolation.
+
+As was found with correction magnitudes in Chapter \ref{paper1}, sUHI magnitudes are not strongly correlated with water vapor content as, in the absence of fog, humidity near the surface does not strongly influence surface heating or cooling rates in and of itself.
+Atmospheric transmittance of longwave radiation decreases sharply with small increases in water vapor when the atmosphere is relatively dry - shown in Figure \ref{humtest} - with decreasing variation in atmospheric transmittance as water vapor content increases from zero.
+Therefore, typical variations in urban and rural humidities have little effect on the fraction of outgoing available for heating of the canopy layer air volume through absorption of outgoing longwave radiation and for heating of the surface by atmospheric emission towards the surface.
+As a result, water vapor content has little effect on sUHI.
+Wind velocity displays a strong negative relationship with sUHI.
+sUHI is greatest under calm winds when near-surface mixing is minimal, fostering strong microscale and urban-rural contrasts in T\textsubscript{surf}.
+As wind velocity increases, sUHI is suppressed from increased turbulent heat losses.
+
+\begin{table}[H]
+	\centering
+	\caption{A summary of the relationship between sUHI, meteorological variables, and clUHI.}
+	\label{meteo_cont}
+	\begin{tabular}{lc}
+		\toprule 
+		An increase in $x$, has a/an... & $y$ effect on sUHI magnitude. \\
+		\midrule
+		$\Delta$T\textsubscript{hem, r - air} &\textit{increasing} \\
+		&\\
+		T\textsubscript{hem, r} & \textit{increasing} \\
+		&\\
+		T\textsubscript{air} & \textit{increasing} \\
+		&\\
+		clUHI &\textit{increasing} \\
+		&\\
+		Incoming solar radiation & \textit{increasing}  \\
+		& \\
+		& \\
+		Water vapor content & \textit{negligible} \\
+	 	&\\
+		Wind velocity &\textit{decreasing} \\
+		\bottomrule
+	\end{tabular} 
+\end{table}
+
+clUHI and sUHI develop at different times of day and, thus, have a complex diurnal relationship.
+By day, variation in clUHI is minimal and shows a weak positive correlation with sUHI.
+By night, clUHI shows greater variation and a stronger positive relationship with sUHI.
+Increasing sUHI with clUHI is largest on clear sky nights when urban-rural contrasts in air and surface cooling rates are strongest.
+This results in similar sUHI and clUHI development in the hours after sunset under clear sky conditions.
+
+\section{Conclusion}
+
+This chapter presents the results of a time-continuous eight month climatological analysis of the surface urban heat island effect in Basel, Switzerland derived using the method presented in Chapter \ref{paper1}.
+Hemispherical sUHI magnitudes show significant seasonal and diurnal variation and are largest in summer months during the late afternoon during which solar heating of the surface is most intense, forcing the development of strong microscale and urban-rural contrasts in T\textsubscript{surf}.
+Variation in afternoon sUHI is largely dependent on incoming solar radiation, which displays a strong positive relationship with sUHI magnitudes.
+Similarly, sUHI displays positive relationships with $\Delta$T\textsubscript{hem, r} and nighttime clUHI and a negative relationship with wind velocity.
+In the winter months, a lower overall solar input and increased cloud cover frequency suppresses daytime sUHI development day-to-day variation in sUHI.
+
+ To elucidate the effect of surface-sensor geometry on remote sensed sUHI patterns, sUHI was calculated using weighted facet T\textsubscript{surf} to compute complete and plan representations of sUHI against sUHI calculated from T\textsubscript{hem, r}

### 2017-04-13 paper1.tex

@@
-\chapter{A method to correct upwelling longwave radiation to estimate hemispherical urban surface temperature}\label{paper1}
+\chapter{A hemispherical atmospheric correction method to derive time-continuous urban surface temperature}\label{paper1}
 
@@
 However, in spite of it's widespread usage, several questions concerning the use and validity of urban remote thermal remote sensing, first posed in Roth et al.
-1989 \cite{Roth1989}, have yet to be sufficiently answered, viz,
+1989 \citet{Roth1989}, have yet to be sufficiently answered, viz,
 

### 2017-04-13 paper2.tex

@@
 atmospheric and emissivity effects).
-Meta-analysis of air temperature (T\textsubscript{air}) UHI literature shows that these goals are rarely satisfied \cite{Stewart2011}.
+Meta-analysis of air temperature (T\textsubscript{air}) UHI literature shows that these goals are rarely satisfied \citep{Stewart2011}.
 Given the relative difficulty in retrieving accurate, representative urban T\textsubscript{surf}, similar conclusions are likely for sUHI analysis.
-In spite of this fact, and the short period over which large-scale, generalizable methods for measuring urban T\textsubscript{surf} have been available, study of sUHI via TIR remote sensing has expanded significantly in the last twenty years \cite{Peng2012,Voogt2003}.
-Using the method described in \ref{paper1} we derive an 8-month climatology of hemispherical radiometric urban T\textsubscript{surf} (T\textsubscript{hem, r}) from time continuous near-ground upwelling TIR as measured from an inverted pyrgeometer.
+In spite of this fact, and the short period over which large-scale, generalizable methods for measuring urban T\textsubscript{surf} have been available, study of sUHI via TIR remote sensing has expanded significantly in the last twenty years \citep{Peng2012,Voogt2003}.
+Using the method described in Chapter \ref{paper1} we derive an 8-month climatology of hemispherical radiometric urban T\textsubscript{surf} (T\textsubscript{hem, r}) from time continuous near-ground upwelling TIR as measured from an inverted pyrgeometer.
 The method was developed to address and overcome biases inherent in traditional methods for urban T\textsubscript{surf} retrieval by providing temporally continuous, geometrically representative\footnote{A hemispherical view is not perfectly geometrically representative of urban surface geometry.
@@
 The resulting microscale spatiotemporal contrasts in urban T\textsubscript{surf} create a directional dependence in observed urban T\textsubscript{surf} when measured from conventional narrow-field-of-view (FOV) TIR remote sensing platforms.
-Thus, remote sensed urban T\textsubscript{surf} varies based on sensor FOV, viewing angle and direction, and sun-surface geometry - this directional dependence of urban surface temperature is termed "effective thermal anisotropy" \cite{Voogt1998a}.
-Traditional satellite or airborne remote sensing platforms, by viewing the surface in the nadir, sample only a fraction of the complete urban surface and fail to capture this effect - leading to spatiotemporally variant directional biases of up to 10 \si{\kelvin} in observed urban T\textsubscript{surf} \cite{Voogt1995}.
-In general, geometric undersampling by a remote sensor in the nadir manifests as an overestimation of daytime T\textsubscript{surf} and an underestimation of nighttime T\textsubscript{surf} \cite{Adderley2015}.
+Thus, remote sensed urban T\textsubscript{surf} varies based on sensor FOV, viewing angle and direction, and sun-surface geometry - this directional dependence of urban surface temperature is termed "effective thermal anisotropy" \citep{Voogt1998a}.
+Traditional satellite or airborne remote sensing platforms, by viewing the surface in the nadir, sample only a fraction of the complete urban surface and fail to capture this effect - leading to spatiotemporally variant directional biases of up to 10 \si{\kelvin} in observed urban T\textsubscript{surf} \citep{Voogt1995}.
+In general, geometric undersampling by a remote sensor in the nadir manifests as an overestimation of daytime T\textsubscript{surf} and an underestimation of nighttime T\textsubscript{surf} \citep{Adderley2015}.
 However, the magnitude and diurnal and seasonal behaviors of this bias are dependent on sensor viewing geometry and unique site characteristics (canyon height-to-width ratio, canyon orientation and materials, vegetation coverage, etc.).
@@
 Thus, analysis of sUHI patterns and magnitudes across cities and instrument platforms is difficult.
-At smaller time scales still, time-continuous analysis of urban T\textsubscript{surf} shows significant microscale (second to minute) fluctuations in temperature \cite{Christen2012}.
+At smaller time scales still, time-continuous analysis of urban T\textsubscript{surf} shows significant microscale (second to minute) fluctuations in temperature \citep{Christen2012}.
 Most thermal remote sensors provide instantaneous T\textsubscript{surf} (rather than temporally averaged) and are potentially contaminated by high-frequency microscale fluctuations in urban T\textsubscript{surf}.
@@
 
-First, surface-to-sensor path length and view factor geometries are calculated with the surface-sensor-sun urban model (SUM) \cite{Soux2004}.
+First, surface-to-sensor path length and view factor geometries are calculated with the surface-sensor-sun urban model (SUM) \citep{Soux2004}.
 Using a digital building model, SUM calculates path lengths from the surface to the sensor and angular view factors by projecting the pyrgeometer FOV onto a simplified (orthogonal) representation of the surface, determining which points in the DBM are 'seen' by the sensor, and calculating the distance from each 'seen' point to the sensor.
@@
 
-Second, version 4.1 of the MODTRAN radiative transfer code \cite{Berk1987} is used to model spectral at-sensor radiances for each angular path length.
+Second, version 4.1 of the MODTRAN radiative transfer code \citep{Berk1987} is used to model spectral at-sensor radiances for each angular path length.
 Runs are initialized using profiles of T\textsubscript{air} and water vapor content collected concurrently at each time step.
@@
 
-T\textsubscript{hem, r} are retrieved from radiation and meteorological data collected from December 2001 through July 2002 as a part of the Basel Urban Boundary Layer Experiment (BUBBLE) campaign conducted in Basel, Switzerland \cite{Rotach2005}.
+T\textsubscript{hem, r} are retrieved from radiation and meteorological data collected from December 2001 through July 2002 as a part of the Basel Urban Boundary Layer Experiment (BUBBLE) campaign conducted in Basel, Switzerland \citep{Rotach2005}.
 Urban data were observed from a tower located near city center at the "Basel Sperrstrasse" site.
@@
 	\centering
-	\caption{A description of morphological and measured variables for urban and rural sites. Modified from Rotach et al., 2005 to include only relevant parameters \cite{Rotach2005}.}
+	\caption{A description of morphological and measured variables for urban and rural sites. Modified from \citet{Rotach2005} to include only relevant parameters.}
 	\label{morphbspr}
@@
 		\raggedright
-		\textsuperscript{2} Morphological parameters for the Sperrstrasse site were calculated for a 250 \si{\meter} circular area surrounding the study sites using the method described in \cite{Grimmond1999}.
+		\textsuperscript{2} Morphological parameters for the Sperrstrasse site were calculated for a 250 \si{\meter} circular area surrounding the study sites using the method described in \citet{Grimmond1999}.
 $z_H$: average building height, $\sigma_H $: standard deviation of building height, $\lambda_P $: plan aspect ratio, $\lambda_C $: complete aspect ratio, H/W: local canyon height to width ratio, $\alpha$: surface albedo.
@@
 
-Figure \ref{heatsuhi} shows seasonal variance in diurnal patterns of mean monthly hemispherical sUHI magnitudes at 30 minute intervals for the eight month study period.
-sUHI displays significant seasonal variance, particularly by day, and is stronly controlled by day length and solar angle.
+Figure \ref{heatsuhi} shows seasonal variations in diurnal patterns of mean monthly hemispherical sUHI magnitudes at 30 minute intervals for the eight month study period.
+sUHI displays significant seasonal variation, particularly by day, and is strongly controlled by day length and solar angle.
 Nighttime sUHI does not vary significantly across seasons.
 
-Figure \ref{heatcluhi} shows seasonal variance in diurnal patterns of mean monthly clUHI calculated at 30 minute intervals for the eight month study period.
+Figure \ref{heatcluhi} shows seasonal variations in diurnal patterns of mean monthly clUHI calculated at 30 minute intervals for the eight month study period.
 clUHI development is strongly controlled by the timing of sunset/sunrise cycles and displays the largest seasonal variance in the hours immediately after sunset.
 Daytime clUHI does not vary significantly across seasons.
-
 \begin{figure}[H]
@@
 	\includegraphics[width=15cm,height=7in,keepaspectratio]{meteo_kdown}
-	\caption{sUHI magnitude versus incoming solar radiation binned for clear sky (when the sum of daytime K\textsubscript{down} \textgreater 10000 \si{\watt \per \square \meter})and cloudy sky days (when the sum of daytime K\textsubscript{down} \textless 10000 \si{\watt \per \square \meter})}
+	\caption{sUHI magnitude versus incoming solar radiation binned for clear sky (when the sum of daytime K\textsubscript{down} \textgreater 10000 \si{\watt \per \square \meter}) and cloudy sky days (when the sum of daytime K\textsubscript{down} \textless 10000 \si{\watt \per \square \meter})}
 	\label{meteo_kdown}
@@
 	\includegraphics[width=15cm,height=7in,keepaspectratio]{meteo_hum}
-	\caption{sUHI magnitude versus atmospheric water vapor content measured at 2 \si{\meter}. Similar patterns are observed when rural water vapor content is substituted.}
+	\caption{sUHI magnitude versus atmospheric water vapor content measured at 2 \si{\meter} at the Sperrstrasse site. Similar patterns are observed when rural water vapor content is substituted.}
 	\label{meteo_hum}
@@
 	\includegraphics[width=15cm,height=7in,keepaspectratio]{meteo_wv}
-	\caption{Bees nest plot}
+	\caption{sUHI magnitude versus rural wind velocity measured at approximately 10 \si{\meter} above ground at the Lange Erlen site.}
 	\label{meteo_wv}
@@
 	\includegraphics[width=15cm,height=7in,keepaspectratio]{meteo_atuhi}
-	\caption{Bees nest plot}
+	\caption{sUHI magnitude versus clUHI magnitude. clUHI is calculated as the difference between urban and rural T\textsubscript{air} measured at 2 \si{\meter} above ground level.}
 	\label{meteo_atuhi}
@@
 A late afternoon peak in sUHI is likely the result of the urban-rural contrasts in thermal admittance (and by relation, urban-rural soil moisture contrasts).
-The thermal admittance of many urban surfaces, particularly rooftops, tends to be lower than rural and vegetated environments \cite{Spronken-Smith1998} – particularly in temperate climates where rural soil moisture is largely maintained through summer.
+The thermal admittance of many urban surfaces, particularly rooftops, tends to be lower than rural and vegetated environments \citep{Spronken-Smith1998} – particularly in temperate climates where rural soil moisture is largely maintained through summer.
 Lower urban thermal admittance manifests in a larger diurnal thermal amplitude and thus, greater daytime surface heating and nighttime surface cooling.
-Nadir and hemispherical views are biased towards low thermal admittance rooftop facets, and thus display large daytime peaks in sUHI.
+Nadir and hemispherical views are biased towards low thermal admittance rooftop facets, and thus display large late afternoon peaks in sUHI.
 
@@
 In this case, nighttime sUHI is likely the result of canyon radiation trapping from obstructed canyon sky-view-factors.
-The second peak sUHI is largest on following clear sky conditions when the potential for significant differentials in urban-rural nocturnal cooling rates is maximized.
+The second peak sUHI is largest following clear sky conditions when the potential for significant differentials in urban-rural nocturnal cooling rates is maximized.
 A second peak is observed across all representations of urban T\textsubscript{surf} but is most apparent in complete and nadir representations of sUHI, which, for the Sperrstrasse site, sample T\textsubscript{road} to a greater extent than the hemispherical view.
-A weak sUHI is maintained through the nighttime as urban and rural cooling rates reach parity in the hours after sunset \cite{Oke2017}.
+A weak sUHI is maintained through the nighttime as urban and rural cooling rates reach parity in the hours after sunset \citep{Oke2017}.
 Minimum sUHI is observed just after sunrise, where sUHI magnitudes are occasionally negative.
@@
 Hemispherical sUHI displays significant seasonality, both in terms of its diurnal pattern and overall magnitude.
-Seasonal variance in sUHI is greatest by day, where large seasonal contrasts in cloud cover frequency and solar angle exert strong control on daytime heating rates and sUHI development.
+Seasonal variations in sUHI are largest by day, where large seasonal contrasts in cloud cover frequency and solar angle exert strong control on daytime heating rates and sUHI development.
 Peak sUHI occurs near the summer solstice, when days are longest and isolation on clear sky days is most intense.
 In winter, daytime sUHI is suppressed from a lower solar input and frequent cloud cover.
-This shifts and weakens peak sUHI in winter to shortly after sunset.
-Nighttime sUHI is relatively consistent across seasons, largely a product of contrasts in urban-rural cooling rates just after sunset resulting from radiation trapping by canyon geometry.
-Similar patterns are observed in seasonal variance of canopy layer UHI, albeit concentrated at different times of day.
+This weakens wintertime sUHI and shifts its peak to shortly after sunset.
+Nighttime sUHI is relatively consistent across seasons.
+Similar seasonality is observed in clUHI, albeit concentrated at different times of day.
 
@@
 Peak clUHI occurs just after sunset across seasons, differing only in terms of magnitude, with a larger peak clUHI in summer.
-However, in this study, it should be note that summertime clUHI magnitudes in the hours just after sunset are slightly higher than expected \cite{Runnalls2000} – possibly the result of cold air draining from hilly terrain west of the rural site.
-Seasonality in clUHI results from seasonal contrasts in cloud cover frequency.
+However, in this study, it should be note that summertime clUHI magnitudes in the hours just after sunset are slightly higher than expected \citep{Runnalls2000} - possibly the result of cold air draining from hilly terrain west of the rural site.
+As is the case with sUHI seasonality in clUHI results from seasonal contrasts in solar input and cloud cover frequency.
 Increased wintertime cloud cover reduces radiative losses by the canyon air volume after sunset, suppressing mean wintertime clUHI magnitudes.
@@
 Thus, although sUHI and clUHI are controlled by similar climatic controls and have a similar micro meteorological genesis, they do not display the same seasonal patterns.
-Thus, \textit{a surface urban heat island is not simply a stronger, more pronounced, canopy layer urban heat island.}
+A surface urban heat island is not simply a stronger, more pronounced, canopy layer urban heat island.
 
@@
 As such, incoming shortwave radiation and wind velocity display positive and negative relationships with sUHI magnitudes respectively.
-Under clear sky conditions, strong solar heating of the surfaec combined with differences in urban and rural thermal properties fosters the development of large daytime urban-rural thermal contrasts.
+Under clear sky conditions, strong solar heating of the surface combined with differences in urban and rural thermal properties fosters the development of large daytime urban-rural thermal contrasts.
 Thus, sUHI\textsubscript{max} coincides with insolation\textsubscript{max}.
@@
 As was found with correction magnitudes in Chapter \ref{paper1}, sUHI magnitudes are not strongly correlated with water vapor content as, in the absence of fog, humidity near the surface does not strongly influence surface heating or cooling rates in and of itself.
-Atmospheric transmittance of longwave radiation decreases sharply with small increases in water vapor when the atmosphere is relatively dry - shown in Figure \ref{humtest} - with decreasing variation in atmospheric transmittance as water vapor content increases from zero.
-Therefore, typical variations in urban and rural humidities have little effect on the fraction of outgoing available for heating of the canopy layer air volume through absorption of outgoing longwave radiation and for heating of the surface by atmospheric emission towards the surface.
+Atmospheric transmittance of longwave radiation decreases sharply with small increases in water vapor when the atmosphere is very dry - shown in Figure \ref{humtest}.
+As water vapor content increases from zero, variation in atmospheric transmittance decrease.
+Therefore, typical variations in urban and rural humidities have little effect on the fraction of outgoing available for heating of the canopy layer air volume through absorption of outgoing longwave radiation and for heating of the surface by atmospheric emission.
 As a result, water vapor content has little effect on sUHI.
@@
 As wind velocity increases, sUHI is suppressed from increased turbulent heat losses.
+Meteorological controls on sUHI are summarized in Table \ref{meteo_cont}.
 
@@
 		T\textsubscript{air} & \textit{increasing} \\
-		&\\
-		clUHI &\textit{increasing} \\
+
 		&\\
@@
 		Wind velocity &\textit{decreasing} \\
+		&\\
+		clUHI &\textit{increasing} \\
 		\bottomrule
@@
 By day, variation in clUHI is minimal and shows a weak positive correlation with sUHI.
-By night, clUHI shows greater variation and a stronger positive relationship with sUHI.
-Increasing sUHI with clUHI is largest on clear sky nights when urban-rural contrasts in air and surface cooling rates are strongest.
-This results in similar sUHI and clUHI development in the hours after sunset under clear sky conditions.
+By night, clUHI shows greater variation and a stronger, more positive relationship with sUHI.
+The relationship between clUHI and sUHI is most evident on clear sky nights when urban-rural contrasts in air and surface cooling rates are strongest.
+Under these conditions sUHI and clUHI  show similar development in the hours after sunset.
 
@@
 This chapter presents the results of a time-continuous eight month climatological analysis of the surface urban heat island effect in Basel, Switzerland derived using the method presented in Chapter \ref{paper1}.
-Hemispherical sUHI magnitudes show significant seasonal and diurnal variation and are largest in summer months during the late afternoon during which solar heating of the surface is most intense, forcing the development of strong microscale and urban-rural contrasts in T\textsubscript{surf}.
-Variation in afternoon sUHI is largely dependent on incoming solar radiation, which displays a strong positive relationship with sUHI magnitudes.
-Similarly, sUHI displays positive relationships with $\Delta$T\textsubscript{hem, r} and nighttime clUHI and a negative relationship with wind velocity.
-In the winter months, a lower overall solar input and increased cloud cover frequency suppresses daytime sUHI development day-to-day variation in sUHI.
+Hemispherical sUHI magnitudes display significant seasonal and diurnal variation and are largest in summer months during the late afternoon during which solar heating of the surface is most intense, forcing the development of strong microscale and urban-rural contrasts in T\textsubscript{surf}.
+Seasonality in late afternoon sUHI is largely dependent on incoming solar radiation, which displays a strong positive relationship with sUHI magnitudes.
+In the winter months, a reduced overall solar input and increased cloud cover frequency suppresses daytime sUHI development.
+resulting in less day-to-day variation in sUHI, and lower overall sUHI magnitudes.
 
- To elucidate the effect of surface-sensor geometry on remote sensed sUHI patterns, sUHI was calculated using weighted facet T\textsubscript{surf} to compute complete and plan representations of sUHI against sUHI calculated from T\textsubscript{hem, r}
+Remote sensed sUHI is highly dependent on sensor-surface-sun geometry.
+This results in significant variation in observed urban T\textsubscript{surf} and sUHI when observed from sensors with different spatial, geometric, and temporal sampling regimes.
+Thus, traditional methods for thermal remote sensing of urban areas do not capture the true temporal and geometric nature of the urban effect on land T\textsubscript{surf}.
+Urban effective anisotropy results in large geometric biases in remote sensing of urban areas as traditional satellite and aerial remote sensors sample only a fraction of the urban surface.
+These geometric biases manifest not only in deviation between observed sUHI magnitudes and the 'true' complete sUHI, but also impact characterizations of diurnal and seasonal patterns of sUHI and urban T\textsubscript{surf} by changing the time of day of sUHI\textsubscript{max}.
+Geometric biases in thermal remote sensing of urban area are compounded by myriad temporal biases across a range of temporal scales, as traditional methods for urban thermal remote sensing are not time-continuous and only operable under clear sky conditions.
+Both geometric and temporal biases affect the vast majority of sUHI study, most of which is performed using only a small number of satellite remote sensors, to which these biases are inherent and largely unavoidable.
+
+This study provides the first, long-term, temporally continuous, geometrically representative analysis of urban T\textsubscript{surf} and sUHI.
+By sampling the urban surface in 3-dimensions and time-continuously, the method utilized in this chapter overcomes and quantifies spatial and temporal biases in the satellite thermal remote sensing record to better understand the geometric and temporal effects of a city on the surface climate.
+Further application of the method in other urban areas can help to facilitate a better understanding of the true nature of sUHI across myriad time scales and help to quantify the magnitude of geometric and temporal biases inherent in the satellite sUHI record.

### 2017-04-17 introduction.tex
Add files via upload

@@
 \chapter{Introduction}
+Urban development drastically changes the character of the surface.
+Replacement of natural terrain with roads, buildings, parks, and other urban features modifies the geometry of the surface as well as its thermal, radiative, moisture, and aerodynamic properties.
+These modifications, combined with anthropogenic emissions of heat, result in a distinct urban climate; one which is often warmer, a phenomenon termed the urban heat island effect (UHI).
+As both the Earth's population and the urbanized fraction of that population increase \citep{Nations2014}, cities grow and more people are exposed to urban-modified atmospheres.
+Thus, understanding how cities interact with the climate across spatiotemporal scales has important implications for human health, energy efficiency, and for informing future urban development.
+This, in part, has prompted significant expansion in study of the urban effect on climates, particularly in the decades preceding this thesis.
 
-\section{Defining the urban surface}
- 	Sometimes I have no idea what I am doing.
+The notion that urban areas have elevated air and surface temperatures (T\textsubscript{air} and T\textsubscript{surf} respectively) is not new, with the first formal studies of the urban effect on T\textsubscript{air} dating back to \citet{Howard1833}'s identification of "artificial warming" and "urban contamination" in his characterization of spatial patterns of T\textsubscript{air} in early 19\textsuperscript{th} century London.
+Conceptually, contemporary studies of urban climate do not stray far from \citet{Howard1833}, by seeking to isolate and quantify the urban signal in measured or modeled observation of a given meteorological variable.
+However, as \citet{Howard1833} identified nearly two centuries ago, the sheer number of factors influencing direct measurement of climates makes experimental control nearly impossible in observational study.
+Myriad urban and non urban signals make it difficult to determine what is urban and what is not, degrading interstudy comparison.
+Even in the modeling domain, where experimental control is possible, the inherent complexity in simulating flows of energy, momentum, and mass at the scales necessary to represent physical phenomena over realistic urban environments is, as of yet, not possible.
+Notwithstanding implementation of accurate, physically based schemes to represent urban climates in regional or global climate models.
+Thus, accurate measurement of the urban effect on climate is a deceptively challenging task.
+To address this, \citet{Lowry1977} presents a framework to isolate and analyze the urban effect in observation of a given meteorological variable.
+In his framework - termed the "Lowry method" - measurement of a meteorological variable ($V_M$) is the sum of forcings from the background climate and synoptic conditions ($V_B$), local topography ($V_L$), and the urban landscape ($V_H$)
 
- \section{The urban radiation budget}
+\begin{equation}
+	V_M = V_B + V_L + V_H
+\end{equation}
+
+Thus, isolation of the urban effect on observed $V_M$ simply requires the removal of $V_B$ and $V_L$.
+This can be achieved by subtracting some urban affected observation ($V_{M, U}$) from a non-urban observation ($V_{M, R}$) where $V_{H, R}$ = 0,
+
+\begin{equation}
+V_{M, U} - V_{M, R} = (V_B + V_L + V_H) - (V_B + V_L) = V_H
+\end{equation}
+
+\noindent simplified as,
+
+\begin{equation}
+	V_H = V_{M, U} - V_{M, R}
+\end{equation}
+
+Study of the urban heat island effect tacitly falls into such a framework, by replacing $V_{M, U} - V_{M, R}$ with $\Delta$T\textsubscript{air, U-R}.
+However, in spite of its conceptual simplicity, meta-analysis of observational air temperature UHI (aUHI) study in \citet{Stewart2011} found that nearly fifty percent of aUHI studies published between 1950 and 2007 are "scientifically indefensible" based on an analysis of relevant criteria\footnote{Each study was given a "passing" or "failing" grade and ranked using a points-based scheme to assess methodological quality in the following criteria: conceptual model, operational definitions, instrument specification, site metadata, site representativeness, number of replicates, weather control, surface control, and synchronicity.}, suggesting a significant gulf between conceptual simplicity and practical realities in aUHI analysis.
+Of the 190 surveyed studies, only 13\% were derived from field sites that were sufficiently representative of the local-scale environment or lacked the meta-data to make such a determination, leading to potential biases from micro-scale terrain.
+This is particularly discouraging as it is not the result of an evolution in measurement techniques or improved instrumentation - in fact, \citet{Stewart2011} suggests the opposite: a disproportionately large number of studies deemed scientifically sound were published early in the study purview.
+Thus, in study of the urban effect on variables for which significant methodological shifts have or will occur, the warnings implicit in \citet{Stewart2011} are particularly salient, as inherent difficulties found in translating "Lowry"-esque conceptual models to real world observations may be compounded by significant methodological changes.
+In the five years since its publication, aUHI study has seen further expansion and, as such, it is difficult to assess whether its critical analysis has prompted a shift towards more thoughtful, careful, and methodical study of aUHI, notwithstanding its adoption in study of the urban effect on different meteorological variables.
+
+Study of the urban effect on land T\textsubscript{surf}, compared to study of urban T\textsubscript{air}, is relatively new and is rapidly expanding and evolving, with a vast majority of urban T\textsubscript{surf} study focusing on passive remote sensing of upwelling TIR.
+In the decades preceding this thesis, a combination of factors has lead to significant expansion of TIR remote sensed study of urban T\textsubscript{surf} including:
+
+\begin{itemize}
+	\item The proliferation of satellite and aerial thermal infrared (TIR) remote sensors and "openness" in data access policies.
+	\item Improvements in sensor spatial, spatial, and radiometric resolutions. 
+	\item An increased focus on the importance of the surface in both determining and understanding key near-ground micro-meteorological phenomena.
+\end{itemize}
+
+Remote sensed study of the urban effect on land T\textsubscript{surf} has been instrumental in characterizing spatial and temporal patterns of sUHI \citep{Peng2012, Streutker2003, Imhoff2010}.
+However, given the relative novelty of TIR remote sensing methods, the wide (and broadening) range of remote sensing instruments and methods used for urban T\textsubscript{surf} retrieval, and the difficulty observed in maintaining scientific integrity in aUHI study, similar conclusions to those in \citet{Stewart2011} are likely for study of sUHI.
+In addition, traditional methods for urban T\textsubscript{surf} measurement are also subject to a suite of geometric and temporal biases.
+Geometric biases inherent in thermal remote sensing of the urban surface are the results of urban modification of surface structure and thermal and radiative processes.
+The convoluted 3-dimensional structure of the urban surface modifies sunlit/shading regimes and creates large micro-scale spatiotemporal contrasts in urban T\textsubscript{surf} depending on surface-sun geometry, can be amplified by significant contrasts in the thermal admittance of common urban materials.
+Thus, when urban T\textsubscript{surf} is viewed from a narrow-field-of-view (FOV) remote sensor, urban T\textsubscript{surf} is directionally dependent - an "effective anisotropy" of urban T\textsubscript{surf}.
+Urban effective anisotropy can reach up to 10 \si{\kelvin} and is highly dependent on surface-sensor geometry, surface structure, and urban materials \citep{Krayenhoff2016, Voogt1997}.
+Temporal biases in urban thermal remote sensing occur across multiple scales including:
+
+\begin{itemize}
+	\item Contamination by turbulence forced, high frequency fluctuations in urban T\textsubscript{surf}.
+	\item Discontinuity in satellite overpass cycles for remote sensors with sufficient spatial resolutions.
+	\item Diurnal and seasonal biases towards clear sky "satellite friendly" conditions, under which the atmosphere is relatively transparent to TIR.
+\end{itemize}
+
+\begin{table}[H]
+	\centering
+	\caption{Sources of temporal bias in urban thermal remote sensing.}
+	\label{tbias}
+	\begin{tabular*}{\textwidth}{p{6cm}p{1.5cm}p{3cm}p{3cm}}
+		\toprule 
+		Bias & Temporal scale & Magnitude & Citation  \\  
+		\midrule
+		High-frequency turbulence forced fluctuations in urban T\textsubscript{surf} & Micro- & 1 - 3 \si{\kelvin} & \citep{Christen2012} \\
+		&&&\\
+		Discontinuity in satellite overpass cycles & Micro- to seasonal- & unknown & \\
+		&&&\\
+		Clear sky bias & Diurnal- to seasonal- & unknown & \\
+		\bottomrule
+	\end{tabular*} 
+\end{table}
+
+\section{Research questions and objectives}
+
+Thermal-infrared (TIR) remote sensing of urban surface temperatures (Tsurf) has emerged as a primary research focus in the field of urban climatology as researchers seek to elucidate diurnal and seasonal Tsurf regimes and understand coupling between urban surfaces and the ambient air in complex, 3-d urban environments (Voogt  Oke, 1997, 2003).
+Recent technological improvements in thermal imaging of earth’s surface temperatures have fostered significant progress in characterizations of the surface urban heat island effect (sUHI) and its function in driving urban/suburban air temperature regimes (Peng et al., 2012; Rigo  Parlow, 2005; Roth et al., 1989).
+However, these improvements have not mitigated the geometric and temporal biases inherent in satellite/airborne remote sensing of complex urban surface climates.
+
+Deriving urban Tsurf from conventional remote sensing platforms (i.e., satellites, airborne sensors) presents three main geometrical and temporal deficiencies.
+First, conventional satellite remote sensing represents the urban form as a two-dimensional plane – a ‘bird’s eye view’.
+The resulting measurements undersample urban surfaces by ignoring and/or underrepresenting vertical (wall) facets (Voogt, 2000).
+This is particularly salient in dense urban areas where canyon height is often greater than canyon width (Offerle et al., 2003).
+Second, clouds absorb thermal infrared radiation (TIR) – thus satellite thermal imaging of the surface requires clear sky conditions.
+Third, satellite overpass cycles are temporally discontinuous.
+Consequently, satellite assessment of surface temperature either sacrifices spatial resolution for a daily repeat cycle (MODIS) or sacrifices temporal resolution for a higher spatial resolution (ASTER or Landsat).
+As a result of these three shortcomings, the satellite record of urban Tsurf is discontinuous, has a significant clear sky bias, and fails to accurately sample the complete, 3-d urban form.
+Prompted by these shortcomings the proposed research introduces novel analyses of urban/rural Tsurf regimes to elucidate the following:
+
+\begin{itemize}
+	\item What is the nature of urban surface temperature when viewed from a hemispherical downward-facing radiometer?
+And how does it relate to urban temperatures derived from other methods for urban surface temperature retrieval?
+	\item What is the diurnal and seasonal nature of the surface urban heat island effect?
+\end{itemize}
+
+\noindent In an attempt to answer these questions, I did the following:
+
+\begin{enumerate}
+	\item Developed and evaluated a method to retrieve atmospherically correction hemispherical radiometric urban surface temperatures from time-continuous measurements of upwelling longwave radiation.
+	\item Compared urban surface temperatures and surface urban heat island magnitudes retrieved using the method to common remote sensed representations of the urban surface.
+	\item Derived an eight month climatology of hemispherical urban surface temperatures to observe seasonal and diurnal patterns of the surface urban heat island effect.
+\end{enumerate}
 
 \section{The structure of the following sections}
+
+The following two papers make up the body of the thesis, read them you fuck.

### 2017-04-17 paper1.tex
Add files via upload

@@
 
-Within urban climatology, a combination of satellite, aerial, and ground-based thermal remote sensors have been integral in elucidating the spatial \cite{Roth1989}, temporal \cite{Peng2012}, and geometric \cite{Voogt1997} effects of the built environment on land T\textsubscript{surf}; in evaluating and partitioning urban surface energy balances \cite{BastiaanssenW.G.M.1998, Yamaguchi2005} and; in characterizations of the relationship between surface and \gls{blayer} air temperatures (T\textsubscript{air}) \cite{Stoll1992}.
+Within urban climatology, a combination of satellite, aerial, and ground-based thermal remote sensors have been integral in elucidating the spatial \citep{Roth1989}, temporal \citep{Peng2012}, and geometric \citep{Voogt1997} effects of the built environment on land T\textsubscript{surf}; in evaluating and partitioning urban surface energy balances \citep{BastiaanssenW.G.M.1998, Yamaguchi2005} and; in characterizations of the relationship between surface and \gls{blayer} air temperatures (T\textsubscript{air}) \citep{Stoll1992}.
 These advances have been aided by substantial improvements in sensor spatial, spectral, and radiometric resolutions, and by the proliferation of both large-scale public satellite remote sensing campaigns and low-cost aerial and near-ground thermography.
-However, in spite of it's widespread usage, several questions concerning the use and validity of urban remote thermal remote sensing, first posed in Roth et al.
-1989 \citet{Roth1989}, have yet to be sufficiently answered, viz,
+However, in spite of it's widespread usage, several questions concerning the use and validity of urban remote thermal remote sensing, first posed in \citet{Roth1989}, have yet to be sufficiently answered, viz,
 
@@
 Spectral radiance received by a sensor at height ($z$) can be described by a function deviating from a Planck curve defined by T\textsubscript{surf} based on the spectral transmittance of the intervening atmosphere, with the magnitude of that deviation determined by the difference between Planck curves calculated from T\textsubscript{surf} and ambient T\textsubscript{air}.
-A radiative transfer code, such as MODTRAN \cite{Berk1987}, can be used to model the effect of the atmosphere on a remote sensed TIR signal.
-Radiative transfer simulations run in MODTRAN version 4.1 for two path lengths are shown in figure \ref{spectairtsurf}, for spectral radiance emitted from a planar surface at 300 \si{kelvin} through an isothermal atmosphere at 290 \si{\kelvin}, water vapor content of 12 \si{\gram\per\meter\cubed}, and aerosol and trace gas profiles from the mid-latitude summer standard atmosphere \cite{Kantor1962}.
+A radiative transfer code, such as MODTRAN \citep{Berk1987}, can be used to model the effect of the atmosphere on a remote sensed TIR signal.
+Radiative transfer simulations run in MODTRAN version 4.1 for two path lengths are shown in figure \ref{spectairtsurf}, for spectral radiance emitted from a planar surface at 300 \si{kelvin} through an isothermal atmosphere at 290 \si{\kelvin}, water vapor content of 12 \si{\gram\per\meter\cubed}, and aerosol and trace gas profiles from the mid-latitude summer standard atmosphere \citep{Kantor1962}.
 At both short and long path lengths, radiance in each band shows some deviation from the ground emitted signal.
@@
 	\includegraphics[width=\textwidth]{spectairtsurf}
-	\caption{At-sensor spectral directional radiances computed in MODTRAN 4.1 \cite{Berk1987} for short (30 \si{m}) and long (3000 \si{m}) path lengths (z) with Planck curves indicating spectral radiances at T\textsubscript{surf} = 300 \si{K} and T\textsubscript{air} = 290 \si{K}.}
+	\caption{At-sensor spectral directional radiances computed in MODTRAN 4.1 \citep{Berk1987} for short (30 \si{m}) and long (3000 \si{m}) path lengths (z) with Planck curves indicating spectral radiances at T\textsubscript{surf} = 300 \si{K} and T\textsubscript{air} = 290 \si{K}.}
 	\label{spectairtsurf}
@@
 
-Atmospheric effects can lead to differences between the 'true' radiometric T\textsubscript{surf} and the remote sensed T\textsubscript{surf} of over 10 \si{K} for satellite platforms \cite{Cooper1989} and over 6 \si{K} for near-ground sensors \cite{Meier2011}.
+Atmospheric effects can lead to differences between the 'true' radiometric T\textsubscript{surf} and the remote sensed T\textsubscript{surf} of over 10 \si{K} for satellite platforms \citep{Cooper1989} and over 6 \si{K} for near-ground sensors \citep{Meier2011}.
 Moreover, because atmospheric effects are a function of non-uniform and spatiotemporally variant surface and atmospheric properties, their associated errors change depending on instrument type, surface-sensor geometry, study location, and ambient conditions.
@@
 
-Spectral transmission of longwave radiation through a given layer of atmosphere is dependent on total column absorber content (the principal broadband TIR absorbers are H\textsubscript{2}O, CO\textsubscript{2}, and to a lesser extent O\textsubscript{3}, N\textsubscript{2}O, CO, CH\textsubscript{4}, and O\textsubscript{2} \cite{Miskolczi1993}).
+Spectral transmission of longwave radiation through a given layer of atmosphere is dependent on total column absorber content (the principal broadband TIR absorbers are H\textsubscript{2}O, CO\textsubscript{2}, and to a lesser extent O\textsubscript{3}, N\textsubscript{2}O, CO, CH\textsubscript{4}, and O\textsubscript{2} \citep{Miskolczi1993}).
 Holding vertical absorber content constant through a layer of atmosphere, variation in band-by-band TIR transmittance with path length is greatest at urban scales (approximately 1 to 50 \si{\meter}), where transparent spectral bands can quickly become opaque with small changes in path length or absorber content - illustrated in figure \ref{spectransheight}.
@@
 
-Radiation passing through a layer of atmosphere from an emitting surface towards a sensor can be described in a number of ways: \textit{Spectral directional radiance} $R^\uparrow_z (\lambda, \theta, \phi)$, at height $ z $, wavelength $\lambda$ and from a direction defined by viewingzenith angle \(\theta\), and azimuth angle \(\phi\) is commonly written as
+Radiation passing through a layer of atmosphere from an emitting surface towards a sensor can be described in a number of ways: \textit{Spectral directional radiance} $R^\uparrow_z (\lambda, \theta, \phi)$, at height $ z $, wavelength $\lambda$ and from a direction defined by viewing zenith angle \(\theta\), and azimuth angle \(\phi\) is commonly written as
 
@@
 Thus, proper terminology must be attached to land T\textsubscript{surf} inferred from TIR.
-Definitions and nomenclature conventions for multiple methods for T\textsubscript{surf} retrieval are discussed at length in Norman \& Becker, (1995) \cite{Norman1995}.
+Definitions and nomenclature conventions for multiple methods for T\textsubscript{surf} retrieval are discussed at length in \citet{Norman1995}.
 
@@
 A large number of correction routines have been developed to remove atmospheric and emissivity effects from aerial and satellite TIR signals and derive accurate T\textsubscript{rad}.
-Methods range from simple mono- \cite{Qin2001} and split-window \cite{Wan1996} routines for single- and multi-channel remote sensors, to schemes that integrate a radiative transfer code to isolate the surface emitted signal from interfering signals.
+Methods range from simple mono-window \citep{Qin2001} and split-window \citep{Wan1996} routines for single-channel and multi-channel remote sensors, to schemes that integrate a radiative transfer code to isolate the surface emitted signal from interfering signals.
 Boundary conditions are standard across most correction methods: generally requiring vertical profiles of T\textsubscript{air}, humidity, pressure, and aerosol content to remove atmospheric effects, and surface radiative properties to correct for emissivity effects.
@@
 
-Several multi-LOS correction routines have been developed for to remove atmospheric effects on near-ground remote sensed TIR: Meier et al., 2011 describes a correction method for oblique angled thermal imagery of urban terrain \cite{Meier2011}.
+Several multi-LOS correction routines have been developed for to remove atmospheric effects on near-ground remote sensed TIR: \citet{Meier2011} describes a correction method for oblique angled thermal imagery of urban terrain.
 In the method, spatially distributed LOS geometries were calculated by linking each image pixel to the corresponding 3-d coordinates of its viewpoint on a digital building model (DBM).
@@
 
-Kotani \& Sugita, 2009 describes a method for correction of wide-FOV (pyrgeometer) TIR irradiances \cite{Kotani2009a} over a homogeneous planar surface.
+\citet{Kotani2009a} describes a method for correction of wide-FOV (pyrgeometer) TIR irradiances over a homogeneous planar surface.
 In this method, path lengths were calculated for six sensor heights.
@@
 
-To calculate surface-sensor geometries, the Surface-Sensor-Sun Urban Model (SUM) \cite{Soux2004} is initialized with a simplified, orthogonal 3-dimensional DBM representing surface geometry of the area surrounding the sensor.
+To calculate surface-sensor geometries, the Surface-Sensor-Sun Urban Model (SUM) \citep{Soux2004} is initialized with a simplified, orthogonal 3-dimensional DBM representing surface geometry of the area surrounding the sensor.
 SUM uses a four-dimensional array to represent surface morphology with three spatial dimensions ($x$, $y$, and $z$), with $z$ representing height above the $x$, $y$ plane.
@@
 \subsection{Modeling hemispherical irradiances}
-With path length geometries calculated in SUM, irradiances are modeled for each time step using version 4.1 of the MODerate resolution atmospheric TRANsmission radiative transfer code (MODTRAN) \cite{Berk1987}.
+With path length geometries calculated in SUM, irradiances are modeled for each time step using version 4.1 of the MODerate resolution atmospheric TRANsmission radiative transfer code (MODTRAN) \citep{Berk1987}.
 At each time-step, a range of potential temperatures is specified (T\textsubscript{spec}) based on T\textsubscript{hem, b} calculated from the measured irradiance with
@@
 Profiles of T\textsubscript{air} and humidity are retrieved from 30-minute averages of conditions observed at the Sperrstrasse site.
-Aerosol, trace gas absorber, and above-sensor T\textsubscript{air} and humidity conditions are defined by the mid-latitude summer standard atmosphere when daytime T\textsubscript{air, max} \textgreater 10 \si{\degreeCelsius} (the mid-latitude winter profile is substituted on days where T\textsubscript{air, max} \textless 10 \si{\degreeCelsius}) \cite{Kantor1962}.
+Aerosol, trace gas absorber, and above-sensor T\textsubscript{air} and humidity conditions are defined by the mid-latitude summer standard atmosphere when daytime T\textsubscript{air, max} \textgreater 10 \si{\degreeCelsius} (the mid-latitude winter profile is substituted on days where T\textsubscript{air, max} \textless 10 \si{\degreeCelsius}) \citep{Kantor1962}.
 
@@
 
-MODTRAN has been shown to effectively model near-ground radiative transfer at urban-scale path lengths in 2-dimensions \cite{Hoch2005} \cite{Hoch2007}.
+MODTRAN has been shown to effectively model near-ground radiative transfer at urban-scale path lengths in 2-dimensions \citep{Hoch2005, Hoch2007}.
 However, the method described in this paper accounts for complex 3-dimensional surface geometry, which can exert significant influence on atmospheric transmittance.
@@
 For each time step, T\textsubscript{hem, b}, calculated from the lowest $L_z$ measurement, was used to model irradiances at 10 \si{m} and 30 \si{m} at 30-minute intervals using concurrent profiles of T\textsubscript{air} and humidity.
-Daytime T\textsubscript{air} observed at the highest measurement level was anomalously high compared to typical summertime lapse rates above planar vegetated terrain \cite{Oke1987}.
+Daytime T\textsubscript{air} observed at the highest measurement level was anomalously high compared to typical summertime lapse rates above planar vegetated terrain \citep{Oke1987}.
 Thus, the T\textsubscript{air} profile was modified to be more representative of the neutral or weakly unstable daytime conditions and weakly stable nighttime conditions.
 It is likely that these modifications better represent the T\textsubscript{air} profile in the layer between 10 and 30 \si{\meter} than a simple interpolation between the 10 \si{\meter} T\textsubscript{air} and 30 \si{\meter} T\textsubscript{air} measurements.
-Contrary to the findings in Kotani \& Sugita (2009), an isothermal atmospheric profile was not sufficient to accurately model upwelling fluxes over planar terrain.
-Although thermal stratification is relatively small by day in an urban canyon \cite{Nakamura1988} - where strong microscale contrasts in T\textsubscript{surf} foster strong canyon mixing and neutral stability - the large daytime T\textsubscript{surf} - T\textsubscript{air} differential and the path-length/transmittance gradient can create large differences in upwelling longwave forced by inter-canyon thermal contrasts - a phenomenon termed radiative divergence.
+Contrary to the findings in \citet{Kotani2009a}, an isothermal atmospheric profile was not sufficient to accurately model upwelling fluxes over planar terrain.
+Although thermal stratification is relatively small by day in an urban canyon \citep{Nakamura1988} - where strong microscale contrasts in T\textsubscript{surf} foster strong canyon mixing and neutral stability - the large daytime T\textsubscript{surf} - T\textsubscript{air} differential and the path-length/transmittance gradient can create large differences in upwelling longwave forced by inter-canyon thermal contrasts - a phenomenon termed radiative divergence.
 By underestimating near-surface air temperatures and overestimating roughness-layer T\textsubscript{air}, positive divergences by day are underestimated by an isothermal profile.
@@
 		\textit{R\textsuperscript{2}} & 0.995 & 0.996 & 0.863 & \\ 
-		MAE,  (\si{\watt\per\meter\squared}) & 1.781 & 1.325 & 1.191 & \cite{Willmott1985a} \\ 
-		RMSE,  (\si{\watt\per\meter\squared}) & 2.204 & 1.690 & 1.423 & \cite{Willmott1985a} \\ 
-		RMSE\textit{\textsubscript{s}},  (\si{\watt\per\meter\squared}) & 1.444 & 0.851 & 1.028 & \cite{Willmott1985a} \\ 
-		RMSE\textit{\textsubscript{u}},  (\si{\watt\per\meter\squared}) & 1.666 & 1.460 & 0.985 & \cite{Willmott1985a} \\ 
-		\textit{d} & 0.999 & 0.999 & 0.959 & \cite{Willmott2012} \\
+		MAE,  (\si{\watt\per\meter\squared}) & 1.781 & 1.325 & 1.191 & \citep{Willmott1985a} \\ 
+		RMSE,  (\si{\watt\per\meter\squared}) & 2.204 & 1.690 & 1.423 & \citep{Willmott1985a} \\ 
+		RMSE\textit{\textsubscript{s}},  (\si{\watt\per\meter\squared}) & 1.444 & 0.851 & 1.028 & \citep{Willmott1985a} \\ 
+		RMSE\textit{\textsubscript{u}},  (\si{\watt\per\meter\squared}) & 1.666 & 1.460 & 0.985 & \citep{Willmott1985a} \\ 
+		\textit{d}, agreement index & 0.999 & 0.999 & 0.959 & \citep{Willmott2012} \\
 		\bottomrule
@@
 Figure \ref{season_fit} shows the effect of seasonality and cloud cover on the relationship between correction magnitudes and $\Delta$T\textsubscript{hem, r - air}.
-The relationship is strongest and most clear in summer when variations in cloud cover can lead to large day-to-day contrasts in solar input and $\Delta$T\textsubscript{surf - air}, which translates into a wide range of correction magnitudes.
+The relationship is strongest and most evident in summer when variations in cloud cover can lead to large day-to-day contrasts in solar input and $\Delta$T\textsubscript{surf - air}, which translates into a wide range of correction magnitudes.
 
-Figure \ref{nofit} illustrates the the relationship between correction magnitude and meteorological variables commonly measured as a part of climatological assessments of urban environments.
+Figure \ref{nofit} illustrates the relationship between correction magnitude and meteorological variables commonly measured as a part of climatological assessments of urban environments.
 A strong positive relationship is observed between correction magnitudes and incoming solar radiation.
@@
 	height=7in,
-	keepaspectratio]{bxsac_season_clima_pres}
-	\caption{Atmospheric correction magnitudes calculated as the difference between T\textsubscript{hem, r} and T\textsubscript{bright} for the duration of the study period.
-Grey shading indicates nighttime.}
+	keepaspectratio]{bx_ac1}
+	\caption{Atmospheric correction magnitudes calculated as the difference between T\textsubscript{hem, r} and T\textsubscript{hem, b} for the duration of the study period.
+Grey shading indicates nighttime.
+The red line indicates mean correction magnitude at each time step.}
 	\label{bx_ac}
@@
 	keepaspectratio]{nofit}
-	\caption{Correction magnitude versus meteorological variables for the daytime hours of the study period. }
+	\caption{Correction magnitude versus meteorological variables for the daytime hours of the study period.}
 	\label{nofit}
@@
 
-%\begin{figure}[H]
-%	\centering
-%	\includegraphics[width=15cm,
-%	height=7in,
-%	keepaspectratio]{corr_sd}
-%	\caption{Correction magnitude versus the T\textsubscript{hem, r}-T\textsubscript{air} differential for summer months of the study period (April 15\textsuperscript{th} - July 7\textsuperscript{th}).
-Contour shading indicates frequency (higher frequencies in red, lower in yellow).}
-%	\label{corr_sd}
-%\end{figure}
-%
-%\begin{figure}[H]
-%	\centering
-%	\includegraphics[width=15cm,
-%	height=7in,
-%	keepaspectratio]{corr_wd}
-%	\caption{Correction magnitude versus the T\textsubscript{hem, r}-T\textsubscript{air} differential for winter and spring months of the study period (Dec 1\textsuperscript{st} - April 14\textsuperscript{th}).
-Contour shading indicates frequency (higher frequencies in red, lower in yellow).}
-%	\label{corr_wd}
-%\end{figure}
-%
-%\begin{figure}[H]
-%	\centering
-%	\includegraphics[width=15cm,
-%	height=7in,
-%	keepaspectratio]{corr_sds}
-%	\caption{Correction magnitude versus the T\textsubscript{hem, r}-T\textsubscript{air} differential for clear sky summertime days over the study period.
-Contour shading indicates frequency (higher frequencies in red, lower in yellow).}
-%	\label{corr_sds}
-%\end{figure}
-%
-%\begin{figure}[H]
-%	\centering
-%	\includegraphics[width=15cm,
-%	height=7in,
-%	keepaspectratio]{corr_sdc}
-%	\caption{Correction magnitude versus the T\textsubscript{hem, r}-T\textsubscript{air} differential for partly cloudy and overcast summertime days over the study period.
-Contour shading indicates frequency (higher frequencies in red, lower in yellow).}
-%	\label{corr_sdc}
-%\end{figure}
-%
-%\begin{figure}[H]
-%	\centering
-%	\includegraphics[width=15cm,
-%	height=7in,
-%	keepaspectratio]{corr_kdown}
-%	\caption{Correction magnitude versus incoming shortwave radiation for the daytime hours of the study period. Contour shading indicates frequency (higher frequencies in red, lower in yellow).}
-%	\label{corr_kdown}
-%\end{figure}
-%
-%\begin{figure}[H]
-%	\centering
-%	\includegraphics[width=15cm,
-%	height=7in,
-%	keepaspectratio]{corr_them}
-%	\caption{Correction magnitude versus T\textsubscript{hem, r} over the duration of the study period. Contour shading indicates frequency (higher frequencies in red, lower in yellow).}
-%	\label{corr_them}
-%\end{figure}
-%
-%\begin{figure}[H]
-%	\centering
-%	\includegraphics[width=15cm,
-%	height=7in,
-%	keepaspectratio]{corr_tair}
-%	\caption{Correction magnitude versus T\textsubscript{air} over the duration of the study period. Contour shading indicates frequency (higher frequencies in red, lower in yellow).}
-%	\label{corr_tair}
-%\end{figure}
-%
-%\begin{figure}[H]
-%	\centering
-%	\includegraphics[width=15cm,
-%	height=7in,
-%	keepaspectratio]{corr_abshum}
-%	\caption{Correction magnitude versus atmospheric water vapor content for the daytime hours of the study period. Contour shading indicates frequency (higher frequencies in red, lower in yellow).}
-%	\label{corr_abshum}
-%\end{figure}
-
 \section{Comparing T\textsubscript{surf} from different sensor geometries}
-
-%\begin{figure}[H]
-%	\centering
-%	\includegraphics[width=15cm,
-%	height=7in,
-%	keepaspectratio]{tc_quarts2}
-%	\caption{A comparison of T\textsubscript{surf} from sensor geometries described in Table \ref{weightings}. T\textsubscript{surf} is normalized against complete %T\textsubscript{surf}.}
-%	\label{tc_quarts}
-%\end{figure}
 
@@
 
-Mean normalized T\textsubscript{hem, r}, T\textsubscript{plan}, and $\Delta$T\textsubscript{plan - hem, r} are shown in figure \ref{tc_mean_314} along with clear sky and overcast sky case days.
+Mean normalized T\textsubscript{hem, r}, T\textsubscript{plan}, and $\Delta$T\textsubscript{plan - hem, r} are shown in figure \ref{tc_mean} along with clear sky and overcast sky case days.
 While mean T\textsubscript{hem, r} and T\textsubscript{plan} display similar overestimations of T\textsubscript{comp}, under clear sky conditions, overestimation by T\textsubscript{plan} is much larger than that by T\textsubscript{hem, r}.
@@
 \end{figure}
-
-%As shown in figure \ref{tc_mean},
-
-%\begin{figure}[H]
-%	\centering
-%	\includegraphics[width=15cm,
-%	height=7in,
-%	keepaspectratio]{tc_mean3}
-%	\caption{Normalized T\textsubscript{hem, r}, T\textsubscript{plan}, and T\textsubscript{plan} minus T\textsubscript{hem, r} averaged over the IOP and for clear sky and overcast case days.}
-%	\label{tc_mean}
-%\end{figure}
 
@@
 Correction magnitudes are smallest at temperatures between approximately $10$ and $20$ \si{\degreeCelsius} - largely made up of morning and cloudy observations during the summer months where thermal contrasts between T\textsubscript{surf} and T\textsubscript{air} are small or weakly negative, forcing small correction magnitudes.
-During the winter months, a negative relationship between T\textsubscript{hem, r}/T\textsubscript{air} and correction magnitudes is observed, likely a result of the effect of shifting spectral emittance curves combined with non-uniform sensor response - discussed in section \ref{Persistence in correction magnitudes}.
+During the winter months, a negative relationship between T\textsubscript{hem, r} and T\textsubscript{air} and correction magnitudes is observed, likely a result of the effect of shifting spectral emittance curves combined with non-uniform sensor response - discussed in section \ref{Persistence in correction magnitudes}.
 On clear sky summer days, positive relationships between surface and air temperatures and correction magnitudes is observed as an increased solar input foster higher overall temperatures and large contrasts between T\textsubscript{surf} and T\textsubscript{air}.
@@
 Results show significant seasonal, day to day, and diurnal variation in correction magnitudes.
-The range of observed correction magnitudes is largest in summer, when a wide range of mid latitude synoptic conditions result in significant contrasts in daytime solar input and similarly large variance in the T\textsubscript{surf} - T\textsubscript{air} differential.
+The range of observed correction magnitudes is largest in summer, when a wide range of mid latitude synoptic conditions result in significant contrasts in daytime solar input and consequently large variations in the T\textsubscript{surf} - T\textsubscript{air} differential.
 Maximum correction magnitudes over the study period (often approaching $7$ to $8$ \si{\kelvin}) occur frequently near solar noon on clear, hot, humid days, during which intense solar heating of the surface produces a large T\textsubscript{surf} - T\textsubscript{air} differential.
-Minimum correction magnitudes near $-1$ \si{\kelvin} occurs consistently on clear, calm nights following hot days, when T\textsubscript{surf} can dip below T\textsubscript{air}.
-In winter, decreased variance in overall solar input and lower solar angles result in less intense solar heating of the surface and smaller T\textsubscript{surf} - T\textsubscript{air} contrasts.
+Minimum correction magnitudes near $-1$ \si{\kelvin} occur consistently on clear, calm nights following hot days, when T\textsubscript{surf} can dip below T\textsubscript{air}, even in urban areas where canyon trapping of radiation reduces cooling rates for both T\textsubscript{surf} and T\textsubscript{air}.
+In winter, decreased variability in overall solar input results in less intense solar heating of the surface and smaller T\textsubscript{surf} - T\textsubscript{air} contrasts.
 Thus, winter correction magnitudes are smaller and more consistent with a mean of approximately $2$ \si{\kelvin}.
@@
 A pyrgeometer's dome transmittance is spectrally non-uniform.
-Thus, the "dome effect" a pyrgeometer's exerts on a given incoming TIR signal changes as a function of emission temperature (T\textsubscript{emiss}), as the shape of a given spectral radiance curve is temperature dependent.
-This is illustrated in Figure \ref{nonuniform} by comparing the relative peaks and shape of a pyrgeometer dome transmittance curve to a Planckian spectral radiance curve (R\textsubscript{Planck}) at T = 300 \si{\kelvin} and the same curve shifted to smaller wavenumber to represent the spectral radiance peak of a curve at T = 270 \si{\kelvin} (R\textsubscript{Planck, s}).
-When integrated, the two curves have the same radiance, but different spectral peaks and different resulting dome effects.
+Thus, the "dome effect" a pyrgeometer's exerts on a given incoming TIR signal changes as a function of emission temperature (T\textsubscript{emiss}) because the shape of a given spectral radiance curve (R\textsubscript{$\nu$} (T)) is temperature dependent.
+This is illustrated in Figure \ref{nonuniform} by comparing the relative peaks and shape of a pyrgeometer dome transmittance curve for several R\textsubscript{$\nu$} (T) for a range of typical urban T\textsubscript{emiss} (T\textsubscript{emiss} = 260 - 320 \si{\kelvin}, at 5 \si{\kelvin} increments).
+As each curve interacts with a different portion of the spectral dome transmittance curve (\textit{r\textsubscript{$\nu$}}), and \textit{r\textsubscript{$\nu$}} is spectrally non-uniform, the distinct spectral shapes of R\textsubscript{$\nu$} (T) results in differential dome effects for each T\textsubscript{emiss}.
 
@@
 
-Spectrally non-uniform pyrgeometer dome transmittance results in differential dome effects on the two spectral radiance curves in Figure \ref{nonuniform} and two different radiance signals "seen" by the pyrgeometer.
-Thus, calibration routines designed to remove dome effects and retrieve accurate irradiance values are temperature specific and subject to errors based on the difference between T\textsubscript{emiss} and the temperature at which the instrument was calibrated (T\textsubscript{calib}).
-The magnitude of this error can be expressed as differences Planck weighted mean sensor response ($\overline{r} $) between T\textsubscript{emiss}, computed from equation \ref{meansens} - where lower $\overline{r} $ indicates a larger dome effect on the signal received by the sensor.
-In addition, this error can be expressed as a temperature bias (T\textsubscript{bias}) computed as the difference between a reference T\textsubscript{calib} and a dome adjusted temperature (T\textsubscript{adj}).
-T\textsubscript{adj} is computed by applying a spectral shift to R\textsubscript{Planck} calculated for T = 300 \si{\kelvin} so that its peak spectral radiance matches the wavenumber of peak spectral emission for the target T\textsubscript{emiss}\footnote{In this sensitivity test, the waveband was truncated from 1 - 2500 \si{cm^{-1}} to 80 - 2500 \si{cm^{-1}} to accommodate shorter wavenumber when shifting R\textsubscript{Planck} calculated at the reference T\textsubscript{calib} to lower temperatures where peak spectral emission occurs at smaller wavenumber.
+Spectrally non-uniform pyrgeometer dome transmittance results in differential dome effects on the R\textsubscript{$\nu$} (T) for different T\textsubscript{emiss}.
+Thus, temperature specific calibration routines designed to remove dome effects and retrieve accurate irradiance values are subject to errors based on the difference between T\textsubscript{emiss} and the temperature at which the instrument was calibrated (T\textsubscript{calib}).
+The magnitude of this error can be expressed as differentials in Planck weighted mean sensor response ($\overline{r} $) calculated via equation \ref{meansens} for different T\textsubscript{emiss} - where lower $\overline{r} $ indicates a larger dome effect on the signal received by the sensor.
+In addition, this error can be expressed as a dome adjusted temperature bias (T\textsubscript{bias}) computed as the difference between a reference T\textsubscript{calib} and a dome adjusted temperature (T\textsubscript{adj}).
+With R\textsubscript{$\nu$} (T\textsubscript{adj}) representing a spectral curve with the same integrated radiance as R\textsubscript{$\nu$} (T\textsubscript{calib}) shifted to represent the dome effect incurred at the target T\textsubscript{emiss}.
+For example, to calculate T\textsubscript{bias} for T\textsubscript{emiss} = 270 \si{\kelvin} at T\textsubscript{calib} = 300 \si{\kelvin}, a spectral shift is applied to R\textsubscript{$\nu$} (300 \si{\kelvin}) so that its peak spectral radiance matches the wavenumber of peak spectral emission for R\textsubscript{$\nu$} (270 \si{\kelvin}).
+When integrated, the two curves have the same radiance, but interact with \textit{r} at over different wavebands.
+Following the spectral shift, spectral radiances are convolved by the dome transmittance function, integrated over the waveband and hemisphere, and used to derive T\textsubscript{adj} via equation \ref{stefb1} with $\overline{r}$ at T\textsubscript{calib}.\footnote{In this sensitivity test, the waveband was truncated from 1 - 2500 \si{cm^{-1}} to 80 - 2500 \si{cm^{-1}} to accommodate shorter wavenumber when shifting R\textsubscript{$\nu$} (T\textsubscript{calib}) to lower temperatures where peak spectral emission occurs at smaller wavenumber.
 This results in a slight underestimation of the total radiance at a given T\textsubscript{emiss}, and thus, slightly negative T\textsubscript{bias} at T\textsubscript{calib}, where no spectral shift has been applied.
-When computed from total in band radiance, T\textsubscript{bias} at the reference T\textsubscript{calib} is zero.} - a similar shift is illustrated for a reference T\textsubscript{calib} = 300 \si{\kelvin} and T\textsubscript{emiss} = 270 \si{\kelvin} in Figure \ref{nonuniform}.
-Following the spectral shift, spectral radiances were convolved by the dome transmittance function, integrated over the waveband and hemisphere, and used to derive T\textsubscript{adj} via equation \ref{stefb1} with $\overline{r}$ at T\textsubscript{calib}.
-Table \ref{nonuniformsen} shows $\overline{r}$ and T\textsubscript{bias} and for a range of T\textsubscript{emiss}.
+When computed from total in band radiance, T\textsubscript{bias} at the reference T\textsubscript{calib} is zero.} Table \ref{nonuniformsen} shows $\overline{r}$ and T\textsubscript{bias} and for a range of T\textsubscript{emiss}.
 
@@
 	\centering
-	\caption{Planck weighted mean sensor response and T\textsubscript{bias} for a suite of common urban T\textsubscript{surf}.}
+	\caption{Planck weighted mean sensor response and T\textsubscript{bias} for a suite of common urban emission temperatures.}
 	\label{nonuniformsen}
-	\begin{tabular}{lp{2cm}c}
+	\begin{tabular}{p{1.5cm}p{1.5cm}c}
 		\toprule 
-		Emission & & T\textsubscript{bias}, \si{\kelvin} \\
-		temperature, \si{\kelvin}& $\overline{r}$, \% & computed for T\textsubscript{calib} = 300 \si{\kelvin} \\ \midrule
+		& & T\textsubscript{bias}, \si{\kelvin} \\
+		T\textsubscript{emiss}, \si{\kelvin}& $\overline{r}$, \% & computed for T\textsubscript{calib} = 300 \si{\kelvin} \\ \midrule
 		%Emission temperature, \si{\kelvin}& $\overline{r}$, \% & T\textsubscript{adj}, \si{\kelvin} \\ \midrule
@@
 The magnitude of T\textsubscript{bias} increases as the difference between the reference T\textsubscript{calib} and T\textsubscript{emiss} increases, with approximately equal rates of change with increasing and decreasing T\textsubscript{emiss}.
-Thus, dome effects have a larger effect on the pyrgeometer received signal when comparing irradiances or T\textsubscript{hem} over a wide range of temperatures, or when T\textsubscript{emiss} is significantly different from T\textsubscript{calib}.
-As pyrgeometers used in the BUBBLE campaign were calibrated outdoors under clear-sky, mid-latitude springtime conditions \cite{Christen2005}, the dome effect present in uncorrected BUBBLE $L_z$ measurements is likely smaller in summer than in winter, and is well represented by Table \ref{nonuniformsen}.
+Dome effects have a significant effect on the pyrgeometer received signal when comparing irradiances or T\textsubscript{hem} over a wide range of temperatures, or when T\textsubscript{emiss} is significantly different from T\textsubscript{calib}.
+Pyrgeometers used in the BUBBLE campaign were calibrated outdoors under clear-sky, mid-latitude springtime conditions \citep{Christen2005}.
+Thus, dome effects inherent in uncorrected $L_z$ measurements in this study are likely smaller in summer than in winter (as T\textsubscript{emiss} is likely closer to T\textsubscript{calib} in summer than it is in winter), and are well represented by Table \ref{nonuniformsen}.
 
 By convolving modeled spectral radiances by a dome transmittance curve, the correction method presented in this study accounts for dome effects.
-Thus, correction magnitudes include both atmospheric and dome effects in aggregate (and, crudely, emissivity effects, as modeled radiances are calculated with $\epsilon$ = 0.95 and include a reflected L\textsubscript{sky} component).
+Thus, correction magnitudes include both atmospheric and effects from spectrally non-uniform dome transmittance (and, crudely, emissivity effects, as modeled radiances are calculated with $\epsilon$ = 0.95 and include a reflected L\textsubscript{sky} component).
 While the individual effects are difficult to separate, temperature dependent differential dome effects are visible in correction magnitudes as a persistent minor positive correction in winter.
@@
 As a result, above-sensor T\textsubscript{air} and humidity values are likely underestimated by the standard wintertime atmosphere, reducing the modeled at-sensor reflected component of downwelling longwave in winter and the overall modeled at-sensor signal.
-This results in an increase in both derived T\textsubscript{hem, r} and correction magnitudes as measured irradiances are matched with irradiances modeled at slightly higher T\textsubscript{surf} to make up for the underestimated at-sensor L\textsubscript{sky} signal.
+This results in an increase in derived T\textsubscript{hem, r} and correction magnitudes as measured irradiances are matched with irradiances modeled at slightly higher T\textsubscript{surf} to make up for the underestimated at-sensor L\textsubscript{sky} signal.
 
@@
 Overestimation of T\textsubscript{comp} under clear sky conditions is much larger for T\textsubscript{plan} than T\textsubscript{hem, r}.
-Representing the surface from the nadir, over half of the T\textsubscript{plan} signal from the Sperrstrasse canyon is generated by rooftops with both a large diurnal T\textsubscript{surf} amplitude and the highest daytime facet T\textsubscript{max}.
-Undersampling of wall facets by T\textsubscript{plan} also has a significant effect on the daytime clear-sky warming bias, as walls show much cooler daytime T\textsubscript{max} than T\textsubscript{roof} or T\textsubscript{road} and have a moderating effect on T\textsubscript{comp} and T\textsubscript{hem, r} by day, particularly on hot clear sky days.
+Over half of the T\textsubscript{plan} signal from the Sperrstrasse canyon is generated by rooftops that have both a large diurnal T\textsubscript{surf} amplitude and the highest daytime facet T\textsubscript{max}.
+Undersampling of sloped facets and neglect of wall facets by T\textsubscript{plan} also has a significant effect on the daytime clear-sky warming bias, as walls show much cooler daytime T\textsubscript{max} than T\textsubscript{roof} or T\textsubscript{road} and have a moderating effect on T\textsubscript{comp} and T\textsubscript{hem, r} by day, particularly on hot clear sky days.
 Mean daytime overestimations by T\textsubscript{plan} just after solar noon are approximately $4$ \si{K} over the IOP, approaching $8$ \si{K} on the clear sky case day.
 T\textsubscript{hem, r} shows smaller overestimations of approximately $3$ \si{K} over the IOP and $5$ \si{K} on the clear sky case day.
-Urban T\textsubscript{surf} and the sUHI are often measured from a satellite TIR remote sensors that retrieve T\textsubscript{plan} exclusively under clear-sky conditions and often sample in the nadir.
+Urban T\textsubscript{surf} and the sUHI are often measured from satellite TIR remote sensors that retrieve T\textsubscript{plan} exclusively under clear-sky conditions and often sample in the nadir.
 Thus, overestimation of daytime urban T\textsubscript{comp} inherent in the satellite T\textsubscript{surf} record is best represented by T\textsubscript{plan} on the clear sky case day.
@@
 Nighttime underestimations of T\textsubscript{comp} by nadir and hemispherical views are a result of an undersampling of wall and road facets respectively.
-T\textsubscript{hem, r}, derived from a pyrgeometer skewed towards the south-facing canyon wall oversamples the nearest roof and underestimates the canyon floor compared to T\textsubscript{comp}.
-This results in a large underestimation of nighttime T\textsubscript{comp}.
+T\textsubscript{hem, r} derived from $L_z$ measured at the Sperrstrasse canyon oversamples the nearest roof and underestimates the canyon floor compared to T\textsubscript{comp} as it's location along the canyon axis is skewed towards the south facing wall (rather than the center of the canyon) and its height is approximately 2.17 times mean building height.
+Non-optimal sensor placement in the Sperrstrasse canyon results in a large underestimation of nighttime T\textsubscript{comp} by T\textsubscript{hem, r}.
 However, nighttime underestimation - and to some extent, daytime overestimation - by T\textsubscript{hem, r} is the combined result of a sampling bias inherent in a hemispherical view of the surface and improper sensor placement, discussed in greater detail in section \ref{Methodological limitations and considerations}.
@@
 
- When measured from a near ground downward facing remote sensor, upwelling TIR is sensitive to sensor height and position relative to surface geometry \cite{Adderley2015, Roberts2010}.
-Below 3 times mean roof level, remote sensed TIR varies from a 'true', geometrically representative, TIR signal based on a comparison between view factor proportions of and those of the urban surface.
-For an idealized, orthogonal, block-like urban array, Roberts 2010 found that facet view factor proportions - and thus measured TIR - varied significantly with sensor position and height.
+ When measured from a near ground downward facing remote sensor, upwelling TIR is sensitive to sensor height and position relative to surface geometry \citep{Adderley2015, Roberts2010}.
+Below 3 times mean roof level, remote sensed TIR varies from a "true", geometrically representative, TIR signal based on a comparison between facet view factor proportions of the sensor and those of the urban surface.
+A perfectly geometrically representative remote sensed TIR signal is one measured from a sensor with facet view factor proportions equaling those of the urban surface.
+I.e.
+for an urban surface made up of 50\% wall, 25\% road, and 25\% roof, a perfectly geometrically representative remote sensed signal would be one "seen" by a sensor with the same fractional facet view factor proportions.
+For an idealized, orthogonal, block-like urban array, \citet{Roberts2010} found that facet view factor proportions - and thus measured TIR - varied significantly with sensor position and height.
 To investigate the effect of sensor placement on T\textsubscript{hem, r} for a simplified representation of the Sperrstrasse canyon site, we used SUM to vary sensor position over a simplified DBM and calculated normalized view factor proportions for a 160\si{\degree} FOV sensor.
@@
  Normalized facet view factors show significant variation based on sensor position at both test heights.
-At the lower test height, in the horizontal direction (moving perpendicular to the canyon axis), sensor positions at the canyon center (positions 4, 5, and 6) are most representative of actual facet view factors, with decreasing representivity at the edge of the canyon (positions 1, 2, and 3), and over the rooftops (positions 7, 8, and 9).
-In the vertical direction (moving parallel to the canyon axis), positions slightly offset from the roof edge (positions 2, 5, and 8) agree most closely with actual canyon surface area proportions, with decreasing representivity in positions further from the roof edge.
+At the lower test height, varying sensor position perpendicular to the canyon axis, positions at the canyon center (positions 4, 5, and 6) are most representative of actual facet view factors, with decreasing representivity at the edge of the canyon (positions 1, 2, and 3), and over the rooftops (positions 7, 8, and 9).
+Moving the sensor parallel to the canyon axis, positions slightly offset from the roof edge (positions 2, 5, and 8) agree most closely with actual canyon surface area proportions, with decreasing representivity in positions further from the roof edge.
 Roof and road view factor proportions are most strongly affected by sensor placement.
@@
 	height=7in,
-	keepaspectratio]{placement1}
+	keepaspectratio]{placement2}
 	\caption{A plan view of the simplified DBM showing the nine test sensor placements. Actual sensor placement at the Sperrstrasse canyon is approximated by number 3.}
@@
 
-Results in Roberts 2010 and Adderley et al.
-2015 show that sensitivity to sensor placement decreases with increasing sensor height  Above a height of approximately 3 to 5 times mean building height, depending on surface geometry, a TIR signal becomes spatially invariant and view factor proportions are constant across all sensor placements.
-Results from the higher test height support findings in Roberts 2010 and Adderley et al.
-2015, with decreased spatial variance in facet view factor proportions across the nine sensor positions.
+Results in \citet{Adderley2015} and \citet{Roberts2010} show that sensitivity to sensor placement decreases with increasing sensor height.
+Depending on surface geometry, sensor view factor proportions and the TIR signal "seen" by a sensor become spatially invariant sensor at heights above approximately 3 to 5 times mean building height.
+Results from the higher sensor test height support findings in \citet{Adderley2015} and \citet{Roberts2010}, with decreased spatial variance in facet view factor proportions across the nine sensor positions.
 However increased spatial stability in view factor proportions does not necessarily entail increased geometric representativity.
@@
 
-To visualize how sensor placement affects inferred surface temperatures, T\textsubscript{hem, r} is calculated using view factors as weights in an average of T\textsubscript{wall}, T\textsubscript{road}, and T\textsubscript{roof}.
-Temperatures are normalized by T\textsubscript{comp} and averaged at 30-minute intervals over the IOP, the results of which are shown in figures \ref{vf_all} and \ref{vf_all45} for sensor heights of 2.17 and 3 times mean building height respectively.
-As with view factor proportions, calculated T\textsubscript{hem, r} is highly dependent on relative to T\textsubscript{comp}.
-For the lower sensor height, as predicted by view factor proportions, positions at the center of the canyon (positions 4, 5, and 6) and edge of the building (positions 1, 4, and 7) follow most closely to T\textsubscript{comp}.
+To visualize how sensor placement affects inferred surface temperatures, T\textsubscript{hem, r} is calculated using view factors to calculate a weighted average of T\textsubscript{wall}, T\textsubscript{road}, and T\textsubscript{roof} for each sensor position.
+Temperatures are normalized by T\textsubscript{comp} and averaged at 30-minute intervals over the IOP, the results of which are shown in Figures \ref{vf_all} and \ref{vf_all45} for sensor heights of 2.17 and 3 times mean building height respectively.
+As is the case with view factor proportions, calculated T\textsubscript{hem, r} is highly dependent on sensor placement.
+For the lower sensor height, as predicted by view factor proportions, positions at the center of the canyon (positions 4, 5, and 6) and edge of the building (positions 1, 4, and 7) follow T\textsubscript{comp} most closely.
 By day, all sensor positions overestimate T\textsubscript{comp}, from a systemic bias towards rooftop facets.
@@
 At the higher test height, T\textsubscript{hem, r} is more representative of T\textsubscript{comp} for positions 2, 3, and 5 - 9.
-T\textsubscript{hem, r} from positions 1 and 4 displays a slight systemic overestimation of T\textsubscript{comp}.
+T\textsubscript{hem, r} from positions 1 and 4 display a slight systematic overestimation of T\textsubscript{comp}.
 For these positions, although roof view factor proportion is well represented, road view factor is significantly overestimated, oversampling warmer nighttime road temperatures.
 For all other positions daytime overestimation and nighttime underestimation of T\textsubscript{comp} is reduced.
-As is the case with view factor proportions, T\textsubscript{hem, r} is more spatially stable based on sensor placement.
+As is the case with view factor proportions, T\textsubscript{hem, r} is more spatially stable with respect to sensor placement.
 
@@
  	keepaspectratio]{vf_all}
- 	\caption{Mean T\textsubscript{hem, r} normalized against T\textsubscript{comp} for each sensor position the 14-day IOP. Shaded area indicates quartiles one through three.}
+ 	\caption{Mean T\textsubscript{hem, r} normalized against T\textsubscript{comp} for each sensor position over the 14-day IOP for a sensor height of 2.17 times mean building height.
+Shaded area indicates quartiles one through three.}
  	\label{vf_all}
@@
 	keepaspectratio]{vf_all45}
-	\caption{Mean T\textsubscript{hem, r} normalized against T\textsubscript{comp} for each sensor position the 14-day IOP. Shaded area indicates quartiles one through three.}
+	\caption{Mean T\textsubscript{hem, r} normalized against T\textsubscript{comp} for each sensor position over the 14-day IOP for a sensor height of 3 times mean building height.
+Shaded area indicates quartiles one through three.}
 	\label{vf_all45}
@@
 
-Both Roberts 2010 and Adderley et al.
-2015 did not find a perfectly representative sensor position for typical low- and mid-rise urban street canyons, this is supported by sensitivity testing performed in this study.
+Both \citet{Roberts2010} and \citet{Adderley2015} did not find a perfectly representative sensor position for typical low- and mid-rise urban street canyons; this is supported by sensitivity testing performed in this study.
 As such, under conditions with strong microscale spatial contrasts in T\textsubscript{surf}, T\textsubscript{hem, r} is unlikely to be equal T\textsubscript{comp} regardless of sensor placement.
@@
 This paper details a method for urban T\textsubscript{surf} retrieval using atmospherically corrected hemispherical radiometric surface temperatures derived from continuous near-ground measurements of thermal infrared radiation.
-Atmospherically corrected, hemispherical urban T\textsubscript{surf} is derived using a method which combines a sensor view model (SUM) to represent urban surface geometry and a point-to-point radiative transfer model (MODTRAN 4.1) to model irradiances upwelling from complex surface terrain in 3-dimensions.
+Atmospherically corrected, hemispherical urban T\textsubscript{surf} is derived using a method which combines a sensor view model (SUM) to represent urban surface geometry and a point-to-point radiative transfer code (MODTRAN 4.1) to model irradiances upwelling from complex surface terrain in 3-dimensions.
 At each time step, irradiances are modeled using profiles of T\textsubscript{air} and water vapor content for a range of possible T\textsubscript{hem, r}.
 Irradiance - T\textsubscript{hem, r} pairings are aggregated into a LUT and the measured irradiance is matched with the closest modeled irradiance to retrieve T\textsubscript{hem, r} for a given time step.
-Repeated at 30-minute intervals, this method is used to derive an 8-month climatology of urban T\textsubscript{hem, r} for the Basel Sperrstrasse canyon, instrument as a part of the BUBBLE campaign.
-T\textsubscript{hem, r} at each time step is a view factor weighted temperature representative of the  at which the surfaces contained within the sensor FOV are emitting longwave radiation.
+Repeated at 30-minute intervals, this method is used to derive an 8-month climatology of urban T\textsubscript{hem, r} from urban irradiances measured via a pyrgeometer at 2.17 times mean building height in the Sperrstrasse canyon instrument as a part of the BUBBLE campaign in Basel, Switzerland.
+T\textsubscript{hem, r} at each time step is a view factor weighted temperature representative of the surfaces contained within the sensor FOV.
 
-Atmospheric corrections are large compared to typical sUHI magnitudes \cite{Peng2012} and vary significantly based on season, time of day, and synoptic conditions.
+Atmospheric corrections are large and vary significantly based on sensor-surface-sun geometry and synoptic conditions.
+Thus, for accurate intersite comparison, T\textsubscript{hem} derived from irradiances measured at different heights or above different surface geometries may have different atmospheric or dome effects that can confound results.
+This fact makes clear the need for robust atmospheric correction of T\textsubscript{hem} for sUHI analysis.
 Correction magnitudes are largest on hot, clear sky summertime days - with large solar input and high T\textsubscript{hem, r} and T\textsubscript{air} - and smallest by night and under overcast conditions, during which solar input is suppressed and T\textsubscript{hem, r} and T\textsubscript{air} are low.

### 2017-04-17 paper2.tex

@@
 Urban modification of surface geometry and thermal, radiative, moisture, and aerodynamic properties results in differential surface heating and cooling patterns and strong microscale spatiotemporal variations in urban surface temperature (T\textsubscript{surf}).
-Integrated up to larger scales, urban areas tend to store and generate more heat relative to non-built surroundings which manifests in elevated T\textsubscript{surf} and T\textsubscript{air} --- a phenomenon termed the urban heat island effect (UHI).
+Integrated up to larger scales, urban areas tend to store and generate more heat relative to non-built surroundings which manifests in elevated T\textsubscript{surf} and T\textsubscript{air} a phenomenon termed the urban heat island effect (UHI).
 To foster a more complete understanding of the effect of urban areas climates across scales, accurate, spatiotemporally continuous and geometrically representative characterization of T\textsubscript{surf} in cities has long been a goal in urban climatology.
@@
 	\centering
-	\includegraphics[width=15cm,height=7in,keepaspectratio]{bx_suhi_compare}
+	\includegraphics[width=15cm,height=7in,keepaspectratio]{bxsuhi1}
 	\caption{A comparison of sUHI magnitudes from complete, hemispherical, and nadir remote sensed representations of the Sperrstrasse canyon over the IOP.
@@
 		T\textsubscript{air} & \textit{increasing} \\
-
 		&\\
 		Incoming solar radiation & \textit{increasing}  \\
-		& \\
 		& \\

### 2017-04-19 introduction.tex
Add files via upload

--- 2017-02-01 introduction.tex
@@
 \chapter{Introduction}
+Urban development drastically changes the character of the surface.
+Replacement of natural terrain with roads, buildings, parks, and other urban features modifies the geometry of the surface as well as its thermal, radiative, moisture, and aerodynamic properties.
+These modifications, combined with anthropogenic emissions of heat, result in a distinct urban climate; one which is often warmer, a phenomenon termed the urban heat island effect (UHI).
+As both the Earth's population and the urbanized fraction of that population increase \citep{Nations2014}, cities grow and more people are exposed to urban-modified atmospheres.
+Thus, understanding how cities interact with the climate across spatiotemporal scales has important implications for human health, energy efficiency, and for informing future urban development.
+This, in part, has prompted significant expansion in study of the urban effect on climates, particularly in the decades preceding this thesis.
 
-\section{Defining the urban surface}
- 	Sometimes I have no idea what I am doing.
+The notion that urban areas have elevated air and surface temperatures (T\textsubscript{air} and T\textsubscript{surf} respectively) is not new, with the first formal studies of the urban effect on T\textsubscript{air} dating back to \citet{Howard1833}'s identification of "artificial warming" and "urban contamination" in his characterization of spatial patterns of T\textsubscript{air} in early 19\textsuperscript{th} century London.
+Conceptually, contemporary studies of urban climate do not stray far from \citet{Howard1833}, by seeking to isolate and quantify the urban signal in measured or modeled assessment of a given meteorological variable.
+However, as \citet{Howard1833} identified nearly two centuries ago, the sheer number of factors influencing direct measurement of climates makes experimental control nearly impossible in observational study.
+In any given observation of climate, myriad urban and non urban signals make it difficult to determine what is urban and what is not, degrading interstudy comparison.
+Even in the modeling domain, where experimental control is possible, the inherent complexity in simulating flows of energy, momentum, and mass at the scales necessary to represent physical phenomena over realistic urban environments is, as of yet, not possible.
+Notwithstanding implementation of accurate, physically based schemes to represent urban climates in regional or global climate models.
+Thus, accurate assessment how a city affects the climate is a deceptively challenging task.
+To address this, \citet{Lowry1977} presents a framework to isolate the urban effect in observation of a given meteorological variable.
+In his framework - termed the "Lowry method" - measurement of a meteorological variable ($V_M$) is the sum of forcings from the background climate and synoptic conditions ($V_B$), local topography ($V_L$), and the urban landscape ($V_H$)
 
- \section{The urban radiation budget}
+\begin{equation}
+	V_M = V_B + V_L + V_H
+\end{equation}
+
+Thus, isolation of the urban effect on observed $V_M$ simply requires the removal of $V_B$ and $V_L$.
+This can be achieved by subtracting some urban affected observation ($V_{M, U}$) from a non-urban observation ($V_{M, R}$) where $V_{H, R}$ = 0,
+
+\begin{equation}
+V_{M, U} - V_{M, R} = (V_B + V_L + V_H) - (V_B + V_L) = V_H
+\end{equation}
+
+\noindent simplified as,
+
+\begin{equation}
+	V_H = V_{M, U} - V_{M, R}
+\end{equation}
+
+Study of the urban heat island effect tacitly falls into such a framework, by replacing $V_{M, U} - V_{M, R}$ with $\Delta$T\textsubscript{air, U-R}.
+However, in spite of its conceptual simplicity through the lens of the Lowry method, meta-analysis of observational air temperature UHI (aUHI) study suggests that that simplicty may not translate well to real world analysis.
+In his analysis of the aUHI literature, \citet{Stewart2011} found that nearly fifty percent of aUHI studies published between 1950 and 2007 are "scientifically indefensible" based on an analysis of relevant criteria\footnote{Each study was given a "passing" or "failing" grade and ranked using a points-based scheme to assess methodological quality in the following criteria: conceptual model, operational definitions, instrument specification, site metadata, site representativeness, number of replicates, weather control, surface control, and synchronicity.}, suggesting a significant gulf between conceptual simplicity and practical realities in aUHI analysis.
+Of the 190 surveyed studies, only 13\% were derived from field sites that were sufficiently representative of the local-scale environment or lacked the meta-data to make such a determination, leading to potential biases from micro-scale terrain.
+This is particularly discouraging as it is not the result of an evolution in measurement techniques or improved instrumentation, as techniques to observe and analyze aUHI did not see significant evolution since 1950 - in fact, \citet{Stewart2011} suggests the opposite: a disproportionately large number of studies deemed scientifically sound were published early in the study purview.
+Thus, in study of the urban effect on variables for which significant methodological shifts have or will occur, the warnings implicit in \citet{Stewart2011} are particularly salient, as inherent difficulties found in translating "Lowry"-esque conceptual models to real world observations may be compounded by significant methodological changes.
+In the five years since its publication, aUHI study has seen further expansion and, as such, it is difficult to assess whether its critical analysis has prompted a shift towards more thoughtful, careful, and methodical study of aUHI, notwithstanding its relevance in study of the urban effect on different meteorological variables.
+
+In contrast to study of urban T\textsubscript{air}, the introduction of new methods, instruments, and analytical techniques in TIR remote sensing of urban environments has lead to rapid evolution in study of the urban effect on land T\textsubscript{surf} - a vast majority of which has focused on passive remote sensing of upwelling TIR.
+In the decades preceding this thesis, a combination of factors has lead to an increased focus on remote sensed study of urban T\textsubscript{surf}:
+
+\begin{itemize}
+	\item The proliferation of satellite and aerial thermal infrared (TIR) remote sensors and "openness" in data access policies.
+	\item Improvements in sensor spatial, spatial, and radiometric resolutions. 
+	\item An increased focus on the importance of the surface in both determining and understanding key near-ground micro-meteorological phenomena.
+\end{itemize}
+
+Remote sensed study of the urban effect on land T\textsubscript{surf} has been instrumental in characterizing large-scale spatial and temporal patterns of sUHI \citep{Peng2012, Streutker2003, Imhoff2010}.
+However, the methodological and conceptual concerns raised in \citet{Stewart2011} in combination with the wide (and broadening) range of methodologies for remote sensed urban T\textsubscript{surf} retrieval should prompt critical assessment of the remote sensed sUHI literature.
+Difficulties suggested in \citet{Stewart2011} may be amplified by methodological shifts and a variety of methods for analysis.
+At the time of writing, no systemic or critical reviews of the remote sensed sUHI literature are available.
+An increasingly dogmatic focus on improving spatial and radiometric resolutions has left other inherent biases in remote sensing of urban environments largely ignored in sUHI literature.
+This suggests that similar conclusions to those found of aUHI literature in \citet{Stewart2011} are possible (if not likely) for study of sUHI.
+
+Traditional methods for urban T\textsubscript{surf} measurement are subject to a suite of geometric and temporal biases.
+These biases are largely inherent to the discipline and contribute an additional source of error on top of potential concerns carried over from meta-analysis of conceptual flaws in aUHI literature.
+
+Geometric biases in TIR remote sensing of the urban surface are the result of urban modification of surface structure and thermal and radiative properties.
+Relative to typical non-urban and rural terrain, urban development adds significant 3-dimensionality to the surface.
+This added complexity results in undersampling of the surface by narrow-FOV remote sensors.
+A narrow-FOV sensor, by sampling surface 3-dimensionality in 2-dimensions, "sees" a different assemblage of urban surfaces depending on its viewing direction, resulting in a geometric bias towards some arrangement of facets.
+I.e.
+viewing from the nadir, a narrow-FOV sensor sees the urban surface as an array of rooftops and roads, the same sensor viewing perpendicular to the surface sees wall facets exclusively, and one viewing from some off nadir angle sees some array of walls, rooftops, and roads depending on its orientation.
+
+In addition, the convoluted 3-dimensional structure of the urban surface modifies sunlit/shading regimes, resulting in strong micro-scale spatiotemporal contrasts in urban T\textsubscript{surf}, which depend on surface-sun geometry.
+These microscale variations in urban T\textsubscript{surf} are often amplified by significant directional contrasts in the thermal admittance of common urban materials, which determines the diurnal amplitude of T\textsubscript{surf} for a given facet.
+Thus, when viewed from a narrow-FOV remote sensor with some inherent geometric bias, urban T\textsubscript{surf} is directionally dependent - constituting an "effective anisotropy" of urban T\textsubscript{surf}, shown in Figure \ref{anisot}.
+The qualifier "effective" is used to differentiate directional contrasts in urban T\textsubscript{surf} that arise from a city's 3-dimensional surface structure from those resulting from the non-Lambertian nature of individual urban facets (e.g.
+walls, rooftops, or roads).
+The magnitude of urban effective anisotropy can reach up to 10 \si{\kelvin} and is highly dependent on surface-sensor-sun geometry, surface structure, and urban materials \citep{Krayenhoff2016, Voogt1997}.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=15cm,
+	height=7in,
+	keepaspectratio]{anisot}
+	\caption{A narrow-FOV sensor viewing an idealized urban surface from two angles.
+Left: viewing the surface approximately perpendicular to the sun's angle.
+Right: viewing the surface approximately parallel to the sun's angle.
+The two viewing angles described in the left and right panels yield different remote sensed T\textsubscript{surf} by sampling different arrangements of sunlit and shaded features.
+Both will deviate from an area weighted "complete" urban T\textsubscript{surf}.}
+	\label{anisot}
+\end{figure}
+
+Geometric undersampling by narrow-FOV remote sensors results in remote sensed urban T\textsubscript{surf} changing based on a sensor's viewing direction and the assemblage facets included within the sensor FOV.
+A sensor viewing the surface perpendicular (parallel) to the sun's angle will tend to underestimate (overestimate) the "true" complete urban surface temperature - often calculated as an area weighted average of wall, rooftop, and road T\textsubscript{surf} - by differentially biasing sunlit or shaded facets.
+Similarly, a sensor in the nadir will tend to overestimate daytime urban T\textsubscript{surf} and underestimate nighttime T\textsubscript{surf}, from a bias towards differentially hot (cool) rooftop facets by day (night) and a neglect of wall facets which are differentially cooler (warmer) by day (night).
+As the effect of geometric biases is highly dependent on surface-sensor-sun geometry and requires significant instrumentation for quantification, its influence in the remote sensed sUHI record is presently unknown.
+
+Temporal biases in thermal remote sensing of urban areas occur across multiple time scales including:
+
+\noindent---\textit{Contamination by turbulence forced, high frequency fluctuations in urban T\textsubscript{surf}}.
+
+Using time-sequential thermography \citet{Christen2012} found that many common urban fabric types (e.g.
+rooftops, walls, roads, and vegetation) display large micro-scale (second to minute) fluctuations in T\textsubscript{surf}.
+The magnitude of which is inversely related to surface thermal admittance and is, as of writing, poorly understood.
+Most thermal remote sensors provide instantaneous, rather than temporally averaged, T\textsubscript{surf} and thus are subject to contamination by high-frequency fluctuations.
+These biases are difficult to estimate in urban environments, where a large variety of fabric materials can produce significant directional contrasts in thermal admittance - and thus spatial variations in the magnitude of microscale fluctuations depending on the facet material types viewed by the sensor.
+
+\noindent---\textit{Discontinuity in satellite overpass cycles.}
+
+Satellite overpass cycles are discontinuous.
+Assessment of land T\textsubscript{surf} either sacrifices spatial resolution for a daily repeat cycle - as is the case with MODIS  - or sacrifices temporal resolution for a higher spatial resolution - ASTER or Landsat.
+Geostationary satellites do not currently have sufficient spatial resolution to resolve coherent urban pixels.
+
+\noindent---\textit{Clear-sky bias.}
+
+Clouds absorb TIR – thus satellite TIR remote sensing of the surface requires clear sky conditions.
+Although the urban effect on T\textsubscript{surf} is most evident under "satellite friendly" clear sky conditions (manifesting as large sUHI magnitudes), a clear sky bias likely entails overestimation of "all-sky" sUHI and further adds to discontinuities in the remote sensed urban T\textsubscript{surf} record.
+
+As is the case with geometric biases, the magnitude of these temporal biases in remote sensed urban T\textsubscript{surf} is difficult to quantify without long-term ground truthing campaigns or significant interpolation.
+Thus, the true temporal and geometric nature of urban effect of T\textsubscript{surf} is presently unknown.
+
+\section{Research questions and objectives}
+
+Prompted by these shortcomings in the remote sensed urban T\textsubscript{surf} record, this thesis introduces a method to provide geometrically representative, temporally continuous urban T\textsubscript{surf} for sUHI analysis to address the following questions:
+
+\begin{itemize}
+	\item What is the nature of urban surface temperature when viewed from a hemispherical downward-facing radiometer?
+And how does it relate to urban temperatures derived from other methods for urban surface temperature retrieval?
+	\item What is the diurnal and seasonal nature of the surface urban heat island effect?
+\end{itemize}
+
+\noindent In an attempt to answer these questions, I did the following:
+
+\begin{enumerate}
+	\item Developed and evaluated a method to retrieve atmospherically correction hemispherical radiometric urban surface temperatures from time-continuous measurements of upwelling longwave radiation.
+	\item Compared urban surface temperatures and surface urban heat island magnitudes retrieved using the method to common remote sensed representations of the urban surface.
+	\item Derived an eight month climatology of hemispherical urban surface temperatures to observe seasonal and diurnal patterns of the surface urban heat island effect.
+\end{enumerate}
 
 \section{The structure of the following sections}
+
+The following two papers make up the body of the thesis, read them you fuck.

### 2017-04-28 introduction.tex
Add files via upload

@@
+\begin{bibunit}
+
+\rhead{\thepage}
+
 \chapter{Introduction}
@@
 I.e.
-viewing from the nadir, a narrow-FOV sensor sees the urban surface as an array of rooftops and roads, the same sensor viewing perpendicular to the surface sees wall facets exclusively, and one viewing from some off nadir angle sees some array of walls, rooftops, and roads depending on its orientation.
+viewing from the nadir, a narrow-FOV sensor sees the urban surface as an array of rooftops and roads, the same sensor viewing the from an oblique angle sees some array of walls, rooftops, and roads depending on its orientation.
+For a sensor in the nadir, increasing FOV introduces sampling of vertical features.
+To illustrate how sensor-surface geometry influences sampling of the urban surface by narrow- and wide-FOV sensors, Figure \ref{model} shows FOVs for the three sensor orientation/type combinations.The vast majority of urban thermal remote sensing is done via a narrow-FOV sensor in the nadir.
 
-In addition, the convoluted 3-dimensional structure of the urban surface modifies sunlit/shading regimes, resulting in strong micro-scale spatiotemporal contrasts in urban T\textsubscript{surf}, which depend on surface-sun geometry.
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=15cm,
+	height=7in,
+	keepaspectratio]{model}
+	\caption{Projected FOVs for a narrow-FOV sensor at nadir and oblique angles, and a wide-FOV hemispherical sensor viewing an idealized urban array with building width T\textit{H}, height 2\textit{H}, and inter element spacing textit{H}.
+Figure from \cite{Adderley2015}.}
+	\label{model}
+\end{figure}
+
+In addition to modifying how a remote sensor samples the surface, the convoluted 3-dimensional structure of the urban surface modifies sunlit/shading regimes, resulting in strong micro-scale spatiotemporal contrasts in urban T\textsubscript{surf}, which depend on surface-sun geometry.
 These microscale variations in urban T\textsubscript{surf} are often amplified by significant directional contrasts in the thermal admittance of common urban materials, which determines the diurnal amplitude of T\textsubscript{surf} for a given facet.
@@
 Right: viewing the surface approximately parallel to the sun's angle.
-The two viewing angles described in the left and right panels yield different remote sensed T\textsubscript{surf} by sampling different arrangements of sunlit and shaded features.
+The two viewing angles yield different remote sensed T\textsubscript{surf} by sampling different arrangements of sunlit and shaded features.
 Both will deviate from an area weighted "complete" urban T\textsubscript{surf}.}
@@
 
-Prompted by these shortcomings in the remote sensed urban T\textsubscript{surf} record, this thesis introduces a method to provide geometrically representative, temporally continuous urban T\textsubscript{surf} for sUHI analysis to address the following questions:
+Prompted by these shortcomings in the remote sensed urban T\textsubscript{surf} record, and in an effort to better understand the temporal and geometric aspects of sUHI, this thesis introduces a method to provide hemispherical, temporally continuous urban T\textsubscript{surf} for sUHI analysis to address the following questions:
 
-\begin{itemize}
+\begin{enumerate}
 	\item What is the nature of urban surface temperature when viewed from a hemispherical downward-facing radiometer?
@@
 	\item What is the diurnal and seasonal nature of the surface urban heat island effect?
-\end{itemize}
+\end{enumerate}
 
@@
 
-\begin{enumerate}
+\begin{itemize}
 	\item Developed and evaluated a method to retrieve atmospherically correction hemispherical radiometric urban surface temperatures from time-continuous measurements of upwelling longwave radiation.
@@
 	\item Derived an eight month climatology of hemispherical urban surface temperatures to observe seasonal and diurnal patterns of the surface urban heat island effect.
-\end{enumerate}
+\end{itemize}
 
-\section{The structure of the following sections}
+The following two chapters address questions one and two respectively.
+Chapter \ref{paper1} details and evaluates a method to retrieve atmospherically corrected, hemispherical urban T\textsubscript{surf} from time-continuous measurements of upwelling longwave radiation.
+Chapter \ref{paper2} explores the temporal and geometric nature of sUHI through an eight month climatological analysis of hemispherical sUHI and compares sUHI magnitudes from several representations of the urban surface.
 
-The following two papers make up the body of the thesis, read them you fuck.
+\cleardoublepage 
+\phantomsection  
+\renewcommand*{\bibname}{References}
+\addcontentsline{toc}{section}{\textbf{References}}
+
+\putbib
+\end{bibunit}

### 2017-04-28 paper1.tex
Add files via upload

@@
-\chapter{A hemispherical atmospheric correction method to derive time-continuous urban surface temperature}\label{paper1}
+\begin{bibunit}
+
+\rhead{\thepage}
+
+\chapter{An atmospheric correction method to derive hemispherical time-continuous urban surface temperature}\label{paper1}
 
@@
 
+\section{A practical parameterization}
+
+To save computational time and facilitate broader use of the method, a parameterization scheme was developed to simplify the correction process.
+In the parameterization, a suite of radiative transfer simulations is run prior to correction (as opposed to at each time step) to retrieve $\tau$ for a range of humidities and path lengths.
+T\textsubscript{surf} and T\textsubscript{air} are held constant in these simulations as they have a relatively small effect on $\tau$.
+Simulations for path lengths from 1 to 100 \si{\meter} and water vapor mass densities from 0.1 to 23.5 \si{\gram\per\meter\cubed} are included in Appendix \ref{appendixa}.
+Sensor-surface geometry is approximated as a single view factor weighted, azimuthally averaged surface-sensor path length calculated using the surface-sensor-sun urban model \citep{Soux2004} and a simplified DBM.
+This path length represents the average distance from the sensor to the surface as "seen" by the sensor.
+For each time step, measured humidity is used to infer a hemispherical transmittance ($\overline{\tau_{\Phi}}$) for the approximated surface-sensor geometry.
+A parameterized T\textsubscript{hem, r} can then be calculated by using measured (in this case, profile averaged) T\textsubscript{air} to decompose measured L\textsubscript{up} into irradiance received by the sensor from the surface $L_{surf}^{at-sensor}$ and from the atmosphere $L_{atm}^{at-sensor}$ via,
+
+\begin{equation}
+	\label{para}
+	T_{hem, ~para} =\left(\displaystyle\frac{\biggl(\cfrac{\displaystyle L_{surf}^{at-sensor}}{ \displaystyle\tau}\biggr)}{ \displaystyle\sigma}\right)^{0.25}
+\end{equation}
+\noindent where $\tau$ is calculated using a radiative transfer code and a we and $L_{surf}^{at-sensor}$ is the fraction of ,
+
+\begin{equation}
+	\label{para2}
+	\displaystyle L_{surf}^{at-sensor} = L_{total}^{at-sensor} - (1-\tau)\sigma T_{air}^4
+\end{equation}
+
+Using this parameterization, a climatology of T\textsubscript{hem, para} for a given surface geometry can be retrieved from approximately 100 simulations as opposed to the many millions needed for full hemispherical radiative transfer simulation.
+A variety of online and offline, open and closed source resources are available for radiative transfer simulation \citep{Gastellu-Etchegorry1996,Berk1987,Buehler2005}, some of which are available for free or at a low cost (Spectral-Calc), all of which are likely to provide similar results to those included in Appendix \ref{appendixa}.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=15cm,
+	height=7in,
+	keepaspectratio]{para_eval}
+	\caption{Modeled T\textsubscript{hem, r} versus T\textsubscript{hem, r} derived via the parameterization scheme.}
+	\label{para_eval}
+\end{figure}
+
+\begin{table}[H]
+ 	\centering
+ 	\caption{Statistical performance of T\textsubscript{hem} derived using the parameterization scheme relative to modeled T\textsubscript{hem,r}.
+RMSE\textit{\textsubscript{s}} and RMSE\textit{\textsubscript{u}} represent the systemic and unsystematic RMSE respectively.
+n = 853}
+ 	\label{parastats}
+ 	\begin{tabular}{lcl}
+ 		\toprule 
+ 		Statistic & Parameterization & Citation \\  \midrule
+ 		Slope & 1.058 &  \\ 
+ 		Intercept, (\si{\kelvin}) & -1.488 &   \\ 
+ 		\textit{R\textsuperscript{2}} & 0.991 & \\ 
+ 		MAE,  (\si{\kelvin}) & 0.607& \citep{Willmott1985a} \\ 
+ 		RMSE,  (\si{\kelvin}) & 0.757 & \citep{Willmott1985a} \\ 
+ 		RMSE\textit{\textsubscript{s}},  (\si{\kelvin}) & 0.476  & \citep{Willmott1985a} \\ 
+ 		RMSE\textit{\textsubscript{u}},  (\si{\kelvin}) & 0.588  & \citep{Willmott1985a} \\ 
+ 		\textit{d}, agreement index & 0.998 & \citep{Willmott2012} \\
+ 		\bottomrule
+ 	\end{tabular} 
+\end{table}
+
  \section{Conclusions}
@@
 However, the measures used to derive T\textsubscript{hem, r} are common to most energy balance assessments and constitute an untapped, but promising, resource to improve understanding of how the built environment affects land T\textsubscript{surf} and to quantify the geometric and temporal biases inherent in satellite TIR remote sensing of complex terrain.
+
+\cleardoublepage 
+\phantomsection  
+\renewcommand*{\bibname}{References}
+\addcontentsline{toc}{section}{\textbf{References}}
+
+\putbib
+\end{bibunit}

### 2017-04-28 paper2.tex
Add files via upload

@@
+\begin{bibunit}
+
+\rhead{\thepage}
+
 \chapter{A climatology of urban surface heat islands derived from hemispherical radiometric surface temperatures}
+\label{paper2}
 
@@
 Urban modification of surface geometry and thermal, radiative, moisture, and aerodynamic properties results in differential surface heating and cooling patterns and strong microscale spatiotemporal variations in urban surface temperature (T\textsubscript{surf}).
-Integrated up to larger scales, urban areas tend to store and generate more heat relative to non-built surroundings which manifests in elevated T\textsubscript{surf} and T\textsubscript{air} a phenomenon termed the urban heat island effect (UHI).
-To foster a more complete understanding of the effect of urban areas climates across scales, accurate, spatiotemporally continuous and geometrically representative characterization of T\textsubscript{surf} in cities has long been a goal in urban climatology.
+Integrated up to larger scales, urban areas tend to store more heat relative to non-built surroundings; traffic, HVAC systems, and industrial processes contribute to an additional anthropogenic heat input.
+These manifest in elevated T\textsubscript{surf} and T\textsubscript{air} in cities, a phenomenon termed the urban heat island effect (UHI).
+To foster a more complete understanding of the effect of urban areas on climates across a range of scales, accurate, spatiotemporally continuous and geometrically representative characterization of T\textsubscript{surf} in cities has long been a goal in urban climatology.
 The proliferation of satellite and aerial thermal infrared (TIR) remote sensing has enabled spatially-extensive characterizations of surface climates at ever improving spatial and spectral resolutions.
-Such campaigns have elucidated urban T\textsubscript{surf} and surface urban heat island (sUHI) patterns globally at large spatial scales.
-However, technological improvements in TIR remote sensing have yet to address a three potential sources of error when applied in urban areas:
+Such campaigns have elucidated urban T\textsubscript{surf} and surface urban heat island (sUHI) patterns globally at large spatial scales \cite{Peng2012,Zhao2014}.
+However, technological improvements in TIR remote sensing have yet to address three potential sources of error that arise when applied in urban areas:
 
 \begin{enumerate}
-	\item Geometric undersampling of 3-dimensional terrain
-	\item Temporal discontinuity in overpass cycles and sensor sampling regimes
-	\item Clear-sky bias
+	\item Geometric undersampling of 3-dimensional surface structure
+	\item Satellite overpass cycles and sensor sampling regimes that result in temporal discontinuity of the surface temperature time series
+	\item Clear-sky bias that contribute to an overestimation of "all-sky" surface temperatures
 \end{enumerate}
@@
 atmospheric and emissivity effects).
-Meta-analysis of air temperature (T\textsubscript{air}) UHI literature shows that these goals are rarely satisfied \citep{Stewart2011}.
+Meta-analysis of the air temperature (T\textsubscript{air}) UHI literature shows that these goals are rarely satisfied \citep{Stewart2011}.
 Given the relative difficulty in retrieving accurate, representative urban T\textsubscript{surf}, similar conclusions are likely for sUHI analysis.
 In spite of this fact, and the short period over which large-scale, generalizable methods for measuring urban T\textsubscript{surf} have been available, study of sUHI via TIR remote sensing has expanded significantly in the last twenty years \citep{Peng2012,Voogt2003}.
-Using the method described in Chapter \ref{paper1} we derive an 8-month climatology of hemispherical radiometric urban T\textsubscript{surf} (T\textsubscript{hem, r}) from time continuous near-ground upwelling TIR as measured from an inverted pyrgeometer.
+Using the method described in Chapter \ref{paper1} an 8-month climatology of hemispherical radiometric urban T\textsubscript{surf} (T\textsubscript{hem, r}) was derived from time continuous near-ground upwelling TIR as measured from an inverted pyrgeometer.
 The method was developed to address and overcome biases inherent in traditional methods for urban T\textsubscript{surf} retrieval by providing temporally continuous, geometrically representative\footnote{A hemispherical view is not perfectly geometrically representative of urban surface geometry.
@@
 The resulting microscale spatiotemporal contrasts in urban T\textsubscript{surf} create a directional dependence in observed urban T\textsubscript{surf} when measured from conventional narrow-field-of-view (FOV) TIR remote sensing platforms.
-Thus, remote sensed urban T\textsubscript{surf} varies based on sensor FOV, viewing angle and direction, and sun-surface geometry - this directional dependence of urban surface temperature is termed "effective thermal anisotropy" \citep{Voogt1998a}.
-Traditional satellite or airborne remote sensing platforms, by viewing the surface in the nadir, sample only a fraction of the complete urban surface and fail to capture this effect - leading to spatiotemporally variant directional biases of up to 10 \si{\kelvin} in observed urban T\textsubscript{surf} \citep{Voogt1995}.
+Thus, remote sensed urban T\textsubscript{surf} varies based on sensor FOV, viewing angle and direction, and sun-surface geometry.
+This directional dependence of urban surface temperature is termed "effective thermal anisotropy" \citep{Voogt1998a}.
+Traditional satellite or airborne remote sensing platforms, by viewing the surface in the nadir, sample only a fraction of the complete urban surface, leading to spatiotemporally variant directional biases in remote sensed urban T\textsubscript{surf} of up to 10 \si{\kelvin} \citep{Voogt1995}.
 In general, geometric undersampling by a remote sensor in the nadir manifests as an overestimation of daytime T\textsubscript{surf} and an underestimation of nighttime T\textsubscript{surf} \citep{Adderley2015}.
@@
 In addition to undersampling the urban surface, most thermal remote sensing platforms yield an instantaneous ‘snap-shot’ and cannot characterize time-continuous T\textsubscript{surf} patterns without sacrificing ground resolution.
-Temporal discontinuities in thermal remote sensing result in myriad of potential sources of bias over a wide range of time scales.
+Temporal discontinuities in thermal remote sensing result in myriad potential sources of bias over a wide range of time scales.
 Aerial and satellite thermal remote sensing require clear sky conditions (clouds are opaque with respect to thermal infrared radiation).
@@
 Thus, analysis of sUHI patterns and magnitudes across cities and instrument platforms is difficult.
-At smaller time scales still, time-continuous analysis of urban T\textsubscript{surf} shows significant microscale (second to minute) fluctuations in temperature \citep{Christen2012}.
+At shorter time scales still, time-continuous analysis of urban T\textsubscript{surf} shows significant microscale (second to minute) fluctuations in temperature \citep{Christen2012}.
 Most thermal remote sensors provide instantaneous T\textsubscript{surf} (rather than temporally averaged) and are potentially contaminated by high-frequency microscale fluctuations in urban T\textsubscript{surf}.
-This is particularly salient in urban environments, where a large variety of fabric materials can produce significant directional contrasts in thermal admittance - and thus spatial variations in the magnitude of microscale fluctuations depending on the facet material types viewed by the sensor.
+This is particularly salient in urban environments, where a large variety of fabric materials can produce significant spatial contrasts in thermal admittance (for instance, roof materials tend to have much lower thermal admittances than wall or road materials.
+Such large spatial contrasts in thermal admittance are generally not present in non-urban or rural environments).
+As a result, the magnitude of potential biases in remote sensed urban T\textsubscript{surf} from microscale fluctuations depends on the facet material types viewed by the sensor and can change with viewing direction and sensor FOV.
 The effect of this phenomenon on thermal remote sensing has not been extensively studied, however, the magnitude of microscale fluctuations in T\textsubscript{surf} is significant relative to a typical sUHI signal and thus constitutes a potentially large source of bias.
@@
 The magnitude of these biases has not been extensively studied, particularly from a long term, climatological perspective.
-This study presents the first time-continuous, climatological analysis of sUHI retrieved from radiometric hemispherical urban T\textsubscript{surf} retrieved via the correction method described in Chapter \ref{paper1}.
+These biases lead us to ask: What is the nature of the true sUHI?
+How does it vary temporally and over a full range of weather conditions?
+To address these questions, this study presents the first time-continuous, climatological analysis of sUHI retrieved from radiometric hemispherical urban T\textsubscript{surf} retrieved via the correction method described in Chapter \ref{paper1}.
 These measures are used to assess the magnitude of and overcome geometric and temporal biases inherent in urban TIR remote sensing.
@@
 
-This section describes the study area and provides a brief overview of the multiple line-of-sight atmospheric correction method used to derive a time-continuous eight month climatology urban T\textsubscript{hem, r} for sUHI analysis.
+This section describes the study area and provides a brief overview of the multiple line-of-sight atmospheric correction method used to derive a time-continuous eight month climatology of urban T\textsubscript{hem, r} for sUHI analysis.
 A thorough discussion of the method is included in Chapter \ref{paper1}.
@@
 First, surface-to-sensor path length and view factor geometries are calculated with the surface-sensor-sun urban model (SUM) \citep{Soux2004}.
-Using a digital building model, SUM calculates path lengths from the surface to the sensor and angular view factors by projecting the pyrgeometer FOV onto a simplified (orthogonal) representation of the surface, determining which points in the DBM are 'seen' by the sensor, and calculating the distance from each 'seen' point to the sensor.
+Using a digital building model, SUM calculates path lengths from the surface to the sensor and angular view factors by projecting the pyrgeometer FOV onto a simplified representation of the surface, determining which points in the DBM are 'seen' by the sensor, and calculating the distance from each 'seen' point to the sensor.
 Path lengths are binned at 5\si{\degree} intervals and averaged over the azimuth angle.
@@
 Spectral radiances are modeled over a bandpass of 1 to 2500 \si{cm^{-1}} at an emissivity of 0.95 and convolved by an extended dome transmittance function to ensure that modeled radiances replicate the actual remote sensed signal.
-Consultation with radiometer manufacturers made clear the need to extend the modeled bandpass beyond a typical longwave bandpass (approximately 250 - 2500 \si{cm^{-1}}), as silicone domed pyrgeometers are transmissive of radiation in much shorter wavenumber (longer wavelengths) than 250 \si{cm^{-1}} (40 \si{\micro\meter}).
+Consultation with radiometer manufacturers made clear the need to extend the modeled bandpass beyond a typical longwave bandpass (approximately 250 - 2500 \si{cm^{-1}}), as silicone domed pyrgeometers transmit radiation at much smaller wavenumbers (longer wavelengths) than 250 \si{cm^{-1}} (40 \si{\micro\meter}).
 
-Third, radiances for each angular path length are integrated over the waveband and multiplied their respective normalized view factors.
+Third, radiances for each angular path length are integrated over the waveband and multiplied by their respective normalized view factors.
 Weighted and corrected angular radiances are integrated over the hemisphere to yield a modeled at-sensor irradiance as ‘seen’ by the pyrgeometer for the target T\textsubscript{hem, r}.
 
-For each time step, the above steps are repeated at an interval of 0.5 \si{\kelvin} for a predefined range of potential T\textsubscript{hem, r}, with results aggregated into a lookup table relating modeled irradiances to corrected T\textsubscript{hem, r} for the observed ambient T\textsubscript{air} and humidities.
+For each time step, these steps are repeated at an interval of 0.5 \si{\kelvin} for a predefined range of potential T\textsubscript{hem, r}, with results aggregated into a lookup table relating modeled irradiances to corrected T\textsubscript{hem, r} for the observed ambient T\textsubscript{air} and humidities.
 The measured irradiance is then matched with the closest modeled irradiance to return the associated corrected T\textsubscript{hem, r} for the target time step.
@@
 
-\noindent where $L_z^\uparrow$ is at-sensor upwelling longwave radiation, $L_{sky}^\downarrow$ is at-sensor reflected downwelling longwave radiation, and emissivity ($\epsilon$) = 0.98.
+\noindent where $L_z^\uparrow$ is at-sensor upwelling longwave radiation, $L_{sky}^\downarrow$ is at-sensor reflected downwelling longwave radiation, and emissivity ($\epsilon$) = 0.94.
 
@@
 Rural irradiances in this study were measured from approximately 2 \si{\meter} above ground, thus rural T\textsubscript{hem, b} is approximately equal to T\textsubscript{hem, r}.
+
+\subsection{Evaluation and parameterization}
+
+To evaluate the method
 
@@
 Plan and complete aspect ratios for the Sperrstrasse canyon are used to compute weighting schemes to retrieve complete and plan surface temperatures (T\textsubscript{comp} and T\textsubscript{plan} respectively) at 30 minute intervals over the IOP.
-T\textsubscript{comp} represents an area weighted average complete urban T\textsubscript{surf}, while T\textsubscript{plan} represents the Sperrstrasse site as viewed by a satellite or aerial remote sensor in the nadir.
+T\textsubscript{comp} represents a complete urban T\textsubscript{surf}, where facet temperatures are weighted and averaged based on their areal proportion of the complete urban surface area.
+T\textsubscript{plan} represents the Sperrstrasse site as viewed by a satellite or aerial remote sensor in the nadir.
 Weighting schemes for T\textsubscript{plan} and T\textsubscript{comp} are included in Table \ref{weightings}.
-Urban T\textsubscript{surf} derived from complete, plan, and hemispherical representations of the Sperrstrasse site are used to investigate the effect of sensor-surface geometry on remote sensed T\textsubscript{surf} and sUHI, and to quantify the magnitude of geometric biases time-continuously over a range of synoptic conditions.
+Urban T\textsubscript{surf} derived from complete, plan, and hemispherical representations of the Sperrstrasse site are used to investigate the effect of sensor-surface geometry on remote sensed T\textsubscript{surf} and sUHI, and to quantify the magnitude of geometric biases over a range of time scales and synoptic conditions.
 For comparison of sUHI to air temperature UHI, canopy layer air temperature UHI (clUHI) magnitudes are calculated as the difference between urban and rural T\textsubscript{air} measured from 2 \si{\meter} above ground level.
@@
 Figure \ref{heatsuhi} shows seasonal variations in diurnal patterns of mean monthly hemispherical sUHI magnitudes at 30 minute intervals for the eight month study period.
-sUHI displays significant seasonal variation, particularly by day, and is strongly controlled by day length and solar angle.
+The sUHI displays significant seasonal variation, largest in the daytime hours, and is strongly controlled by day length and solar angle.
 Nighttime sUHI does not vary significantly across seasons.
@@
 	\includegraphics[width=15cm,height=7in,keepaspectratio]{heatsuhi}
-	\caption{A heatmap of mean half hourly hemispherical sUHI for each month calculated at 30-minute intervals over the eight month study period.}
+	\caption{A heatmap of mean half hourly hemispherical sUHI for each month calculated at 30 \si{\min} intervals over the eight month study period. Results are interpolated between months.}
 	\label{heatsuhi}
@@
 	\includegraphics[width=15cm,height=7in,keepaspectratio]{heatatuhi}
-	\caption{A heatmap of mean half hourly canopy layer UHI for each month calculated at 30-minute intervals from T\textsubscript{air} measured at approximately 2 \si{\meter} above ground level over the eight month study period.}
+	\caption{A heatmap of mean half hourly canopy layer UHI for each month calculated at 30 \si{\min} intervals from T\textsubscript{air} measured at approximately 2 \si{\meter} above ground level over the eight month study period.
+Results are interpolated between months.}
 	\label{heatcluhi}
@@
 Further application of the method in other urban areas can help to facilitate a better understanding of the true nature of sUHI across myriad time scales and help to quantify the magnitude of geometric and temporal biases inherent in the satellite sUHI record.
+
+\cleardoublepage 
+\phantomsection  
+\renewcommand*{\bibname}{References}
+\addcontentsline{toc}{section}{\textbf{References}}
+
+\putbib
+\end{bibunit}

### 2017-04-28 uw-ethesis-frontpgs.tex
Add files via upload

@@
 \pagestyle{plain}
-\setcounter{page}{2}
+\setcounter{page}{3}
 
@@
 % In a two-sided printing style, it also makes the next page a right-hand (odd-numbered) page, producing a blank page if necessary.
-
-% D E C L A R A T I O N   P A G E
-% -------------------------------
-  % The following is a sample Delaration Page as provided by the GSO
-  % December 13th, 2006.  It is designed for an electronic thesis.
-  \noindent
-I hereby declare that I am the sole author of this thesis. This is a true copy of the thesis, including any required final revisions, as accepted by my examiners.
-
-  \bigskip
-
-  \noindent
-I understand that my thesis may be made electronically available to the public.
-
-\cleardoublepage
 
@@
 
-\begin{center}\textbf{Acknowledgements}\end{center}
+\begin{center}\textbf{Acknowledgments}\end{center}
 
@@
 
-% Change page numbering back to Arabic numerals
+% Change page numbering back to Arabic numerals and have it in the upper right corner
 \pagenumbering{arabic}
+\renewcommand{\headrulewidth}{0pt}
+\pagestyle{fancy}
+\fancyhf{}
+\rhead{\thepage}
+\makeatletter
+\let\ps@plain\ps@fancy % plain style = fancy style
+\makeatother

### 2017-05-03 conclusion.tex
Add files via upload

\begin{bibunit}

\rhead{\thepage}

\chapter{Concluding statements}

This study presents the first long-term, time-continuous analysis of urban T\textsubscript{surf} to contribute to a more complete understanding of the temporal and geometric nature of the urban effect on land T\textsubscript{surf} and sUHI.

Chapter \ref{paper1} details a hemispherical atmospheric correction method that uses version 4.1 of the MODTRAN radiative transfer code \citep{Berk1987} in conjunction with the Surface-Sensor-Sun Urban Model \citep{Soux2004} to derive atmospherically correct hemispherical radiometric surface temperatures from upwelling longwave radiation measured from a downward facing pyrgeometer. The correction method was developed to be readily adaptable to other sensor types and study sites, these measurements are commonly made as a part of urban energy balance assessment. By accounting for surface geometry and spectrally non-uniform sensor response, the correction method is readily adaptable to other sensor types and study sites. Atmospheric effects on inferred T\textsubscript{hem} are significant relative to typical sUHI magnitudes and show significant diurnal and seasonal variation. Correction magnitudes are particularly sensitive to $\Delta$T\textsubscript{surf-air}. Although water vapor is the main absorber of broadband TIR radiation, it does not exert significant influence on correction magnitudes, as atmospheric transmittance shows little variation at typical humidities.

In addition, Chapter \ref{paper1} explores the effect of surface geometry on remote sensed urban T\textsubscript{surf} for three representations of the urban surface (complete, plan, and hemispherical). Analysis shows significant overestimation (underestimation) of T\textsubscript{comp} by T\textsubscript{plan} by day (night), particularly under clear sky conditions. By day, T\textsubscript{hem} overestimates T\textsubscript{comp}, but to a lesser degree than T\textsubscript{plan}. Daytime overestimation of T\textsubscript{comp} by T\textsubscript{hem} is more consistent under a wide variety of synoptic conditions, whereas overestimation by T\textsubscript{plan} fluctuates significantly, resulting in poor performance under clear-sky "satellite friendly" conditions. As make up the bulk of the satellite urban T\textsubscript{surf} record, our understanding of urban T\textsubscript{surf} likely includes a significant warming bias. These biases are reflected in sUHI analysis in \ref{paper2}.

Analysis of an eight month, time continuous climatology of hemispherical sUHI is included in Chapter \ref{paper2}.

\cleardoublepage 
\phantomsection  
\renewcommand*{\bibname}{References}
\addcontentsline{toc}{section}{\textbf{References}}

\putbib
\end{bibunit}

### 2017-05-03 introduction.tex
Add files via upload

@@
 Chapter \ref{paper1} details and evaluates a method to retrieve atmospherically corrected, hemispherical urban T\textsubscript{surf} from time-continuous measurements of upwelling longwave radiation.
-Chapter \ref{paper2} explores the temporal and geometric nature of sUHI through an eight month climatological analysis of hemispherical sUHI and compares sUHI magnitudes from several representations of the urban surface.
+Chapter \ref{paper2} explores the temporal and geometric nature of sUHI through long-term climatological analysis of hemispherical sUHI and compares sUHI magnitudes from several representations of the urban surface.
 

### 2017-05-03 paper1.tex
Add files via upload

@@
 Simulations for path lengths from 1 to 100 \si{\meter} and water vapor mass densities from 0.1 to 23.5 \si{\gram\per\meter\cubed} are included in Appendix \ref{appendixa}.
-Sensor-surface geometry is approximated as a single view factor weighted, azimuthally averaged surface-sensor path length calculated using the surface-sensor-sun urban model \citep{Soux2004} and a simplified DBM.
+Sensor-surface geometry is approximated as a single view factor weighted, azimuthally averaged, surface-to-sensor path length calculated using the surface-sensor-sun urban model \citep{Soux2004} and a simplified DBM.
 This path length represents the average distance from the sensor to the surface as "seen" by the sensor.
 For each time step, measured humidity is used to infer a hemispherical transmittance ($\overline{\tau_{\Phi}}$) for the approximated surface-sensor geometry.
-A parameterized T\textsubscript{hem, r} can then be calculated by using measured (in this case, profile averaged) T\textsubscript{air} to decompose measured L\textsubscript{up} into irradiance received by the sensor from the surface $L_{surf}^{at-sensor}$ and from the atmosphere $L_{atm}^{at-sensor}$ via,
+A parameterized T\textsubscript{hem, r} (T\textsubscript{hem, para}) can then be calculated by using measured (in this case, profile averaged) T\textsubscript{air} to decompose measured L\textsubscript{up} into irradiance received by the sensor from the surface $L_{surf}^{at-sensor}$ and from the atmosphere $L_{atm}^{at-sensor}$ via,
 
@@
 
-Using this parameterization, a climatology of T\textsubscript{hem, para} for a given surface geometry can be retrieved from approximately 100 simulations as opposed to the many millions needed for full hemispherical radiative transfer simulation.
-A variety of online and offline, open and closed source resources are available for radiative transfer simulation \citep{Gastellu-Etchegorry1996,Berk1987,Buehler2005}, some of which are available for free or at a low cost (Spectral-Calc), all of which are likely to provide similar results to those included in Appendix \ref{appendixa}.
+Using this parameterization, a climatology of T\textsubscript{hem, para} for a given surface geometry can be retrieved from approximately 100 simulations as opposed to the many thousands needed for full hemispherical radiative transfer simulation.
+A variety of online and offline, open and closed source resources are available for radiative transfer simulation \citep{Gastellu-Etchegorry1996,Berk1987,Buehler2005}, some of which are available for free or at a low cost (Spectral-Calc), all of which are likely to provide similar transmittances to those included in Appendix \ref{appendixa}.
+Figure \ref{para_eval} and Table \ref{parastats} summarize parameterization performance relative to corrected T\textsubscript{hem, r}.
+Errors are largest during the daytime hours (shaded in grey), where the profile averaged T\textsubscript{air} overestimates above canyon T\textsubscript{air} and underestimates near-surface T\textsubscript{air}.
+Neutral stability in the nighttime and early morning hours reduces these errors significantly as the canyon T\textsubscript{air} profile is approximately isothermal.
 

### 2017-05-03 paper2.tex
Add files via upload

@@
 In spite of this fact, and the short period over which large-scale, generalizable methods for measuring urban T\textsubscript{surf} have been available, study of sUHI via TIR remote sensing has expanded significantly in the last twenty years \citep{Peng2012,Voogt2003}.
-Using the method described in Chapter \ref{paper1} an 8-month climatology of hemispherical radiometric urban T\textsubscript{surf} (T\textsubscript{hem, r}) was derived from time continuous near-ground upwelling TIR as measured from an inverted pyrgeometer.
-The method was developed to address and overcome biases inherent in traditional methods for urban T\textsubscript{surf} retrieval by providing temporally continuous, geometrically representative\footnote{A hemispherical view is not perfectly geometrically representative of urban surface geometry.
+Using the method and parameterization described in Chapter \ref{paper1} twl long-term climatologies of hemispherical radiometric urban T\textsubscript{surf} (T\textsubscript{hem, r}) were derived from time continuous near-ground upwelling TIR as measured from an inverted pyrgeometer.
+The method, and an associated parameterization scheme, was developed to address and overcome biases inherent in traditional methods for urban T\textsubscript{surf} retrieval by providing temporally continuous, geometrically representative\footnote{A hemispherical view is not perfectly geometrically representative of urban surface geometry.
 This is best illustrated by visualizing an urban area from the perspective of a downward facing ‘fish-eye’ camera.
@@
 \subsection{A method to retrieve hemispherical radiometric urban T\textsubscript{surf}}
+\label{method2}
 
@@
 The measured irradiance is then matched with the closest modeled irradiance to return the associated corrected T\textsubscript{hem, r} for the target time step.
-sUHI magnitudes for the study area are calculated as the difference between urban T\textsubscript{hem, r} and rural T\textsubscript{hem, b} with rural T\textsubscript{hem, b} calculated via
-
-\begin{equation}
-\label{ruralt}
-T_{hem,~ b} = \sqrt[4]{\frac{\epsilon L_z^\uparrow + (1 - \epsilon)L_{sky}^\downarrow}{\sigma}}
-\end{equation}.
-
-\noindent where $L_z^\uparrow$ is at-sensor upwelling longwave radiation, $L_{sky}^\downarrow$ is at-sensor reflected downwelling longwave radiation, and emissivity ($\epsilon$) = 0.94.
-
-Sensitivity tests included in Section \ref{sensitivity} of the companion paper show that screen level irradiances ($ z $ = ~2 \si{\meter}) are not subject to significant effects from the intervening atmosphere and do not require atmospheric correction.
-Rural irradiances in this study were measured from approximately 2 \si{\meter} above ground, thus rural T\textsubscript{hem, b} is approximately equal to T\textsubscript{hem, r}.
 
@@
 
-To evaluate the method
+Although MODTRAN has been show to accurately model radiative transfer of broadband longwave radiation at urban scale path lengths \citep{Hoch2005, Hoch2007}, the added complexity inherent in simulating fluxes of radiation hemispherically over complex urban terrain prompted an independent evaluation of the correction method using observations of upwelling longwave radiation (measured from 2 \si{\meter}, 10 \si{\meter}, and 30 \si{\meter} above ground level) over a planar surface for a continuous 16-day period.
+In the evaluation, a modified version of the method uses T\textsubscript{hem, b} inferred from the lowest L\textsubscript{up} measurement along with profiles of T\textsubscript{air} and humidity to model irradiances at 10 \si{\meter} and 30 \si{\meter} at each time step.
+Modeled and measured irradiances match well, over wide range of mid-latitude atmospheric conditions, with results summarized in Table \ref{evalstats}.
+
+By virtue of requiring nearly 500 radiative transfer simulations per time step and significant post-processing, the method described in \ref{method2} is computationally expensive.
+This prompted the development of a parameterization scheme to facilitate more widespread usage of urban T\textsubscript{hem, r} for climatological and comparative sUHI analysis.
+In the parametrization, to reduce computational time, radiative transfer simulations are run for a single view factor weighted average path length representing the average distance from the surface to the sensor as "seen" by the sensor over its FOV.
+This view factor weighted "hemispherical" path length can be calculated using a sensor view model (such as SUM) or approximated by hand using a digital building model (although care should be taken to ensure the sensor cosine response is accurately represented).
+Using the view factor weighted path length, a suite of radiative transfer simulations is run prior to correction to retrieve "hemispherical" broadband atmospheric transmittances for a range of potential humidities.
+Humidity-transmittance pairings are aggregated into a lookup table (LUT), where the measured humidity at each time step can be matched with the associated hemispherical transmittance.
+Using the hemispherical transmittance as the fraction of surface emitted L\textsubscript{up} reaching the sensor, measured L\textsubscript{up} can then be decomposed into its surface and atmospheric emitted components, from which a parameterized T\textsubscript{hem, r} (T\textsubscript{hem, para}) is inferred.
+Comparison of T\textsubscript{hem, r} derived from the correction method and the parameterization show good agreement, the results of which are included in Figure \ref{para_eval} and Table \ref{parastats}.
 
@@
 
-T\textsubscript{hem, r} are retrieved from radiation and meteorological data collected from December 2001 through July 2002 as a part of the Basel Urban Boundary Layer Experiment (BUBBLE) campaign conducted in Basel, Switzerland \citep{Rotach2005}.
+This study uses data from two long-term climatological campaigns: the Basel Urban Boundary Layer Experiment (BUBBLE) conducted over December 2001 through July 2002 in Basel, Switzerland \citep{Rotach2005}, and the Environmental Prediction in Canadian Cities (EPiCC) (CITE?) campaign in Vancouver, Canada.
+Using the method described in section \ref{method2}, an eight month climatology of T\textsubscript{hem, r} are retrieved from radiation and meteorological data collected as a part of the BUBBLE campaign.
 Urban data were observed from a tower located near city center at the "Basel Sperrstrasse" site.
-Rural reference data were observed at the "Lange Erlen" site approximately 6 \si{\kilo \meter} to the north east in the outskirts of the city of Basel.
-Urban and rural site morphologies and observed variables are described in Table \ref{morphbspr}.
+Rural reference data were observed at the "Lange Erlen" site approximately 6 \si{\kilo\meter} to the north east in the outskirts of the city of Basel.
+
+During an intensive observation period (IOP) in late June/early July 2002, the Sperrstrasse canyon was instrumented with an array of infrared thermometers (IRT) to view individual canyon facet T\textsubscript{surf} to sample representative road, roof, and wall surface temperatures (T\textsubscript{road}, T\textsubscript{roof}, and T\textsubscript{wall} respectively).
+Plan and complete aspect ratios for the Sperrstrasse canyon are used to compute weighting schemes to retrieve complete and plan surface temperatures (T\textsubscript{comp} and T\textsubscript{plan} respectively) at 30 minute intervals over the IOP.
+T\textsubscript{comp} represents a complete urban T\textsubscript{surf}, where facet temperatures are weighted and averaged based on their areal proportion of the complete urban surface area.
+T\textsubscript{plan} represents the Sperrstrasse site as viewed by a satellite or aerial remote sensor in the nadir.
+Weighting schemes for T\textsubscript{plan} and T\textsubscript{comp} are included in Table \ref{weightings}.
+Urban T\textsubscript{surf} derived from complete, plan, and hemispherical representations of the Sperrstrasse site are used to investigate the effect of sensor-surface geometry on remote sensed T\textsubscript{surf} and sUHI, and to quantify the magnitude of geometric biases over a range of time scales and synoptic conditions.
+In addition, during the IOP, a second rural reference site ("Village Neuf"), located approximately 5 \si{\kilo\meter} SSE of the Sperrstrasse site, was used to investigate the influence of rural site characteristics on sUHI.
+
+To supplement data from the BUBBLE campaign, radiation and meteorological variables observed as a part of the EPiCC project are used to derive a climatology of T\textsubscript{hem, r} for the year long period spanning October 2008 through September 2009 using the parameterization scheme.
+These data provide a second, long term, time continuous climatology of sUHI for comparison of seasonal and diurnal patterns, to understand the influence of urban and rural site selection and characteristics on sUHI, and to aid in developing a generalized model for sUHI development over multiple time scales.
+
+Site morphologies and observed variables for BUBBLE and EPiCC sites are described in Tables \ref{morphbspr} and \ref{morphepicc}.
 
@@
 	\centering
-	\caption{A description of morphological and measured variables for urban and rural sites. Modified from \citet{Rotach2005} to include only relevant parameters.}
+	\caption{A description of morphological parameters and measured variables for BUBBLE urban and rural sites. Modified from \citet{Rotach2005} to include only relevant parameters.}
 	\label{morphbspr}
@@
 		& & & & \\
-		Lange Erlen \newline \textit{Rural parkland} \newline LCZ: B & 47.59\si{\degree} N \newline 7.65\si{\degree} E \newline 240 \si{\meter} a.s.l.
+		Lange Erlen \newline \textit{Rural parkland} \newline LCZ: B/D & 47.59\si{\degree} N \newline 7.65\si{\degree} E \newline 275 \si{\meter} a.s.l.
 & $\alpha$ = 21.4\%  & T\textsubscript{air} [4] \newline H [4] \newline WV [3] \newline WD [1] &  L\textsubscript{up} [1] \newline L\textsubscript{down} [1] \newline S\textsubscript{up} [1] \newline S\textsubscript{down} [1]  \\
+		&&&&\\
+		Village Neuf \newline \textit{Rural farmland} \newline LCZ: D/F & 47.57\si{\degree} N \newline 7.56\si{\degree} E \newline 240 \si{\meter} a.s.l. & $\alpha$ = 19.2\%
+		& T\textsubscript{air} [3] \newline H [3] \newline WV [2] \newline WD [1] &  L\textsubscript{up} [1] \newline L\textsubscript{down} [1] \newline
+		S\textsubscript{up} [2] \newline S\textsubscript{down} [2]  \\ 
 		\bottomrule
@@
 
-During an intensive observation period (IOP) in late June/early July 2002, the Sperrstrasse canyon was instrumented with an array of infrared thermometers (IRT) to view individual canyon facet T\textsubscript{surf} to sample representative road, roof, and wall surface temperatures (T\textsubscript{road}, T\textsubscript{roof}, and T\textsubscript{wall} respectively).
-Plan and complete aspect ratios for the Sperrstrasse canyon are used to compute weighting schemes to retrieve complete and plan surface temperatures (T\textsubscript{comp} and T\textsubscript{plan} respectively) at 30 minute intervals over the IOP.
-T\textsubscript{comp} represents a complete urban T\textsubscript{surf}, where facet temperatures are weighted and averaged based on their areal proportion of the complete urban surface area.
-T\textsubscript{plan} represents the Sperrstrasse site as viewed by a satellite or aerial remote sensor in the nadir.
-Weighting schemes for T\textsubscript{plan} and T\textsubscript{comp} are included in Table \ref{weightings}.
-Urban T\textsubscript{surf} derived from complete, plan, and hemispherical representations of the Sperrstrasse site are used to investigate the effect of sensor-surface geometry on remote sensed T\textsubscript{surf} and sUHI, and to quantify the magnitude of geometric biases over a range of time scales and synoptic conditions.
-For comparison of sUHI to air temperature UHI, canopy layer air temperature UHI (clUHI) magnitudes are calculated as the difference between urban and rural T\textsubscript{air} measured from 2 \si{\meter} above ground level.
+\begin{table}[H]
+	\centering
+	\caption{A description of morphological parameters and measured variables for EPiCC urban and rural sites.}
+	\label{morphepicc}
+	\begin{tabular*}{\textwidth}{p{3.75cm} p{2.25cm}p{3.5cm}p{2.75cm}p{2.75cm}}
+		\toprule 
+		Site & Location & Morphological & Meteorological & Radiation \\ 
+		& Height & Characteristics\footnote{} & Variables & [No of levels] \\ 	\midrule
+
+		Sunset Tower \newline \textit{Urban/Suburban neighborhood} \newline LCZ: 6 & 49.23\si{\degree} N \newline 123.07\si{\degree} W \newline 2.0 \si{\meter} a.s.l.
+& $z_H$ = 5.3 \si{\meter} \newline $\lambda_B $ = 0.29 \newline $\lambda_T $ = 0.15\newline $\lambda_V$ = 0.22 \newline $\lambda_I$ = 0.37 \newline $\alpha$ = 14.7\% & T\textsubscript{air} [1] \newline H [1] \newline WV [1] \newline WD [1] \newline P [1] & L\textsubscript{up} [1] \newline L\textsubscript{down} [1] \newline S\textsubscript{up} [1] \newline S\textsubscript{down} [1] \\
+		& & & & \\
+		Westham Island \newline \textit{Rural, unmanaged} \newline LCZ: D & 49.09\si{\degree} N \newline 123.18\si{\degree} W \newline 91.4 \si{\meter} a.s.l.
+& $\alpha$ = 20.6\%  & T\textsubscript{air} [2] \newline H [2] \newline WV [2] \newline WD [2] &  L\textsubscript{up} [1] \newline L\textsubscript{down} [1] \newline S\textsubscript{up} [1] \newline S\textsubscript{down} [1]  \\
+		\bottomrule
+	\end{tabular*} 
+	\textsuperscript{5}$\lambda_B $: Plan area of buildings, $\lambda_T $: Plan area of trees, $\lambda_V $: Plan area of ground vegetation, $\lambda_I $: Plan area of impervious ground.
+	\raggedright
+\end{table}
+
+\subsection{sUHI analysis}
+
+sUHI and canopy later UHI (clUHI) magnitudes were calculated from T\textsubscript{hem, r} and T\textsubscript{air} (measured from 2 \si{\meter} above ground level) as $\Delta$T\textsubscript{(urban-rural)} at 30 \si{\min} intervals.
+Rural T\textsubscript{hem, b} was calculated via
+
+\begin{equation}
+\label{ruralt}
+T_{hem,~ b} = \sqrt[4]{\frac{\epsilon L_z^\uparrow + (1 - \epsilon)L_{sky}^\downarrow}{\sigma}}
+\end{equation}
+
+\noindent where $L_z^\uparrow$ is at-sensor upwelling longwave radiation, $L_{sky}^\downarrow$ is at-sensor reflected downwelling longwave radiation, and emissivity ($\epsilon$) = 0.94.
+Sensitivity tests included in Section \ref{sensitivity} of the companion paper show that screen level irradiances ($ z $ = ~2 \si{\meter}) are not subject to significant effects from the intervening atmosphere and do not require atmospheric correction.
+Rural irradiances in this study were measured from approximately 2 \si{\meter} above ground, thus rural T\textsubscript{hem, b} is approximately equal to T\textsubscript{hem, r}.
+
+To compare seasonal and diurnal patterns of sUHI and clUHI development over the BUBBLE campaign, mean monthly UHI magnitudes were normalized via,
+
+\begin{equation}
+\text{UHI}_{\text{norm}} = \frac{(\text{UHI} - \text{UHI}_{\text{min}})}{(\text{UHI}_{\text{max}} - \text{UHI}_{\text{min}})}
+\label{uhinorm}
+\end{equation}
+
+\noindent with UHI\textsubscript{min} and UHI\textsubscript{max} referring to minimum and maximum mean monthly sUHI and clUHI values over the climatology.
+This normalization restricts mean monthly UHI magnitudes to a range of values between zero (mean monthly UHI\textsubscript{min} over the climatology) and one (mean monthly UHI\textsubscript{max} over the climatology).
+
+To investigate the influence of rural site characteristics on Basel's sUHI, sUHI magnitudes are calculated using T\textsubscript{hem, b} inferred for a different rural site ("Village Neuf") over the summertime IOP.
+Observations at Village Neuf were intermittent and often discontinous.
+As such, a subset of nine partially-complete days was selected to represent T\textsubscript{hem, b} and sUHI over a range of cloud coverages, to investigate the influence of rural site characteristics and T\textsubscript{surf} regimes on sUHI.
+As sUHI develop, preference was given to days with continuois observations in early afternoon through night.
+For each of the nine days sUHI\textsubscript{max} was used to infer a normalized sUHI magnitude for both Lange Erlen and Village Neuf sites at 30- \si{\minute} intervals, to remove day-to and compare patterns of sUHI development for the two sites.
+Mean normalized sUHI was then calculated at each time step over the nine day period.
+A normalized T\textsubscript{hem, b} was also calculated (using maximum T\textsubscript{hem, b} for each day) to compare T\textsubscript{surf} regimes between the two sites.
 
@@
 
-Figure \ref{heatsuhi} shows seasonal variations in diurnal patterns of mean monthly hemispherical sUHI magnitudes at 30 minute intervals for the eight month study period.
-The sUHI displays significant seasonal variation, largest in the daytime hours, and is strongly controlled by day length and solar angle.
+Figure \ref{heatsuhi} shows seasonal variations in diurnal patterns of mean monthly hemispherical sUHI magnitudes at 30 minute intervals for the BUBBLE campaign.
+The sUHI in Basel displays significant seasonal variation, largest in the daytime hours, and is strongly controlled by day length and solar angle.
 Nighttime sUHI does not vary significantly across seasons.
 
-Figure \ref{heatcluhi} shows seasonal variations in diurnal patterns of mean monthly clUHI calculated at 30 minute intervals for the eight month study period.
-clUHI development is strongly controlled by the timing of sunset/sunrise cycles and displays the largest seasonal variance in the hours immediately after sunset.
-Daytime clUHI does not vary significantly across seasons.
 \begin{figure}[H]
@@
 	\includegraphics[width=15cm,height=7in,keepaspectratio]{heatsuhi}
-	\caption{A heatmap of mean half hourly hemispherical sUHI for each month calculated at 30 \si{\min} intervals over the eight month study period. Results are interpolated between months.}
+	\caption{A heatmap of mean half hourly hemispherical sUHI for each month calculated at 30 \si{\min} intervals for the BUBBLE campaign. Results are interpolated between months.}
 	\label{heatsuhi}
+\end{figure}
+
+Figure \ref{heatcluhi} shows seasonal variations in diurnal patterns of mean monthly clUHI calculated for the same study period as Figure \ref{heatsuhi}.
+clUHI development is strongly controlled by the timing of sunset/sunrise cycles and displays the largest seasonal variance in the hours immediately after sunset.
+Daytime clUHI does not vary significantly across seasons.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=15cm,height=7in,keepaspectratio]{heatatuhi}
+	\caption{A heatmap of mean half hourly canopy layer UHI for each month calculated at 30 \si{\min} intervals for the BUBBLE campaign. Results are interpolated between months.}
+	\label{heatcluhi}
 \end{figure}
@@
 	\centering
-	\includegraphics[width=15cm,height=7in,keepaspectratio]{heatatuhi}
-	\caption{A heatmap of mean half hourly canopy layer UHI for each month calculated at 30 \si{\min} intervals from T\textsubscript{air} measured at approximately 2 \si{\meter} above ground level over the eight month study period.
+	\includegraphics[width=15cm,height=7in,keepaspectratio]{heat_uhi_diff}
+	\caption{A heatmap of the difference between mean half hourly hemispherical normalized sUHI and canopy layer UHI for each month calculated at 30 \si{\min} intervals over the BUBBLE campaign.
 Results are interpolated between months.}
-	\label{heatcluhi}
+	\label{heat_uhi_diff}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=15cm,height=7in,keepaspectratio]{vanc_suhi}
+	\caption{A heatmap of mean half hourly hemispherical sUHI for each month calculated at 30 \si{\min} intervals for a year-long subset of the EPiCC campaign. Results are interpolated between months.}
+	\label{vanc_suhi}
 \end{figure}
@@
 Figure \ref{bx_suhi_compare} shows sUHI magnitudes calculated from urban T\textsubscript{comp}, T\textsubscript{plan}, and T\textsubscript{hem}.
-Compared to complete sUHI; nadir and hemispherical views of the surface overestimate sUHI by day and underestimate sUHI by night.
-sUHI from a nadir view shows the greatest diurnal variance in sUHI, particularly under clear sky 'satellite friendly' conditions.
+Compared to sUHI viewed from a complete representation of the urban surface (sUHI\textsubscript{comp}); nadir (sUHI\textsubscript{nadir}) and hemispherical (sUHI\textsubscript{hem}) views of the surface overestimate sUHI\textsubscript{comp} by day and underestimate sUHI\textsubscript{comp} by night.
+sUHI\textsubscript{nadir} from a nadir view shows the greatest diurnal range of magnitudes, particularly under clear sky 'satellite friendly' conditions.
 
@@
 	\centering
-	\includegraphics[width=15cm,height=7in,keepaspectratio]{bxsuhi1}
-	\caption{A comparison of sUHI magnitudes from complete, hemispherical, and nadir remote sensed representations of the Sperrstrasse canyon over the IOP.
+	\includegraphics[width=15cm,height=7in,keepaspectratio]{bxsUHI_iop_bubble}
+	\caption{A comparison of sUHI magnitudes from complete (sUHI\textsubscript{comp}), hemispherical (sUHI\textsubscript{hem}), and nadir (sUHI\textsubscript{nadir}) remote sensed representations of the Sperrstrasse canyon over the IOP.
 Each plot includes mean sUHI, as well case days representing sUHI under clear sky and clouded sky conditions.}
@@
 
-Figures \ref{meteo_kdown}, \ref{meteo_hum}, \ref{meteo_wv}, and \ref{meteo_atuhi} show sUHI magnitudes as a function of incoming solar radiation, water vapor content, wind velocity, and clUHI.
+Figures \ref{meteo_kdown}, \ref{meteo_hum}, \ref{meteo_wv}, and \ref{meteo_atuhi} show hemispherical sUHI magnitudes as a function of incoming solar radiation, water vapor content, wind velocity, and clUHI measured over the BUBBLE campaign.
 Winter hours are omitted in Figures \ref{meteo_hum}, and \ref{meteo_wv}, and \ref{meteo_atuhi} as sUHI is small and displays minimal variation in winter.
@@
 	\includegraphics[width=15cm,height=7in,keepaspectratio]{meteo_kdown}
-	\caption{sUHI magnitude versus incoming solar radiation binned for clear sky (when the sum of daytime K\textsubscript{down} \textgreater 10000 \si{\watt \per \square \meter}) and cloudy sky days (when the sum of daytime K\textsubscript{down} \textless 10000 \si{\watt \per \square \meter})}
+	\caption{sUHI magnitude versus incoming solar radiation binned for clear sky (when the sum of daytime K\textsubscript{down} \textgreater 18 \si{\mega\joule\per\meter\squared}) and cloudy sky days (when the sum of daytime K\textsubscript{down} \textless 18 \si{\mega\joule\per\meter\squared})}
 	\label{meteo_kdown}
@@
 
+\pagebreak
+
+\subsection{The effect of rural/non-urban characteristics on sUHI}
+
+Figure \ref{rur_uhi_means} shows urban T\textsubscript{hem, r} from the Sperrstrasse site and rural T\textsubscript{hem, b} for the Lange Erlen and Village Neuf sites as well as their associated hemispherical sUHI magnitudes for a representative, summertime clear sky day.
+In addition, the figure shows mean normalized hemispherical sUHI and T\textsubscript{hem, b} (calculated from maximum sUHI and T\textsubscript{hem, b} values for each day) for the Lange Erlen and Village Neuf stations.
+Days were chosen to represent a wide range of summertime conditions and cloud coverages, however, observations at the Village Neuf site were intermittent and n = 9 only applies to hours after 12:00 LST when data could be continuously sampled.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=15cm,height=7in,keepaspectratio]{rur_uhi_means1}
+	\caption{Top: T\textsubscript{hem} from the Sperrstrasse urban and Lange Erlen and Village Neuf sites and sUHI magnitudes calculated from the two rural stations for a representative summer clear sky day.
+Middle and bottom: Mean normalized hemispherical sUHI and T\textsubscript{hem, b} calculated for the Lange Erlen and Village Neuf sites.
+Data at Village Neuf before noon were intermittent, thus n = 9 only applies to values after 12:00 LST.
+Both sites are sampled over the same truncated period on days with incomplete data.}
+	\label{rur_uhi_means}
+\end{figure}
+
 \section{Discussion}
+\citet{Oke2017} presents a generalized conceptual model for sUHI development for multiple representations of the surface.
+However, the results in \citet{Oke2017} are based on simulations of an idealized urban area without trees, complex non-orthogonal surface geometry, or sloped facets and have not been supported observationally.
+The results in this paper provides an ideal platform to test conceptual understanding of sUHI development and to understand the diurnal and seasonal nature of sUHI.
 
 \subsection{Diurnal patterns of sUHI}
+\label{gen}
 
-sUHI in Basel from complete, nadir, and hemispherical representations of T\textsubscript{surf} displays two diurnal peaks, the first in the late afternoon and a second just after sunset.
-Clear sky conditions amplify this phenomena.
-A late afternoon peak in sUHI is likely the result of the urban-rural contrasts in thermal admittance (and by relation, urban-rural soil moisture contrasts).
-The thermal admittance of many urban surfaces, particularly rooftops, tends to be lower than rural and vegetated environments \citep{Spronken-Smith1998} – particularly in temperate climates where rural soil moisture is largely maintained through summer.
-Lower urban thermal admittance manifests in a larger diurnal thermal amplitude and thus, greater daytime surface heating and nighttime surface cooling.
-Nadir and hemispherical views are biased towards low thermal admittance rooftop facets, and thus display large late afternoon peaks in sUHI.
+sUHI development is the result of differences in urban and rural heating and cooling rates.
+A variety of factors modify urban heating and cooling rates relative to non-urban spaces, most notably urban-rural contrasts in sky-view-factor, impervious surface fraction, and thermal admittance (and by relation, urban-rural soil moisture contrasts).
+These differences provide the genesis for diurnal sUHI development.
+Rural surfaces have more open (less obstructed) sky view factors, and respond quickly to insolation a low solar angles in the early morning hours.
+Shading by canyon features causes urban T\textsubscript{surf} to respond more slowly, resulting in a small sUHI\textsubscript{min} in the hours just after sunset.
+This is observed in Basel across all representations of urban T\textsubscript{surf} at similar magnitudes.
+Through the late morning hours, urban heating rates increase relative to rural rates as the solar angle increases and low thermal admittance facets (particularly rooftops) experience more direct insolation and quickly heat up.
+Increases in urban T\textsubscript{surf} are amplified by the abundance of impervious surfaces, which result in a smaller fraction of insolation partitioned to latent heat relative to vegetated non urban spaces.
+After the initial spike in rural heating in the morning hours, urban heating rates overtake rural and the sUHI slowly grows through midday.
+In the hours after midday, the solar angle decreases and both urban and rural heating rates decline and begin to converge, as shading by canyon geometry causes some facets to cool.
+sUHI reaches a peak at approximately two to three hours hours after solar noon.
 
-A second peak in sUHI after sunset is the result of differences in relative cooling rates of urban and rural surfaces.
-In this case, nighttime sUHI is likely the result of canyon radiation trapping from obstructed canyon sky-view-factors.
+In the afternoon hours, both urban and rural surfaces transition to cooling, with approximately equal rates.
+The late afternoon and evening decline in sUHI is strongly controlled by urban surface structure.
+Relative to non-urban spaces, cooling rates for wall and road facets lag slightly as a result of canyon radiation trapping, this is offset by increased cooling rates for low thermal admittance rooftop facets.
+These two factors control whether sUHI grows or shrinks in the hours after sUHI\textsubscript{max}.
+In urban areas with relatively unobstructed sky-view-factors or large rooftop plan area fractions, strong radiative cooling can lead to a sharp reduction in sUHI.
+The opposite is true of urban areas with obstructed sky view factors, and tall street canyons with large wall surface area fractions, where canyon radiative trapping reduces radiative cooling and sUHI is maintained through the evening.
+For the Sperrstrasse canyon, the combination of a narrow but relatively low street canyon results in a slight reduction in Basel's sUHI as strong radiative cooling of rooftops is slightly offset by canyon radiation trapping.
+
+After sunset, rural cooling rates spike relative to urban, as an unobstructed rural sky-view-factor allows for strong radiative cooling of the surface, leading to a slight jump in sUHI.
+However, variations in rural thermal admittance, surface geometry, land cover and use, and/or influence from surrounding terrain can result in significant contrasts in rural cooling rates.
+All of these factors show strong spatiotemporal variation, resulting in large differences in sUHI development depending on rural site characteristics.
+This is evident in characterization of sUHI in Basel, where the two rural sites entail different rural cooling rates, particularly in the hours just after sunset, and different patterns of post-sunset sUHI development.
+Although similar in terms of surface coverage and site geometry, BUBBLE rural sites are subject to different after sunset air masses that strongly effect sUHI (and clUHI) development.
+Hilly terrain surrounding the Lange Erlen site results in cold air drainage and an increased rural cooling rate, particularly following hot, clear sky days.
+This effect is not present at the Village Neuf site, where nocturnal cooling rates are likely driven primarily driven by radiative losses.
+The "extra", largely non-radiative cooling at the Lange Erlen site results in a strong second peak in sUHI that is not present in sUHI calculated using the Village Neuf site.
+This makes clear the need for proper characterization of the radiative and thermal properties, surface geometry and coverage of rural sites in sUHI analysis, as well as the surrounding topography to ensure characterizations of sUHI development are free of confounding influence.
+After initial nighttime divergence, sUHI in Basel is relatively constant over the nighttime hours, with slight growth observed on clear nights where canyon radiation trapping retards urban nocturnal cooling rates relative to rural.
+
+%However, has not been observed sUHI in Basel from complete, nadir, and hemispherical representations of T\textsubscript{surf} displays two diurnal peaks, the first in the late afternoon and a second just after sunset.
+Clear sky conditions amplify this phenomena.Contrasting urban-rural heating rates manifest in  The thermal admittance of many urban surfaces, particularly rooftops, tends to be lower than rural and vegetated environments \citep{Spronken-Smith1998} – particularly in temperate climates where rural soil moisture is largely maintained through summer.
+Lower urban thermal admittance manifests in a larger diurnal thermal amplitude and greater daytime surface heating.
+Thus, sensors biased towards low thermal admittance rooftop facets (as is the case with a narrow-FOV sensor and the hemispherical sensor at the Sperrstrasse site), will observe large late afternoon peaks in sUHI, which are not necessarily reflected in sUHI from a complete representation of the surface.
+
+%A second peak in sUHI after sunset is the result of differences in relative cooling rates of urban and rural surfaces.
+In this case, nighttime sUHI is likely the result of a reduction in urban cooling rates from radiation trapping by canyon walls.
+However, as is the case with clUHI, post sunset sUHI development is highly dependent on rural site characteristics.
+Factors such as soil moisture, land cover/use, vegetation types, cold air drainage all result in different after sunset cooling rates (and patterns of sUHI development).
+These factors can vary significantly over small areas.
 The second peak sUHI is largest following clear sky conditions when the potential for significant differentials in urban-rural nocturnal cooling rates is maximized.
 A second peak is observed across all representations of urban T\textsubscript{surf} but is most apparent in complete and nadir representations of sUHI, which, for the Sperrstrasse site, sample T\textsubscript{road} to a greater extent than the hemispherical view.
-A weak sUHI is maintained through the nighttime as urban and rural cooling rates reach parity in the hours after sunset \citep{Oke2017}.
+
+%A weak sUHI is maintained through the nighttime as urban and rural cooling rates reach parity in the hours after sunset \citep{Oke2017}.
 Minimum sUHI is observed just after sunrise, where sUHI magnitudes are occasionally negative.
-At low solar angles, shading by 3-dimensional urban geometry suppresses heating rates for a large portion of urban surface area, reducing overall urban heating rates in early morning.
-As solar angles increase through the morning and more facets experience direct isolation, sUHI develops, with nadir and hemispherical representations displaying more rapid intensification than complete sUHI from a bias towards low thermal admittance and directly sunlit rooftop surfaces.
+At low solar elevations angles, shading by 3-dimensional urban geometry suppresses heating rates for a large portion of urban surface area, reducing overall urban heating rates in early morning.
+As the solar elevation angle increases through the morning and more facets experience direct insolation, sUHI develops, with nadir and hemispherical representations displaying more rapid intensification than complete sUHI from a bias towards low thermal admittance and directly sunlit rooftop surfaces.
 
@@
 
-Hemispherical sUHI displays significant seasonality, both in terms of its diurnal pattern and overall magnitude.
-Seasonal variations in sUHI are largest by day, where large seasonal contrasts in cloud cover frequency and solar angle exert strong control on daytime heating rates and sUHI development.
-Peak sUHI occurs near the summer solstice, when days are longest and isolation on clear sky days is most intense.
+Hemispherical sUHI in Basel displays significant seasonality, both in terms of its diurnal pattern and overall magnitude.
+Seasonal variations in sUHI are most evident during the daytime hours, where large seasonal contrasts in cloud cover frequency and solar angle exert strong control on daytime heating rates and sUHI development.
+Nighttime sUHI is consistent across seasons.
+Peak seasonal sUHI occurs near the summer solstice, when days are longest and isolation on clear sky days is most intense, forcing significant contrasts in urban-rural heating/cooling rates.
 In winter, daytime sUHI is suppressed from a lower solar input and frequent cloud cover.
-This weakens wintertime sUHI and shifts its peak to shortly after sunset.
-Nighttime sUHI is relatively consistent across seasons.
-Similar seasonality is observed in clUHI, albeit concentrated at different times of day.
+These reduce urban-rural contrasts in heating/cooling rates and weaken wintertime sUHI.
+In Basel, wintertime sUHI has a small peak shortly after sunset, when cold air drainage amplifies rural cooling rates and increases sUHI.
+This, the small after sunset peak in sUHI present in both summer and winter observations is likely not generalizable and is specific to the Lange Erlen site.
 
-In contrast to sUHI, clUHI is greatest just after sunset and is strongly controlled by sunset/sunrise timing.
-Peak clUHI occurs just after sunset across seasons, differing only in terms of magnitude, with a larger peak clUHI in summer.
-However, in this study, it should be note that summertime clUHI magnitudes in the hours just after sunset are slightly higher than expected \citep{Runnalls2000} - possibly the result of cold air draining from hilly terrain west of the rural site.
-As is the case with sUHI seasonality in clUHI results from seasonal contrasts in solar input and cloud cover frequency.
-Increased wintertime cloud cover reduces radiative losses by the canyon air volume after sunset, suppressing mean wintertime clUHI magnitudes.
-Daytime clUHI is consistently small across all seasons.
-Thus, although sUHI and clUHI are controlled by similar climatic controls and have a similar micro meteorological genesis, they do not display the same seasonal patterns.
-A surface urban heat island is not simply a stronger, more pronounced, canopy layer urban heat island.
+These results are supported by similar seasonal patterns of hemispherical sUHI observed in Vancouver, where seasonal sUHI\textsubscript{max} (sUHI\textsubscript{min}) is observed near the summer (winter) solstice.
+Observations of sUHI in Vancouver show similar seasonal patterns of development to those in Basel, with sUHI with a larger diurnal sUHI amplitude in summer months and suppressed sUHI development in winter.
+Cooling rates at the Westham island rural site are not affected by cold air drainage and, as a result, sUHI observed in Vancouver does not display two diurnal peaks in summer nor a coherent peak after sunset in winter months.
+
+Although sUHI and clUHI are controlled by similar climatic controls and have a similar micro meteorological genesis, they do not display the same seasonal patterns.
+In contrast to sUHI, clUHI in Basel is greatest just after sunset following closely to sunset/sunrise timing.
+Peak clUHI occurs just after sunset across seasons, differing in terms of magnitude but not time of day, with a larger peak clUHI in summer\footnote{As is the case with sUHI observed in Basel, clUHI magnitudes are effected by cold air drainage at the Lange Erlen site and are slightly higher than expected \citep{Runnalls2000}.}.
+Daytime clUHI is small and displays little seasonal variation.
+Thus, a surface urban heat island is not simply a stronger, more pronounced, canopy layer urban heat island.
 
@@
 Sensor-surface geometry exerts significant influence on observed sUHI magnitudes.
-By undersampling vertical and sloped urban surfaces and oversampling low thermal admittance fabric materials, nadir sUHI overestimates mean afternoon complete sUHI by 3.41 \si{\kelvin} and up to 5.96 \si{\kelvin} under clear sky “satellite friendly” conditions.
-Thus, peak nadir sUHI overestimates the peak 'true' complete sUHI by nearly double under clear sky conditions.
-As clear sky conditions are the only conditions under which satellite TIR remote sensors can operate, and make up a large fraction of the urban T\textsubscript{surf} literature, these biases cannot be ignored.
-By sampling the urban surface in 3-dimensions, hemispherical sUHI yields a much smaller mean daytime overestimation (0.54 \si{\kelvin}), particularly under clear sky conditions (0.93 \si{\kelvin}), providing a more geometrically representative depiction of complete urban T\textsubscript{surf} and sUHI, in addition to allowing for time-continuous analysis.
-Cloud cover suppresses the influence of sensor-surface geometry on remote sensed sUHI by reducing microscale geometric and spatiotemporal contrasts in urban T\textsubscript{surf}, resulting in less facet-scale heterogeneity in urban T\textsubscript{surf} and reducing the overall magnitude of sUHI.
+By undersampling vertical and sloped urban surfaces and oversampling low thermal admittance fabric materials, sUHI\textsubscript{nadir} overestimates mean afternoon sUHI\textsubscript{comp} by 3.41 \si{\kelvin} and up to 5.96 \si{\kelvin} under clear sky “satellite friendly” conditions.
+Thus, peak sUHI\textsubscript{nadir} overestimates the 'true' sUHI\textsubscript{comp} by nearly double under clear sky conditions.
+As clear sky conditions are the only conditions under which satellite TIR remote sensors can operate, and make up a large fraction of the urban T\textsubscript{surf} literature, these biases are pervasive and cannot be ignored.
+In contrast, by sampling the urban surface in 3-dimensions, sUHI\textsubscript{hem} yields a much smaller mean daytime overestimation (0.54 \si{\kelvin}), particularly under clear sky conditions (0.93 \si{\kelvin}), providing a significantly more representative depiction of urban T\textsubscript{comp} and sUHI\textsubscript{comp}, in addition to allowing for time-continuous analysis.
+Cloud cover suppresses the influence of sensor-surface geometry on remote sensed sUHI by reducing microscale geometric and spatiotemporal contrasts in urban T\textsubscript{surf}, resulting in less facet-scale heterogeneity in urban T\textsubscript{surf} and reducing the overall magnitude of sUHI for all definitions.
 
@@
 
-sUHI is highly dependent on meteorological conditions, which account for much of the variation in daytime sUHI magnitudes.
+sUHI is highly dependent on meteorological conditions, which account for much of the day to day variation sUHI magnitudes.
 Conditions that foster the largest microscale contrasts in urban T\textsubscript{surf} also entail the largest sUHI magnitudes.
@@
 As such, incoming shortwave radiation and wind velocity display positive and negative relationships with sUHI magnitudes respectively.
-Under clear sky conditions, strong solar heating of the surface combined with differences in urban and rural thermal properties fosters the development of large daytime urban-rural thermal contrasts.
+Under clear sky conditions, strong solar heating of the surface combined with differences in urban and rural thermal properties and sunlit shading geometries fosters the development of large daytime urban-rural contrasts in T\textsubscript{surf} and T\textsubscript{air} and the development of strong sUHI by day and clUHI by night.
 Thus, sUHI\textsubscript{max} coincides with insolation\textsubscript{max}.
-Under cloudy conditions, decreased direct solar heating of the surface suppresses the development of urban-rural thermal contrasts, resulting in a reduced sUHI.
+However, significanyUnder cloudy conditions, decreased direct solar heating of the surface suppresses the development of urban-rural thermal contrasts, resulting in a reduced sUHI.
 $\Delta$T\textsubscript{hem, r - air} has a similar positive effect on sUHI, owing to its dependence on solar insolation.
@@
 Atmospheric transmittance of longwave radiation decreases sharply with small increases in water vapor when the atmosphere is very dry - shown in Figure \ref{humtest}.
-As water vapor content increases from zero, variation in atmospheric transmittance decrease.
-Therefore, typical variations in urban and rural humidities have little effect on the fraction of outgoing available for heating of the canopy layer air volume through absorption of outgoing longwave radiation and for heating of the surface by atmospheric emission.
+As water vapor content increases from zero, variation in spectral atmospheric transmittance decreases.
+Therefore, typical variations in urban and rural humidities have little effect on the fraction of outgoing longwave radiation available for heating the canopy layer air volume through absorption and for heating of the surface by atmospheric emission.
 As a result, water vapor content has little effect on sUHI.
@@
 As wind velocity increases, sUHI is suppressed from increased turbulent heat losses.
+During the nighttime hours, increased wind velocities destroy the surface based thermal inversion often observed over rural terrain.
+This further limits nighttime cooling rates relative to urban and decreases sUHI magnitudes.
 Meteorological controls on sUHI are summarized in Table \ref{meteo_cont}.
@@
 		\midrule
-		$\Delta$T\textsubscript{hem, r - air} &\textit{increasing} \\
+		$\Delta$T\textsubscript{surf - air} &\textit{increasing} \\
 		&\\
-		T\textsubscript{hem, r} & \textit{increasing} \\
+		T\textsubscript{surf} & \textit{increasing} \\
 		&\\
@@
 
-This chapter presents the results of a time-continuous eight month climatological analysis of the surface urban heat island effect in Basel, Switzerland derived using the method presented in Chapter \ref{paper1}.
+This chapter presents the results of a time-continuous eight month climatological analysis of the surface urban heat island in Basel, Switzerland derived using the method presented in Chapter \ref{paper1}.
 Hemispherical sUHI magnitudes display significant seasonal and diurnal variation and are largest in summer months during the late afternoon during which solar heating of the surface is most intense, forcing the development of strong microscale and urban-rural contrasts in T\textsubscript{surf}.
 Seasonality in late afternoon sUHI is largely dependent on incoming solar radiation, which displays a strong positive relationship with sUHI magnitudes.
-In the winter months, a reduced overall solar input and increased cloud cover frequency suppresses daytime sUHI development.
-resulting in less day-to-day variation in sUHI, and lower overall sUHI magnitudes.
+In the winter months, a reduced overall solar input and increased cloud cover frequency suppresses daytime sUHI development resulting in less day-to-day variation in sUHI, and lower overall sUHI magnitudes.
 
 Remote sensed sUHI is highly dependent on sensor-surface-sun geometry.
-This results in significant variation in observed urban T\textsubscript{surf} and sUHI when observed from sensors with different spatial, geometric, and temporal sampling regimes.
+This results in significant variation in urban T\textsubscript{surf} and sUHI when observed from sensors with different spatial, geometric, and temporal sampling regimes.
 Thus, traditional methods for thermal remote sensing of urban areas do not capture the true temporal and geometric nature of the urban effect on land T\textsubscript{surf}.

### 2017-05-04 conclusion.tex
Add files via upload

@@
 
-Chapter \ref{paper1} details a hemispherical atmospheric correction method that uses version 4.1 of the MODTRAN radiative transfer code \citep{Berk1987} in conjunction with the Surface-Sensor-Sun Urban Model \citep{Soux2004} to derive atmospherically correct hemispherical radiometric surface temperatures from upwelling longwave radiation measured from a downward facing pyrgeometer.
-The correction method was developed to be readily adaptable to other sensor types and study sites, these measurements are commonly made as a part of urban energy balance assessment.
-By accounting for surface geometry and spectrally non-uniform sensor response, the correction method is readily adaptable to other sensor types and study sites.
+Chapter \ref{paper1} details a hemispherical atmospheric correction method that uses version 4.1 of the MODTRAN radiative transfer code \citep{Berk1987} in conjunction with the Surface-Sensor-Sun Urban Model \citep{Soux2004} to derive atmospherically corrected hemispherical radiometric surface temperatures from upwelling longwave radiation measured from a downward facing pyrgeometer.
+These temperatures are representative of the temperature at which the urban surface is radiating.
+The correction method was developed to be readily adaptable to other sensor types and study sites by accounting for surface geometry and non uniform spectral sensor response.
+These measurements are commonly made as a part of urban energy balance assessment and constitute a hitherto untapped resource for urban T\textsubscript{surf} and sUHI analysis.
 Atmospheric effects on inferred T\textsubscript{hem} are significant relative to typical sUHI magnitudes and show significant diurnal and seasonal variation.
-Correction magnitudes are particularly sensitive to $\Delta$T\textsubscript{surf-air}.
+Relationships between correction magnitude and meteorological variables are complex.
 Although water vapor is the main absorber of broadband TIR radiation, it does not exert significant influence on correction magnitudes, as atmospheric transmittance shows little variation at typical humidities.
+Correction magnitudes show strong positive relationships with to $\Delta$T\textsubscript{surf-air} and incoming solar radiation.
+To further facilitate method usability, a parameterization was developed incorporating a small number of generalizable radiative transfer simulations (included in \ref{appendixa}) and information about site geometry to retrieve radiometric hemispherical T\textsubscript{surf} from measurements of upwelling longwave en masse.
 
@@
 Daytime overestimation of T\textsubscript{comp} by T\textsubscript{hem} is more consistent under a wide variety of synoptic conditions, whereas overestimation by T\textsubscript{plan} fluctuates significantly, resulting in poor performance under clear-sky "satellite friendly" conditions.
-As make up the bulk of the satellite urban T\textsubscript{surf} record, our understanding of urban T\textsubscript{surf} likely includes a significant warming bias.
+As these measures make up the bulk of the satellite urban T\textsubscript{surf} record, our understanding of urban T\textsubscript{surf} likely includes a significant warming bias.
 These biases are reflected in sUHI analysis in \ref{paper2}.
 
-Analysis of an eight month, time continuous climatology of hemispherical sUHI is included in Chapter \ref{paper2}.
+Chapter \ref{paper2} includes an analysis of two long term, time continuous climatologies of hemispherical sUHI, one in Basel, Switzerland and another in Vancouver, Canada.
+sUHI for both study sites shows significant diurnal, day-to-day, and seasonal variability.
+sUHI is highly dependent on meteorological conditions, particularly accumulated solar radiation over a day and wind velocity.
+However, as sUHI is the result of differences in heating and cooling rates between the city and its non urban surroundings, its relationship to meteorological controls is complex.
+Analysis of sUHI development under a range of mid-latitude synoptic conditions lead to the development of a data supported generalizable model for summertime sUHI development presented in section \ref{gen}.
+
+Analysis of the directional dependence of sUHI is also included in Chapter \ref{paper2}.
+Similar results are observed as those found in Chapter \ref{paper1}, with significant overestimation of sUHI\textsubscript{comp} by T\textsubscript{plan}, particularly during hot clear sky conditions.
+Not only did directional biases in remote sensed sUHI change observed magnitudes, these biases also changed the diurnal character of sUHI, informing patterns of sUHI in the hours during near sunset.
+Failure, in satelite based assessment of the diurnal pattern of sUHI may lead to misrepresentations of its 'true' diurnal character.
+
+\section{Final remarks}
+
+Broadly, the method, parameterization, and analysis presented in this study not only provides for novel analysis of the urban effect on T\textsubscript{surf}, but also aims to prompt further critical assessment of the urban T\textsubscript{surf} and sUHI record.
+Indeed, if similar conceptual flaws and shortcomings to those found for the aUHI in \citet{Stewart2011} are likely in sUHI analysis, quantification of instrument and methodological biases and shortcomings in remote sensing of urban T\textsubscript{surf} and the introduction of new methods to overcome these shortcomings may bring us closer to and to addressing the warnings raised in \citet{Stewart2011}.
 

### 2017-05-04 introduction.tex
Add files via upload

@@
 Thus, understanding how cities interact with the climate across spatiotemporal scales has important implications for human health, energy efficiency, and for informing future urban development.
-This, in part, has prompted significant expansion in study of the urban effect on climates, particularly in the decades preceding this thesis.
+This, in part, has prompted significant expansion in study of the urban effect on climates, with particular focus on characterization of the urban effect on surface (T\textsubscript{surf}) and air (T\textsubscript{air}) temperatures.
 
-The notion that urban areas have elevated air and surface temperatures (T\textsubscript{air} and T\textsubscript{surf} respectively) is not new, with the first formal studies of the urban effect on T\textsubscript{air} dating back to \citet{Howard1833}'s identification of "artificial warming" and "urban contamination" in his characterization of spatial patterns of T\textsubscript{air} in early 19\textsuperscript{th} century London.
+\pagebreak
+
+The temperature of a surface (urban or otherwise) at its interface with the atmosphere is defined as the net result of the surface energy balance \citep{Oke2017}, wherein change in temperature at time $t$ for a surface with thickness $z$ and heat capacity $C$ can be solved as,
+
+\begin{equation}
+	C~\frac{\partial \text{T\textsubscript{surf}}}{\partial t}~z = Q^*+ Q_F - Q_H - Q_E - \Delta Q_A - \Delta Q_G
+\end{equation}
+
+\noindent where $Q^*$ is net radiation at the surface, $Q_F$, $Q_H$, and $Q_E$ are fluxes of anthropogenic, sensible, and latent heat, $\Delta Q_A$ is net advection of surface heat to or from the volume of air above the surface, and $\Delta Q_G$ is net storage of heat by the substrate below.
+A change in T\textsubscript{surf} is translated to change in T\textsubscript{air} as the surface exchanges energy with the atmosphere above it, primarily via convection and radiation.
+As such, T\textsubscript{surf} and T\textsubscript{are} are inextricably linked, particularly near to the ground, and a city's altered surface characteristics and energy balance manifest in modifications to both surface and air temperature regimes.
+
+%While most study of urban effects on climate is concentrated on analysis of urban air temperatures, the linkage between the two, and an increased focus on the importance of the surface in boundary-layer climatology, has also  Knowledge of how cities interface with and modify climates can come through study of the urban effect on both surface and air temperatures.
+
+%As accurate characterization of all of the terms of the energy balance is difficult and plagued by uncertainties and closure problems, characterization of how cities affect the \textit{net result} of energy balances at and near the surface - via characterization of the urban effect on air and surface temperatures - provides a viable alternative for understanding how cities interface with and modify climate.
+
+%The relationship between the temperature of the surface and the temperature of the air is complex, particularly in cities, and poorly understood.
+
+The notion that urban modifications to the environment produce elevated T\textsubscript{air} and T\textsubscript{surf} is not new, with the first formal studies of the urban effect on T\textsubscript{air} dating back to \citet{Howard1833}'s identification of "artificial warming" and "urban contamination" in his characterization of spatial patterns of T\textsubscript{air} in early 19\textsuperscript{th} century London.
 Conceptually, contemporary studies of urban climate do not stray far from \citet{Howard1833}, by seeking to isolate and quantify the urban signal in measured or modeled assessment of a given meteorological variable.
 However, as \citet{Howard1833} identified nearly two centuries ago, the sheer number of factors influencing direct measurement of climates makes experimental control nearly impossible in observational study.
-In any given observation of climate, myriad urban and non urban signals make it difficult to determine what is urban and what is not, degrading interstudy comparison.
-Even in the modeling domain, where experimental control is possible, the inherent complexity in simulating flows of energy, momentum, and mass at the scales necessary to represent physical phenomena over realistic urban environments is, as of yet, not possible.
+In any given observation of climate, myriad urban and non urban signals make it difficult to reliably differentiate urban influenced measures from those free of urban influence, degrading interstudy comparison.
+Even in the modeling domain, where experimental control is possible, the inherent complexity in simulating flows of energy, momentum, and mass at the scales necessary to represent physical phenomena over realistic urban environments is, as of yet, not feasible.
 Notwithstanding implementation of accurate, physically based schemes to represent urban climates in regional or global climate models.
-Thus, accurate assessment how a city affects the climate is a deceptively challenging task.
+Thus, accurate assessment of how a city affects the climate is a deceptively challenging task.
 To address this, \citet{Lowry1977} presents a framework to isolate the urban effect in observation of a given meteorological variable.
@@
 Thus, isolation of the urban effect on observed $V_M$ simply requires the removal of $V_B$ and $V_L$.
-This can be achieved by subtracting some urban affected observation ($V_{M, U}$) from a non-urban observation ($V_{M, R}$) where $V_{H, R}$ = 0,
+This can be achieved by subtracting some urban affected observation ($V_{M, \ U}$) from a non-urban observation ($V_{M, \ R}$) where $V_{H, \ R}$ = 0,
 
@@
 \begin{equation}
-	V_H = V_{M, U} - V_{M, R}
+	V_H = V_{M, \ U} - V_{M, \ R}
 \end{equation}
 
-Study of the urban heat island effect tacitly falls into such a framework, by replacing $V_{M, U} - V_{M, R}$ with $\Delta$T\textsubscript{air, U-R}.
-However, in spite of its conceptual simplicity through the lens of the Lowry method, meta-analysis of observational air temperature UHI (aUHI) study suggests that that simplicty may not translate well to real world analysis.
-In his analysis of the aUHI literature, \citet{Stewart2011} found that nearly fifty percent of aUHI studies published between 1950 and 2007 are "scientifically indefensible" based on an analysis of relevant criteria\footnote{Each study was given a "passing" or "failing" grade and ranked using a points-based scheme to assess methodological quality in the following criteria: conceptual model, operational definitions, instrument specification, site metadata, site representativeness, number of replicates, weather control, surface control, and synchronicity.}, suggesting a significant gulf between conceptual simplicity and practical realities in aUHI analysis.
+Study of the UHI with respect to T\textsubscript{air} (aUHI) and T\textsubscript{surf} (sUHI) tacitly falls into such a framework, by replacing $V_{M, \ U} - V_{M, \ R}$ with $\Delta$T\textsubscript{U - R}.
+However, in spite of its conceptual simplicity through the lens of the Lowry method, meta-analysis of observational aUHI studies suggests that that simplicity may not translate well to real world analysis.
+In a points based assessment of 190 aUHI studies published between 1950 and 2007, \citet{Stewart2011} found that nearly 50\% of studies are "scientifically indefensible".
+In his analysis, each study was given a "passing" or "failing" grade and ranked based on methodological quality in the following criteria: conceptual model, operational definitions, instrument specification, site metadata, site representativeness, number of replicates, weather control, surface control, and synchronisity.
 Of the 190 surveyed studies, only 13\% were derived from field sites that were sufficiently representative of the local-scale environment or lacked the meta-data to make such a determination, leading to potential biases from micro-scale terrain.
-This is particularly discouraging as it is not the result of an evolution in measurement techniques or improved instrumentation, as techniques to observe and analyze aUHI did not see significant evolution since 1950 - in fact, \citet{Stewart2011} suggests the opposite: a disproportionately large number of studies deemed scientifically sound were published early in the study purview.
+89\% lacked meta-data altogether.
+These findings suggest a significant gap between conceptual simplicity and practical realities in UHI analysis.
+
+Conceptual problems in the aUHI literature are particularly discouraging as they are not the result of an evolution in measurement techniques or improved instrumentation.
+Techniques to observe and analyze aUHI did not see significant evolution in the period between 1950 and 2007 - in fact, \citet{Stewart2011} suggests the opposite: a disproportionately large number of studies deemed scientifically sound were published early in the study purview.
+In the five years since its publication, aUHI study has seen further expansion and, as such, it is difficult to assess whether its critical analysis has prompted a shift towards more thoughtful, careful, and methodical study of aUHI, notwithstanding its relevance in study of the urban effect on other meteorological variables.
 Thus, in study of the urban effect on variables for which significant methodological shifts have or will occur, the warnings implicit in \citet{Stewart2011} are particularly salient, as inherent difficulties found in translating "Lowry"-esque conceptual models to real world observations may be compounded by significant methodological changes.
-In the five years since its publication, aUHI study has seen further expansion and, as such, it is difficult to assess whether its critical analysis has prompted a shift towards more thoughtful, careful, and methodical study of aUHI, notwithstanding its relevance in study of the urban effect on different meteorological variables.
+What is clear in light of these findings, however, is that assessment of the urban effect on climate is \textit{fundamentally difficult} in spite of its conceptual simplicity.
 
-In contrast to study of urban T\textsubscript{air}, the introduction of new methods, instruments, and analytical techniques in TIR remote sensing of urban environments has lead to rapid evolution in study of the urban effect on land T\textsubscript{surf} - a vast majority of which has focused on passive remote sensing of upwelling TIR.
+%In the context of assessment of the urban effect on T\textsubscript{surf} assessment, where new techniques, instruments, and add significant complexity, critical assessment of both conceptual and methodological shortcomings should accompany assessment of biases in different measurement and modeling techniques.
+
+In contrast to urban T\textsubscript{air} and aUHI assessment, where measurement techniques have not undergone significant evolution in recent memory and methodological problems arise in large part from conceptual shortcomings and simple carelessness, study of the urban effect on land T\textsubscript{surf} is relatively new and significantly more complex.
+As such, methodological problems in sUHI study can arise from both conceptual flaws \textit{and} instrument biases.
+In spite of these difficulties, the introduction of new methods, instruments, and analytical techniques in TIR remote sensing of urban environments has lead to rapid proliferation in study of urban T\textsubscript{surf} and the sUHI - much of which is focused on remote sensing of upwelling TIR.
 In the decades preceding this thesis, a combination of factors has lead to an increased focus on remote sensed study of urban T\textsubscript{surf}:
@@
 	\item The proliferation of satellite and aerial thermal infrared (TIR) remote sensors and "openness" in data access policies.
-	\item Improvements in sensor spatial, spatial, and radiometric resolutions. 
+	\item Improvements in sensor spatial, spectral, and radiometric resolutions. 
 	\item An increased focus on the importance of the surface in both determining and understanding key near-ground micro-meteorological phenomena.
@@
 
-Remote sensed study of the urban effect on land T\textsubscript{surf} has been instrumental in characterizing large-scale spatial and temporal patterns of sUHI \citep{Peng2012, Streutker2003, Imhoff2010}.
+Remote sensed study of the urban effect on land T\textsubscript{surf} has been critical in characterizing large-scale spatial and temporal patterns of sUHI \citep{Peng2012, Streutker2003, Imhoff2010}.
 However, the methodological and conceptual concerns raised in \citet{Stewart2011} in combination with the wide (and broadening) range of methodologies for remote sensed urban T\textsubscript{surf} retrieval should prompt critical assessment of the remote sensed sUHI literature.
-Difficulties suggested in \citet{Stewart2011} may be amplified by methodological shifts and a variety of methods for analysis.
 At the time of writing, no systemic or critical reviews of the remote sensed sUHI literature are available.
-An increasingly dogmatic focus on improving spatial and radiometric resolutions has left other inherent biases in remote sensing of urban environments largely ignored in sUHI literature.
+In addition, an increasingly dogmatic focus on improving spatial and radiometric resolutions has left other biases inherent in remote sensing of urban environments largely ignored in sUHI literature.
 This suggests that similar conclusions to those found of aUHI literature in \citet{Stewart2011} are possible (if not likely) for study of sUHI.
+Assessment of these biases is, therefore, paramount to ensure assessments of the urban effect on T\textsubscript{surf} are robust, methodologically valid, and comparable across techniques, space, and time.
+
+\section{Biases and shortcomings in thermal remote sensing of urban T\textsubscript{surf}}
 
 Traditional methods for urban T\textsubscript{surf} measurement are subject to a suite of geometric and temporal biases.
-These biases are largely inherent to the discipline and contribute an additional source of error on top of potential concerns carried over from meta-analysis of conceptual flaws in aUHI literature.
+Geometric biases are the result of urban modification of surface structure and thermal and radiative properties combined with the 2-dimensional, narrow field-of-view (FOV) viewing geometry of conventional remote sensors:
 
-Geometric biases in TIR remote sensing of the urban surface are the result of urban modification of surface structure and thermal and radiative properties.
+%These biases are largely inherent to the discipline and contribute an additional source of error on top of potential concerns carried over from meta-analysis of conceptual flaws in aUHI literature.
+
+\noindent---\textit{Undersampling of surface 3-dimensionality by narrow-FOV remote sensors}.
+
 Relative to typical non-urban and rural terrain, urban development adds significant 3-dimensionality to the surface.
 This added complexity results in undersampling of the surface by narrow-FOV remote sensors.
-A narrow-FOV sensor, by sampling surface 3-dimensionality in 2-dimensions, "sees" a different assemblage of urban surfaces depending on its viewing direction, resulting in a geometric bias towards some arrangement of facets.
-I.e.
-viewing from the nadir, a narrow-FOV sensor sees the urban surface as an array of rooftops and roads, the same sensor viewing the from an oblique angle sees some array of walls, rooftops, and roads depending on its orientation.
+A narrow-FOV sensor, by sampling surface 3-dimensionality in 2-dimensions, "sees" a different assemblage of urban surfaces depending on its viewing direction, resulting in a geometric bias towards some arrangement of facets; i.e.
+viewing from the nadir, a narrow-FOV sensor sees the urban surface as an array of rooftops and roads, the same sensor viewing the surface from an oblique angle sees some array of walls, rooftops, and roads depending on its orientation.
 For a sensor in the nadir, increasing FOV introduces sampling of vertical features.
@@
 	keepaspectratio]{model}
-	\caption{Projected FOVs for a narrow-FOV sensor at nadir and oblique angles, and a wide-FOV hemispherical sensor viewing an idealized urban array with building width T\textit{H}, height 2\textit{H}, and inter element spacing textit{H}.
+	\caption{Projected FOVs for a narrow-FOV sensor at nadir and oblique angles, and a wide-FOV hemispherical sensor viewing an idealized urban array with building width \textit{H}, height 2\textit{H}, and inter element spacing \textit{H}.
 Figure from \cite{Adderley2015}.}
@@
 
+\noindent---\textit{Directional dependence in remote sensed urban T\textsubscript{surf}}.
+
 In addition to modifying how a remote sensor samples the surface, the convoluted 3-dimensional structure of the urban surface modifies sunlit/shading regimes, resulting in strong micro-scale spatiotemporal contrasts in urban T\textsubscript{surf}, which depend on surface-sun geometry.
-These microscale variations in urban T\textsubscript{surf} are often amplified by significant directional contrasts in the thermal admittance of common urban materials, which determines the diurnal amplitude of T\textsubscript{surf} for a given facet.
+These microscale variations in urban T\textsubscript{surf} are often amplified by significant geometric and spatial contrasts in the thermal admittance of common urban materials, which determines the diurnal amplitude of T\textsubscript{surf} for a given facet.
 Thus, when viewed from a narrow-FOV remote sensor with some inherent geometric bias, urban T\textsubscript{surf} is directionally dependent - constituting an "effective anisotropy" of urban T\textsubscript{surf}, shown in Figure \ref{anisot}.
@@
 
-Geometric undersampling by narrow-FOV remote sensors results in remote sensed urban T\textsubscript{surf} changing based on a sensor's viewing direction and the assemblage facets included within the sensor FOV.
+Geometric undersampling by narrow-FOV remote sensors results in remote sensed urban T\textsubscript{surf} changing based on a sensor's viewing direction and the assemblage of facets included within the sensor FOV.
 A sensor viewing the surface perpendicular (parallel) to the sun's angle will tend to underestimate (overestimate) the "true" complete urban surface temperature - often calculated as an area weighted average of wall, rooftop, and road T\textsubscript{surf} - by differentially biasing sunlit or shaded facets.
@@
 
-\noindent---\textit{Contamination by turbulence forced, high frequency fluctuations in urban T\textsubscript{surf}}.
+\noindent---\textit{ Contamination by turbulence forced, high frequency fluctuations in urban T\textsubscript{surf}}.
 
@@
 rooftops, walls, roads, and vegetation) display large micro-scale (second to minute) fluctuations in T\textsubscript{surf}.
-The magnitude of which is inversely related to surface thermal admittance and is, as of writing, poorly understood.
+The magnitude of these is inversely related to surface thermal admittance and is, as of writing, poorly understood.
 Most thermal remote sensors provide instantaneous, rather than temporally averaged, T\textsubscript{surf} and thus are subject to contamination by high-frequency fluctuations.
-These biases are difficult to estimate in urban environments, where a large variety of fabric materials can produce significant directional contrasts in thermal admittance - and thus spatial variations in the magnitude of microscale fluctuations depending on the facet material types viewed by the sensor.
+These biases are difficult to estimate in urban environments, where a large variety of fabric materials can produce significant geometric and spatial contrasts in thermal admittance - and thus directional and spatial variations in the magnitude of microscale fluctuations depending on the facet material types viewed by the sensor.
 
@@
 
-Satellite overpass cycles are discontinuous.
+Overpass cycles for polar orbiting sensors are discontinuous.
+These sensors provide data for the vast majority of urban T\textsubscript{surf} study.
 Assessment of land T\textsubscript{surf} either sacrifices spatial resolution for a daily repeat cycle - as is the case with MODIS  - or sacrifices temporal resolution for a higher spatial resolution - ASTER or Landsat.
-Geostationary satellites do not currently have sufficient spatial resolution to resolve coherent urban pixels.
 
@@
 
-Clouds absorb TIR – thus satellite TIR remote sensing of the surface requires clear sky conditions.
+Clouds absorb TIR - thus satellite TIR remote sensing of the surface requires clear sky conditions.
 Although the urban effect on T\textsubscript{surf} is most evident under "satellite friendly" clear sky conditions (manifesting as large sUHI magnitudes), a clear sky bias likely entails overestimation of "all-sky" sUHI and further adds to discontinuities in the remote sensed urban T\textsubscript{surf} record.
@@
 As is the case with geometric biases, the magnitude of these temporal biases in remote sensed urban T\textsubscript{surf} is difficult to quantify without long-term ground truthing campaigns or significant interpolation.
-Thus, the true temporal and geometric nature of urban effect of T\textsubscript{surf} is presently unknown.
+As such, the true temporal and geometric nature of urban effect of T\textsubscript{surf} is presently unknown.
 

### 2017-05-08 paper1.tex
Add files via upload

@@
 A radiative transfer code, such as MODTRAN \citep{Berk1987}, can be used to model the effect of the atmosphere on a remote sensed TIR signal.
-Radiative transfer simulations run in MODTRAN version 4.1 for two path lengths are shown in figure \ref{spectairtsurf}, for spectral radiance emitted from a planar surface at 300 \si{kelvin} through an isothermal atmosphere at 290 \si{\kelvin}, water vapor content of 12 \si{\gram\per\meter\cubed}, and aerosol and trace gas profiles from the mid-latitude summer standard atmosphere \citep{Kantor1962}.
+Radiative transfer simulations run in MODTRAN version 4.1 for two path lengths are shown in Figure \ref{spectairtsurf}, for spectral radiance emitted from a planar surface at 300 \si{kelvin} through an isothermal atmosphere at 290 \si{\kelvin}, water vapor content of 12 \si{\gram\per\meter\cubed}, and aerosol and trace gas profiles from the mid-latitude summer standard atmosphere \citep{Kantor1962}.
 At both short and long path lengths, radiance in each band shows some deviation from the ground emitted signal.
@@
 By convolving modeled spectral radiances by a dome transmittance curve, the correction method presented in this study accounts for dome effects.
-Thus, correction magnitudes include both atmospheric and effects from spectrally non-uniform dome transmittance (and, crudely, emissivity effects, as modeled radiances are calculated with $\epsilon$ = 0.95 and include a reflected L\textsubscript{sky} component).
+Thus, correction magnitudes include both atmospheric effects and effects from spectrally non-uniform dome transmittance (and, crudely, emissivity effects, as modeled radiances are calculated with $\epsilon$ = 0.95 and include a reflected L\textsubscript{sky} component).
 While the individual effects are difficult to separate, temperature dependent differential dome effects are visible in correction magnitudes as a persistent minor positive correction in winter.

### 2017-05-08 paper2.tex

@@
 
-\subsection{Seasonality in sUHI magnitudes}
+\subsection{Diurnal and seasonal variability in sUHI magnitudes}
 
@@
 
-Figure \ref{heatcluhi} shows seasonal variations in diurnal patterns of mean monthly clUHI calculated for the same study period as Figure \ref{heatsuhi}.
+Figure \ref{heatcluhi} shows seasonal variation in diurnal patterns of mean monthly clUHI calculated for the same study period as Figure \ref{heatsuhi}.
 clUHI development is strongly controlled by the timing of sunset/sunrise cycles and displays the largest seasonal variance in the hours immediately after sunset.
@@
 
+Figure \ref{heat_uhi_diff} shows seasonal variation in the difference between mean monthly hemispherical sUHI\textsubscript{norm} and clUHi\textsubscript{norm} for the BUBBLE campaign.
+By restricting UHI values to a range between one and zero, diurnal patterns of sUHI and clUHI can be compared directly.
+Comparison of the two in summer shows a peak sUHI approximately two hours after solar noon (when clUHI is small) and peak clUHI approximately two hours after sunset (when sUHI is small).
+Values near zero in the winder months indicate little variation in both sUHI and clUHI.
+
 \begin{figure}[H]
@@
 
+Figure \ref{vanc_suhi} shows seasonal variations in diurnal patterns of mean monthly hemispherical sUHI magnitudes at 30 minute intervals for the EPiCC campaign.
+The sUHI in Vancouver shows similar seasonal and diurnal patterns to the sUHI in Basel, but does not persist as far into the late evening hours.
+
 \begin{figure}[H]
@@
 	\label{vanc_suhi}
+\end{figure}
+
+Figure \ref{bxsUHI_season_bubble} shows hemispherical sUHI magnitudes for the eight month BUBBLE campaign and binned for winter (December through February) and summer (April through July) seasons.
+Diurnal variability in sUHI magnitudes is concentrated in summer months, as well as seasonal sUHI\textsubscript{max}.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=15cm,height=7in,keepaspectratio]{bxsUHI_season_bubble}
+	\caption{Hemispherical sUHI magnitudes for the eight month BUBBLE climatology (top) and binned for winter (middle) and summer (bottom) seasons.
+Grey shading indicates nighttime hours averaged over the bin interval.}
+	\label{bxsUHI_season_bubble}
 \end{figure}
@@
 
-Figures \ref{meteo_kdown}, \ref{meteo_hum}, \ref{meteo_wv}, and \ref{meteo_atuhi} show hemispherical sUHI magnitudes as a function of incoming solar radiation, water vapor content, wind velocity, and clUHI measured over the BUBBLE campaign.
-Winter hours are omitted in Figures \ref{meteo_hum}, and \ref{meteo_wv}, and \ref{meteo_atuhi} as sUHI is small and displays minimal variation in winter.
-Both winter and nighttime hours are omitted in Figure \ref{meteo_kdown} to examine daytime sUHI under clear sky and clouded sky conditions.
+Figure \ref{meteo_wv_sol} panel (a) shows mean and maximum sUHI magnitudes as a function of solar radiant exposure over each day of the summer months of the BUBBLE campaign.
+Point shading indicates the month (white is April and black is July).
+A clear positive relationship is observed between accumulated solar radiation and both sUHI\textsubscript{max} and sUHI\textsubscript{mean}.
+Panel (b) of Figure \ref{meteo_wv_sol} shows mean and maximum hemispherical sUHI magnitudes as a function of mean wind velocity for the duration of the eight month BUBBLE climatology.
+Colors indicate solar radiant exposure accumulated over each day.
+Similar results are observed when rural winds are substituted.
+
+Figures \ref{meteo_hum} and \ref{meteo_atuhi} show hemispherical sUHI magnitudes as a function of water vapor content and clUHI measured over the BUBBLE campaign.
+Winter hours are omitted in both as sUHI is small and displays minimal variation in winter.
+Jamie: still working on getting color palettes for these.
+Will merge into one figure.
 
@@
 	\centering
-	\includegraphics[width=15cm,height=7in,keepaspectratio]{meteo_kdown}
-	\caption{sUHI magnitude versus incoming solar radiation binned for clear sky (when the sum of daytime K\textsubscript{down} \textgreater 18 \si{\mega\joule\per\meter\squared}) and cloudy sky days (when the sum of daytime K\textsubscript{down} \textless 18 \si{\mega\joule\per\meter\squared})}
-	\label{meteo_kdown}
+	\includegraphics[width=15cm,height=7in,keepaspectratio]{meteo_wv_sol}
+	\caption{(a) Maximum and mean 24 \si{\hour} hemispherical sUHI magnitude for each day of April through July versus integrated solar radiant exposure measured over the same 24 \si{\hour} period.
+Shading indicates month of the year, with dark points indicating later months.
+(b) Maximum and mean 24 \si{\hour} hemispherical sUHI magnitude for each day in the eight month climatology versus mean wind velocity measured at approximately 2 \si{\meter} above ground at the Sperrstrasse site.
+Coloring indicates integrated solar radiant exposure over each day.}
+	\label{meteo_wv_sol}
 \end{figure}
@@
 	\label{meteo_hum}
-\end{figure}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=15cm,height=7in,keepaspectratio]{meteo_wv}
-	\caption{sUHI magnitude versus rural wind velocity measured at approximately 10 \si{\meter} above ground at the Lange Erlen site.}
-	\label{meteo_wv}
 \end{figure}
@@
 However, the results in \citet{Oke2017} are based on simulations of an idealized urban area without trees, complex non-orthogonal surface geometry, or sloped facets and have not been supported observationally.
-The results in this paper provides an ideal platform to test conceptual understanding of sUHI development and to understand the diurnal and seasonal nature of sUHI.
+The results in this chapter provide an ideal platform to test conceptual understanding of sUHI development, to understand the diurnal and seasonal nature of sUHI and the meteorological factors that drive its development.
 
@@
 
-sUHI development is the result of differences in urban and rural heating and cooling rates.
-A variety of factors modify urban heating and cooling rates relative to non-urban spaces, most notably urban-rural contrasts in sky-view-factor, impervious surface fraction, and thermal admittance (and by relation, urban-rural soil moisture contrasts).
-These differences provide the genesis for diurnal sUHI development.
-Rural surfaces have more open (less obstructed) sky view factors, and respond quickly to insolation a low solar angles in the early morning hours.
-Shading by canyon features causes urban T\textsubscript{surf} to respond more slowly, resulting in a small sUHI\textsubscript{min} in the hours just after sunset.
+sUHI development is the result of urban modification to surface geometry and thermal and radiative properties, most notably sky-view-factor, impervious surface fraction, and thermal admittance (and by relation, urban-rural soil moisture contrasts).
+These differences alter urban heating/cooling rates relative to non-urban spaces and provide the genesis for diurnal sUHI development.
+In the early morning hours, rural surfaces respond quickly to insolation at low solar angles, and rural T\textsubscript{surf} rapidly increases.
+In contrast, shading by urban features causes urban T\textsubscript{surf} to respond more slowly, resulting in a small sUHI\textsubscript{min} in the hours just after sunrise.
 This is observed in Basel across all representations of urban T\textsubscript{surf} at similar magnitudes.
@@
 Increases in urban T\textsubscript{surf} are amplified by the abundance of impervious surfaces, which result in a smaller fraction of insolation partitioned to latent heat relative to vegetated non urban spaces.
-After the initial spike in rural heating in the morning hours, urban heating rates overtake rural and the sUHI slowly grows through midday.
-In the hours after midday, the solar angle decreases and both urban and rural heating rates decline and begin to converge, as shading by canyon geometry causes some facets to cool.
-sUHI reaches a peak at approximately two to three hours hours after solar noon.
+After the initial spike in rural heating in the early morning hours, urban heating rates overtake rural approximately three to four hours after sunrise and the sUHI slowly grows through midday to a peak at approximately two to three hours hours after solar noon.
+In the hours after UHI\textsubscript{max}, the solar angle decreases and both urban and rural heating rates decline and begin to converge, as shading by canyon geometry causes some facets to begin to cool.
 
-In the afternoon hours, both urban and rural surfaces transition to cooling, with approximately equal rates.
-The late afternoon and evening decline in sUHI is strongly controlled by urban surface structure.
+In the afternoon hours, both urban and rural surfaces transition to cooling, this can lead to several different patterns in evening sUHI development and is strongly controlled by urban surface structure.
 Relative to non-urban spaces, cooling rates for wall and road facets lag slightly as a result of canyon radiation trapping, this is offset by increased cooling rates for low thermal admittance rooftop facets.
@@
 In urban areas with relatively unobstructed sky-view-factors or large rooftop plan area fractions, strong radiative cooling can lead to a sharp reduction in sUHI.
-The opposite is true of urban areas with obstructed sky view factors, and tall street canyons with large wall surface area fractions, where canyon radiative trapping reduces radiative cooling and sUHI is maintained through the evening.
+The opposite is true of urban areas with restricted sky view factors and tall street canyons with large wall surface area fractions, where canyon radiative trapping reduces radiative cooling and sUHI is maintained through the evening.
 For the Sperrstrasse canyon, the combination of a narrow but relatively low street canyon results in a slight reduction in Basel's sUHI as strong radiative cooling of rooftops is slightly offset by canyon radiation trapping.
+The dependence of sUHI development on canyon geometry is illustrated well by comparing nadir and complete representations of sUHI observed in Basel.
+A nadir view, by virtue of oversampling rapidly cooling rooftops, observes a strong evening decline in sUHI.
+This is roughly equivalent to sUHI calculated from an urban site with low canyon height to width ratio and a relatively unobstructed sky view factor.
+In contrast, by sampling vertical facets that cool more gradually, this decline is less evident in sUHI calculated from a complete view of Basel T\textsubscript{surf}.
+Thus, as expected, remote sensed sUHI analysis is also directionally dependent and subject to anisotropic effects when viewed from a narrow-FOV sensor.
+These effects not only affect the magnitude of remote sensed sUHI, but may also alter perceived diurnal patterns of sUHI.
 
-After sunset, rural cooling rates spike relative to urban, as an unobstructed rural sky-view-factor allows for strong radiative cooling of the surface, leading to a slight jump in sUHI.
+After sunset, the unobstructed rural sky-view-factor allows for strong radiative cooling of the surface and rural cooling rates spike relative to urban, leading to a slight jump in sUHI.
 However, variations in rural thermal admittance, surface geometry, land cover and use, and/or influence from surrounding terrain can result in significant contrasts in rural cooling rates.
 All of these factors show strong spatiotemporal variation, resulting in large differences in sUHI development depending on rural site characteristics.
-This is evident in characterization of sUHI in Basel, where the two rural sites entail different rural cooling rates, particularly in the hours just after sunset, and different patterns of post-sunset sUHI development.
-Although similar in terms of surface coverage and site geometry, BUBBLE rural sites are subject to different after sunset air masses that strongly effect sUHI (and clUHI) development.
-Hilly terrain surrounding the Lange Erlen site results in cold air drainage and an increased rural cooling rate, particularly following hot, clear sky days.
+This is illustrated by two different patterns of post-sunset sUHI development observed for the two rural sites in Basel.
+Although similar in terms of surface coverage and site geometry, hilly terrain surrounding the Lange Erlen site results in cold air drainage and an increased rural cooling rate, particularly following hot, clear sky days.
 This effect is not present at the Village Neuf site, where nocturnal cooling rates are likely driven primarily driven by radiative losses.
 The "extra", largely non-radiative cooling at the Lange Erlen site results in a strong second peak in sUHI that is not present in sUHI calculated using the Village Neuf site.
-This makes clear the need for proper characterization of the radiative and thermal properties, surface geometry and coverage of rural sites in sUHI analysis, as well as the surrounding topography to ensure characterizations of sUHI development are free of confounding influence.
-After initial nighttime divergence, sUHI in Basel is relatively constant over the nighttime hours, with slight growth observed on clear nights where canyon radiation trapping retards urban nocturnal cooling rates relative to rural.
+Following site dependent effects on sUHI in the hours after sunset, sUHI in Basel is relatively constant over the nighttime hours, with slight growth observed on clear nights where canyon radiation trapping retards urban nocturnal cooling rates relative to rural.
+
+% This makes clear the need for proper characterization of the radiative and thermal properties, surface geometry and coverage of rural sites in sUHI analysis, as well as the surrounding topography to ensure characterizations of sUHI development are free of confounding influence.
 
@@
 
-Hemispherical sUHI in Basel displays significant seasonality, both in terms of its diurnal pattern and overall magnitude.
+Hemispherical sUHI in both Basel and Vancouver displays significant seasonality, both in terms of its diurnal pattern and overall magnitude.
 Seasonal variations in sUHI are most evident during the daytime hours, where large seasonal contrasts in cloud cover frequency and solar angle exert strong control on daytime heating rates and sUHI development.
-Nighttime sUHI is consistent across seasons.
-Peak seasonal sUHI occurs near the summer solstice, when days are longest and isolation on clear sky days is most intense, forcing significant contrasts in urban-rural heating/cooling rates.
-In winter, daytime sUHI is suppressed from a lower solar input and frequent cloud cover.
-These reduce urban-rural contrasts in heating/cooling rates and weaken wintertime sUHI.
-In Basel, wintertime sUHI has a small peak shortly after sunset, when cold air drainage amplifies rural cooling rates and increases sUHI.
-This, the small after sunset peak in sUHI present in both summer and winter observations is likely not generalizable and is specific to the Lange Erlen site.
+Mean monthly sUHI\textsubscript{max} is observed near the summer solstice in both study areas, when days are longest and insolation on clear sky days is most intense, forcing significant contrasts in urban-rural heating/cooling rates and the development of a strong sUHI.
+In winter, lower overall solar input and frequent cloud cover suppress the potential for urban-rural contrasts in heating/cooling rates and weakens wintertime sUHI development.
+In Basel, wintertime sUHI has a small peak shortly after sunset, when cold air drainage anomalously amplifies rural cooling rates, increasing the sUHI.
+This small, after-sunset peak in sUHI observed in both summer and winter observations is specific to the Lange Erlen site and is not generalizable.
+Cooling rates at the Vancouver Westham island rural site are not affected by cold air drainage and, as a result, sUHI does not display two diurnal peaks in summer nor a coherent peak after sunset in winter months.
+Thus, observations of seasonal variations in hemispherical sUHI in Vancouver are likely highly generalizable to other mid-latitude cities with distinct winter and summer seasons.
 
-These results are supported by similar seasonal patterns of hemispherical sUHI observed in Vancouver, where seasonal sUHI\textsubscript{max} (sUHI\textsubscript{min}) is observed near the summer (winter) solstice.
-Observations of sUHI in Vancouver show similar seasonal patterns of development to those in Basel, with sUHI with a larger diurnal sUHI amplitude in summer months and suppressed sUHI development in winter.
-Cooling rates at the Westham island rural site are not affected by cold air drainage and, as a result, sUHI observed in Vancouver does not display two diurnal peaks in summer nor a coherent peak after sunset in winter months.
-
-Although sUHI and clUHI are controlled by similar climatic controls and have a similar micro meteorological genesis, they do not display the same seasonal patterns.
-In contrast to sUHI, clUHI in Basel is greatest just after sunset following closely to sunset/sunrise timing.
-Peak clUHI occurs just after sunset across seasons, differing in terms of magnitude but not time of day, with a larger peak clUHI in summer\footnote{As is the case with sUHI observed in Basel, clUHI magnitudes are effected by cold air drainage at the Lange Erlen site and are slightly higher than expected \citep{Runnalls2000}.}.
-Daytime clUHI is small and displays little seasonal variation.
+The sUHI and the clUHI have similar micro meteorological genesis, climatic controls, and display similar seasonal variability, albeit concentrated at different times of the day.
+This is evident when comparing normalized mean monthly hemispherical sUHI and clUHI - where the range of both types of sUHI is restricted to values between zero and one.
+In winter, low solar input and frequent cloud cover inhibit variability in both sUHI and clUHI and both are of similar normalized magnitudes.
+In summer, larger solar input fosters contrasts in urban-rural heating and cooling rates and both clUHI and sUHI are most variable, with peak sUHI in the early afternoon clUHI is smallest, and peak clUHI in the hours after sunset when sUHI is relatively low.
+These results suggest the following generalization: when conditions that foster the formation of strong urban-rural contrasts in T\textsubscript{air} and T\textsubscript{surf} are present (most notably a large solar input), sUHI peaks approximately two hours after solar noon and clUHI peaks approximately two hours after sunset.
 Thus, a surface urban heat island is not simply a stronger, more pronounced, canopy layer urban heat island.
@@
 Sensor-surface geometry exerts significant influence on observed sUHI magnitudes.
-By undersampling vertical and sloped urban surfaces and oversampling low thermal admittance fabric materials, sUHI\textsubscript{nadir} overestimates mean afternoon sUHI\textsubscript{comp} by 3.41 \si{\kelvin} and up to 5.96 \si{\kelvin} under clear sky “satellite friendly” conditions.
-Thus, peak sUHI\textsubscript{nadir} overestimates the 'true' sUHI\textsubscript{comp} by nearly double under clear sky conditions.
+2-dimensional, complete, and hemispherical observations of sUHI not only yield different magnitudes but also different diurnal sUHI regimes and peak timings.
+Variations in observed sUHI across different sensor sampling regimes are largest under clear sky, hot conditions, coinciding with typical seasonal and diurnal sUHI\textsubscript{max}.
+In general, as urban T\textsubscript{surf} is most often viewed from the top down, remote sensors preferentially sample horizontally oriented facets with unobstructed sky view factors and undersample or ignore vertical and sloped facets with more obstructed sky view factors.
+Over/undersampling of the urban surface changes both the magnitude and character of remote sensed urban T\textsubscript{surf} and sUHI.
+
+In an analysis of Basel sUHI from complete, nadir, and hemispherical representations of the surface, significant variation in remote sensed sUHI was observed across sensor sampling regimes at all times of the day.
+Variations in observed sUHI are most evident on clear sky, hot days in the hours after solar noon, when microscale spatiotemporal contrasts in urban T\textsubscript{surf} are largest.
+Cloud cover suppresses the influence of sensor-surface geometry on remote sensed sUHI by reducing microscale geometric and spatiotemporal contrasts in urban T\textsubscript{surf}, resulting in less facet-scale heterogeneity in urban T\textsubscript{surf} and reducing the overall magnitude of sUHI for all definitions.
+
+By undersampling vertical and sloped urban surfaces and oversampling low thermal admittance materials, sUHI\textsubscript{nadir} overestimates mean afternoon sUHI\textsubscript{comp} by approximately 3.5 \si{\kelvin} and up to 6.0 \si{\kelvin} under clear sky “satellite friendly” conditions.
+This constitutes a significant bias in remote sensed sUHI\textsubscript{max} as sUHI\textsubscript{nadir} overestimates the 'true' sUHI\textsubscript{comp} by nearly double under clear sky conditions.
 As clear sky conditions are the only conditions under which satellite TIR remote sensors can operate, and make up a large fraction of the urban T\textsubscript{surf} literature, these biases are pervasive and cannot be ignored.
-In contrast, by sampling the urban surface in 3-dimensions, sUHI\textsubscript{hem} yields a much smaller mean daytime overestimation (0.54 \si{\kelvin}), particularly under clear sky conditions (0.93 \si{\kelvin}), providing a significantly more representative depiction of urban T\textsubscript{comp} and sUHI\textsubscript{comp}, in addition to allowing for time-continuous analysis.
-Cloud cover suppresses the influence of sensor-surface geometry on remote sensed sUHI by reducing microscale geometric and spatiotemporal contrasts in urban T\textsubscript{surf}, resulting in less facet-scale heterogeneity in urban T\textsubscript{surf} and reducing the overall magnitude of sUHI for all definitions.
+In contrast to sUHI\textsubscript{nadir}, sUHI\textsubscript{hem} samples the urban surface in 3-dimensions ("seeing" a wider range of facet types and orientations) and yields a much smaller mean daytime overestimation (0.5 \si{\kelvin}), particularly under clear sky conditions (1.0 \si{\kelvin}), providing a significantly more representative depiction of urban T\textsubscript{comp} and sUHI\textsubscript{comp}, in addition to allowing for time-continuous analysis.
 
@@
 sUHI is highly dependent on meteorological conditions, which account for much of the day to day variation sUHI magnitudes.
-Conditions that foster the largest microscale contrasts in urban T\textsubscript{surf} also entail the largest sUHI magnitudes.
-Contra, conditions that suppress microscale contrasts in T\textsubscript{surf} show similarly reduced sUHI magnitudes.
-As such, incoming shortwave radiation and wind velocity display positive and negative relationships with sUHI magnitudes respectively.
-Under clear sky conditions, strong solar heating of the surface combined with differences in urban and rural thermal properties and sunlit shading geometries fosters the development of large daytime urban-rural contrasts in T\textsubscript{surf} and T\textsubscript{air} and the development of strong sUHI by day and clUHI by night.
-Thus, sUHI\textsubscript{max} coincides with insolation\textsubscript{max}.
-However, significanyUnder cloudy conditions, decreased direct solar heating of the surface suppresses the development of urban-rural thermal contrasts, resulting in a reduced sUHI.
-$\Delta$T\textsubscript{hem, r - air} has a similar positive effect on sUHI, owing to its dependence on solar insolation.
+Conditions that foster the largest microscale spatiotemporal contrasts in urban T\textsubscript{surf} also entail the largest sUHI magnitudes.
+Contra, conditions that suppress microscale contrasts in T\textsubscript{surf} show reduced sUHI magnitudes.
+As such, sUHI development is strongly tied to integrated solar radiant exposure accumulated over a day and wind velocity, which display positive and negative relationships with sUHI magnitudes respectively.
+Meteorological controls on sUHI are summarized in Table \ref{meteo_cont}.
+
+The relationship between solar radiation and sUHI is relatively simple but easily misinterpreted.
+Surfaces (both urban and otherwise) change in temperature based on accumulated radiative or convective gains or losses - as such, instantaneous changes in insolation have little effect on sUHI magnitudes.
+Thus, to say that sUHI is a function of measured incoming solar radiation (in \si{\watt\per\square\meter}) does not accurately characterize the relationship and produces a significant amount of unrepresentative scatter.
+Strong accumulated solar heating of the surface combined with differences in urban and rural thermal properties and sunlit shading geometries fosters the development of large daytime urban-rural contrasts in T\textsubscript{surf} and T\textsubscript{air} and the development of strong sUHI by day and clUHI by night.
+Thus, sUHI\textsubscript{max} often coincides with insolation\textsubscript{max}.
+Under cloudy conditions, reduced accumulated solar heating of the surface suppresses the development of urban-rural thermal contrasts, resulting in reduced sUHI magnitudes.
+Owing to its dependence on accumulated solar heating, $\Delta$T\textsubscript{hem, r - air} has a similar positive effect on sUHI magnitudes.
+
+Wind velocity has a similarly nuanced relationship with sUHI.
+Winds affect surface heating and cooling rates by facilitating convective cooling (or, under less common circumstances, convective heating).
+In general, as wind velocity increases, convective heat losses increase and T\textsubscript{surf} decreases.
+This suppresses microscale spatiotemporal variations in T\textsubscript{surf} as hot sunlit surfaces experience greater rates of convective cooling under high wind conditions than cooler shaded surfaces.
+This reduces overall urban and rural T\textsubscript{surf} and inhibits sUHI development.
+Nighttime sUHI is also affected by wind velocity as the formation of a near surface thermal inversion, which increases radiative cooling in non-urban areas, requires calm winds.
+As such, increased nighttime wind velocity decreases nocturnal sUHI.
 
@@
 As a result, water vapor content has little effect on sUHI.
-Wind velocity displays a strong negative relationship with sUHI.
-sUHI is greatest under calm winds when near-surface mixing is minimal, fostering strong microscale and urban-rural contrasts in T\textsubscript{surf}.
-As wind velocity increases, sUHI is suppressed from increased turbulent heat losses.
-During the nighttime hours, increased wind velocities destroy the surface based thermal inversion often observed over rural terrain.
-This further limits nighttime cooling rates relative to urban and decreases sUHI magnitudes.
-Meteorological controls on sUHI are summarized in Table \ref{meteo_cont}.
 
@@
 		\midrule
-		$\Delta$T\textsubscript{surf - air} &\textit{increasing} \\
+		$\Delta$T\textsubscript{surf - air}, \si{\kelvin} &\textit{increasing} \\
 		&\\
-		T\textsubscript{surf} & \textit{increasing} \\
+		T\textsubscript{surf}, \si{\degreeCelsius}  & \textit{increasing} \\
 		&\\
-		T\textsubscript{air} & \textit{increasing} \\
+		T\textsubscript{air}, \si{\degreeCelsius} & \textit{increasing} \\
 		&\\
-		Incoming solar radiation & \textit{increasing}  \\
+		Insolation, \si{\mega\joule\per\meter\squared\per\day} & \textit{increasing}  \\
 		& \\
-		Water vapor content & \textit{negligible} \\
+		Water vapor content, \si{\gram\per\meter\squared} & \textit{negligible} \\
 		&\\
-		Wind velocity &\textit{decreasing} \\
+		Wind velocity, \si{\meter\per\second} &\textit{decreasing} \\
 		&\\
-		clUHI &\textit{increasing} \\
+		clUHI, \si{\kelvin} &\textit{increasing} \\
 		\bottomrule
@@
 clUHI and sUHI develop at different times of day and, thus, have a complex diurnal relationship.
-By day, variation in clUHI is minimal and shows a weak positive correlation with sUHI.
-By night, clUHI shows greater variation and a stronger, more positive relationship with sUHI.
-The relationship between clUHI and sUHI is most evident on clear sky nights when urban-rural contrasts in air and surface cooling rates are strongest.
-Under these conditions sUHI and clUHI  show similar development in the hours after sunset.
+By day, variation in clUHI is minimal and shows weak positive correlation with sUHI.
+Occasionally, under clear sky conditions following hot days, strong evening radiative cooling forces clUHI development slightly before sunset.
+This causes a slight bifurcation in clUHI versus sUHI.
+By night, clUHI shows greater variation and a stronger, more positive relationship with sUHI, most evident on clear sky nights when urban-rural contrasts in air and surface cooling rates are strongest.
 
@@
 
-This chapter presents the results of a time-continuous eight month climatological analysis of the surface urban heat island in Basel, Switzerland derived using the method presented in Chapter \ref{paper1}.
-Hemispherical sUHI magnitudes display significant seasonal and diurnal variation and are largest in summer months during the late afternoon during which solar heating of the surface is most intense, forcing the development of strong microscale and urban-rural contrasts in T\textsubscript{surf}.
-Seasonality in late afternoon sUHI is largely dependent on incoming solar radiation, which displays a strong positive relationship with sUHI magnitudes.
-In the winter months, a reduced overall solar input and increased cloud cover frequency suppresses daytime sUHI development resulting in less day-to-day variation in sUHI, and lower overall sUHI magnitudes.
+This chapter presents the results of an analysis of two time-continuous long-term climatologies of the surface urban heat island effect in Basel, Switzerland and Vancouver, Canada derived using the method presented in Chapter \ref{paper1}.
+Hemispherical sUHI magnitudes display significant seasonal and diurnal variation and are largest in summer months during the late afternoon during which solar heating of the surface is most intense, forcing the development of strong urban-rural contrasts in T\textsubscript{surf}.
+Seasonality in late afternoon sUHI is largely dependent on accumulated radiative exposure over the day which displays a strong positive relationship with sUHI magnitudes.
+In the winter months, a reduced overall solar input and increased cloud cover frequency suppresses daytime sUHI development resulting in less variability in sUHI, and lower overall sUHI magnitudes.
 
 Remote sensed sUHI is highly dependent on sensor-surface-sun geometry.
-This results in significant variation in urban T\textsubscript{surf} and sUHI when observed from sensors with different spatial, geometric, and temporal sampling regimes.
+This results in significant variation in the character of urban T\textsubscript{surf} and sUHI when observed from sensors with different spatial, geometric, and temporal sampling regimes.
 Thus, traditional methods for thermal remote sensing of urban areas do not capture the true temporal and geometric nature of the urban effect on land T\textsubscript{surf}.
 Urban effective anisotropy results in large geometric biases in remote sensing of urban areas as traditional satellite and aerial remote sensors sample only a fraction of the urban surface.
-These geometric biases manifest not only in deviation between observed sUHI magnitudes and the 'true' complete sUHI, but also impact characterizations of diurnal and seasonal patterns of sUHI and urban T\textsubscript{surf} by changing the time of day of sUHI\textsubscript{max}.
+These geometric biases manifest not only in deviation between observed sUHI magnitudes and the 'true' complete sUHI, but also impact characterizations of diurnal and seasonal patterns of sUHI and urban T\textsubscript{surf}.
 Geometric biases in thermal remote sensing of urban area are compounded by myriad temporal biases across a range of temporal scales, as traditional methods for urban thermal remote sensing are not time-continuous and only operable under clear sky conditions.
-Both geometric and temporal biases affect the vast majority of sUHI study, most of which is performed using only a small number of satellite remote sensors, to which these biases are inherent and largely unavoidable.
+Both geometric and temporal biases affect the vast majority of sUHI study, most of which is performed using only a small number of satellite remote sensors, to which these biases are inherent.
 

### 2017-05-09 paper1.tex
Add files via upload

@@
 \section{Introduction}
+
 Thermal infrared (TIR) remote sensing of land surface temperature (T\textsubscript{surf}) has emerged as a primary research focus in climatology, as researchers seek to better describe spatiotemporal patterns of T\textsubscript{surf} globally and better understand how anthropogenic modification of earth's surface influences land T\textsubscript{surf} and impacts climate at various scales.
 Over the last two decades, the application of thermal remote sensing to study surface climates has expanded significantly.
-Thermal remote sensing of earth's surface has applications over a wide range of disciplines: from informing micro-, urban-, and global-scale climate models, to aiding decision making and mitigation praxis with respect to climate change and the \gls{uhi}.
+Thermal remote sensing of earth's surface has applications over a wide range of disciplines: from informing micro-, urban-, and global-scale climate models, to aiding decision making and mitigation praxis with respect to climate change and the urban heat island effect.
 
-Within urban climatology, a combination of satellite, aerial, and ground-based thermal remote sensors have been integral in elucidating the spatial \citep{Roth1989}, temporal \citep{Peng2012}, and geometric \citep{Voogt1997} effects of the built environment on land T\textsubscript{surf}; in evaluating and partitioning urban surface energy balances \citep{BastiaanssenW.G.M.1998, Yamaguchi2005} and; in characterizations of the relationship between surface and \gls{blayer} air temperatures (T\textsubscript{air}) \citep{Stoll1992}.
+Within urban climatology, a combination of satellite, aerial, and ground-based thermal remote sensors have been integral in elucidating the spatial \citep{Roth1989}, temporal \citep{Peng2012}, and geometric \citep{Voogt1997} effects of the built environment on land T\textsubscript{surf}; in evaluating and partitioning urban surface energy balances \citep{BastiaanssenW.G.M.1998, Yamaguchi2005,Frey2007} and; in characterizations of the relationship between surface and boundary-layer air temperatures (T\textsubscript{air}) \citep{Stoll1992}.
 These advances have been aided by substantial improvements in sensor spatial, spectral, and radiometric resolutions, and by the proliferation of both large-scale public satellite remote sensing campaigns and low-cost aerial and near-ground thermography.
@@
 
-\noindent In this paper, we seek to examine question two by introducing and evaluating a method for atmospheric correction of near-ground hemispherical TIR --- measured via \gls{pyrg} --- for hemispherical radiometric temperature (T\textsubscript{hem, r}) retrieval.
-These measures are common to most urban energy balance assessments and thus constitute a hitherto untapped method for urban T\textsubscript{surf} analysis.
-A companion paper responds to question one through an analysis of an 8-month climatology of T\textsubscript{hem, r} and derived surface UHI magnitudes (sUHI), to quantify geometric and temporal biases across multiple methods for remote sensing of urban T\textsubscript{surf}.
+\noindent In this paper, we seek to examine question two by introducing and evaluating a method for atmospheric correction of near-ground hemispherical TIR - measured via pyrgeometer - for hemispherical radiometric temperature (T\textsubscript{hem, r}) retrieval.
+These measures are common to most urban energy balance assessments because they are made as a part of the net radiation measurement and thus constitute a hitherto untapped method for urban T\textsubscript{surf} analysis.
+A companion paper responds to question one through an analysis of two long term climatologies of T\textsubscript{hem, r} and derived surface UHI magnitudes (sUHI), to quantify geometric and temporal biases across multiple methods for remote sensing of urban T\textsubscript{surf}.
 
@@
 
-Although most thermal remote sensors operate within one of the \gls{atmwind}[s] - where atmospheric effects are minimal - virtually any remote sensed TIR signal is subject to radiative effects from the layer of atmosphere between the surface and the sensor.
-Over much of the thermal infrared \gls{waveband} the atmosphere emits radiation and absorbs a fraction of radiation emitted by the surface.
-Thus, a remote sensed TIR signal almost certainly not equal to the ground emitted signal.
+Although most thermal remote sensors operate within one of the atmospheric windows - where atmospheric effects are minimal - virtually any remote sensed TIR signal is subject to radiative effects from the layer of atmosphere between the surface and the sensor.
+Over much of the thermal infrared waveband the atmosphere emits radiation and absorbs a fraction of radiation emitted by the surface.
+Thus, a remote sensed TIR signal is almost certainly not equal to the ground emitted signal.
 Spectral radiance received by a sensor at height ($z$) can be described by a function deviating from a Planck curve defined by T\textsubscript{surf} based on the spectral transmittance of the intervening atmosphere, with the magnitude of that deviation determined by the difference between Planck curves calculated from T\textsubscript{surf} and ambient T\textsubscript{air}.
 A radiative transfer code, such as MODTRAN \citep{Berk1987}, can be used to model the effect of the atmosphere on a remote sensed TIR signal.
-Radiative transfer simulations run in MODTRAN version 4.1 for two path lengths are shown in Figure \ref{spectairtsurf}, for spectral radiance emitted from a planar surface at 300 \si{kelvin} through an isothermal atmosphere at 290 \si{\kelvin}, water vapor content of 12 \si{\gram\per\meter\cubed}, and aerosol and trace gas profiles from the mid-latitude summer standard atmosphere \citep{Kantor1962}.
+Radiative transfer simulations run in MODTRAN version 4.1 for two path lengths are shown in Figure \ref{spectairtsurf}, for spectral radiance emitted from a planar surface at 300 \si{\kelvin} through an isothermal atmosphere at 290 \si{\kelvin}, water vapor content of 12 \si{\gram\per\meter\cubed}, and aerosol and trace gas profiles from the mid-latitude summer standard atmosphere \citep{Kantor1962}.
 At both short and long path lengths, radiance in each band shows some deviation from the ground emitted signal.
@@
 Spectral transmission of longwave radiation through a given layer of atmosphere is dependent on total column absorber content (the principal broadband TIR absorbers are H\textsubscript{2}O, CO\textsubscript{2}, and to a lesser extent O\textsubscript{3}, N\textsubscript{2}O, CO, CH\textsubscript{4}, and O\textsubscript{2} \citep{Miskolczi1993}).
-Holding vertical absorber content constant through a layer of atmosphere, variation in band-by-band TIR transmittance with path length is greatest at urban scales (approximately 1 to 50 \si{\meter}), where transparent spectral bands can quickly become opaque with small changes in path length or absorber content - illustrated in figure \ref{spectransheight}.
+Holding vertical absorber content constant through a layer of atmosphere, variation in band-by-band TIR transmittance with path length is greatest at the scale of an urban canyon (approximately 1 to 50 \si{\meter}), where transparent spectral bands can quickly become opaque with small changes in path length or absorber content - illustrated in Figure \ref{spectransheight}.
 Thus, transmittance of TIR near the surface is highly dependent on surface-sensor geometry, instrument spectral response, and atmospheric absorber content.
@@
 	\caption{Spectral transmission of water vapor as a function of height.
-Run in MODTRAN 4.1 for radiance emitted from a planar surface at 300 \si{\kelvin} through an atmosphere with water vapor content of 8 \si{\gram\per\meter\cubed}.
+Model results from MODTRAN 4.1 for radiance emitted from a planar surface at 300 \si{\kelvin} through an atmosphere with water vapor content of 8 \si{\gram\per\meter\cubed}.
 Sampled at heights of 1, 5, 10, 15, 20, 25, and 30 \si{\meter}.}
@@
 \noindent where \(\epsilon\)\textsubscript{\( \lambda \)} is spectral surface emissivity, \(\tau\)\textsubscript{\( \lambda \)} is spectral "slab" transmittance through the layer between the emitting surface $( z  = 0) $ and $ z $.
-Directional radiances upwelling from the atmosphere $R^\uparrow_{atm}$, and the surface $R^\uparrow_0$, and downwelling from the sky $R^\downarrow_{sky}$, can be described spectrally or integrated over a waveband bounded by \(\lambda_1\) and \(\lambda_2\) as
+Directional radiances upwelling from the atmosphere $R^\uparrow_{atm}$, and the surface $R^\uparrow_0$, and downwelling from the sky $R^\downarrow_{sky}$, can be described spectrally or integrated over a waveband bounded by \(\lambda_1\) and \(\lambda_2\) as Planck's law
 
@@
 \label{stefb1}
-T_{hem,~ b} = \sqrt[4]{\frac{L_z}{\overline{r}\sigma}}
+T_{hem,~ b} = \sqrt[4]{\frac{L_z}{\sigma}}
 \end{equation}
 
-\noindent where $ \sigma $ is the Stefan-Boltzmann constant and $ \overline{r} $ is Planck weighted mean broadband sensor response computed as,
+\noindent where $ \sigma $ is the Stefan-Boltzmann constant.
 
-\begin{equation}
-\label{meansens}
-\overline{r} = \cfrac{\int R(\lambda) ~ r(\lambda)~d\lambda}{\int R(\lambda) ~ d\lambda}
-\end{equation}
-
-\noindent with $ R(\lambda) $ computed from a Planck function at an approximated emitter temperature.
-
-A \textit{directional brightness temperature} T\textsubscript{bright} $(\theta, \phi)$ from some viewing able described $\theta$ and $\phi$ by can be inferred from directional radiance via equation \ref{stefb1} by replacing $ L_z $ with $L_z'$ multiplied by a constant.
+A \textit{directional brightness temperature} T\textsubscript{bright} $(\theta, \phi)$ from some viewing angle described by $\theta$ and $\phi$ can be inferred from directional radiance via equation \ref{stefb1} by replacing $ L_z $ with $L_z'$ multiplied by a constant.
 This method is commonly used to infer T\textsubscript{bright} $(\theta, \phi)$ from infrared thermometers (IRT) operating over the atmospheric window - where atmospheric effects are minimal and T\textsubscript{bright} is a reasonably accurate approximation of T\textsubscript{surf}.
@@
 Since $ L_z $ is unlikely to be equal to $ L_0 $ and, by extension,  $L_z'$ is unlikely to be equal to $L_0'$, T\textsubscript{hem, b} at $ z = 0 $ and T\textsubscript{hem, b} at $ z $ often show significant deviation based on sensor characteristics, ambient conditions, and surface-sensor geometry.
-Hence, T\textsubscript{bright} and T\textsubscript{hem, b} are generally considered only a rough approximation representation of radiometric, kinetic, or thermodynamic T\textsubscript{surf}.
+Hence, T\textsubscript{bright} and T\textsubscript{hem, b} are generally considered only a rough approximation representation of radiometric T\textsubscript{surf}.
 
@@
 However, correction methods are often instrument (or, at the very least, least platform) specific and difficult to generalize across sensor and platform types.
-Few methods exist for correction of ground based remote sensed TIR --- none of which are robust enough to correct irradiances measured from a near-ground wide-\gls{fov} (FOV) radiometer mounted to view rough terrain.
+Few methods exist for correction of ground based remote sensed TIR --- none of which are robust enough to correct irradiances measured from a near-ground wide-field-of-view (FOV) radiometer mounted to view rough terrain.
 
@@
 
-Several multi-LOS correction routines have been developed for to remove atmospheric effects on near-ground remote sensed TIR: \citet{Meier2011} describes a correction method for oblique angled thermal imagery of urban terrain.
+Several multi-LOS correction routines have been developed to remove atmospheric effects on near-ground remote sensed TIR: \citet{Meier2011} describes a correction method for oblique angled thermal imagery of urban terrain.
 In the method, spatially distributed LOS geometries were calculated by linking each image pixel to the corresponding 3-d coordinates of its viewpoint on a digital building model (DBM).
@@
 In this method, path lengths were calculated for six sensor heights.
-Radiances were then modeled using LOWTRAN initialized at $5^{\circ}$ intervals and integrated over the hemisphere to retrieve irradiances for a suite of T\textsubscript{surf}, and ambient T\textsubscript{air} and humidities.
+Radiances were then modeled using the LOWTRAN \citep{Kneizys1988} radiative transfer code initialized at $5^{\circ}$ intervals and integrated over the hemisphere to retrieve irradiances for a suite of T\textsubscript{surf}, and ambient T\textsubscript{air} and humidities.
 The resulting lookup table (LUT) of values is then used to correct $L_z$ and quantify atmospheric effects on remote sensed $L_z$ measured from several sensor heights.
-Methods outlined in Meier et al., 2011 and Kotani \& Sugita, 2009 are limited to narrow-FOV thermal imagers and planar terrain respectively.
-A method which combines Meier et al., 2011's representation of complex surface geometry and Kotani \& Sugita, 2009's broadband hemispherical integration is needed for atmospheric correction of urban TIR measured via downward facing pyrgeometer.
+Methods outlined in \citet{Meier2011} and \citet{Kotani2009a} are limited to narrow-FOV thermal imagers and planar terrain respectively.
+A method which combines \citet{Meier2011}'s representation of complex surface geometry and \citep{Kotani2009a}'s broadband hemispherical integration is needed for atmospheric correction of urban TIR measured via downward facing pyrgeometer.
 
@@
 The site, instrumented as a part of the Basel Urban Boundary Layer Experiment (BUBBLE) \cite{Rotach2005}, has an approximately northeast-southwest orientation and site morphology representative of local climate zone (LCZ) 2\footnote{Site surroundings can be described by the following morphological parameters: mean building height: 14.6m, plan aspect ratio: 0.54, complete aspect ratio: 1.92, local canyon aspect ratio: 1.0, and average shortwave albedo: 0.11 \cite{Rotach2005}.} \cite{Stewart2012}.
-LCZ classification was based on an assessment of surface characteristics in a 250m circular area extending from the Sperrstrasse tower using a 1 \si{m} raster digital building model (DBM).
+LCZ classification was based on an assessment of surface characteristics in a 250 \si{\meter} circular area extending from the Sperrstrasse tower using a 1 \si{m} raster digital building model (DBM).
 Thus, the morphological parameters identified in \cite{Rotach2005} are representative of a majority of the pyrgeometer footprint.
@@
 Upwelling and downwelling short/longwave fluxes were obtained from radiometers mounted at the lowest and highest measurement levels, with an additional downward facing pyrgeometer mounted at roof level near the center of the street canyon.
-In addition, during a summertime intense operation period (IOP) an array of narrow-FOV IRTs was installed to sample representative individual facet surface temperatures (T\textsubscript{facet}).
+In addition, during a summertime intensive observation period (IOP) an array of narrow-FOV IRTs was installed to sample representative individual facet surface temperatures (T\textsubscript{facet}).
 
-The BUBBLE Sperrstrasse site was chosen for two primary reasons: 1.
+The BUBBLE Sperrstrasse site was chosen here for two primary reasons: 1.
 The site provided a long-term climatology of radiation and meteorological variables for a representative mid-latitude city.
@@
 2.
-In inclusion of T\textsubscript{facet} over the IOP allows for investigation of the effect of sensor-surface geometry on remote sensed urban T\textsubscript{surf} for common urban T\textsubscript{surf} measurement strategies.
+Inclusion of T\textsubscript{facet} over the IOP allows for investigation of the effect of sensor-surface geometry on remote sensed urban T\textsubscript{surf} for common urban T\textsubscript{surf} measurement strategies.
 This directional dependence of urban T\textsubscript{surf} - termed "effective anisotropy" - refers to the notion that a TIR signal upwelling from rough terrain viewed from a remote sensor changes based on the combined effects of surface-sensor-sun geometry and the convoluted, 3-dimensional structure of the urban surface \cite{Voogt1998a}.
 Facet surface temperatures measured during the summertime IOP allow for direct climatological comparison of T\textsubscript{hem, r} to common remote sensed representations of the urban surface constructed from weighted averages of wall (T\textsubscript{wall}), road (T\textsubscript{road}), and roof (T\textsubscript{roof}) temperatures.
-Table \ref{weightings} describes T\textsubscript{facet} weightings for nadir remote sensed (T\textsubscript{plan}) and complete (T\textsubscript{comp}) representations of urban surface temperature.
-At each time step, both T\textsubscript{hem, r} and T\textsubscript{plan} were divided by T\textsubscript{comp} and averaged over the IOP to yield normalized mean T\textsubscript{hem, r} and T\textsubscript{plan} at 30-minute intervals.
+Plan and complete aspect ratios are used to derive weightings for nadir remote sensed (T\textsubscript{plan}) and complete (T\textsubscript{comp}) representations of urban surface temperature - described in Table \ref{weightings}.
+At each time step, T\textsubscript{hem, r} and T\textsubscript{plan} were divided by T\textsubscript{comp} and averaged over the IOP to yield normalized mean T\textsubscript{hem, r} and T\textsubscript{plan} at 30-minute intervals.
 Through comparison of T\textsubscript{hem, r} to T\textsubscript{plan} and T\textsubscript{comp} we investigate the effect of sensor-surface geometry on remote sensed urban T\textsubscript{surf} and quantify directional biases in common urban T\textsubscript{surf} measurements.
@@
 With path length geometries calculated in SUM, irradiances are modeled for each time step using version 4.1 of the MODerate resolution atmospheric TRANsmission radiative transfer code (MODTRAN) \citep{Berk1987}.
-At each time-step, a range of potential temperatures is specified (T\textsubscript{spec}) based on T\textsubscript{hem, b} calculated from the measured irradiance with
+At each time-step, a range of possible temperatures is specified (T\textsubscript{spec}) based on T\textsubscript{hem, b} calculated from the measured irradiance with
 
@@
 After the LUT is defined, at-sensor spectral radiances for each path length/zenith angle are modeled at an emissivity of 0.95 over a waveband of 0 - 2300 \si{cm^{-1}} for each T\textsubscript{spec}.
-Profiles of T\textsubscript{air} and humidity are retrieved from 30-minute averages of conditions observed at the Sperrstrasse site.
+Profiles of T\textsubscript{air} and humidity are retrieved from 30 \si{\minute} averages of conditions observed at the Sperrstrasse site.
 Aerosol, trace gas absorber, and above-sensor T\textsubscript{air} and humidity conditions are defined by the mid-latitude summer standard atmosphere when daytime T\textsubscript{air, max} \textgreater 10 \si{\degreeCelsius} (the mid-latitude winter profile is substituted on days where T\textsubscript{air, max} \textless 10 \si{\degreeCelsius}) \citep{Kantor1962}.
@@
 In this method, a "typical" longwave bandpass, approximately 250 - 2300 \si{\centi\meter^{-1}} (4 - 42 \si{\micro\meter}), was extended to include much smaller wavenumber (longer wavelengths).
-This was done for two reasons: 1) To accurately represent spectral longwave curves for typical urban T\textsubscript{surf}, which show significant emittance in wavenumber smaller than 250 \si{cm^{-1}} (wavelengths much longer than 42 \si{\micro\meter}).
+This was done for two reasons: 1) To accurately represent broadband spectral longwave emission curves, which show significant emittance in wavenumber smaller than 250 \si{cm^{-1}} (wavelengths much longer than 42 \si{\micro\meter}).
 2) To replicate the spectral signal "seen" by a silicone-domed pyrgeometer, which continues to transmit radiation at wavenumber \textless 250 \si{cm^{-1}} (wavelengths \textgreater 42 \si{\micro\meter}).
@@
 	\includegraphics[width=\textwidth]{dometrans}
-	\caption{A comparison of dome transmittance for a Kipp \& Zonen silicone domed pyrgeometer and at-sensor radiance (T\textsubscript{surf} = 300 \si{K}, T\textsubscript{air} = 300 \si{K}, $z$ = 30 \si{m})  for "typical" and extended bandpasses.}
+	\caption{A comparison of dome transmittance for a Kipp \& Zonen silicone domed pyrgeometer (data supplied by Kipp \& Zonen \textit{pers.
+comm.}) and at-sensor radiance (T\textsubscript{surf} = 300 \si{K}, T\textsubscript{air} = 300 \si{K}, $z$ = 30 \si{m})  for "typical" and extended bandpasses.}
 	\label{dometrans}
@@
 
-\noindent to yield a directional radiance $ L'_z $ for each $\theta$, with $\overline{r}$ as mean dome transmittance over a waveband of $\nu_1 $= 0 \si{cm^{-1}} to $\nu_2$ = 2300 \si{cm^{-1}}.
+\noindent to yield a directional radiance $ L'_z $ for each $\theta$ over a waveband of $\nu_1 $= 0 \si{\per\centi\meter} to $\nu_2$ = 2300 \si{cm^{-1}}.
+$ \overline{r} $ is Planck weighted mean broadband sensor response computed as,
+
+\begin{equation}
+\label{meansens}
+\overline{r} = \cfrac{\int R(\lambda) ~ r(\lambda)~d\lambda}{\int R(\lambda) ~ d\lambda}
+\end{equation}
+
+\noindent where $ R(\lambda) $ is spectral radiance computed from a Planck function at an approximated emitter temperature and $r (\theta)$ is spectral sensor response.
+
 $ L'_z (\theta) $ are then multiplied by their associated angular view factor weighting ($\Phi$) and integrated over the hemisphere to yield an irradiance $ L_z $ for the target T\textsubscript{hem, r}
@@
 Irradiance - T\textsubscript{hem, r} pairings are then aggregated into a LUT.
-Finally, the measured irradiance from BUBBLE is matched with its closest modeled irradiance to yield an atmospherically corrected, radiometric hemispherical surface temperature for the given time step.
-The workflow is repeated at 30-minute intervals to yield a continuous time series of T\textsubscript{hem, r}.
-Atmospheric correction magnitudes can then be calculated as the difference between T\textsubscript{hem, r} and T\textsubscript{hem, b}, with T\textsubscript{hem, b} calculated via:
+Finally, the measured irradiance is matched with its closest modeled irradiance to yield an atmospherically corrected, radiometric hemispherical surface temperature for the given time step.
+The workflow is repeated at 30 \si{\minute} intervals to yield a continuous time series of T\textsubscript{hem, r}.
+Atmospheric correction magnitudes can then be calculated as the difference between T\textsubscript{hem, r} and T\textsubscript{hem, b}, with T\textsubscript{hem, b} calculated via equation \ref{stefb1}.
 
-\begin{equation}
-	\label{stefboltz2}
-	T_{hem,~ b} = \sqrt[4]{\frac{L_z}{\sigma}}
-\end{equation}.
+\section{Evaluation of the method using profiles of upwelling longwave radiation over a homogeneous flat surface} \label{sensitivity}
 
-\section{Evaluation of the method using profiles of upwelling longwave radiation over a homogeneous planar surface} \label{sensitivity}
-
-MODTRAN has been shown to effectively model near-ground radiative transfer at urban-scale path lengths in 2-dimensions \citep{Hoch2005, Hoch2007}.
+MODTRAN has been shown to effectively model near-ground radiative transfer at urban canyon scale path lengths in 2-dimensions \citep{Hoch2005, Hoch2007}.
 However, the method described in this paper accounts for complex 3-dimensional surface geometry, which can exert significant influence on atmospheric transmittance.
 In addition, the method includes significant post-processing to weight and integrate point-to-point spectral directional radiances to derive 3-dimensional, hemispherical irradiances.
-Thus, prior to deriving a climatology of urban T\textsubscript{hem, r}, the method was evaluated using profiles of upwelling longwave irradiances measured over a simple planar surface.
+Thus, prior to deriving a climatology of urban T\textsubscript{hem, r}, the method was evaluated using profiles of upwelling longwave irradiances measured over a simple flat surface.
 
@@
 To evaluate the method, we used a modified version of the workflow described in Figure \ref{flow}.
-First, azimuthally averaged path lengths over a planar surface were calculated in SUM at 5\si{\degree} increments of zenith angle over the sensor FOV for the 10 \si{\meter} and 30 \si{\meter} sensors.
-For each time step, T\textsubscript{hem, b}, calculated from the lowest $L_z$ measurement, was used to model irradiances at 10 \si{m} and 30 \si{m} at 30-minute intervals using concurrent profiles of T\textsubscript{air} and humidity.
-Daytime T\textsubscript{air} observed at the highest measurement level was anomalously high compared to typical summertime lapse rates above planar vegetated terrain \citep{Oke1987}.
-Thus, the T\textsubscript{air} profile was modified to be more representative of the neutral or weakly unstable daytime conditions and weakly stable nighttime conditions.
-It is likely that these modifications better represent the T\textsubscript{air} profile in the layer between 10 and 30 \si{\meter} than a simple interpolation between the 10 \si{\meter} T\textsubscript{air} and 30 \si{\meter} T\textsubscript{air} measurements.
-Contrary to the findings in \citet{Kotani2009a}, an isothermal atmospheric profile was not sufficient to accurately model upwelling fluxes over planar terrain.
+First, azimuthally averaged path lengths over a flat surface were calculated in SUM at 5\si{\degree} increments of zenith angle over the sensor FOV for the 10 \si{\meter} and 30 \si{\meter} sensors.
+For each time step, T\textsubscript{hem, b}, calculated from the lowest $L_z$ measurement, was used to model irradiances at 10 \si{m} and 30 \si{m} at 30 \si{\minute} intervals using concurrent profiles of T\textsubscript{air} and humidity.
+Daytime T\textsubscript{air} observed at the highest measurement level was anomalously high at all times of day compared to typical summertime lapse rates above flat vegetated terrain \citep{Oke1987}.
+This warm bias in T\textsubscript{air} measured at 30 \si{\meter} is observed even under high wind velocities, where one would expect little vertical difference in temperature not a strong temperature inversion.
+The fact that the observed warming bias remains constant over the 14-day study period prompted a slight modification to ensure the T\textsubscript{air} profile between 10 \si{\meter} and 30 \si{\meter} was representative of the neutral or weakly unstable daytime conditions and weakly stable nighttime conditions.
+It is likely that these modifications better represent the T\textsubscript{air} profile in the layer between 10 and 30 \si{\meter} than a simple interpolation between the 10 \si{\meter} T\textsubscript{air} measurement and the anomalously warm 30 \si{\meter} T\textsubscript{air} measurement.
+Contrary to the findings in \citet{Kotani2009a}, an isothermal atmospheric profile was not sufficient to accurately model upwelling fluxes over flat terrain.
 Although thermal stratification is relatively small by day in an urban canyon \citep{Nakamura1988} - where strong microscale contrasts in T\textsubscript{surf} foster strong canyon mixing and neutral stability - the large daytime T\textsubscript{surf} - T\textsubscript{air} differential and the path-length/transmittance gradient can create large differences in upwelling longwave forced by inter-canyon thermal contrasts - a phenomenon termed radiative divergence.
-By underestimating near-surface air temperatures and overestimating roughness-layer T\textsubscript{air}, positive divergences by day are underestimated by an isothermal profile.
-As such, we suggest using a full canyon T\textsubscript{air} and humidity profile to most accurately model near ground fluxes.
+By underestimating near-surface air temperatures and overestimating roughness-layer T\textsubscript{air}, daytime divergences are underestimated by an isothermal profile.
+As such, a full canyon T\textsubscript{air} and humidity profile is prefered to most accurately model near ground fluxes.
 
 Modeled and measured upwelling longwave at 10 \si{m} and 30 \si{m} fluxes and divergences - calculated as the difference between 30 \si{\meter} and 10 \si{\meter} fluxes - are compared in Figure \ref{fluxes}.
-Both fluxes and divergences show strong correlation, thus we conclude: 1.
-Irradiances measured near screen level (approximately 2 \si{\meter} above ground level) do not include significant atmospheric influence - as T\textsubscript{bright} inferred from $L_{z=2~\si{\meter}}$ was sufficient for modeling irradiances at 10 \si{m} and 30 \si{m}.
+Both fluxes and divergences show strong correlation, thus we conclude: First, that Irradiances measured near screen level (approximately 2 \si{\meter} above ground level) do not include significant atmospheric influence - as T\textsubscript{bright} inferred from $L_{z=2~\si{\meter}}$ was sufficient for modeling irradiances at 10 \si{m} and 30 \si{m}.
 As such, T\textsubscript{bright} and T\textsubscript{hem, r} are reasonably equal when $z$ = 2 \si{m}.
 However, it should be noted that in urban areas a 2 \si{m} sensor is not sufficiently representative of canyon geometry and should not be used to derive urban T\textsubscript{hem, r}.
-2.
-MODTRAN can accurately model longwave fluxes and divergences above a flat, homogeneous surface over a wide range of temperatures, humidities, and cloud coverages and large resulting contrasts in near-ground stability.
-At 30 \si{\meter} under typical summertime profiles of humidity, CO\textsubscript{2}, and O\textsubscript{3}, view factor weighted hemispherical transmittance over a planar surface is approximately 55\% - meaning 45\% of a broadband hemispherical TIR signal at 30 \si{\meter} is emitted by the atmosphere, rather than the target surface.
+Second, MODTRAN can accurately model longwave fluxes and divergences above a flat, homogeneous surface over a wide range of temperatures, humidities, and cloud coverages and large resulting contrasts in near-ground stability.
+At 30 \si{\meter} under typical summertime profiles of humidity, CO\textsubscript{2}, and O\textsubscript{3}, view factor weighted hemispherical transmittance over a flat surface is approximately 55\% - meaning 45\% of a broadband hemispherical TIR signal at 30 \si{\meter} is emitted by the atmosphere, rather than the target surface.
 As such, it is imperative that T\textsubscript{air} profiles are well represented in MODTRAN to accurately model the atmospheric component of remote sensed TIR signal.
@@
 
-Atmospheric correction magnitudes binned for season and time of day are show in Figure \ref{bx_ac}.
+Atmospheric correction magnitudes binned for season and time of day are shown in Figure \ref{bx_ac}.
 Summertime correction magnitudes show significant day-to-day and diurnal variation.
@@
 
-Figure \ref{humtest} shows hemispherical atmospheric transmittance as a function of vapor pressure for pyrgeometers at 10 \si{\meter} and 30 \si{\meter} above a planar surface.
+Figure \ref{humtest} shows hemispherical atmospheric transmittance as a function of vapor pressure for pyrgeometers at 10 \si{\meter} and 30 \si{\meter} above a flat surface.
 Hemispherical transmittances are modeled in MODTRAN 4.1 as the fraction of surface emission reaching the sensor for the path length and view factor geometries calculated in the method evaluation in Section \ref{sensitivity}.
@@
 	keepaspectratio]{season_fit}
-	\caption{Correction magnitude versus $\Delta$T\textsubscript{plan - hem, r} binned first for season and the then for cloud cover in the summer months.}
+	\caption{Correction magnitude versus $\Delta$T\textsubscript{plan - hem, r} binned first for season and then for cloud cover in the summer months.}
 	\label{season_fit}
@@
 	keepaspectratio]{hum_test}
-	\caption{Hemispherical at-sensor atmospheric transmittance as a function of vapor pressure for two pyrgeometer heights (10 \si{\meter} and 30 \si{\meter}) over a planar surface.
+	\caption{Hemispherical at-sensor atmospheric transmittance as a function of vapor pressure for two pyrgeometer heights (10 \si{\meter} and 30 \si{\meter}) over a flat surface.
 Calculated as the fraction of surface emission reaching the sensor height using MODTRAN 4.1 with T\textsubscript{surf} = T\textsubscript{air} = 300 \si{\kelvin} with water vapor as the sole atmospheric absorber.}
@@
 
-Figure \ref{tc_nadhem} shows a comparison of mean normalized T\textsubscript{hem, r} and T\textsubscript{plan} calculated as T\textsubscript{x} divided by T\textsubscript{comp}, where T\textsubscript{x} is T\textsubscript{hem, r} or T\textsubscript{plan}, and averaged over the IOP.
+To illustrate the effect of sensor-surface geometry on remote sensed T\textsubscript{surf}, Figure \ref{tc_nadhem} shows a comparison of mean normalized T\textsubscript{hem, r} and T\textsubscript{plan} calculated as T\textsubscript{x} divided by T\textsubscript{comp}, where T\textsubscript{x} is T\textsubscript{hem, r} or T\textsubscript{plan}, and averaged over the IOP.
+In the figure, overestimation of T\textsubscript{comp} is observed at values above $1$ and underestimation by values below $1$.
 Both T\textsubscript{hem, r} and T\textsubscript{plan} overestimate T\textsubscript{comp} by day and underestimate T\textsubscript{comp} by night.
 Mean daytime overestimation of T\textsubscript{comp} by T\textsubscript{hem, r} is smaller and less variable than that by T\textsubscript{plan}.
-In the figure, overestimation of T\textsubscript{comp} is observed at values above $1$ and underestimation by values below $1$.
 
-Mean normalized T\textsubscript{hem, r}, T\textsubscript{plan}, and $\Delta$T\textsubscript{plan - hem, r} are shown in figure \ref{tc_mean} along with clear sky and overcast sky case days.
+Mean normalized T\textsubscript{hem, r}, T\textsubscript{plan}, and $\Delta$T\textsubscript{plan - hem, r} are shown in Figure \ref{tc_mean2} along with clear sky and overcast sky case days.
 While mean T\textsubscript{hem, r} and T\textsubscript{plan} display similar overestimations of T\textsubscript{comp}, under clear sky conditions, overestimation by T\textsubscript{plan} is much larger than that by T\textsubscript{hem, r}.
@@
 Atmospheric correction magnitudes are highly dependent on factors that control $\Delta$T\textsubscript{surf - air}.
-With a nonzero $\Delta$T\textsubscript{surf - air}, TIR emitted from the surface and absorbed by the atmosphere is re-emitted at a different temperature, modifying the at-sensor TIR signal.
+With a non-zero $\Delta$T\textsubscript{surf - air}, TIR emitted from the surface and absorbed by the atmosphere is re-emitted at a different temperature, modifying the at-sensor TIR signal.
 Thus, the strength of the T\textsubscript{surf} to T\textsubscript{air} differential is directly related to the magnitude of atmospheric effects on a remote sensed TIR signal.
@@
 A pyrgeometer's dome transmittance is spectrally non-uniform.
-Thus, the "dome effect" a pyrgeometer's exerts on a given incoming TIR signal changes as a function of emission temperature (T\textsubscript{emiss}) because the shape of a given spectral radiance curve (R\textsubscript{$\nu$} (T)) is temperature dependent.
+Thus, the "dome effect" a pyrgeometer exerts on a given incoming TIR signal changes as a function of emission temperature (T\textsubscript{emiss}) because the shape of a given spectral radiance curve (R\textsubscript{$\nu$} (T)) is temperature dependent.
 This is illustrated in Figure \ref{nonuniform} by comparing the relative peaks and shape of a pyrgeometer dome transmittance curve for several R\textsubscript{$\nu$} (T) for a range of typical urban T\textsubscript{emiss} (T\textsubscript{emiss} = 260 - 320 \si{\kelvin}, at 5 \si{\kelvin} increments).
@@
 For example, to calculate T\textsubscript{bias} for T\textsubscript{emiss} = 270 \si{\kelvin} at T\textsubscript{calib} = 300 \si{\kelvin}, a spectral shift is applied to R\textsubscript{$\nu$} (300 \si{\kelvin}) so that its peak spectral radiance matches the wavenumber of peak spectral emission for R\textsubscript{$\nu$} (270 \si{\kelvin}).
-When integrated, the two curves have the same radiance, but interact with \textit{r} at over different wavebands.
+When integrated, the two curves have the same radiance, but interact with \textit{r} over different wavebands.
 Following the spectral shift, spectral radiances are convolved by the dome transmittance function, integrated over the waveband and hemisphere, and used to derive T\textsubscript{adj} via equation \ref{stefb1} with $\overline{r}$ at T\textsubscript{calib}.\footnote{In this sensitivity test, the waveband was truncated from 1 - 2500 \si{cm^{-1}} to 80 - 2500 \si{cm^{-1}} to accommodate shorter wavenumber when shifting R\textsubscript{$\nu$} (T\textsubscript{calib}) to lower temperatures where peak spectral emission occurs at smaller wavenumber.
@@
 Additionally, modeled irradiances include a reflected downwelling component, which, although small, is sensitive to T\textsubscript{air} and humidity profiles above the sensor height.
-Because this data was not consistently measured, mid-latitude summer and winter standard atmospheric profiles were supplemented.
+Because these data were not consistently measured, mid-latitude summer and winter standard atmospheric profiles supplemented the measured data.
 The climate of Basel is characterized by mild, humid winters (K\"oppen classification: Cfb).
@@
 However, over/underestimations from geometric effects are highly dependent on synoptic conditions, particularly for a nadir view of the urban surface.
-Overestimation of T\textsubscript{comp} under clear sky conditions is much larger for T\textsubscript{plan} than T\textsubscript{hem, r}.
+T\textsubscript{plan} is much greater than T\textsubscript{comp} under clear sky conditions, with T\textsubscript{hem, r} showing a much smaller positive bias.
 Over half of the T\textsubscript{plan} signal from the Sperrstrasse canyon is generated by rooftops that have both a large diurnal T\textsubscript{surf} amplitude and the highest daytime facet T\textsubscript{max}.
@@
 Non-optimal sensor placement in the Sperrstrasse canyon results in a large underestimation of nighttime T\textsubscript{comp} by T\textsubscript{hem, r}.
-However, nighttime underestimation - and to some extent, daytime overestimation - by T\textsubscript{hem, r} is the combined result of a sampling bias inherent in a hemispherical view of the surface and improper sensor placement, discussed in greater detail in section \ref{Methodological limitations and considerations}.
-Thus, both of these bias can likely be reduced significantly by adjusting sensor placement to best represent surface geometry.
+However, nighttime underestimation --- and to some extent, daytime overestimation --- by T\textsubscript{hem, r} is the combined result of a sampling bias inherent in a hemispherical view of the surface and improper sensor placement, discussed in greater detail in section \ref{Methodological limitations and considerations}.
+Thus, both of these biases can likely be reduced significantly by adjusting sensor placement to best represent surface geometry.
 
@@
 for an urban surface made up of 50\% wall, 25\% road, and 25\% roof, a perfectly geometrically representative remote sensed signal would be one "seen" by a sensor with the same fractional facet view factor proportions.
-For an idealized, orthogonal, block-like urban array, \citet{Roberts2010} found that facet view factor proportions - and thus measured TIR - varied significantly with sensor position and height.
+For an idealized, orthogonal, block-like urban array, \citet{Roberts2010} found that facet view factor proportions --- and thus measured TIR --- varied significantly with sensor position and height.
 To investigate the effect of sensor placement on T\textsubscript{hem, r} for a simplified representation of the Sperrstrasse canyon site, we used SUM to vary sensor position over a simplified DBM and calculated normalized view factor proportions for a 160\si{\degree} FOV sensor.
-For nine sensor positions, show in in Figure \ref{placement}, facet view factor proportions normalized to unity are shown in Figure \ref{bars1} for a sensor height of 2.17 times mean building height and in Figure \ref{bars45} for a sensor height of 3 times mean building height.
+For nine sensor positions, show in Figure \ref{placement}, facet view factor proportions normalized to unity are shown in Figure \ref{bars1} for a sensor height of 2.17 times mean building height and in Figure \ref{bars45} for a sensor height of 3 times mean building height.
 Approximated actual facet surface area proportions for the Sperrstrasse canyon are included for reference.
@@
 
-Results in \citet{Adderley2015} and \citet{Roberts2010} show that sensitivity to sensor placement decreases with increasing sensor height.
+Results in \citet{Adderley2015} and \citet{Roberts2010} show that sensitivity to sensor placement decreases with increasing height.
 Depending on surface geometry, sensor view factor proportions and the TIR signal "seen" by a sensor become spatially invariant sensor at heights above approximately 3 to 5 times mean building height.
 Results from the higher sensor test height support findings in \citet{Adderley2015} and \citet{Roberts2010}, with decreased spatial variance in facet view factor proportions across the nine sensor positions.
-However increased spatial stability in view factor proportions does not necessarily entail increased geometric representativity.
+However increased spatial stability in view factor proportions does not necessarily entail increased geometric representivity.
 For all sensor placements, road view factor proportion is increased and wall view factor is reduced.
@@
 Simulations for path lengths from 1 to 100 \si{\meter} and water vapor mass densities from 0.1 to 23.5 \si{\gram\per\meter\cubed} are included in Appendix \ref{appendixa}.
-Sensor-surface geometry is approximated as a single view factor weighted, azimuthally averaged, surface-to-sensor path length calculated using the surface-sensor-sun urban model \citep{Soux2004} and a simplified DBM.
+Sensor-surface geometry is approximated as a single view factor weighted, azimuthally averaged, surface-to-sensor path length calculated using the SUM \citep{Soux2004} and a simplified DBM.
 This path length represents the average distance from the sensor to the surface as "seen" by the sensor.
@@
 \end{equation}
-\noindent where $\tau$ is calculated using a radiative transfer code and a we and $L_{surf}^{at-sensor}$ is the fraction of ,
+\noindent where $\tau$ is calculated using a radiative transfer code and $L_{surf}^{at-sensor}$ is an estimation of the fraction of at sensor measured irradiance emitted by the surface calculated as,
 
@@
  	\label{parastats}
- 	\begin{tabular}{lcl}
+ 	\begin{tabular}{lcr}
  		\toprule 
@@
 This fact makes clear the need for robust atmospheric correction of T\textsubscript{hem} for sUHI analysis.
-Correction magnitudes are largest on hot, clear sky summertime days - with large solar input and high T\textsubscript{hem, r} and T\textsubscript{air} - and smallest by night and under overcast conditions, during which solar input is suppressed and T\textsubscript{hem, r} and T\textsubscript{air} are low.
+Correction magnitudes are largest on hot, clear sky summertime days --- with large solar input and high T\textsubscript{hem, r} and T\textsubscript{air} --- and smallest by night and under overcast conditions, during which solar input is suppressed and T\textsubscript{hem, r} and T\textsubscript{air} are low.
 Correction magnitudes are most strongly correlated with the T\textsubscript{surf} - T\textsubscript{air} differential and incoming solar radiation and weakly correlated with T\textsubscript{hem, r} and T\textsubscript{air}.

### 2017-05-09 paper2.tex
Add files via upload

@@
 
-\chapter{A climatology of urban surface heat islands derived from hemispherical radiometric surface temperatures}
+\chapter{A climatology of sUHI derived from hemispherical radiometric surface temperatures}
 \label{paper2}
@@
 Integrated up to larger scales, urban areas tend to store more heat relative to non-built surroundings; traffic, HVAC systems, and industrial processes contribute to an additional anthropogenic heat input.
+In addition, a reduced pervious (and vegetative) cover in urban areas shifts convective fluxes towards sensible and away from latent.
 These manifest in elevated T\textsubscript{surf} and T\textsubscript{air} in cities, a phenomenon termed the urban heat island effect (UHI).
@@
 Geometric biases in remote sensing of the urban surface are a result of its 3-dimensional, convoluted structure.
-Compared to flat terrain, complex urban surface geometry modifies receipt of incoming solar radiation and traps a portion of reflected solar and outgoing terrestrial radiation.
+Compared to flat terrain, complex urban surface geometry modifies receipt of incoming solar radiation and traps a portion of reflected solar and outgoing longwave radiation.
 The resulting microscale spatiotemporal contrasts in urban T\textsubscript{surf} create a directional dependence in observed urban T\textsubscript{surf} when measured from conventional narrow-field-of-view (FOV) TIR remote sensing platforms.
@@
 This directional dependence of urban surface temperature is termed "effective thermal anisotropy" \citep{Voogt1998a}.
-Traditional satellite or airborne remote sensing platforms, by viewing the surface in the nadir, sample only a fraction of the complete urban surface, leading to spatiotemporally variant directional biases in remote sensed urban T\textsubscript{surf} of up to 10 \si{\kelvin} \citep{Voogt1995}.
+Traditional satellite or airborne remote sensing platforms, by viewing the surface in the nadir, sample only a fraction of the complete urban surface, leading to spatiotemporally variant directional biases in remote sensed urban T\textsubscript{surf} of up to 10 to 12 \si{\kelvin} \citep{Voogt1995,Lagouarde2012}.
 In general, geometric undersampling by a remote sensor in the nadir manifests as an overestimation of daytime T\textsubscript{surf} and an underestimation of nighttime T\textsubscript{surf} \citep{Adderley2015}.
@@
 At diurnal scales, satellite overpass cycles rarely coincide with sUHI maximums and are not standard across cities or platforms.
-Thus, analysis of sUHI patterns and magnitudes across cities and instrument platforms is difficult.
+Thus, analysis of sUHI patterns and magnitudes across instrument platforms is difficult.
 At shorter time scales still, time-continuous analysis of urban T\textsubscript{surf} shows significant microscale (second to minute) fluctuations in temperature \citep{Christen2012}.
@@
 
-Both geometric and temporal shortcomings limit the representativity of traditional remote sensed evaluations of urban T\textsubscript{surf} and sUHI.
+Both geometric and temporal shortcomings limit the representivity of traditional remote sensed evaluations of urban T\textsubscript{surf} and sUHI.
 The magnitude of these biases has not been extensively studied, particularly from a long term, climatological perspective.
-These biases lead us to ask: What is the nature of the true sUHI?
+These biases lead us to the question: What is the nature of the true sUHI?
 How does it vary temporally and over a full range of weather conditions?
@@
 Thus, urban T\textsubscript{hem, r} in this study are derived using a multiple line-of-sight correction routine which accounts for the 3-dimensionality of the urban surface, changing atmospheric conditions, and spectrally non-uniform sensor response.
-For each time step, the method generates of modeled at-sensor irradiances for a range of potential T\textsubscript{hem, r} using profiles of measured T\textsubscript{air} and humidity.
+For each time step, the method generates a LUT of modeled at-sensor irradiances for a range of potential T\textsubscript{hem, r} using profiles of measured T\textsubscript{air} and humidity.
 A corrected T\textsubscript{hem, r} for the target time step is then retrieved by matching the measured irradiance value to the closed modeled irradiance - T\textsubscript{hem, r} pairing.
@@
 First, surface-to-sensor path length and view factor geometries are calculated with the surface-sensor-sun urban model (SUM) \citep{Soux2004}.
-Using a digital building model, SUM calculates path lengths from the surface to the sensor and angular view factors by projecting the pyrgeometer FOV onto a simplified representation of the surface, determining which points in the DBM are 'seen' by the sensor, and calculating the distance from each 'seen' point to the sensor.
+Using a digital building model (DBM), SUM calculates path lengths from the surface to the sensor and angular view factors by projecting the pyrgeometer FOV onto a simplified representation of the surface, determining which points in the DBM are 'seen' by the sensor, and calculating the distance from each 'seen' point to the sensor.
 Path lengths are binned at 5\si{\degree} intervals and averaged over the azimuth angle.
@@
 
-Although MODTRAN has been show to accurately model radiative transfer of broadband longwave radiation at urban scale path lengths \citep{Hoch2005, Hoch2007}, the added complexity inherent in simulating fluxes of radiation hemispherically over complex urban terrain prompted an independent evaluation of the correction method using observations of upwelling longwave radiation (measured from 2 \si{\meter}, 10 \si{\meter}, and 30 \si{\meter} above ground level) over a planar surface for a continuous 16-day period.
+Although MODTRAN has been show to accurately model radiative transfer of broadband longwave radiation at urban canyon scale path lengths \citep{Hoch2005, Hoch2007}, the added complexity inherent in simulating fluxes of radiation hemispherically over complex urban terrain prompted an independent evaluation of the correction method using observations of upwelling longwave radiation (measured from 2 \si{\meter}, 10 \si{\meter}, and 30 \si{\meter} above ground level) over a simple, flat surface for a continuous 16-day period.
 In the evaluation, a modified version of the method uses T\textsubscript{hem, b} inferred from the lowest L\textsubscript{up} measurement along with profiles of T\textsubscript{air} and humidity to model irradiances at 10 \si{\meter} and 30 \si{\meter} at each time step.
@@
 Using the view factor weighted path length, a suite of radiative transfer simulations is run prior to correction to retrieve "hemispherical" broadband atmospheric transmittances for a range of potential humidities.
-Humidity-transmittance pairings are aggregated into a lookup table (LUT), where the measured humidity at each time step can be matched with the associated hemispherical transmittance.
+Humidity-transmittance pairings are aggregated into a look-up table (LUT), where the measured humidity at each time step can be matched with the associated hemispherical transmittance.
 Using the hemispherical transmittance as the fraction of surface emitted L\textsubscript{up} reaching the sensor, measured L\textsubscript{up} can then be decomposed into its surface and atmospheric emitted components, from which a parameterized T\textsubscript{hem, r} (T\textsubscript{hem, para}) is inferred.
@@
 
-This study uses data from two long-term climatological campaigns: the Basel Urban Boundary Layer Experiment (BUBBLE) conducted over December 2001 through July 2002 in Basel, Switzerland \citep{Rotach2005}, and the Environmental Prediction in Canadian Cities (EPiCC) (CITE?) campaign in Vancouver, Canada.
+This study uses data from two long-term climatological campaigns: the Basel Urban Boundary Layer Experiment (BUBBLE) conducted over December 2001 through July 2002 in Basel, Switzerland \citep{Rotach2005}, and the Environmental Prediction in Canadian Cities (EPiCC) \citep{Voogt2007} campaign in Vancouver, Canada.
 Using the method described in section \ref{method2}, an eight month climatology of T\textsubscript{hem, r} are retrieved from radiation and meteorological data collected as a part of the BUBBLE campaign.
@@
 To supplement data from the BUBBLE campaign, radiation and meteorological variables observed as a part of the EPiCC project are used to derive a climatology of T\textsubscript{hem, r} for the year long period spanning October 2008 through September 2009 using the parameterization scheme.
+EPiCC urban and rural data were collected from an urban/suburban site in central Vancouver ("Sunset Tower") and from a rural site approximately 25 \si{\kilo\meter} to the SSW ("Westham Island").
 These data provide a second, long term, time continuous climatology of sUHI for comparison of seasonal and diurnal patterns, to understand the influence of urban and rural site selection and characteristics on sUHI, and to aid in developing a generalized model for sUHI development over multiple time scales.
@@
 
-sUHI and canopy later UHI (clUHI) magnitudes were calculated from T\textsubscript{hem, r} and T\textsubscript{air} (measured from 2 \si{\meter} above ground level) as $\Delta$T\textsubscript{(urban-rural)} at 30 \si{\min} intervals.
-Rural T\textsubscript{hem, b} was calculated via
+sUHI and clUHI magnitudes were calculated from T\textsubscript{hem, r} and T\textsubscript{air} (measured from 2 \si{\meter} above ground level) as $\Delta$T\textsubscript{(urban-rural)} at 30 \si{\min} intervals.
+Rural T\textsubscript{hem, b} was calculated an expanded version of equation \ref{stefb1} to remove emissivity effects,
 
@@
 To investigate the influence of rural site characteristics on Basel's sUHI, sUHI magnitudes are calculated using T\textsubscript{hem, b} inferred for a different rural site ("Village Neuf") over the summertime IOP.
-Observations at Village Neuf were intermittent and often discontinous.
-As such, a subset of nine partially-complete days was selected to represent T\textsubscript{hem, b} and sUHI over a range of cloud coverages, to investigate the influence of rural site characteristics and T\textsubscript{surf} regimes on sUHI.
-As sUHI develop, preference was given to days with continuois observations in early afternoon through night.
-For each of the nine days sUHI\textsubscript{max} was used to infer a normalized sUHI magnitude for both Lange Erlen and Village Neuf sites at 30- \si{\minute} intervals, to remove day-to and compare patterns of sUHI development for the two sites.
+Observations at Village Neuf were intermittent and often discontinuous As such, a subset of nine partially-complete days was selected to represent T\textsubscript{hem, b} and sUHI over a range of cloud coverages, to investigate the influence of rural site characteristics and T\textsubscript{surf} regimes on sUHI.
+As sUHI develop, preference was given to days with continuous observations from early afternoon through the night.
+For each of the nine days sUHI\textsubscript{max} was used to infer a normalized sUHI magnitude for both Lange Erlen and Village Neuf sites at 30- \si{\minute} intervals to compare patterns of sUHI development for the two sites.
 Mean normalized sUHI was then calculated at each time step over the nine day period.
@@
 
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=15cm,height=7in,keepaspectratio]{bxsUHI_season_vanc}
+	\caption{Hemispherical sUHI magnitudes for the year-long EPiCC climatology (top) and binned for winter (middle) and summer (bottom) seasons.
+Grey shading indicates nighttime hours averaged over the bin interval.}
+	\label{bxsUHI_season_vanc}
+\end{figure}
+
 \subsection{The effect of sensor-surface geometry on sUHI}
@@
 Compared to sUHI viewed from a complete representation of the urban surface (sUHI\textsubscript{comp}); nadir (sUHI\textsubscript{nadir}) and hemispherical (sUHI\textsubscript{hem}) views of the surface overestimate sUHI\textsubscript{comp} by day and underestimate sUHI\textsubscript{comp} by night.
-sUHI\textsubscript{nadir} from a nadir view shows the greatest diurnal range of magnitudes, particularly under clear sky 'satellite friendly' conditions.
+sUHI\textsubscript{nadir} shows the greatest diurnal range of magnitudes, particularly under clear sky 'satellite friendly' conditions.
 
@@
 	\caption{A comparison of sUHI magnitudes from complete (sUHI\textsubscript{comp}), hemispherical (sUHI\textsubscript{hem}), and nadir (sUHI\textsubscript{nadir}) remote sensed representations of the Sperrstrasse canyon over the IOP.
-Each plot includes mean sUHI, as well case days representing sUHI under clear sky and clouded sky conditions.}
+Each plot includes mean sUHI, as well as case days representing sUHI under clear sky and clouded sky conditions.}
 	\label{bx_suhi_compare}
@@
 In contrast, shading by urban features causes urban T\textsubscript{surf} to respond more slowly, resulting in a small sUHI\textsubscript{min} in the hours just after sunrise.
-This is observed in Basel across all representations of urban T\textsubscript{surf} at similar magnitudes.
-Through the late morning hours, urban heating rates increase relative to rural rates as the solar angle increases and low thermal admittance facets (particularly rooftops) experience more direct insolation and quickly heat up.
+This is observed in Basel across all representations of urban T\textsubscript{surf}.
+Through the late morning hours, urban heating rates increase relative to rural rates as the solar zenith angle decreases and low thermal admittance facets (particularly rooftops) experience more direct insolation and quickly heat up.
 Increases in urban T\textsubscript{surf} are amplified by the abundance of impervious surfaces, which result in a smaller fraction of insolation partitioned to latent heat relative to vegetated non urban spaces.
 After the initial spike in rural heating in the early morning hours, urban heating rates overtake rural approximately three to four hours after sunrise and the sUHI slowly grows through midday to a peak at approximately two to three hours hours after solar noon.
-In the hours after UHI\textsubscript{max}, the solar angle decreases and both urban and rural heating rates decline and begin to converge, as shading by canyon geometry causes some facets to begin to cool.
+In the hours after sUHI\textsubscript{max}, the solar zenith angle increases and both urban and rural heating rates decline and begin to converge, as shading by canyon geometry causes some facets to begin to cool.
 
-In the afternoon hours, both urban and rural surfaces transition to cooling, this can lead to several different patterns in evening sUHI development and is strongly controlled by urban surface structure.
+In the afternoon hours, both urban and rural surfaces transition to cooling; this can lead to several different patterns in evening sUHI development and is strongly controlled by urban surface structure.
 Relative to non-urban spaces, cooling rates for wall and road facets lag slightly as a result of canyon radiation trapping, this is offset by increased cooling rates for low thermal admittance rooftop facets.
@@
 A nadir view, by virtue of oversampling rapidly cooling rooftops, observes a strong evening decline in sUHI.
-This is roughly equivalent to sUHI calculated from an urban site with low canyon height to width ratio and a relatively unobstructed sky view factor.
 In contrast, by sampling vertical facets that cool more gradually, this decline is less evident in sUHI calculated from a complete view of Basel T\textsubscript{surf}.
@@
 However, variations in rural thermal admittance, surface geometry, land cover and use, and/or influence from surrounding terrain can result in significant contrasts in rural cooling rates.
-All of these factors show strong spatiotemporal variation, resulting in large differences in sUHI development depending on rural site characteristics.
+All of these factors show strong spatiotemporal variation, resulting in large differences in sUHI development depending on rural site characteristics, similar to observations of variations in clUHI based on rural siting in \citep{Runnalls2000}.
 This is illustrated by two different patterns of post-sunset sUHI development observed for the two rural sites in Basel.
-Although similar in terms of surface coverage and site geometry, hilly terrain surrounding the Lange Erlen site results in cold air drainage and an increased rural cooling rate, particularly following hot, clear sky days.
+Although similar in terms of surface coverage and site geometry, hilly terrain surrounding the Lange Erlen site results in suspected cold air drainage and an increased rural cooling rate, particularly following hot, clear sky days.
 This effect is not present at the Village Neuf site, where nocturnal cooling rates are likely driven primarily driven by radiative losses.
@@
 
-The sUHI and the clUHI have similar micro meteorological genesis, climatic controls, and display similar seasonal variability, albeit concentrated at different times of the day.
+The sUHI and the clUHI have similar micro-meteorological genesis, climatic controls, and display similar seasonal variability, albeit concentrated at different times of the day.
 This is evident when comparing normalized mean monthly hemispherical sUHI and clUHI - where the range of both types of sUHI is restricted to values between zero and one.
 In winter, low solar input and frequent cloud cover inhibit variability in both sUHI and clUHI and both are of similar normalized magnitudes.
-In summer, larger solar input fosters contrasts in urban-rural heating and cooling rates and both clUHI and sUHI are most variable, with peak sUHI in the early afternoon clUHI is smallest, and peak clUHI in the hours after sunset when sUHI is relatively low.
-These results suggest the following generalization: when conditions that foster the formation of strong urban-rural contrasts in T\textsubscript{air} and T\textsubscript{surf} are present (most notably a large solar input), sUHI peaks approximately two hours after solar noon and clUHI peaks approximately two hours after sunset.
+In summer, larger solar input fosters contrasts in urban-rural heating and cooling rates and both clUHI and sUHI are most variable, with peak sUHI in the early afternoon when clUHI is smallest, and peak clUHI in the hours after sunset when sUHI is relatively low.
+These results suggest the following generalization: when conditions that foster the formation of strong urban-rural contrasts in T\textsubscript{air} and T\textsubscript{surf} are present (most notably a large solar input), sUHI peaks approximately two hours after solar noon and clUHI peaks approximately three to four hours after sunset.
 Thus, a surface urban heat island is not simply a stronger, more pronounced, canopy layer urban heat island.
@@
 Sensor-surface geometry exerts significant influence on observed sUHI magnitudes.
-2-dimensional, complete, and hemispherical observations of sUHI not only yield different magnitudes but also different diurnal sUHI regimes and peak timings.
+Two-dimensional, complete, and hemispherical observations of sUHI not only yield different magnitudes but also different diurnal sUHI regimes and peak timings.
 Variations in observed sUHI across different sensor sampling regimes are largest under clear sky, hot conditions, coinciding with typical seasonal and diurnal sUHI\textsubscript{max}.
@@
 
+Results of sensor placement sensitivity testing in Section \ref{sensitivity} show that T\textsubscript{hem, r} is dependent on sensor position relative to canyon features.
+These effects are translated into sUHI\textsubscript{hem}.
+In general, when increasing sUHI magnitudes.
+
 \subsection{Meteorological controls on sUHI}
@@
 The relationship between solar radiation and sUHI is relatively simple but easily misinterpreted.
-Surfaces (both urban and otherwise) change in temperature based on accumulated radiative or convective gains or losses - as such, instantaneous changes in insolation have little effect on sUHI magnitudes.
+Surfaces (both urban and otherwise) change in temperature based on \textit{accumulated} radiative or convective gains or losses - as such, instantaneous changes in insolation have little effect on sUHI magnitudes.
 Thus, to say that sUHI is a function of measured incoming solar radiation (in \si{\watt\per\square\meter}) does not accurately characterize the relationship and produces a significant amount of unrepresentative scatter.
@@
 Under cloudy conditions, reduced accumulated solar heating of the surface suppresses the development of urban-rural thermal contrasts, resulting in reduced sUHI magnitudes.
-Owing to its dependence on accumulated solar heating, $\Delta$T\textsubscript{hem, r - air} has a similar positive effect on sUHI magnitudes.
 
@@
 Winds affect surface heating and cooling rates by facilitating convective cooling (or, under less common circumstances, convective heating).
-In general, as wind velocity increases, convective heat losses increase and T\textsubscript{surf} decreases.
+In general, as wind velocity increases, convective heat losses increase and both T\textsubscript{surf} and its spatial variability decreases.
 This suppresses microscale spatiotemporal variations in T\textsubscript{surf} as hot sunlit surfaces experience greater rates of convective cooling under high wind conditions than cooler shaded surfaces.

### 2017-05-09 introduction.tex
Add files via upload

@@
 A change in T\textsubscript{surf} is translated to change in T\textsubscript{air} as the surface exchanges energy with the atmosphere above it, primarily via convection and radiation.
-As such, T\textsubscript{surf} and T\textsubscript{are} are inextricably linked, particularly near to the ground, and a city's altered surface characteristics and energy balance manifest in modifications to both surface and air temperature regimes.
+As such, T\textsubscript{surf} and T\textsubscript{air} are inextricably linked, particularly near to the ground, and a city's altered surface characteristics and energy balance manifest in modifications to both surface and air temperature regimes.
 
@@
 In any given observation of climate, myriad urban and non urban signals make it difficult to reliably differentiate urban influenced measures from those free of urban influence, degrading interstudy comparison.
-Even in the modeling domain, where experimental control is possible, the inherent complexity in simulating flows of energy, momentum, and mass at the scales necessary to represent physical phenomena over realistic urban environments is, as of yet, not feasible.
-Notwithstanding implementation of accurate, physically based schemes to represent urban climates in regional or global climate models.
+Development of physically based numerical models to represent physical processes affected by urbanization, and which permit experimental control, have shown promise \citep{Krayenhoff2014,Gastellu-Etchegorry1996,Masson2000,Martilli2002} yet the inherent complexity in simulating flows of energy, momentum and mass at the scales necessary to represent the full range of physical processes over realistic environments remains an on-going challenge.
 Thus, accurate assessment of how a city affects the climate is a deceptively challenging task.
@@
 
-Study of the UHI with respect to T\textsubscript{air} (aUHI) and T\textsubscript{surf} (sUHI) tacitly falls into such a framework, by replacing $V_{M, \ U} - V_{M, \ R}$ with $\Delta$T\textsubscript{U - R}.
+Study of the UHI with respect to T\textsubscript{air} (aUHI) - typically referring to T\textsubscript{air} measured from approximately 2 \si{\meter} above ground level, yielding a "canopy layer" UHI (sUHI) - and T\textsubscript{surf} (sUHI) tacitly falls into such a framework, by replacing $V_{M, \ U} - V_{M, \ R}$ with $\Delta$T\textsubscript{U - R}.
 However, in spite of its conceptual simplicity through the lens of the Lowry method, meta-analysis of observational aUHI studies suggests that that simplicity may not translate well to real world analysis.
@@
 
-In contrast to urban T\textsubscript{air} and aUHI assessment, where measurement techniques have not undergone significant evolution in recent memory and methodological problems arise in large part from conceptual shortcomings and simple carelessness, study of the urban effect on land T\textsubscript{surf} is relatively new and significantly more complex.
+In contrast to urban T\textsubscript{air} and aUHI assessment, where measurement techniques have not undergone significant recent evolution and methodological problems arise in large part from conceptual shortcomings and simple carelessness, study of the urban effect on land T\textsubscript{surf} is relatively new and significantly more complex.
 As such, methodological problems in sUHI study can arise from both conceptual flaws \textit{and} instrument biases.
-In spite of these difficulties, the introduction of new methods, instruments, and analytical techniques in TIR remote sensing of urban environments has lead to rapid proliferation in study of urban T\textsubscript{surf} and the sUHI - much of which is focused on remote sensing of upwelling TIR.
+In spite of these difficulties, the introduction of new methods, instruments, and analytical techniques in TIR remote sensing of urban environments has lead to rapid proliferation in study of urban T\textsubscript{surf} and the sUHI - much of which is focused on remote sensing of upwelling thermal infrared radiation (TIR).
 In the decades preceding this thesis, a combination of factors has lead to an increased focus on remote sensed study of urban T\textsubscript{surf}:
@@
 \begin{itemize}
-	\item The proliferation of satellite and aerial thermal infrared (TIR) remote sensors and "openness" in data access policies.
+	\item The proliferation of satellite and aerial TIR remote sensors and "openness" in data access policies.
 	\item Improvements in sensor spatial, spectral, and radiometric resolutions. 
 	\item An increased focus on the importance of the surface in both determining and understanding key near-ground micro-meteorological phenomena.
+	\item Interest in T\textsubscript{surf} for human thermal comfort, energy conservation, and sUHI mitigation applications. 
 \end{itemize}
@@
 Traditional methods for urban T\textsubscript{surf} measurement are subject to a suite of geometric and temporal biases.
-Geometric biases are the result of urban modification of surface structure and thermal and radiative properties combined with the 2-dimensional, narrow field-of-view (FOV) viewing geometry of conventional remote sensors:
+Geometric biases are the result of urban modification of surface structure and thermal and radiative properties combined with the  narrow field-of-view (FOV) viewing geometry of conventional remote sensors:
 
@@
 
-\noindent---\textit{Undersampling of surface 3-dimensionality by narrow-FOV remote sensors}.
+\subsection*{\textnormal{\textit{Undersampling of surface 3-dimensionality by narrow-FOV remote sensors}}}
 
 Relative to typical non-urban and rural terrain, urban development adds significant 3-dimensionality to the surface.
-This added complexity results in undersampling of the surface by narrow-FOV remote sensors.
-A narrow-FOV sensor, by sampling surface 3-dimensionality in 2-dimensions, "sees" a different assemblage of urban surfaces depending on its viewing direction, resulting in a geometric bias towards some arrangement of facets; i.e.
+This added complexity results in undersampling of the complete urban surface by narrow-FOV remote sensors.
+A narrow-FOV sensor "sees" a different assemblage of urban surfaces depending on its viewing direction, resulting in a geometric bias towards some arrangement of facets; i.e.
 viewing from the nadir, a narrow-FOV sensor sees the urban surface as an array of rooftops and roads, the same sensor viewing the surface from an oblique angle sees some array of walls, rooftops, and roads depending on its orientation.
 For a sensor in the nadir, increasing FOV introduces sampling of vertical features.
-To illustrate how sensor-surface geometry influences sampling of the urban surface by narrow- and wide-FOV sensors, Figure \ref{model} shows FOVs for the three sensor orientation/type combinations.The vast majority of urban thermal remote sensing is done via a narrow-FOV sensor in the nadir.
+To illustrate how sensor-surface geometry influences sampling of the urban surface by narrow- and wide-FOV sensors, Figure \ref{model} shows FOVs for the three sensor orientation/type combinations.The vast majority of urban thermal remote sensing is done via a narrow-FOV sensor viewing from the nadir.
 
@@
 
-\noindent---\textit{Directional dependence in remote sensed urban T\textsubscript{surf}}.
+\subsection*{\textnormal{\textit{Directional dependence in remote sensed urban T\textsubscript{surf}}}}
 
-In addition to modifying how a remote sensor samples the surface, the convoluted 3-dimensional structure of the urban surface modifies sunlit/shading regimes, resulting in strong micro-scale spatiotemporal contrasts in urban T\textsubscript{surf}, which depend on surface-sun geometry.
+In addition to modifying how a remote sensor samples the surface, the convoluted 3-dimensional structure of the urban surface modifies the surface radiation budget, resulting in strong micro-scale spatiotemporal contrasts in urban T\textsubscript{surf}, which depend on surface-sun geometry.
 These microscale variations in urban T\textsubscript{surf} are often amplified by significant geometric and spatial contrasts in the thermal admittance of common urban materials, which determines the diurnal amplitude of T\textsubscript{surf} for a given facet.
@@
 Temporal biases in thermal remote sensing of urban areas occur across multiple time scales including: 
-
-\noindent---\textit{ Contamination by turbulence forced, high frequency fluctuations in urban T\textsubscript{surf}}.
+\subsection*{\textnormal{\textit{Contamination by turbulence forced, high frequency fluctuations in urban T\textsubscript{surf}}}}
 
@@
 rooftops, walls, roads, and vegetation) display large micro-scale (second to minute) fluctuations in T\textsubscript{surf}.
-The magnitude of these is inversely related to surface thermal admittance and is, as of writing, poorly understood.
+The magnitude of these is inversely related to surface thermal admittance and is currently, poorly understood.
 Most thermal remote sensors provide instantaneous, rather than temporally averaged, T\textsubscript{surf} and thus are subject to contamination by high-frequency fluctuations.
-These biases are difficult to estimate in urban environments, where a large variety of fabric materials can produce significant geometric and spatial contrasts in thermal admittance - and thus directional and spatial variations in the magnitude of microscale fluctuations depending on the facet material types viewed by the sensor.
+These biases are difficult to estimate in urban environments, where a large variety of fabric materials can produce significant geometric and spatial contrasts in thermal admittance --- and thus directional and spatial variations in the magnitude of microscale fluctuations depending on the facet material types viewed by the sensor.
 
-\noindent---\textit{Discontinuity in satellite overpass cycles.}
+\subsection*{\textnormal{\textit{Discontinuity in satellite overpass cycles.}}}
 
@@
 Assessment of land T\textsubscript{surf} either sacrifices spatial resolution for a daily repeat cycle - as is the case with MODIS  - or sacrifices temporal resolution for a higher spatial resolution - ASTER or Landsat.
+These satellites cannot be used to assess the temporal development of sUHI without significant interpolation.
+Geostationary satellite sensors overcome this limitation but low spatial resolutions limit applicability in urban areas, especially at mid to high latitudes.
 
-\noindent---\textit{Clear-sky bias.}
+\subsection*{\textnormal{\textit{Clear-sky bias}}}
 
@@
 As is the case with geometric biases, the magnitude of these temporal biases in remote sensed urban T\textsubscript{surf} is difficult to quantify without long-term ground truthing campaigns or significant interpolation.
-As such, the true temporal and geometric nature of urban effect of T\textsubscript{surf} is presently unknown.
+As such, the true temporal and geometric nature of urban effect on T\textsubscript{surf} and thus sUHI is presently unknown.
 
@@
 
-\noindent In an attempt to answer these questions, I did the following:
+\noindent In an attempt to answer these questions, this thesis
 
 \begin{itemize}
-	\item Developed and evaluated a method to retrieve atmospherically correction hemispherical radiometric urban surface temperatures from time-continuous measurements of upwelling longwave radiation.
-	\item Compared urban surface temperatures and surface urban heat island magnitudes retrieved using the method to common remote sensed representations of the urban surface.
-	\item Derived an eight month climatology of hemispherical urban surface temperatures to observe seasonal and diurnal patterns of the surface urban heat island effect.
+	\item Develops and evaluates a method to retrieve atmospherically correction hemispherical radiometric urban surface temperatures from time-continuous measurements of upwelling longwave radiation.
+	\item Compares urban surface temperatures and surface urban heat island magnitudes retrieved using the method to common remote sensed representations of the urban surface.
+	\item Derives two long term climatology of hemispherical urban surface temperatures to observe seasonal and diurnal patterns of the surface urban heat island effect.
 \end{itemize}

### 2017-05-10 conclusion.tex
Add files via upload

--- 2017-05-03 conclusion.tex
@@
 
-\chapter{Concluding statements}
+\chapter{Conclusion}
 
-This study presents the first long-term, time-continuous analysis of urban T\textsubscript{surf} to contribute to a more complete understanding of the temporal and geometric nature of the urban effect on land T\textsubscript{surf} and sUHI.
+This study presents the first long-term, time-continuous analysis of urban T\textsubscript{surf} and sUHI that incorporates representative sampling of the three-dimensional urban surface.
+It is based on hemispheric upwelling longwave radiation measurements that are commonly made as a part of urban energy balance assessment and constitute a hitherto untapped resource for urban T\textsubscript{surf} and sUHI analysis.
+The analysis contributes to a more complete understanding of the temporal and geometric nature of the urban effect on land T\textsubscript{surf} and sUHI.
 
-Chapter \ref{paper1} details a hemispherical atmospheric correction method that uses version 4.1 of the MODTRAN radiative transfer code \citep{Berk1987} in conjunction with the Surface-Sensor-Sun Urban Model \citep{Soux2004} to derive atmospherically correct hemispherical radiometric surface temperatures from upwelling longwave radiation measured from a downward facing pyrgeometer.
-The correction method was developed to be readily adaptable to other sensor types and study sites, these measurements are commonly made as a part of urban energy balance assessment.
-By accounting for surface geometry and spectrally non-uniform sensor response, the correction method is readily adaptable to other sensor types and study sites.
-Atmospheric effects on inferred T\textsubscript{hem} are significant relative to typical sUHI magnitudes and show significant diurnal and seasonal variation.
-Correction magnitudes are particularly sensitive to $\Delta$T\textsubscript{surf-air}.
-Although water vapor is the main absorber of broadband TIR radiation, it does not exert significant influence on correction magnitudes, as atmospheric transmittance shows little variation at typical humidities.
+\section{Summary of results}
+
+Chapter \ref{paper1} details a hemispherical atmospheric correction method that uses version 4.1 of the MODTRAN radiative transfer code \citep{Berk1987} in conjunction with the Surface-Sensor-Sun Urban Model \citep{Soux2004} to derive hemispherical radiometric surface temperatures from longwave radiation upwelling from the urban surface measured from a downward facing pyrgeometer.
+The correction method is necessary for two reasons:
+
+\begin{enumerate}
+	\item To retrieve T\textsubscript{surf} that is representative of the temperature at which the surface is radiating (known as radiometric T\textsubscript{surf}) by removing effects from absorption and emission by the atmosphere in between the surface and the sensor.
+	\item To facilitate intersite analysis of T\textsubscript{surf}, where different ambient conditions, surface-sensor-sun geometries, and instrument types can yield differential atmospheric effects on the remote sensed TIR signal.
+\end{enumerate}
+
+ The correction method was developed to be readily adaptable to other sensor types and study sites by accounting for complex three-dimensional surface geometry and non uniform spectral sensor response.
+
+Analysis of an eight months of corrected T\textsubscript{hem, r} shows that atmospheric effects on inferred T\textsubscript{hem} are large relative to typical sUHI magnitudes and show significant diurnal and seasonal variation.
+Relationships between correction magnitude and meteorological variables are complex.
+Correction magnitudes show strong positive relationships with $\Delta$T\textsubscript{surf-air} and incoming solar radiation.
+Conversely, although water vapor is the main absorber of broadband TIR radiation, it does not exert significant influence on correction magnitudes, as atmospheric transmittance shows little variation at typical humidities.
+To facilitate application of the method, a parameterization was developed incorporating a small number of generalizable radiative transfer simulations (included in Appendix \ref{appendixa}) and information about site geometry to retrieve radiometric hemispherical T\textsubscript{surf} en masse using measurements of upwelling longwave, humidity, and T\textsubscript{air}.
 
@@
 Daytime overestimation of T\textsubscript{comp} by T\textsubscript{hem} is more consistent under a wide variety of synoptic conditions, whereas overestimation by T\textsubscript{plan} fluctuates significantly, resulting in poor performance under clear-sky "satellite friendly" conditions.
-As make up the bulk of the satellite urban T\textsubscript{surf} record, our understanding of urban T\textsubscript{surf} likely includes a significant warming bias.
-These biases are reflected in sUHI analysis in \ref{paper2}.
+As these measures make up the bulk of the satellite urban T\textsubscript{surf} record, our understanding of urban T\textsubscript{surf} likely includes a significant warm bias.
+These biases are reflected in sUHI analysis in Chapter \ref{paper2}.
 
-Analysis of an eight month, time continuous climatology of hemispherical sUHI is included in Chapter \ref{paper2}.
+Chapter \ref{paper2} includes an analysis of two long term, time continuous climatologies of hemispherical sUHI, one in Basel, Switzerland and another in Vancouver, Canada.
+sUHI for both study sites shows significant diurnal, day-to-day, and seasonal variability.
+sUHI is highly dependent on meteorological conditions, particularly accumulated solar radiation over a day and wind velocity.
+However, as sUHI is the result of differences in heating and cooling rates between the city and its non urban surroundings, its relationship to meteorological controls is complex.
+Analysis of sUHI development under a range of mid-latitude synoptic conditions lead to the development of the first observationally supported generalizable model for summertime sUHI development presented in section \ref{gen}.
+
+Analysis of the directional dependence of sUHI is also included in Chapter \ref{paper2}.
+Results are similar to those found in Chapter \ref{paper1}, with significant daytime overestimation of sUHI\textsubscript{comp} by sUHI{plan}, particularly during hot clear sky conditions.
+Not only did directional biases in remote sensed sUHI change observed magnitudes, these biases also changed the diurnal character of sUHI, modifying patterns of sUHI in the hours during near sunset.
+Thus, satellite based assessments of the diurnal pattern of sUHI may lead to misrepresentations of its 'true' diurnal character.
+
+\section{Limitations and future work}
+
+%The convoluted and complex nature of the urban surface makes remote sensing urban T\textsubscript{surf} difficult.
+Although the method detailed in this study provides T\textsubscript{surf} that is time continuous and more representative of the complete urban T\textsubscript{surf} than conventional remote sensing methods - particularly under clear sky conditions - it is not without some limitations.
+
+The complex nature of the urban surface introduces myriad shortcomings and biases into conventional remote sensing of urban thermal climates - this method is no exception.
+Sensor placement sensitivity tests in Section \ref{placement} corroborate findings in \citet{Roberts2010,Adderley2015} and indicate that longwave radiation upwelling from the urban surface as measured from a downward facing radiometer is spatially variant when measured from heights below approximately three to five times mean building height.
+Most pyrgeometers (including those used in this study) are mounted below this height and are subject to a bias towards facets that make up a disproportionately large fraction of the sensor FOV.
+For sensors mounted at less than optimal heights, the effect of this bias on derived T\textsubscript{hem, r} is largely dependent on sensor placement relative to rooftop facets, with better performance further from positions over rooftops.
+Pyrgeometers at the Sperrstrasse and Sunset tower sites are placed far enough from rooftops to render this bias relatively small.
+However, future studies should strive for pyrgeometer heights that are more optimal not only to yield more geometrically representative T\textsubscript{hem, r}, but also to ensure irradiances are representative of the complete urban longwave emission in energy balance assessments.
+
+As the method is fixed and ground based, it is not practical for large scale spatially extensive analysis of sUHI.
+This limits the method to mid- and low-rise neighborhoods.
+However, the concept can be adapted to retrieve spatiotemporally continuous urban T\textsubscript{hem, r} from mobile measurements of upwelling longwave irradiances.
+\citet{Sugawara2006} used near-ground, helicopter-mounted, pyrgeometers and narrow to quantify effects from urban effective anisotropy on upwelling longwave radiation, but did not incorporate a robust atmospheric correction routine to retrieve T\textsubscript{hem, r} and remove spatiotemporally variant atmospheric effects.
+Inherent complexity in determining the distance to the surface for a moving sensor "seeing" a complex, three-dimensional surface make sensor view modeling difficult.
+Ray-tracing methods similar to those used in \citet{Ceamanos2017} could provide a solution, but must be modified to return path lengths in addition to seen and unseen points.
+In addition, sensitivity to ambient conditions (particularly the T\textsubscript{air} profile) will make accurate atmospheric correction computationally expensive in mobile applications of the method.
+Notwithstanding evaluations of method performance in such applications.
+If these problems are solved, mobile application of the method may be useful for determine as hemispherical instrumentation can be run simultaneously with near-ground and satellite based narrow-FOV to further understanding of effective anisotropy on TIR remote sensing of urban areas.
+Such analysis could be used to develop parameterizations to modify satellite retrieved urban T\textsubscript{surf} to better represent the complete urban T\textsubscript{surf}.
+
+Currently, little is known about.
+The simple fact that issues raised in have yet to be addressed should prompt diverse assessment of urban T\textsubscript{surf} via a range of sensors operating at wide range of spatiotemporal scales is necessary to foster better understand the true geometric, temporal, and spatial effects of urban areas on land T\textsubscript{surf}.
+While satellite assessment of sUHI and urban T\textsubscript{surf} has been instrumental in elucidating the spatial character of urban effects on T\textsubscript{surf}, several of the questions posed nearly 30 years ago in \citet{Roth1989} remain unanswered.
+This should prompt not only critical assessment of the urban T\textsubscript{surf} record, but also the development and use of novel methods to fill knowledge gaps, quantify biases, and uncover methodological shortcomings.
+This method, and in particular its parameterization, is intended to aid in such an endeavor.
+According to the Urban Flux Network (\href{http://fluxnet.fluxdata.org/}{Urbanflux}) there are 26 active urban flux towers, with many other existing inactive towers, in addition to datasets not indexed under the Urbanflux umbrella for which this method can be used for urban T\textsubscript{surf} and sUHI analysis.
+These span several continents, a range of climate zones and socioeconomic strata, with many operational over very long (5+ years) time scales.
+The application of this method over these datasets, and as a component of future energy balance assessments, could provide a wealth of potentially important analysis within and outside the purview of urban climatology.
+
+\section{Final remarks}
+
+%Broadly, the method, its parameterization, and the analysis presented in this study not only provides for novel analysis of the urban effect on T\textsubscript{surf}, but also aims to prompt further critical assessment of the urban T\textsubscript{surf} and the sUHI record.
+
+Broadly, the method, its parameterization, and the analysis presented in this study was conducted to provide for novel analyses of the urban effect on T\textsubscript{surf} and to prompt further critical assessment of the urban T\textsubscript{surf} and sUHI record.
+work towards addressing methodological and conceptual questions outlined in \citet{Roth1989}.
+Indeed, if similar conceptual flaws and shortcomings to those found for the aUHI in \citet{Stewart2011} are likely in sUHI analysis, quantification of instrument and methodological biases and shortcomings in remote sensing of urban T\textsubscript{surf} and the introduction of new methods to overcome these shortcomings may bring us closer to uncovering the true geometric, temporal, and spatial effects of cities on land surface temperature and on the climate at large.
 

### 2017-05-10 introduction.tex
Add files via upload

@@
 \chapter{Introduction}
+\label{intro}
+
 Urban development drastically changes the character of the surface.
@@
 
-Study of the UHI with respect to T\textsubscript{air} (aUHI) - typically referring to T\textsubscript{air} measured from approximately 2 \si{\meter} above ground level, yielding a "canopy layer" UHI (sUHI) - and T\textsubscript{surf} (sUHI) tacitly falls into such a framework, by replacing $V_{M, \ U} - V_{M, \ R}$ with $\Delta$T\textsubscript{U - R}.
+Study of the UHI with respect to T\textsubscript{surf} (sUHI) and T\textsubscript{air} (aUHI) - typically referring to T\textsubscript{air} measured from approximately 2 \si{\meter} above ground level, yielding a "canopy layer" UHI (clUHI) - tacitly falls into such a framework, by replacing $V_{M, \ U} - V_{M, \ R}$ with $\Delta$T\textsubscript{U - R}.
 However, in spite of its conceptual simplicity through the lens of the Lowry method, meta-analysis of observational aUHI studies suggests that that simplicity may not translate well to real world analysis.
@@
 As such, methodological problems in sUHI study can arise from both conceptual flaws \textit{and} instrument biases.
-In spite of these difficulties, the introduction of new methods, instruments, and analytical techniques in TIR remote sensing of urban environments has lead to rapid proliferation in study of urban T\textsubscript{surf} and the sUHI - much of which is focused on remote sensing of upwelling thermal infrared radiation (TIR).
+In spite of these difficulties, the introduction of new methods, instruments, and analytical techniques in remote sensing of thermal infrared radiation (TIR) upwelling from urban environments has lead to a rapid proliferation of study of urban T\textsubscript{surf} and the sUHI - much of which is focused on remote sensing of TIR.
 In the decades preceding this thesis, a combination of factors has lead to an increased focus on remote sensed study of urban T\textsubscript{surf}:
@@
 However, the methodological and conceptual concerns raised in \citet{Stewart2011} in combination with the wide (and broadening) range of methodologies for remote sensed urban T\textsubscript{surf} retrieval should prompt critical assessment of the remote sensed sUHI literature.
-At the time of writing, no systemic or critical reviews of the remote sensed sUHI literature are available.
-In addition, an increasingly dogmatic focus on improving spatial and radiometric resolutions has left other biases inherent in remote sensing of urban environments largely ignored in sUHI literature.
+At the time of writing, no systemic or critical reviews of the remote sensed sUHI literature are available and questions raised in critical assessments of remote sensed urban T\textsubscript{surf} in \cite{Roth1989} and \cite{Voogt2003} have gone largely unaddressed.
+An increasingly dogmatic focus on improving spatial and radiometric resolutions has left other biases inherent in remote sensing of urban environments largely ignored in sUHI literature.
 This suggests that similar conclusions to those found of aUHI literature in \citet{Stewart2011} are possible (if not likely) for study of sUHI.
@@
 	\caption{Projected FOVs for a narrow-FOV sensor at nadir and oblique angles, and a wide-FOV hemispherical sensor viewing an idealized urban array with building width \textit{H}, height 2\textit{H}, and inter element spacing \textit{H}.
-Figure from \cite{Adderley2015}.}
+Figure from Adderley et al.
+(2015).}
 	\label{model}

### 2017-05-10 paper1.tex
Add files via upload

@@
 	\includegraphics[width=\textwidth]{spectairtsurf}
-	\caption{At-sensor spectral directional radiances computed in MODTRAN 4.1 \citep{Berk1987} for short (30 \si{m}) and long (3000 \si{m}) path lengths (z) with Planck curves indicating spectral radiances at T\textsubscript{surf} = 300 \si{K} and T\textsubscript{air} = 290 \si{K}.}
+	\caption{At-sensor spectral directional radiances computed in MODTRAN 4.1 (Berk et al., 1987) for short (30 \si{m}) and long (3000 \si{m}) path lengths (z) with Planck curves indicating spectral radiances at T\textsubscript{surf} = 300 \si{K} and T\textsubscript{air} = 290 \si{K}.}
 	\label{spectairtsurf}
@@
 	\includegraphics[width=\textwidth]{dometrans}
-	\caption{A comparison of dome transmittance for a Kipp \& Zonen silicone domed pyrgeometer (data supplied by Kipp \& Zonen \textit{pers.
+	\caption{A comparison of dome transmittance for a Kipp \& Zonen silicone domed pyrgeometer (data supplied by Kipp \& Zonen, \textit{pers.
 comm.}) and at-sensor radiance (T\textsubscript{surf} = 300 \si{K}, T\textsubscript{air} = 300 \si{K}, $z$ = 30 \si{m})  for "typical" and extended bandpasses.}
@@
 T\textsubscript{surf} and T\textsubscript{air} are held constant in these simulations as they have a relatively small effect on $\tau$.
-Simulations for path lengths from 1 to 100 \si{\meter} and water vapor mass densities from 0.1 to 23.5 \si{\gram\per\meter\cubed} are included in Appendix \ref{appendixa}.
+Simulations for path lengths from 1 to 48 \si{\meter} and water vapor mass densities from 0.1 to 23.5 \si{\gram\per\meter\cubed} are included in Appendix \ref{appendixa}.
 Sensor-surface geometry is approximated as a single view factor weighted, azimuthally averaged, surface-to-sensor path length calculated using the SUM \citep{Soux2004} and a simplified DBM.

### 2017-05-10 paper2.tex
Add files via upload

@@
 
-\subsection{Evaluation and parameterization}
+\subsection{Method evaluation and sensitivity testing}
 
@@
 Modeled and measured irradiances match well, over wide range of mid-latitude atmospheric conditions, with results summarized in Table \ref{evalstats}.
+
+Sensor placement sensitivity testing in section \ref{sensitivity} shows that T\textsubscript{hem, r} is spatially variant when measured over urban areas at typical pyrgeometer heights.
+This introduces a potential bias in T\textsubscript{hem, r} as changing sensor position relative to canyon features alters view factor proportions of rooftops, roads, and walls.
+The sensitivity of T\textsubscript{hem, r} to sensor placement decreases with height and T\textsubscript{hem, r} becomes spatially invariant at approximately 3.5 times mean building height \citep{Roberts2010,Adderley2015}.
+As such, these biases can be significantly reduced by mounting downward facing pyrgeometers slightly higher than normal.
+Although pyrgeometers at both the Sperrstrasse and are mounted below the height at which T\textsubscript{hem,r} is spatially invariant, sensitivity tests in section \ref{sensitivity} show that the magnitude of this bias is relatively small for sensor placement at the Sperrstrasse canyon.
+Similar or lesser biases as the sensor is mounted offset from surrounding rooftops (which contribute to large diurnal amplitude in T\textsubscript{hem, r} and larger deviation from T\textsubscript{comp}).
+In addition, T\textsubscript{hem, r} shows good performance relative to T\textsubscript{plan} when compared to T\textsubscript{comp}, particularly under clear sky conditions when microscale spatiotemporal variations in urban T\textsubscript{surf} and sUHI magnitudes are greatest.
+As such, the influence of biases in T\textsubscript{hem} in sUHI analysis for both BUBBLE and EPiCC campaigns is likely small.
+
+\subsection{A parameterization scheme}
 
@@
 Humidity-transmittance pairings are aggregated into a look-up table (LUT), where the measured humidity at each time step can be matched with the associated hemispherical transmittance.
-Using the hemispherical transmittance as the fraction of surface emitted L\textsubscript{up} reaching the sensor, measured L\textsubscript{up} can then be decomposed into its surface and atmospheric emitted components, from which a parameterized T\textsubscript{hem, r} (T\textsubscript{hem, para}) is inferred.
+Using the hemispherical transmittance as the fraction of surface emitted L\textsubscript{up} reaching the sensor, measured L\textsubscript{up} can then be decomposed into its surface and atmospheric emitted components, from which a parameterized T\textsubscript{hem, r} is inferred.
 Comparison of T\textsubscript{hem, r} derived from the correction method and the parameterization show good agreement, the results of which are included in Figure \ref{para_eval} and Table \ref{parastats}.
@@
 	\centering
-	\caption{A description of morphological parameters and measured variables for BUBBLE urban and rural sites. Modified from \citet{Rotach2005} to include only relevant parameters.}
+	\caption{A description of morphological parameters and measured variables for BUBBLE urban and rural sites. Modified from Rotach et al. (2005) to include only relevant parameters.}
 	\label{morphbspr}
@@
 
-sUHI and clUHI magnitudes were calculated from T\textsubscript{hem, r} and T\textsubscript{air} (measured from 2 \si{\meter} above ground level) as $\Delta$T\textsubscript{(urban-rural)} at 30 \si{\min} intervals.
+sUHI and clhUHI magnitudes were calculated from T\textsubscript{hem, r} and T\textsubscript{air} (measured from 2 \si{\meter} above ground level) as $\Delta$T\textsubscript{(urban-rural)} at 30 \si{\min} intervals.
 Rural T\textsubscript{hem, b} was calculated an expanded version of equation \ref{stefb1} to remove emissivity effects,
@@
 
-\noindent where $L_z^\uparrow$ is at-sensor upwelling longwave radiation, $L_{sky}^\downarrow$ is at-sensor reflected downwelling longwave radiation, and emissivity ($\epsilon$) = 0.94.
+\noindent where $L_z^\uparrow$ is at-sensor upwelling longwave radiation, $L_{sky}^\downarrow$ is at-sensor reflected downwelling longwave radiation, and rural emissivity ($\epsilon$) = 0.94.
 Sensitivity tests included in Section \ref{sensitivity} of the companion paper show that screen level irradiances ($ z $ = ~2 \si{\meter}) are not subject to significant effects from the intervening atmosphere and do not require atmospheric correction.
@@
 
-Figure \ref{meteo_wv_sol} panel (a) shows mean and maximum sUHI magnitudes as a function of solar radiant exposure over each day of the summer months of the BUBBLE campaign.
-Point shading indicates the month (white is April and black is July).
+Figure \ref{meteo_sol} shows mean and maximum sUHI magnitudes (sUHI\textsubscript{mean} and sUHI\textsubscript{max}) as a function of solar radiant exposure for each day of the BUBBLE campaign.
+Point shading indicates the season, where black is winter (Nov - Feb), white is spring (Mar - Apr), and red is summer (May - Aug).
 A clear positive relationship is observed between accumulated solar radiation and both sUHI\textsubscript{max} and sUHI\textsubscript{mean}.
-Panel (b) of Figure \ref{meteo_wv_sol} shows mean and maximum hemispherical sUHI magnitudes as a function of mean wind velocity for the duration of the eight month BUBBLE climatology.
-Colors indicate solar radiant exposure accumulated over each day.
+
+Figure \ref{meteo_wv} shows hemispherical sUHI\textsubscript{mean} and sUHI\textsubscript{max} magnitudes as a function of mean wind velocity for each day of the BUBBLE climatology.
+A weak negative relationship is observed between mean wind velocity and sUHI\textsubscript{max}, where larger sUHI\textsubscript{max} occurs under low wind conditions.
+sUHI\textsubscript{mean} yields a slightly more muted relationship.
 Similar results are observed when rural winds are substituted.
+Figure \ref{meteo_hum} shows hemispherical sUHI\textsubscript{mean} and sUHI\textsubscript{max} magnitudes as a function of mean atmospheric water vapor content for each day over the BUBBLE campaign.
+No discernible relationship is observed between atmospheric water vapor content and sUHI\textsubscript{mean} or sUHI\textsubscript{max}.
+In both, colors indicate solar radiant exposure accumulated over each day (yellow is large accumulated insolation, brown is low).
 
-Figures \ref{meteo_hum} and \ref{meteo_atuhi} show hemispherical sUHI magnitudes as a function of water vapor content and clUHI measured over the BUBBLE campaign.
-Winter hours are omitted in both as sUHI is small and displays minimal variation in winter.
-Jamie: still working on getting color palettes for these.
-Will merge into one figure.
+Figured \ref{meteo_atuhi} shows hemispherical sUHI magnitudes as a function of water vapor content and clUHI measured over the summer months of the BUBBLE campaign (May through July 2002).
+Winter hours are omitted as sUHI is small and displays minimal variation in winter.
+The relationship between sUHI and clUHI is complex and highly dependent on time of day - owing largely to similar, but slightly offset patterns of sUHI and clUHI development.
+Coloring in the daytime plot indicates time of day: Red is morning (5:00 - 9:00 LST), gray is midday hours (9:30 - 17:00 LST), and yellow is evening hours (17:30 - 20:00 LST).
+
+\pagebreak
 
@@
 	\centering
-	\includegraphics[width=15cm,height=7in,keepaspectratio]{meteo_wv_sol}
-	\caption{(a) Maximum and mean 24 \si{\hour} hemispherical sUHI magnitude for each day of April through July versus integrated solar radiant exposure measured over the same 24 \si{\hour} period.
-Shading indicates month of the year, with dark points indicating later months.
-(b) Maximum and mean 24 \si{\hour} hemispherical sUHI magnitude for each day in the eight month climatology versus mean wind velocity measured at approximately 2 \si{\meter} above ground at the Sperrstrasse site.
+	\includegraphics[width=15cm,height=7in,keepaspectratio]{meteo_kdown}
+	\caption{Maximum and mean 24 \si{\hour} hemispherical sUHI magnitude for each day of April through July versus integrated solar radiant exposure measured over the same 24 \si{\hour} period.
+Color indicates season: black is winter (Nov - Feb), white is spring (Mar - Apr), and red is summer (May - Aug).
+}
+	\label{meteo_sol}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=15cm,height=7in,keepaspectratio]{meteo_wv}
+	\caption{Maximum and mean 24 \si{\hour} hemispherical sUHI magnitude for each day versus mean wind velocity measured at approximately 2 \si{\meter} above ground at the Sperrstrasse site.
 Coloring indicates integrated solar radiant exposure over each day.}
-	\label{meteo_wv_sol}
+	\label{meteo_wv}
 \end{figure}
@@
 	\includegraphics[width=15cm,height=7in,keepaspectratio]{meteo_hum}
-	\caption{sUHI magnitude versus atmospheric water vapor content measured at 2 \si{\meter} at the Sperrstrasse site. Similar patterns are observed when rural water vapor content is substituted.}
+	\caption{Maximum and mean 24 \si{\hour} sUHI magnitude versus atmospheric water vapor content measured at 2 \si{\meter} at the Sperrstrasse site.
+Similar patterns are observed when rural water vapor content is substituted.
+Coloring indicates integrated solar radiant exposure over each day.}
 	\label{meteo_hum}
@@
 	\includegraphics[width=15cm,height=7in,keepaspectratio]{meteo_atuhi}
-	\caption{sUHI magnitude versus clUHI magnitude. clUHI is calculated as the difference between urban and rural T\textsubscript{air} measured at 2 \si{\meter} above ground level.}
+	\caption{sUHI magnitude versus clUHI magnitude binned for day and night calculated over the summer months of the BUBBLE campaign.
+Coloring in the daytime plot indicates time of day: Red is morning (5:00 - 9:00 LST), gray is midday hours (9:30 - 17:00 LST), and yellow is evening hours (17:30 - 20:00 LST).}
 	\label{meteo_atuhi}
+\end{figure}
+
+%Vancouver
+
+Figure \ref{vanc_sol} shows mean and maximum sUHI magnitudes (sUHI\textsubscript{mean} and sUHI\textsubscript{max}) as a function of solar radiant exposure over each day of the year long EPiCC Vancouver campaign.
+Point shading indicates the season, where black is winter (Nov - Feb), white is spring (Mar - Apr), red is summer (May - Aug), and blue is fall (Sept - Oct).
+A clear positive relationship is observed between accumulated solar radiation and both sUHI\textsubscript{max} and sUHI\textsubscript{mean}, with significant seasonal variation.
+
+Figure \ref{vanc_wv} shows hemispherical sUHI\textsubscript{mean} and sUHI\textsubscript{max} magnitudes as a function of mean wind velocity for each day of the EPiCC Vancouver campaign.
+A weak negative relationship is observed between mean wind velocity and both sUHI\textsubscript{mean} and sUHI\textsubscript{max}, where larger sUHI magnitudes tend to occur under low wind conditions.
+Higher overall sUHI magnitudes and mean wind velocities in Vancouver make the relationship between sUHI and mean wind velocity slightly more evident.
+Similar results are observed when rural winds are substituted.
+Figure \ref{vanc_hum} shows hemispherical sUHI\textsubscript{mean} and sUHI\textsubscript{max} magnitudes as a function of mean atmospheric water vapor content for each day of the EPiCC Vancouver campaign.
+In contrast to findings in the BUBBLE campaign, a weak relationship between sUHI and humidity is observed.
+However, the apparent relationship may be a product of Vancouver's proximity to the coast, as there is significant seasonal variation in both  and may be no actual causal relationship.
+In both, colors indicate solar radiant exposure accumulated over each day (yellow is large accumulated insolation, brown is low).
+
+Figured \ref{vanc_swc} shows hemispherical sUHI magnitudes as a function of water vapor content and clUHI measured for April and May 2009 of the EPiCC Vancouver canpaign.
+These months were selected to show a wide range of soil water contents and sUHI magnitudes.
+A negative relationship is observed between sUHI and soil water vapor content, particularly in May, where larger sUHI magnitudes often coincide with lower soil water vapor contents.
+However, some of this relationship may owe to diurnal patterns of soil saturation, which, contrary to sUHI magnitudes, tend to be highest in early morning and lower throughout the day.
+Color in both plots indicates time of day: shades of blue indicate early morning through midday hours, white is solar noon through early evening hours, and red is late evening through the nighttime.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=15cm,height=7in,keepaspectratio]{vanc_kdown}
+	\caption{Maximum and mean 24 \si{\hour} hemispherical sUHI magnitude for each day of the EPiCC climatology versus integrated solar radiant exposure measured over the same 24 \si{\hour} period.
+Color indicates season: black is winter (Nov - Feb), white is spring (Mar - Apr), red is summer (May - Aug), and blue is fall (Sept - Oct).}
+	\label{vanc_sol}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=15cm,height=7in,keepaspectratio]{vanc_wv}
+	\caption{Maximum and mean 24 \si{\hour} hemispherical sUHI magnitude for each day over the EPiCC climatology versus mean wind velocity measured at approximately 10 \si{\meter} above ground at the Sunset tower site.
+Similar patterns are observed when rural mean wind velocity is substituted.
+Coloring indicates integrated solar radiant exposure over each day.}
+	\label{vanc_wv}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=15cm,height=7in,keepaspectratio]{vanc_hum}
+	\caption{Maximum and mean 24 \si{\hour} sUHI\textsubscript{hemi} magnitude versus mean atmospheric water vapor content measured at 2 \si{\meter} at the Sunset tower urban site in Vancouver, Canada.
+Similar patterns are observed when rural water vapor content is substituted.
+Coloring indicates integrated solar radiant exposure over each day.}
+	\label{vanc_hum}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=15cm,height=7in,keepaspectratio]{vanc_swc}
+	\caption{sUHI\textsubscript{hemi} magnitude versus soil water content measured from 0.05 \si{\meter} below the surface at the Westham island site in Vancouver, Canada for April and May 2009.
+Color indicates time of day: blue is early morning through midday, white is solar noon through early evening, and red is late evening through the night.}
+	\label{vanc_swc}
 \end{figure}
@@
 
-\subsection{The effect of rural/non-urban characteristics on sUHI}
+\subsection{The effect of rural/non-urban characteristics on sUHI and a generalizable case for sUHI}
 
@@
 Days were chosen to represent a wide range of summertime conditions and cloud coverages, however, observations at the Village Neuf site were intermittent and n = 9 only applies to hours after 12:00 LST when data could be continuously sampled.
+
+Figure \ref{gen_case} presents a schematic for summertime sUHI development constructed from mean urban and rural T\textsubscript{hem} over the summer months of the EPiCC Vancouver campaign.
+sUHI shows clear diurnal dependence on differential heating/cooling rates at the urban and rural sites, with greater urban heating rates observed from sunrise to approximately solar noon forcing the development of sUHI and greater urban cooling rates in the afternoon and evening hours forcing the destruction of daytime sUHI.
+These results are highly generalizable to other mid-latitude cities.
 
@@
 
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=15cm,height=7in,keepaspectratio]{gen_case}
+	\caption{Top: Mean T\textsubscript{hem, r} for the Sunset tower (urban) and Westham island (rural) sites in Vancouver, Canada over May - Aug 2009.
+Middle: Mean heading/cooling rates for urban and rural sites.
+Bottom: Mean sUHI\textsubscript{hemi} and hourly rate of change in sUHI\textsubscript{hemi}.
+n = 5905}
+	\label{gen_case}
+\end{figure}
+
 \section{Discussion}
@@
 However, the results in \citet{Oke2017} are based on simulations of an idealized urban area without trees, complex non-orthogonal surface geometry, or sloped facets and have not been supported observationally.
-The results in this chapter provide an ideal platform to test conceptual understanding of sUHI development, to understand the diurnal and seasonal nature of sUHI and the meteorological factors that drive its development.
+The results in this chapter provide an ideal platform to test conceptual understanding of sUHI development, to understand and characterize the diurnal and seasonal nature of sUHI, and to elucidate meteorological and climatological factors that drive its development.
 
@@
 In contrast, shading by urban features causes urban T\textsubscript{surf} to respond more slowly, resulting in a small sUHI\textsubscript{min} in the hours just after sunrise.
-This is observed in Basel across all representations of urban T\textsubscript{surf}.
+This is observed in Basel across all representations of urban T\textsubscript{surf} and in the generalizable case presented for Vancouver sUHI.
 Through the late morning hours, urban heating rates increase relative to rural rates as the solar zenith angle decreases and low thermal admittance facets (particularly rooftops) experience more direct insolation and quickly heat up.
 Increases in urban T\textsubscript{surf} are amplified by the abundance of impervious surfaces, which result in a smaller fraction of insolation partitioned to latent heat relative to vegetated non urban spaces.
-After the initial spike in rural heating in the early morning hours, urban heating rates overtake rural approximately three to four hours after sunrise and the sUHI slowly grows through midday to a peak at approximately two to three hours hours after solar noon.
+After the initial spike in rural heating in the early morning hours, urban heating rates overtake rural approximately three to four hours after sunrise and the sUHI slowly grows through midday to a peak at approximately two hours hours after solar noon.
 In the hours after sUHI\textsubscript{max}, the solar zenith angle increases and both urban and rural heating rates decline and begin to converge, as shading by canyon geometry causes some facets to begin to cool.
@@
 For the Sperrstrasse canyon, the combination of a narrow but relatively low street canyon results in a slight reduction in Basel's sUHI as strong radiative cooling of rooftops is slightly offset by canyon radiation trapping.
+In Vancouver, urban cooling rates quickly overtake rural - owing to strong radiative cooling - and sUHI declines precipitously after sUHI\textsubscript{max}.
 The dependence of sUHI development on canyon geometry is illustrated well by comparing nadir and complete representations of sUHI observed in Basel.
@@
 After sunset, the unobstructed rural sky-view-factor allows for strong radiative cooling of the surface and rural cooling rates spike relative to urban, leading to a slight jump in sUHI.
+This is observed in Vancouver sUHI as a slight reduction in the rate of sUHI decline near 20:00 LST.
 However, variations in rural thermal admittance, surface geometry, land cover and use, and/or influence from surrounding terrain can result in significant contrasts in rural cooling rates.
-All of these factors show strong spatiotemporal variation, resulting in large differences in sUHI development depending on rural site characteristics, similar to observations of variations in clUHI based on rural siting in \citep{Runnalls2000}.
+All of these factors show strong spatiotemporal variation, resulting in large differences in sUHI development depending on rural site characteristics, similar to observations of variation in clUHI based on rural siting in \citep{Runnalls2000}.
 This is illustrated by two different patterns of post-sunset sUHI development observed for the two rural sites in Basel.
@@
 The "extra", largely non-radiative cooling at the Lange Erlen site results in a strong second peak in sUHI that is not present in sUHI calculated using the Village Neuf site.
+In the context of the Lowry method discussed in Chapter \ref{intro}, $V_{M, \ R}$ measured from the Lange Erlen site and $V_{M, \ U}$ are not subject to the same $V_{L}$ and equation does not yield a perfectly isolated $V_{H}$ in the hours just after sunset.
+
 Following site dependent effects on sUHI in the hours after sunset, sUHI in Basel is relatively constant over the nighttime hours, with slight growth observed on clear nights where canyon radiation trapping retards urban nocturnal cooling rates relative to rural.
+In contrast, mean sUHI in Vancouver decreases through the night as the urban area experiences slight cooling.
+Both of these cases demonstrate the dependence of observed sUHI on urban and rural site characteristics and sensor-surface geometry illustrating the need for proper critical analysis of the ensemble of climactic forcings present at a given urban-rural pairing and the potential for influence from sensor or instrument bias.
 
@@
 As such, sUHI development is strongly tied to integrated solar radiant exposure accumulated over a day and wind velocity, which display positive and negative relationships with sUHI magnitudes respectively.
+Other variables also show weakly positive (mean atmospheric water vapor content) and negative (soil water content) relationships with sUHI, however, conclusions about causality are more difficult to make as they too display some dependence on the same variables.
 Meteorological controls on sUHI are summarized in Table \ref{meteo_cont}.
@@
 
-As was found with correction magnitudes in Chapter \ref{paper1}, sUHI magnitudes are not strongly correlated with water vapor content as, in the absence of fog, humidity near the surface does not strongly influence surface heating or cooling rates in and of itself.
-Atmospheric transmittance of longwave radiation decreases sharply with small increases in water vapor when the atmosphere is very dry - shown in Figure \ref{humtest}.
+sUHI magnitudes in the BUBBLE and EPiCC campaigns yield slightly different relationships with atmospheric water vapor content, with no relationship observed in Basel and a weakly positive one observed in Vancouver.
+It is unlikely that these are a result of any significant physical or causal connection between the two variables as, in the absence of fog, humidity near the surface does not strongly influence surface heating or cooling rates in and of itself.
+Atmospheric transmittance of longwave radiation decreases sharply with small increases in water vapor when the atmosphere is very dry, with only marginal subsequent decreases - shown in Figure \ref{humtest}.
 As water vapor content increases from zero, variation in spectral atmospheric transmittance decreases.
 Therefore, typical variations in urban and rural humidities have little effect on the fraction of outgoing longwave radiation available for heating the canopy layer air volume through absorption and for heating of the surface by atmospheric emission.
-As a result, water vapor content has little effect on sUHI.
+As a result, a similar relationship is expected to that found with correction magnitudes in Chapter \ref{paper1} is expected between humidity and sUHI.
+A weak positive correlation between the two is therefore likely a result of coincident seasonality in both variables as both mean daytime sUHI and mean water vapor mass densities are large in summer and small in winter.
+As this seasonality is evident in both Basel and Vancouver, it is likely that a full climatology of sUHI and humidity would yield similar correlation (but not causation) in Basel as is observed in Vancouver.
 
@@
 		& \\
-		Water vapor content, \si{\gram\per\meter\squared} & \textit{negligible} \\
+		Atmospheric water vapor content, \si{\gram\per\meter\cubed} & \textit{negligible} \\
+		&\\
+		Volumetric soil moisture content, \si{\percent} & \textit{decreasing} \\
 		&\\

### 2017-05-10 uw-ethesis-frontpgs.tex
Add files via upload

@@
         \Huge
-        {\bf A Method to Assess the Surface Urban Heat Island Effect Using Hemispherical Radiometric Surface Temperatures}
+        {\bf A method for hemispherical ground based remote sensing of urban surface temperatures}
+        % Hemispherical Remote Sensing of Urban Surface Temperature
+        %Hemispherical Remote Sensing of Urban Thermal Climates
+        %A Method to Assess the Surface Urban Heat Island Effect Using Hemispherical Radiometric Surface Temperatures
 
@@
 
+		%\vspace*{1.0cm}
         \vspace*{3.0cm}
@@
 \pagestyle{plain}
-\setcounter{page}{3}
+\setcounter{page}{2}
 
@@
 
-This is the abstract.
+This thesis presents a method for deriving time-continuous urban surface temperature and heat island assessments from hemispherical ground-based measurements of upwelling thermal radiation.
+The method, developed to overcome geometric and temporal biases inherent in traditional thermal remote sensing of urban surface climates, uses a sensor view model in conjunction with a radiative transfer code to derive atmospherically corrected, hemispherical radiometric urban surface temperatures.
+These are used to derive two long-term climatologies of surface urban heat island (sUHI) magnitudes for Basel, Switzerland and Vancouver, Canada.
+sUHI development shows significant variation based on time-of-day, season, and ambient and synoptic conditions.
+Results also show large differences in remote sensed sUHI from hemispherical, nadir and complete representations of the urban surface, with a nadir view overestimating seasonal sUHI\textsubscript{max} from a complete view by nearly a factor of two.
+In contrast, a hemispherical view provides significantly more representative, time-continuous urban surface temperature and sUHI analysis.
 
@@
 
-These are my Acknowledgments
+To be filled in at a later date.
 \cleardoublepage
@@
 % -----------------------------
-\printglossaries
+\chapter*{List of Symbols}
+\addcontentsline{toc}{chapter}{List of Symbols}
+
+\captionsetup[table]{list=no}
+
+\begin{table}[H]
+	\centering
+	\begin{tabular}{p{1.5cm}p{2cm}p{10cm}}
+		Roman &&\\
+		\toprule
+
+		Symbol & Unit & Property \\
+		\midrule
+		$C$ & \si{\joule} & heat capacity \\
+		FOV & \si{\deg} & sensor field of view \\
+		\textit{H} & \si{\meter} & building width \\
+		$Q*$ & \si{\watt\per\square\meter} & net radiation flux density \\
+		$\Delta Q_A$ & \si{\watt\per\square\meter} & net advective heat flux density \\
+		$\Delta Q_G$ & \si{\watt\per\square\meter} & net heat storage flux density \\
+		$Q_E$ & \si{\watt\per\square\meter} & latent heat flux density \\
+		$Q_F$ & \si{\watt\per\square\meter} & anthropogenic heat flux density \\
+		$Q_H$ & \si{\watt\per\square\meter} & sensible heat flux density \\
+		$t$ & \si{\minute} & time \\
+		T\textsubscript{air} & \si{\kelvin} & air temperature \\
+		T\textsubscript{surf} & \si{\kelvin} & surface temperature \\
+		T\textsubscript{comp} & \si{\kelvin} & complete urban surface temperature\\
+		T\textsubscript{plan} & \si{\kelvin} & urban surface temperature viewed from a sensor in the nadir\\
+		T\textsubscript{hem} & \si{\kelvin} & urban surface temperature viewed from a hemispherical remote sensor\\
+		aUHI & \si{\kelvin} & air temperature urban heat island effect \\
+		clUHI & \si{\kelvin} & canopy layer air temperature urban heat island effect \\
+		sUHI & \si{\kelvin} & surface temperature urban heat island effect \\
+		$V_B$ & - & effects from background climate on observation of a given meteorological variable \\
+		$V_H$ & - & effects from the urban environment on observation of a given meteorological variable \\
+		$V_L$ & - & effects from local topography on observation of a given meteorological variable \\
+		$V_M$ & - & observed meteorological variable \\
+		$V_{M, U} $ & - & observed meteorological variable (urban) \\
+		$V_{M  R}$ & - & observed meteorological variable (rural) \\
+		$z$ & \si{\meter} & layer thickness (air, surface, or subsurface)\\
+		\bottomrule
+	\end{tabular} 
+\end{table}
+
+\begin{table}[H]
+	\centering
+	\begin{tabular}{p{1.5cm}p{2cm}p{10cm}}
+		Roman &&\\
+		\toprule
+
+		Symbol & Unit & Property \\
+		\midrule
+		$C$ & \si{\joule} & heat capacity \\
+		FOV & \si{\deg} & sensor field of view \\
+		\textit{H} & \si{\meter} & building width \\
+		$Q*$ & \si{\watt\per\square\meter} & net radiation flux density \\
+		$\Delta Q_A$ & \si{\watt\per\square\meter} & net advective heat flux density \\
+		$\Delta Q_G$ & \si{\watt\per\square\meter} & net heat storage flux density \\
+		$Q_E$ & \si{\watt\per\square\meter} & latent heat flux density \\
+		$Q_F$ & \si{\watt\per\square\meter} & anthropogenic heat flux density \\
+		$Q_H$ & \si{\watt\per\square\meter} & sensible heat flux density \\
+		$t$ & \si{\minute} & time \\
+		T\textsubscript{air} & \si{\kelvin} & air temperature \\
+		T\textsubscript{surf} & \si{\kelvin} & surface temperature \\
+		T\textsubscript{comp} & \si{\kelvin} & complete urban surface temperature\\
+		T\textsubscript{plan} & \si{\kelvin} & urban surface temperature viewed from a sensor in the nadir\\
+		T\textsubscript{hem} & \si{\kelvin} & urban surface temperature viewed from a hemispherical remote sensor\\
+		aUHI & \si{\kelvin} & air temperature urban heat island effect \\
+		clUHI & \si{\kelvin} & canopy layer air temperature urban heat island effect \\
+		sUHI & \si{\kelvin} & surface temperature urban heat island effect \\
+		$V_B$ & - & effects from background climate on observation of a given meteorological variable \\
+		$V_H$ & - & effects from the urban environment on observation of a given meteorological variable \\
+		$V_L$ & - & effects from local topography on observation of a given meteorological variable \\
+		$V_M$ & - & observed meteorological variable \\
+		$V_{M, U} $ & - & observed meteorological variable (urban) \\
+		$V_{M  R}$ & - & observed meteorological variable (rural) \\
+		$z$ & \si{\meter} & layer thickness (air, surface, or subsurface)\\
+		\bottomrule
+	\end{tabular} 
+\end{table}
+
+\captionsetup[table]{list=yes}
+
+%\printglossaries
 \cleardoublepage

### 2017-05-11 conclusion.tex
Add files via upload

@@
 	\item To retrieve T\textsubscript{surf} that is representative of the temperature at which the surface is radiating (known as radiometric T\textsubscript{surf}) by removing effects from absorption and emission by the atmosphere in between the surface and the sensor.
-	\item To facilitate intersite analysis of T\textsubscript{surf}, where different ambient conditions, surface-sensor-sun geometries, and instrument types can yield differential atmospheric effects on the remote sensed TIR signal.
+	\item To facilitate intersite analysis of T\textsubscript{surf}, where different ambient conditions, surface-sensor-sun geometries, and instrument types can have differential effects remote sensed TIR.
 \end{enumerate}
@@
 In addition, Chapter \ref{paper1} explores the effect of surface geometry on remote sensed urban T\textsubscript{surf} for three representations of the urban surface (complete, plan, and hemispherical).
-Analysis shows significant overestimation (underestimation) of T\textsubscript{comp} by T\textsubscript{plan} by day (night), particularly under clear sky conditions.
+Analysis shows significant overestimation of T\textsubscript{comp} by T\textsubscript{plan} during the day and underestimation at night, particularly under clear sky conditions.
 By day, T\textsubscript{hem} overestimates T\textsubscript{comp}, but to a lesser degree than T\textsubscript{plan}.
 Daytime overestimation of T\textsubscript{comp} by T\textsubscript{hem} is more consistent under a wide variety of synoptic conditions, whereas overestimation by T\textsubscript{plan} fluctuates significantly, resulting in poor performance under clear-sky "satellite friendly" conditions.
-As these measures make up the bulk of the satellite urban T\textsubscript{surf} record, our understanding of urban T\textsubscript{surf} likely includes a significant warm bias.
-These biases are reflected in sUHI analysis in Chapter \ref{paper2}.
+As measurements from narrow-FOV sensors in the nadir make up the bulk of the satellite urban T\textsubscript{surf} record, our understanding of urban T\textsubscript{surf} likely includes a significant warm bias.
+These biases are reflected in analysis of the effect of sensor-surface geometry on sUHI in Chapter \ref{paper2}.
 
-Chapter \ref{paper2} includes an analysis of two long term, time continuous climatologies of hemispherical sUHI, one in Basel, Switzerland and another in Vancouver, Canada.
+Chapter \ref{paper2} includes an analysis of two long term, time-continuous climatologies of hemispherical sUHI, one in Basel, Switzerland and another in Vancouver, Canada.
 sUHI for both study sites shows significant diurnal, day-to-day, and seasonal variability.
@@
 However, as sUHI is the result of differences in heating and cooling rates between the city and its non urban surroundings, its relationship to meteorological controls is complex.
-Analysis of sUHI development under a range of mid-latitude synoptic conditions lead to the development of the first observationally supported generalizable model for summertime sUHI development presented in section \ref{gen}.
+Analysis of sUHI development under a range of mid-latitude synoptic conditions lead to the development of the first observationally-supported, generalizable model for sUHI development presented in Section \ref{gen}.
 
-Analysis of the directional dependence of sUHI is also included in Chapter \ref{paper2}.
-Results are similar to those found in Chapter \ref{paper1}, with significant daytime overestimation of sUHI\textsubscript{comp} by sUHI{plan}, particularly during hot clear sky conditions.
-Not only did directional biases in remote sensed sUHI change observed magnitudes, these biases also changed the diurnal character of sUHI, modifying patterns of sUHI in the hours during near sunset.
+Analysis of the effect of sensor-surface geometry on sUHI is also included in Chapter \ref{paper2}.
+Results are similar to those found in Chapter \ref{paper1}, with significant daytime overestimation of sUHI\textsubscript{comp} by sUHI\textsubscript{plan}, particularly during hot clear sky conditions.
+Directional biases in remote sensed assessment of sUHI change observed magnitudes and the diurnal character of sUHI by modifying patterns of observed sUHI in the hours near sunset.
 Thus, satellite based assessments of the diurnal pattern of sUHI may lead to misrepresentations of its 'true' diurnal character.
@@
 The complex nature of the urban surface introduces myriad shortcomings and biases into conventional remote sensing of urban thermal climates - this method is no exception.
-Sensor placement sensitivity tests in Section \ref{placement} corroborate findings in \citet{Roberts2010,Adderley2015} and indicate that longwave radiation upwelling from the urban surface as measured from a downward facing radiometer is spatially variant when measured from heights below approximately three to five times mean building height.
+Sensor placement sensitivity tests in Section \ref{placement} corroborate findings in \citet{Roberts2010,Adderley2015} and indicate that longwave radiation upwelling from the urban surface as measured from a downward-facing radiometer is spatially variant when measured from heights below approximately three to five times mean building height.
 Most pyrgeometers (including those used in this study) are mounted below this height and are subject to a bias towards facets that make up a disproportionately large fraction of the sensor FOV.
-For sensors mounted at less than optimal heights, the effect of this bias on derived T\textsubscript{hem, r} is largely dependent on sensor placement relative to rooftop facets, with better performance further from positions over rooftops.
+For sensors mounted at less than optimal heights, the effect of this bias on derived T\textsubscript{hem, r} is largely dependent on sensor placement relative to rooftop facets, with better performance when mounted offset from roof areas.
 Pyrgeometers at the Sperrstrasse and Sunset tower sites are placed far enough from rooftops to render this bias relatively small.
@@
 
-As the method is fixed and ground based, it is not practical for large scale spatially extensive analysis of sUHI.
+As the method developed in this study uses upwelling longwave measurements from fixed tower sites, it is not practical for large scale spatially-continuous and spatially extensive analysis of sUHI.
 This limits the method to mid- and low-rise neighborhoods.
-However, the concept can be adapted to retrieve spatiotemporally continuous urban T\textsubscript{hem, r} from mobile measurements of upwelling longwave irradiances.
-\citet{Sugawara2006} used near-ground, helicopter-mounted, pyrgeometers and narrow to quantify effects from urban effective anisotropy on upwelling longwave radiation, but did not incorporate a robust atmospheric correction routine to retrieve T\textsubscript{hem, r} and remove spatiotemporally variant atmospheric effects.
-Inherent complexity in determining the distance to the surface for a moving sensor "seeing" a complex, three-dimensional surface make sensor view modeling difficult.
-Ray-tracing methods similar to those used in \citet{Ceamanos2017} could provide a solution, but must be modified to return path lengths in addition to seen and unseen points.
+However, by combining concepts from other work, the method can be adapted to retrieve spatiotemporally continuous urban T\textsubscript{hem, r} from mobile measurements of upwelling longwave irradiances.
+\citet{Sugawara2006} used helicopter-mounted pyrgeometers and narrow-FOV infrared thermometers in aerial traverses to successfully quantify effects from urban effective anisotropy on upwelling longwave radiation, but did not incorporate a robust atmospheric correction routine to account for spatially variant surface geometry to remove atmospheric effects or retrieve T\textsubscript{hem, r}.
+Determining the distance to the surface for a moving hemispherical sensor viewing a complex, three-dimensional surface is difficult.
+Ray-tracing methods similar to those used in \citet{Ceamanos2017} could provide a solution, but must be modified to return path lengths in addition to seen and unseen points for a mobile aerial sensor.
 In addition, sensitivity to ambient conditions (particularly the T\textsubscript{air} profile) will make accurate atmospheric correction computationally expensive in mobile applications of the method.
 Notwithstanding evaluations of method performance in such applications.
-If these problems are solved, mobile application of the method may be useful for determine as hemispherical instrumentation can be run simultaneously with near-ground and satellite based narrow-FOV to further understanding of effective anisotropy on TIR remote sensing of urban areas.
+If these problems are solved, application of the method in aerial traverses may be useful in quantifying the spatiotemporal character of urban effective anisotropy on remote sensed urban T\textsubscript{surf} as hemispherical instrumentation can be run simultaneously with near-ground and satellite based narrow-FOV sensors.
 Such analysis could be used to develop parameterizations to modify satellite retrieved urban T\textsubscript{surf} to better represent the complete urban T\textsubscript{surf}.
 
-Currently, little is known about.
-The simple fact that issues raised in have yet to be addressed should prompt diverse assessment of urban T\textsubscript{surf} via a range of sensors operating at wide range of spatiotemporal scales is necessary to foster better understand the true geometric, temporal, and spatial effects of urban areas on land T\textsubscript{surf}.
-While satellite assessment of sUHI and urban T\textsubscript{surf} has been instrumental in elucidating the spatial character of urban effects on T\textsubscript{surf}, several of the questions posed nearly 30 years ago in \citet{Roth1989} remain unanswered.
+While satellite assessment of sUHI and urban T\textsubscript{surf} has been essential in elucidating the spatial character of urban effects on T\textsubscript{surf}, several of the questions posed nearly 30 years ago in \citet{Roth1989} remain unanswered.
 This should prompt not only critical assessment of the urban T\textsubscript{surf} record, but also the development and use of novel methods to fill knowledge gaps, quantify biases, and uncover methodological shortcomings.
-This method, and in particular its parameterization, is intended to aid in such an endeavor.
-According to the Urban Flux Network (\href{http://fluxnet.fluxdata.org/}{Urbanflux}) there are 26 active urban flux towers, with many other existing inactive towers, in addition to datasets not indexed under the Urbanflux umbrella for which this method can be used for urban T\textsubscript{surf} and sUHI analysis.
-These span several continents, a range of climate zones and socioeconomic strata, with many operational over very long (5+ years) time scales.
-The application of this method over these datasets, and as a component of future energy balance assessments, could provide a wealth of potentially important analysis within and outside the purview of urban climatology.
+This method, and in particular its parameterization, is intended to aid in such an endeavor as it is readily adapted to other study sites and requires few meteorological inputs.
+According to the Urban Flux Network (\href{http://fluxnet.fluxdata.org/}{Urbanflux}) currently there are 26 operational flux towers in urban areas instrumented for assessment of the net radiation budget.
+In addition, many other datasets of urban energy balances from inactive and decommissioned towers and from campaigns not indexed by Urbanflux are available for analysis via this method.
+These datasets span several continents and climatic zones, and characterize energy balances for cities over a range of developing and developed nations, with many operational over very long (5+ years) time scales.
+Thus, the application of this method over existing untapped datasets, and as a component of future energy balance assessments, can not only could provide a wealth of novel urban T\textsubscript{surf} and sUHI analyses for urban climatology and could also can provide for analysis outside the traditional purview of urban climatology.
+The character of the urban effect on land T\textsubscript{surf} (and climate at large) has complex, multifaceted relationships with the layout, size, population density, and development regimes of a given city as well as factors that are more difficult to observe, including social structure and culture.
+Long term, time continuous analysis of urban T\textsubscript{surf} and sUHI can help to better understand these links.
 
@@
 
-Broadly, the method, its parameterization, and the analysis presented in this study was conducted to provide for novel analyses of the urban effect on T\textsubscript{surf} and to prompt further critical assessment of the urban T\textsubscript{surf} and sUHI record.
-work towards addressing methodological and conceptual questions outlined in \citet{Roth1989}.
+Broadly, the method, its parameterization, and the analysis presented in this study were conducted to provide for novel analyses of the urban effect on T\textsubscript{surf} and to prompt further critical assessment of the urban T\textsubscript{surf} and sUHI record.
+In addition, this method was developed in order to contribute to addressing methodological and conceptual questions outlined in \citet{Roth1989} by filling knowledge gaps in the urban T\textsubscript{surf} and sUHI record.
+These analyses help to understand the temporal and geometric character of urban T\textsubscript{surf} and to facilitate more effective and representative study of the urban effect on T\textsubscript{surf}.
 Indeed, if similar conceptual flaws and shortcomings to those found for the aUHI in \citet{Stewart2011} are likely in sUHI analysis, quantification of instrument and methodological biases and shortcomings in remote sensing of urban T\textsubscript{surf} and the introduction of new methods to overcome these shortcomings may bring us closer to uncovering the true geometric, temporal, and spatial effects of cities on land surface temperature and on the climate at large.

### 2017-05-11 introduction.tex
Add files via upload

@@
 In any given observation of climate, myriad urban and non urban signals make it difficult to reliably differentiate urban influenced measures from those free of urban influence, degrading interstudy comparison.
-Development of physically based numerical models to represent physical processes affected by urbanization, and which permit experimental control, have shown promise \citep{Krayenhoff2014,Gastellu-Etchegorry1996,Masson2000,Martilli2002} yet the inherent complexity in simulating flows of energy, momentum and mass at the scales necessary to represent the full range of physical processes over realistic environments remains an on-going challenge.
+Development of physically based numerical models to represent physical processes affected by urbanization and which permit experimental control have shown promise \citep{Krayenhoff2014,Gastellu-Etchegorry1996,Masson2000,Martilli2002} yet the inherent complexity in simulating flows of energy, momentum, and mass at the scales necessary to represent the full range of physical processes over realistic environments remains an on-going challenge.
 Thus, accurate assessment of how a city affects the climate is a deceptively challenging task.
@@
 V_{M, U} - V_{M, R} = (V_B + V_L + V_H) - (V_B + V_L) = V_H
+\label{lowry}
 \end{equation}
@@
 In a points based assessment of 190 aUHI studies published between 1950 and 2007, \citet{Stewart2011} found that nearly 50\% of studies are "scientifically indefensible".
-In his analysis, each study was given a "passing" or "failing" grade and ranked based on methodological quality in the following criteria: conceptual model, operational definitions, instrument specification, site metadata, site representativeness, number of replicates, weather control, surface control, and synchronisity.
+In the analysis, each study was given a "passing" or "failing" grade and ranked based on methodological quality in the following criteria: conceptual model, operational definitions, instrument specification, site metadata, site representativeness, number of replicates, weather control, surface control, and synchronisity.
 Of the 190 surveyed studies, only 13\% were derived from field sites that were sufficiently representative of the local-scale environment or lacked the meta-data to make such a determination, leading to potential biases from micro-scale terrain.
@@
 Conceptual problems in the aUHI literature are particularly discouraging as they are not the result of an evolution in measurement techniques or improved instrumentation.
-Techniques to observe and analyze aUHI did not see significant evolution in the period between 1950 and 2007 - in fact, \citet{Stewart2011} suggests the opposite: a disproportionately large number of studies deemed scientifically sound were published early in the study purview.
+Techniques to observe and analyze aUHI did not see significant evolution or improvement in the period between 1950 and 2007 - in fact, \citet{Stewart2011} suggests the opposite: a disproportionately large number of studies deemed scientifically sound were published early in the study purview.
 In the five years since its publication, aUHI study has seen further expansion and, as such, it is difficult to assess whether its critical analysis has prompted a shift towards more thoughtful, careful, and methodical study of aUHI, notwithstanding its relevance in study of the urban effect on other meteorological variables.
-Thus, in study of the urban effect on variables for which significant methodological shifts have or will occur, the warnings implicit in \citet{Stewart2011} are particularly salient, as inherent difficulties found in translating "Lowry"-esque conceptual models to real world observations may be compounded by significant methodological changes.
+Thus, in study of the urban effect on variables for which significant methodological shifts have or will occur, the warnings implicit in \citet{Stewart2011} are particularly salient, as difficulties found in translating "Lowry"-esque conceptual models to real world observations may be compounded by significant methodological changes.
 What is clear in light of these findings, however, is that assessment of the urban effect on climate is \textit{fundamentally difficult} in spite of its conceptual simplicity.
@@
 As such, methodological problems in sUHI study can arise from both conceptual flaws \textit{and} instrument biases.
-In spite of these difficulties, the introduction of new methods, instruments, and analytical techniques in remote sensing of thermal infrared radiation (TIR) upwelling from urban environments has lead to a rapid proliferation of study of urban T\textsubscript{surf} and the sUHI - much of which is focused on remote sensing of TIR.
+In spite of these difficulties, the introduction of new methods, instruments, and analytical techniques in remote sensing of upwelling thermal infrared radiation (TIR) has lead to a rapid proliferation of study of urban T\textsubscript{surf} and the sUHI - much of which is focused on remote sensing of TIR.
 In the decades preceding this thesis, a combination of factors has lead to an increased focus on remote sensed study of urban T\textsubscript{surf}:
@@
 	\item An increased focus on the importance of the surface in both determining and understanding key near-ground micro-meteorological phenomena.
-	\item Interest in T\textsubscript{surf} for human thermal comfort, energy conservation, and sUHI mitigation applications. 
+	\item Interest in urban T\textsubscript{surf} for human thermal comfort, energy conservation, and sUHI mitigation applications. 
 \end{itemize}
@@
 However, the methodological and conceptual concerns raised in \citet{Stewart2011} in combination with the wide (and broadening) range of methodologies for remote sensed urban T\textsubscript{surf} retrieval should prompt critical assessment of the remote sensed sUHI literature.
-At the time of writing, no systemic or critical reviews of the remote sensed sUHI literature are available and questions raised in critical assessments of remote sensed urban T\textsubscript{surf} in \cite{Roth1989} and \cite{Voogt2003} have gone largely unaddressed.
-An increasingly dogmatic focus on improving spatial and radiometric resolutions has left other biases inherent in remote sensing of urban environments largely ignored in sUHI literature.
+At the time of writing, no systemic or critical reviews of the remote sensed sUHI literature are available, and questions raised in critical assessments of remote sensed urban T\textsubscript{surf} in \cite{Roth1989} and \cite{Voogt2003} have gone largely unaddressed.
+An increasingly dogmatic focus on improving spatial and radiometric resolutions has left other biases in remote sensing of urban environments largely ignored in sUHI literature.
 This suggests that similar conclusions to those found of aUHI literature in \citet{Stewart2011} are possible (if not likely) for study of sUHI.
@@
 walls, rooftops, or roads).
-The magnitude of urban effective anisotropy can reach up to 10 \si{\kelvin} and is highly dependent on surface-sensor-sun geometry, surface structure, and urban materials \citep{Krayenhoff2016, Voogt1997}.
+The magnitude of urban effective anisotropy can reach up to 10 to 12 \si{\kelvin} and is highly dependent on surface-sensor-sun geometry, surface structure, and urban materials \citep{Krayenhoff2016, Voogt1997,Lagouarde2012}.
 
@@
 
-Geometric undersampling by narrow-FOV remote sensors results in remote sensed urban T\textsubscript{surf} changing based on a sensor's viewing direction and the assemblage of facets included within the sensor FOV.
+Geometric undersampling by narrow-FOV remote sensors results in differences in remote sensed urban T\textsubscript{surf} based on a sensor's viewing direction and the assemblage of facets included within the sensor FOV.
 A sensor viewing the surface perpendicular (parallel) to the sun's angle will tend to underestimate (overestimate) the "true" complete urban surface temperature - often calculated as an area weighted average of wall, rooftop, and road T\textsubscript{surf} - by differentially biasing sunlit or shaded facets.
-Similarly, a sensor in the nadir will tend to overestimate daytime urban T\textsubscript{surf} and underestimate nighttime T\textsubscript{surf}, from a bias towards differentially hot (cool) rooftop facets by day (night) and a neglect of wall facets which are differentially cooler (warmer) by day (night).
-As the effect of geometric biases is highly dependent on surface-sensor-sun geometry and requires significant instrumentation for quantification, its influence in the remote sensed sUHI record is presently unknown.
+Similarly, a sensor in the nadir will tend to overestimate daytime urban T\textsubscript{surf} and underestimate nighttime T\textsubscript{surf}, from a bias towards rooftop facets that are differentially hot by day and cool by night and a neglect of wall facets which are differentially cooler by day and warmer by night.
+As the effect of geometric biases is highly dependent on surface-sensor-sun geometry and requires significant instrumentation for observation, its influence in the remote sensed sUHI record is presently unknown.
 
@@
 
-Clouds absorb TIR - thus satellite TIR remote sensing of the surface requires clear sky conditions.
+Clouds absorb TIR.
+Thus, satellite TIR remote sensing of the surface requires clear sky conditions.
 Although the urban effect on T\textsubscript{surf} is most evident under "satellite friendly" clear sky conditions (manifesting as large sUHI magnitudes), a clear sky bias likely entails overestimation of "all-sky" sUHI and further adds to discontinuities in the remote sensed urban T\textsubscript{surf} record.
@@
 \begin{itemize}
-	\item Develops and evaluates a method to retrieve atmospherically correction hemispherical radiometric urban surface temperatures from time-continuous measurements of upwelling longwave radiation.
+	\item Develops and evaluates a method to retrieve atmospherically corrected hemispherical radiometric urban surface temperatures from time-continuous measurements of upwelling longwave radiation.
 	\item Compares urban surface temperatures and surface urban heat island magnitudes retrieved using the method to common remote sensed representations of the urban surface.
-	\item Derives two long term climatology of hemispherical urban surface temperatures to observe seasonal and diurnal patterns of the surface urban heat island effect.
+	\item Derives two long term climatologies of hemispherical urban surface temperatures to observe seasonal and diurnal patterns of the surface urban heat island effect in Basel, Switerland and Vancouver, Canada.
 \end{itemize}

### 2017-05-11 paper1.tex
Add files via upload

@@
 
-Thermal infrared (TIR) remote sensing of land surface temperature (T\textsubscript{surf}) has emerged as a primary research focus in climatology, as researchers seek to better describe spatiotemporal patterns of T\textsubscript{surf} globally and better understand how anthropogenic modification of earth's surface influences land T\textsubscript{surf} and impacts climate at various scales.
-Over the last two decades, the application of thermal remote sensing to study surface climates has expanded significantly.
+TIR remote sensing of land surface temperature has emerged as a primary research focus in climatology, as researchers seek to better describe spatiotemporal patterns of T\textsubscript{surf} globally and better understand how anthropogenic modification of earth's surface influences land T\textsubscript{surf} and impacts climate at various scales.
+Over the last two decades, application of thermal remote sensing to study surface climates has expanded significantly.
 Thermal remote sensing of earth's surface has applications over a wide range of disciplines: from informing micro-, urban-, and global-scale climate models, to aiding decision making and mitigation praxis with respect to climate change and the urban heat island effect.
@@
 At both short and long path lengths, radiance in each band shows some deviation from the ground emitted signal.
-Larger deviations are observed at longer path lengths, where a significant portion of the resulting signal is emitted from that atmosphere, even in bands with high spectral transmittance.
+Larger deviations are observed at longer path lengths, where a significant portion of the resulting signal is emitted from the atmosphere, even in bands with high spectral transmittance.
 Long path lengths and a less transmissive atmosphere increase the potential for deviation in a remote sensed TIR signal from the ground emitted signal, while the difference between T\textsubscript{air} and T\textsubscript{surf} determines the resulting magnitude of atmospheric influence.
@@
 	\includegraphics[width=\textwidth]{spectairtsurf}
-	\caption{At-sensor spectral directional radiances computed in MODTRAN 4.1 (Berk et al., 1987) for short (30 \si{m}) and long (3000 \si{m}) path lengths (z) with Planck curves indicating spectral radiances at T\textsubscript{surf} = 300 \si{K} and T\textsubscript{air} = 290 \si{K}.}
+	\caption{At-sensor spectral directional radiances computed in MODTRAN 4.1 (Berk et al., 1987) for short (30 \si{m}) and long (3000 \si{m}) path lengths ($z$) with Planck curves indicating spectral radiances at T\textsubscript{surf} = 300 \si{K} and T\textsubscript{air} = 290 \si{K}.}
 	\label{spectairtsurf}
@@
 Inversions of uncorrected $ L_z $ or \textbf{$L_z'$} yield a temperature equal to that of a blackbody emitting the same amount of radiation as detected by the sensor.
-Since $ L_z $ is unlikely to be equal to $ L_0 $ and, by extension,  $L_z'$ is unlikely to be equal to $L_0'$, T\textsubscript{hem, b} at $ z = 0 $ and T\textsubscript{hem, b} at $ z $ often show significant deviation based on sensor characteristics, ambient conditions, and surface-sensor geometry.
+Since $ L_z $ is unlikely to be equal to $ L_0 $ and, by extension,  $L_z'$ is unlikely to be equal to $L_0'$, T\textsubscript{hem, b} at $ z = 0 $ and T\textsubscript{hem, b} at $ z $ often show significant deviation.
 Hence, T\textsubscript{bright} and T\textsubscript{hem, b} are generally considered only a rough approximation representation of radiometric T\textsubscript{surf}.
@@
 Boundary conditions are standard across most correction methods: generally requiring vertical profiles of T\textsubscript{air}, humidity, pressure, and aerosol content to remove atmospheric effects, and surface radiative properties to correct for emissivity effects.
-However, correction methods are often instrument (or, at the very least, least platform) specific and difficult to generalize across sensor and platform types.
+However, correction methods are often instrument (or, at the very least, platform) specific and difficult to generalize across sensor and platform types.
 Few methods exist for correction of ground based remote sensed TIR --- none of which are robust enough to correct irradiances measured from a near-ground wide-field-of-view (FOV) radiometer mounted to view rough terrain.
@@
 Path lengths were then calculated as the distance between each pixel's corresponding location on the DBM and the sensor represented as the vanishing point of a 3-dimensional pyramidic projection from the sensor's location in the DBM.
-A pixel-by-pixel correction was then applied to remove atmospheric effects and retrieve T\textsubscript{rad} for each pixel at 30 minute intervals, resulting in a time-continuous climatology of urban T\textsubscript{rad}.
+A pixel-by-pixel correction was then applied to remove atmospheric effects and retrieve T\textsubscript{rad} for each pixel at 30 minute intervals, resulting in a brief, time-continuous climatology of urban T\textsubscript{rad}.
 However, the method uses thermal images in conjunction with a DBM to calculate path length geometries for each pixel's LOS - a technique not possible with a pyrgeometer, which returns a single integrated value over the sensor FOV.
@@
 The resulting lookup table (LUT) of values is then used to correct $L_z$ and quantify atmospheric effects on remote sensed $L_z$ measured from several sensor heights.
-Methods outlined in \citet{Meier2011} and \citet{Kotani2009a} are limited to narrow-FOV thermal imagers and planar terrain respectively.
-A method which combines \citet{Meier2011}'s representation of complex surface geometry and \citep{Kotani2009a}'s broadband hemispherical integration is needed for atmospheric correction of urban TIR measured via downward facing pyrgeometer.
+
+The two existing methods are limited to narrow-FOV thermal imagers and planar terrain respectively.
+Thus, for atmospheric correction of TIR upwelling from urban terrain measured via downward facing pyrgeometer a method which combines \citet{Meier2011}'s representation of complex surface geometry and \citet{Kotani2009a}'s broadband hemispherical integration is needed.
 
@@
 Finally, for each time step, measured irradiances are matched with the closest modeled irradiances in the LUT to return an atmospherically corrected radiometric hemispherical surface temperature.
-This process is repeated at 30 minute intervals to yield a continuous climatology of urban T\textsubscript{hem, r} for surface urban heat island (sUHI) analysis.
+This process is repeated at 30 minute intervals to yield a continuous climatology of urban T\textsubscript{hem, r}.
 The following sections introduce the study area for which the method was developed and describe the sensor view model, radiative transfer, and post-processing steps of the correction method.
@@
 Thus, the morphological parameters identified in \cite{Rotach2005} are representative of a majority of the pyrgeometer footprint.
-Vegetation was not included in the DBM, and thus is not represented in morphological assessment or the sensor view model, however, the street canyon and surrounding area has little vegetation.
+Vegetation was not included in the DBM, and is not represented in morphological assessment or the sensor view model, however, the street canyon and surrounding area has little vegetation.
 
 For the eight month period between December 2001 and July 2002, a triangular lattice tower was installed within the Sperrstrasse street canyon.
-It's location was skewed towards the southeast facing wall near the along-canyon center of the canyon.
+Its location was skewed towards the southeast facing wall near the along-canyon center of the canyon.
 Instruments to observe a full suite of meteorological variables and fluxes of heat, mass, and momentum are mounted at various levels on the tower.
@@
 2.
-Inclusion of T\textsubscript{facet} over the IOP allows for investigation of the effect of sensor-surface geometry on remote sensed urban T\textsubscript{surf} for common urban T\textsubscript{surf} measurement strategies.
-This directional dependence of urban T\textsubscript{surf} - termed "effective anisotropy" - refers to the notion that a TIR signal upwelling from rough terrain viewed from a remote sensor changes based on the combined effects of surface-sensor-sun geometry and the convoluted, 3-dimensional structure of the urban surface \cite{Voogt1998a}.
-Facet surface temperatures measured during the summertime IOP allow for direct climatological comparison of T\textsubscript{hem, r} to common remote sensed representations of the urban surface constructed from weighted averages of wall (T\textsubscript{wall}), road (T\textsubscript{road}), and roof (T\textsubscript{roof}) temperatures.
+Inclusion of T\textsubscript{facet} over the IOP allows for investigation of the effect of sensor FOV and viewing direction on remote sensed urban T\textsubscript{surf} for common methods for urban T\textsubscript{surf} retrieval.
+This directional dependence of urban T\textsubscript{surf} - termed "effective anisotropy" - refers to the notion that a remote sensed TIR signal can vary based on the combined effects of surface-sensor-sun geometry and the convoluted, 3-dimensional structure of the urban surface \cite{Voogt1998a}.
+Facet surface temperatures measured during the summertime IOP allow for direct climatological comparison of T\textsubscript{hem, r} to common remote sensed representations of the urban surface calculated from weighted averages of wall (T\textsubscript{wall}), road (T\textsubscript{road}), and roof (T\textsubscript{roof}) temperatures.
 Plan and complete aspect ratios are used to derive weightings for nadir remote sensed (T\textsubscript{plan}) and complete (T\textsubscript{comp}) representations of urban surface temperature - described in Table \ref{weightings}.
-At each time step, T\textsubscript{hem, r} and T\textsubscript{plan} were divided by T\textsubscript{comp} and averaged over the IOP to yield normalized mean T\textsubscript{hem, r} and T\textsubscript{plan} at 30-minute intervals.
+T\textsubscript{comp} represents a complete urban T\textsubscript{surf}, where facet temperatures are averaged based on their proportion of the complete urban surface area, while T\textsubscript{plan} represents the Sperrstrasse site as viewed by a narrow-FOV remote sensor in the nadir.
+To facilitate comparison over the IOP, T\textsubscript{hem, r} and T\textsubscript{plan} were divided by T\textsubscript{comp} and averaged over the IOP to yield normalized mean T\textsubscript{hem, r} and T\textsubscript{plan} at 30-minute intervals.
 Through comparison of T\textsubscript{hem, r} to T\textsubscript{plan} and T\textsubscript{comp} we investigate the effect of sensor-surface geometry on remote sensed urban T\textsubscript{surf} and quantify directional biases in common urban T\textsubscript{surf} measurements.
@@
 	\centering
-	\caption{Weights applied to individual facet temperature components in order to calculate complete and nadir temperatures for the Basel Sperrstrasse street canyon.}
+	\caption{Weights applied to individual facet temperature components in to calculate complete and nadir temperatures for the Basel Sperrstrasse street canyon.}
 	\label{weightings}
@@
 In this method, to improve efficiency, radiances are calculated for azimuthally averaged path lengths that represent average surface-sensor geometry for each solid angle "slice" of the sensor FOV.
-Thus, hemispherical radiative transfer is reduced to a 2-dimensional problem as shown in Figure \ref{23radtran}.
+Thus, hemispherical radiative transfer is reduced to a 2-dimensional problem similar to that shown in Figure \ref{23radtran}.
 This greatly reduces the computational time required to model each irradiance - T\textsubscript{hem, r} pairing, as angular radiances can be computed as a function of zenith angle alone and subsequently weighted and integrated 3-dimensionally over the hemisphere.
 
-To calculate surface-sensor geometries, the Surface-Sensor-Sun Urban Model (SUM) \citep{Soux2004} is initialized with a simplified, orthogonal 3-dimensional DBM representing surface geometry of the area surrounding the sensor.
+To calculate surface-sensor geometries, the Surface-Sensor-Sun Urban Model (SUM) \citep{Soux2004} is initialized with a simplified, orthogonal 3-dimensional DBM that represents surface geometry of the surrounding area.
 SUM uses a four-dimensional array to represent surface morphology with three spatial dimensions ($x$, $y$, and $z$), with $z$ representing height above the $x$, $y$ plane.
@@
 Path lengths are binned at 5$^{\circ}$ increments of zenith angle and averaged to return an azimuthally-independent mean path length for each bin.
-In addition, while simulating path length geometries, SUM calculates view factors for each solid angle "slice", which are later used to weight 2-dimensional angular radiances in the hemispherical integration post-processing steps.
+In addition, while retrieving path length geometries, SUM calculates view factors for each solid angle "slice", which are later used to weight angular radiances in the hemispherical integration post-processing steps.
 
@@
 	\caption{Path lengths over an urban surface.
-a) A downward radiometer mounted above a simplified three-dimensional urban surface showing b) the component radiances and path lengths from a vertical slice.
+Left: a downward radiometer mounted above a simplified three-dimensional urban surface (Left) and the component radiances and path lengths from a vertical slice (Right).
 Dotted lines indicate radiation streams subject to absorption by the intervening atmospheric layer.}
@@
 In this method, a "typical" longwave bandpass, approximately 250 - 2300 \si{\centi\meter^{-1}} (4 - 42 \si{\micro\meter}), was extended to include much smaller wavenumber (longer wavelengths).
-This was done for two reasons: 1) To accurately represent broadband spectral longwave emission curves, which show significant emittance in wavenumber smaller than 250 \si{cm^{-1}} (wavelengths much longer than 42 \si{\micro\meter}).
+This was done for two reasons: 1) To accurately represent broadband spectral longwave emission curves, which show significant emittance in wavenumber smaller than 250 \si{cm^{-1}} (wavelengths longer than 42 \si{\micro\meter}).
 2) To replicate the spectral signal "seen" by a silicone-domed pyrgeometer, which continues to transmit radiation at wavenumber \textless 250 \si{cm^{-1}} (wavelengths \textgreater 42 \si{\micro\meter}).
@@
 
-\noindent where $ R(\lambda) $ is spectral radiance computed from a Planck function at an approximated emitter temperature and $r (\theta)$ is spectral sensor response.
+\noindent where $ R(\lambda) $ is spectral radiance computed from a Planck function at an approximated emitter temperature and $r (\lambda)$ is spectral sensor response.
 
@@
 The tower, installed over a cultivated field as a part of the Baseline Surface Radiation Network (BSRN), is located approximately 100 \si{\kilo\meter} southwest of the BUBBLE Sperrstrasse tower and is subject to similar summertime conditions as the study site.
-The study period was chosen to include a comprehensive range of T\textsubscript{surf}, T\textsubscript{air}, humidities, and cloud coverages to represent typical summertime conditions over which atmospheric corrections show the greatest variation.
+The study period was chosen to include a comprehensive range of T\textsubscript{surf}, T\textsubscript{air}, humidities, and cloud coverages to represent typical summertime conditions over which atmospheric effects on TIR can vary significantly.
 To evaluate the method, we used a modified version of the workflow described in Figure \ref{flow}.
 First, azimuthally averaged path lengths over a flat surface were calculated in SUM at 5\si{\degree} increments of zenith angle over the sensor FOV for the 10 \si{\meter} and 30 \si{\meter} sensors.
-For each time step, T\textsubscript{hem, b}, calculated from the lowest $L_z$ measurement, was used to model irradiances at 10 \si{m} and 30 \si{m} at 30 \si{\minute} intervals using concurrent profiles of T\textsubscript{air} and humidity.
-Daytime T\textsubscript{air} observed at the highest measurement level was anomalously high at all times of day compared to typical summertime lapse rates above flat vegetated terrain \citep{Oke1987}.
-This warm bias in T\textsubscript{air} measured at 30 \si{\meter} is observed even under high wind velocities, where one would expect little vertical difference in temperature not a strong temperature inversion.
-The fact that the observed warming bias remains constant over the 14-day study period prompted a slight modification to ensure the T\textsubscript{air} profile between 10 \si{\meter} and 30 \si{\meter} was representative of the neutral or weakly unstable daytime conditions and weakly stable nighttime conditions.
-It is likely that these modifications better represent the T\textsubscript{air} profile in the layer between 10 and 30 \si{\meter} than a simple interpolation between the 10 \si{\meter} T\textsubscript{air} measurement and the anomalously warm 30 \si{\meter} T\textsubscript{air} measurement.
+For each time step, T\textsubscript{hem, b} calculated from the lowest $L_z$ measurement was used to model irradiances at 10 \si{m} and 30 \si{m} at 30 \si{\minute} intervals using concurrent profiles of T\textsubscript{air} and humidity.
+Daytime T\textsubscript{air} observed at the highest measurement level was anomalously high compared to typical summertime lapse rates above flat vegetated terrain \citep{Oke1987}.
+This warm bias in T\textsubscript{air} measured at 30 \si{\meter} is observed even under high wind velocities, where one would expect little vertical difference in temperature and certainly not a strong temperature inversion.
+The fact that the observed warming bias remains relatively constant over the 14-day study period prompted a slight modification to ensure the T\textsubscript{air} profile between 10 \si{\meter} and 30 \si{\meter} was representative of the neutral or weakly unstable daytime conditions.
+It is likely that these modifications better represent the actual T\textsubscript{air} profile in the layer between 10 \si{\meter} and 30 \si{\meter} than a simple interpolation between the 10 \si{\meter} T\textsubscript{air} measurement and the anomalously warm 30 \si{\meter} T\textsubscript{air} measurement.
+
 Contrary to the findings in \citet{Kotani2009a}, an isothermal atmospheric profile was not sufficient to accurately model upwelling fluxes over flat terrain.
-Although thermal stratification is relatively small by day in an urban canyon \citep{Nakamura1988} - where strong microscale contrasts in T\textsubscript{surf} foster strong canyon mixing and neutral stability - the large daytime T\textsubscript{surf} - T\textsubscript{air} differential and the path-length/transmittance gradient can create large differences in upwelling longwave forced by inter-canyon thermal contrasts - a phenomenon termed radiative divergence.
-By underestimating near-surface air temperatures and overestimating roughness-layer T\textsubscript{air}, daytime divergences are underestimated by an isothermal profile.
-As such, a full canyon T\textsubscript{air} and humidity profile is prefered to most accurately model near ground fluxes.
+Although thermal stratification is relatively small by day in an urban canyon \citep{Nakamura1988} --- where strong microscale contrasts in T\textsubscript{surf} foster strong canyon mixing and neutral stability --- the large daytime T\textsubscript{surf} - T\textsubscript{air} differential and the path length/transmittance gradient can create large differences in upwelling longwave forced by inter-canyon thermal contrasts - a phenomenon termed radiative divergence.
+By underestimating near-surface T\textsubscript{air} and overestimating roughness-layer T\textsubscript{air}, daytime divergences are underestimated by an isothermal profile.
+As such, a full canyon T\textsubscript{air} and humidity profile is preferred to most accurately model near ground fluxes.
 
 Modeled and measured upwelling longwave at 10 \si{m} and 30 \si{m} fluxes and divergences - calculated as the difference between 30 \si{\meter} and 10 \si{\meter} fluxes - are compared in Figure \ref{fluxes}.
-Both fluxes and divergences show strong correlation, thus we conclude: First, that Irradiances measured near screen level (approximately 2 \si{\meter} above ground level) do not include significant atmospheric influence - as T\textsubscript{bright} inferred from $L_{z=2~\si{\meter}}$ was sufficient for modeling irradiances at 10 \si{m} and 30 \si{m}.
-As such, T\textsubscript{bright} and T\textsubscript{hem, r} are reasonably equal when $z$ = 2 \si{m}.
+Both fluxes and divergences show strong correlation, thus we conclude: First, irradiances measured near screen level (approximately 2 \si{\meter} above ground level) do not include significant atmospheric influence - as T\textsubscript{hem, b} inferred from $L_{z=2~\si{\meter}}$ was sufficient for modeling irradiances at 10 \si{m} and 30 \si{m}.
+As such, T\textsubscript{hem, b} and T\textsubscript{hem, r} are reasonably equal when $z$ = 2 \si{m}.
 However, it should be noted that in urban areas a 2 \si{m} sensor is not sufficiently representative of canyon geometry and should not be used to derive urban T\textsubscript{hem, r}.
-Second, MODTRAN can accurately model longwave fluxes and divergences above a flat, homogeneous surface over a wide range of temperatures, humidities, and cloud coverages and large resulting contrasts in near-ground stability.
+Second, MODTRAN can accurately model longwave fluxes and divergences above a flat, homogeneous surface over a wide range of temperatures, humidities, and cloud coverages and large contrasts in near-ground stability.
 At 30 \si{\meter} under typical summertime profiles of humidity, CO\textsubscript{2}, and O\textsubscript{3}, view factor weighted hemispherical transmittance over a flat surface is approximately 55\% - meaning 45\% of a broadband hemispherical TIR signal at 30 \si{\meter} is emitted by the atmosphere, rather than the target surface.
 As such, it is imperative that T\textsubscript{air} profiles are well represented in MODTRAN to accurately model the atmospheric component of remote sensed TIR signal.
-As profiles of T\textsubscript{air} in the layer between the surface and the sensor at the Sperrstrasse site are generally subject to neutral or weakly unstable conditions, with only rare stable stratifications, urban longwave divergences are likely to be smaller than those observed at Payerne - and well approximated by MODTRAN.
+As profiles of T\textsubscript{air} in the layer between the surface and the sensor at the Sperrstrasse site are generally subject to neutral or weakly unstable conditions, with only rare stable stratifications, urban longwave divergences are likely to be smaller than those observed in this evaluation - and well approximated by MODTRAN.
 Thus, provided path length geometries are accurately replicated in by the sensor view model, we can reasonably assume that the method has similar effectiveness at the urban site.
@@
 Over both summer and winter seasons, correction magnitudes are strongly correlated with $\Delta$T\textsubscript{hem, r - air}.
-Figure \ref{season_fit} shows the effect of seasonality and cloud cover on the relationship between correction magnitudes and $\Delta$T\textsubscript{hem, r - air}.
+Figures \ref{season_sw} and \ref{season_sc} show the effect of seasonality and cloud cover on the relationship between correction magnitudes and $\Delta$T\textsubscript{hem, r - air}.
 The relationship is strongest and most evident in summer when variations in cloud cover can lead to large day-to-day contrasts in solar input and $\Delta$T\textsubscript{surf - air}, which translates into a wide range of correction magnitudes.
 
-Figure \ref{nofit} illustrates the relationship between correction magnitude and meteorological variables commonly measured as a part of climatological assessments of urban environments.
+Figures \ref{meteo_k_hem} and \ref{meteo_air_hum}
+ illustrate the relationship between correction magnitude and meteorological variables commonly measured as a part of climatological assessments of urban environments.
 A strong positive relationship is observed between correction magnitudes and incoming solar radiation.
@@
 Figure \ref{humtest} shows hemispherical atmospheric transmittance as a function of vapor pressure for pyrgeometers at 10 \si{\meter} and 30 \si{\meter} above a flat surface.
-Hemispherical transmittances are modeled in MODTRAN 4.1 as the fraction of surface emission reaching the sensor for the path length and view factor geometries calculated in the method evaluation in Section \ref{sensitivity}.
+Transmittances are modeled in MODTRAN 4.1 as the fraction of surface emission reaching the sensor for the path length and view factor geometries calculated in the method evaluation in Section \ref{sensitivity}.
 Hemispherical transmittance falls sharply with increasing vapor pressure between 0 and 5 \si{\hecto\pascal}, but is relatively constant with vapor pressure \textgreater 5 \si{\hecto\pascal}.
-This results in the weak correlation observed in Figure \ref{nofit} between humidity and correction magnitude.
+This results in the weak correlation observed in Figure \ref{meteo_air_hum} between humidity and correction magnitude.
 
@@
 	height=7in,
-	keepaspectratio]{season_fit}
-	\caption{Correction magnitude versus $\Delta$T\textsubscript{plan - hem, r} binned first for season and then for cloud cover in the summer months.}
-	\label{season_fit}
+	keepaspectratio]{season_sw}
+	\caption{Correction magnitude versus $\Delta$T\textsubscript{hem, r - air} binned for season.}
+	\label{season_sw}
 \end{figure}
@@
 	height=7in,
-	keepaspectratio]{nofit}
-	\caption{Correction magnitude versus meteorological variables for the daytime hours of the study period.}
-	\label{nofit}
+	keepaspectratio]{season_sc}
+	\caption{Correction magnitude versus $\Delta$T\textsubscript{hem, r - air} binned for clear and clouded days over the summer months.}
+	\label{season_sc}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=15cm,
+	height=7in,
+	keepaspectratio]{meteo_k_hem}
+	\caption{Correction magnitude versus incoming shortwave and T\textsubscript{hem, r} for the daytime hours of the study period.}
+	\label{meteo_k_hem}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=15cm,
+	height=7in,
+	keepaspectratio]{meteo_air_hum}
+	\caption{Correction magnitude versus T\textsubscript{air} and water vapor content for the daytime hours of the study period.}
+	\label{meteo_air_hum}
 \end{figure}
@@
 To illustrate the effect of sensor-surface geometry on remote sensed T\textsubscript{surf}, Figure \ref{tc_nadhem} shows a comparison of mean normalized T\textsubscript{hem, r} and T\textsubscript{plan} calculated as T\textsubscript{x} divided by T\textsubscript{comp}, where T\textsubscript{x} is T\textsubscript{hem, r} or T\textsubscript{plan}, and averaged over the IOP.
-In the figure, overestimation of T\textsubscript{comp} is observed at values above $1$ and underestimation by values below $1$.
+In the figure, overestimation of T\textsubscript{comp} is observed at values above one and underestimation by values below one.
 Both T\textsubscript{hem, r} and T\textsubscript{plan} overestimate T\textsubscript{comp} by day and underestimate T\textsubscript{comp} by night.
@@
 
-Of the meteorological variables tested, results show that correction magnitudes are most strongly dependent on solar input.
+Of the meteorological variables tested, correction magnitudes are most strongly dependent on solar input.
 Solar radiation incident on the surface forces microscale thermal contrasts between the surface and the air, directly controlling the T\textsubscript{surf} - T\textsubscript{air} differential.
 The relationships between correction magnitude and T\textsubscript{hem, r} and T\textsubscript{air} individually are weaker and more complex.
-Correction magnitudes are smallest at temperatures between approximately $10$ and $20$ \si{\degreeCelsius} - largely made up of morning and cloudy observations during the summer months where thermal contrasts between T\textsubscript{surf} and T\textsubscript{air} are small or weakly negative, forcing small correction magnitudes.
+Correction magnitudes are smallest when T\textsubscript{hem, r} and T\textsubscript{air} are between approximately $10$ and $20$ \si{\degreeCelsius} - largely made up of morning and cloudy observations during the summer months where thermal contrasts between T\textsubscript{surf} and T\textsubscript{air} are small or weakly negative, forcing small correction magnitudes.
 During the winter months, a negative relationship between T\textsubscript{hem, r} and T\textsubscript{air} and correction magnitudes is observed, likely a result of the effect of shifting spectral emittance curves combined with non-uniform sensor response - discussed in section \ref{Persistence in correction magnitudes}.
-On clear sky summer days, positive relationships between surface and air temperatures and correction magnitudes is observed as an increased solar input foster higher overall temperatures and large contrasts between T\textsubscript{surf} and T\textsubscript{air}.
+On clear sky summer days, positive relationships between correction magnitudes and T\textsubscript{hem, r} and T\textsubscript{air} are observed as an increased solar input foster higher overall temperatures and large contrasts between T\textsubscript{surf} and T\textsubscript{air}.
 
@@
 A pyrgeometer's dome transmittance is spectrally non-uniform.
-Thus, the "dome effect" a pyrgeometer exerts on a given incoming TIR signal changes as a function of emission temperature (T\textsubscript{emiss}) because the shape of a given spectral radiance curve (R\textsubscript{$\nu$} (T)) is temperature dependent.
-This is illustrated in Figure \ref{nonuniform} by comparing the relative peaks and shape of a pyrgeometer dome transmittance curve for several R\textsubscript{$\nu$} (T) for a range of typical urban T\textsubscript{emiss} (T\textsubscript{emiss} = 260 - 320 \si{\kelvin}, at 5 \si{\kelvin} increments).
-As each curve interacts with a different portion of the spectral dome transmittance curve (\textit{r\textsubscript{$\nu$}}), and \textit{r\textsubscript{$\nu$}} is spectrally non-uniform, the distinct spectral shapes of R\textsubscript{$\nu$} (T) results in differential dome effects for each T\textsubscript{emiss}.
+Thus, the "dome effect" a pyrgeometer exerts on a given incoming TIR signal changes as a function of emission temperature (T\textsubscript{emiss}) because the shape of a given spectral radiance curve (R\textsubscript{$\nu$}) is temperature dependent.
+This is illustrated in Figure \ref{nonuniform} by comparing the relative peaks and shape of a pyrgeometer dome transmittance curve for several R\textsubscript{$\nu$} (T\textsubscript{emiss}) for a range of typical urban T\textsubscript{emiss} (T\textsubscript{emiss} = 260 - 320 \si{\kelvin}, at 5 \si{\kelvin} increments).
+As each curve interacts with a different portion of the spectral dome transmittance curve (\textit{r\textsubscript{$\nu$}}), and \textit{r\textsubscript{$\nu$}} is spectrally non-uniform, the distinct spectral shapes of R\textsubscript{$\nu$} results in differential dome effects for each T\textsubscript{emiss}.
 
@@
 	keepaspectratio]{nonuniform}
-	\caption{Spectral dome transmittance (\textit{r}) for a Kipp \& Zonen pyrgeometer overlaid with a Planckian spectral radiance curve at T = 300 \si{\kelvin} and the same curve shifted to represent the shape and peak of a spectral radiance curve at T = 270 \si{\kelvin}.}
+	\caption{Spectral dome transmittance for a Kipp \& Zonen pyrgeometer overlaid with a Planckian spectral radiance curve at T = 300 \si{\kelvin} and the same curve shifted to represent the shape and peak of a spectral radiance curve at T = 270 \si{\kelvin}.}
 	\label{nonuniform}
@@
 
-Spectrally non-uniform pyrgeometer dome transmittance results in differential dome effects on the R\textsubscript{$\nu$} (T) for different T\textsubscript{emiss}.
+Spectrally non-uniform pyrgeometer dome transmittance results in differential dome effects on the R\textsubscript{$\nu$} for different T\textsubscript{emiss}.
 Thus, temperature specific calibration routines designed to remove dome effects and retrieve accurate irradiance values are subject to errors based on the difference between T\textsubscript{emiss} and the temperature at which the instrument was calibrated (T\textsubscript{calib}).
-The magnitude of this error can be expressed as differentials in Planck weighted mean sensor response ($\overline{r} $) calculated via equation \ref{meansens} for different T\textsubscript{emiss} - where lower $\overline{r} $ indicates a larger dome effect on the signal received by the sensor.
+The magnitude of this error can be expressed as differentials in Planck weighted mean sensor response ($\overline{r} $) calculated via equation \ref{meansens} for different T\textsubscript{emiss} --- where lower $\overline{r} $ indicates a larger dome effect on the signal received by the sensor.
 In addition, this error can be expressed as a dome adjusted temperature bias (T\textsubscript{bias}) computed as the difference between a reference T\textsubscript{calib} and a dome adjusted temperature (T\textsubscript{adj}).
@@
 While the individual effects are difficult to separate, temperature dependent differential dome effects are visible in correction magnitudes as a persistent minor positive correction in winter.
-Relative to T\textsubscript{calib}, colder T\textsubscript{emiss} in winter results in a larger dome effects.
+Relative to T\textsubscript{calib}, lower T\textsubscript{emiss} in winter results in a larger dome effects.
 This manifests in artificially low T\textsubscript{hem, b} and increased wintertime correction magnitudes.
+
 Additionally, modeled irradiances include a reflected downwelling component, which, although small, is sensitive to T\textsubscript{air} and humidity profiles above the sensor height.
@@
 Nighttime underestimations of T\textsubscript{comp} by nadir and hemispherical views are a result of an undersampling of wall and road facets respectively.
-T\textsubscript{hem, r} derived from $L_z$ measured at the Sperrstrasse canyon oversamples the nearest roof and underestimates the canyon floor compared to T\textsubscript{comp} as it's location along the canyon axis is skewed towards the south facing wall (rather than the center of the canyon) and its height is approximately 2.17 times mean building height.
+T\textsubscript{hem, r} derived from $L_z$ measured at the Sperrstrasse canyon oversamples the nearest roof and underestimates the canyon floor compared to T\textsubscript{comp} as its location along the canyon axis is skewed towards the south facing wall (rather than the center of the canyon) and its height is approximately 2.17 times mean building height.
 Non-optimal sensor placement in the Sperrstrasse canyon results in a large underestimation of nighttime T\textsubscript{comp} by T\textsubscript{hem, r}.
@@
 
- When measured from a near ground downward facing remote sensor, upwelling TIR is sensitive to sensor height and position relative to surface geometry \citep{Adderley2015, Roberts2010}.
-Below 3 times mean roof level, remote sensed TIR varies from a "true", geometrically representative, TIR signal based on a comparison between facet view factor proportions of the sensor and those of the urban surface.
+ When measured from a near ground downward facing pyrgeometer, upwelling TIR is sensitive to sensor height and position relative to surface geometry \citep{Adderley2015, Roberts2010}.
+Below 3 times mean roof level, measured TIR varies from a "true", geometrically representative, TIR signal based on a comparison between facet view factor proportions of the sensor and those of the urban surface.
 A perfectly geometrically representative remote sensed TIR signal is one measured from a sensor with facet view factor proportions equaling those of the urban surface.
@@
 For an idealized, orthogonal, block-like urban array, \citet{Roberts2010} found that facet view factor proportions --- and thus measured TIR --- varied significantly with sensor position and height.
-To investigate the effect of sensor placement on T\textsubscript{hem, r} for a simplified representation of the Sperrstrasse canyon site, we used SUM to vary sensor position over a simplified DBM and calculated normalized view factor proportions for a 160\si{\degree} FOV sensor.
+To investigate the effect of sensor placement on T\textsubscript{hem, r} for a simplified representation of the Sperrstrasse canyon site, we used SUM to vary sensor position over a simplified DBM and calculate normalized view factor proportions for a 160\si{\degree} FOV sensor.
 For nine sensor positions, show in Figure \ref{placement}, facet view factor proportions normalized to unity are shown in Figure \ref{bars1} for a sensor height of 2.17 times mean building height and in Figure \ref{bars45} for a sensor height of 3 times mean building height.
@@
 
-Results in \citet{Adderley2015} and \citet{Roberts2010} show that sensitivity to sensor placement decreases with increasing height.
+Results in \citet{Adderley2015} and \citet{Roberts2010} show that sensitivity to sensor placement decreases as sensor height increases.
 Depending on surface geometry, sensor view factor proportions and the TIR signal "seen" by a sensor become spatially invariant sensor at heights above approximately 3 to 5 times mean building height.
@@
 
-To visualize how sensor placement affects inferred surface temperatures, T\textsubscript{hem, r} is calculated using view factors to calculate a weighted average of T\textsubscript{wall}, T\textsubscript{road}, and T\textsubscript{roof} for each sensor position.
-Temperatures are normalized by T\textsubscript{comp} and averaged at 30-minute intervals over the IOP, the results of which are shown in Figures \ref{vf_all} and \ref{vf_all45} for sensor heights of 2.17 and 3 times mean building height respectively.
-As is the case with view factor proportions, calculated T\textsubscript{hem, r} is highly dependent on sensor placement.
+To visualize how sensor placement affects T\textsubscript{hem, r} retrieved via the method detailed in this study, T\textsubscript{hem, r} is inferred using wall, roof, and road view factors calculated in SUM for the two sensor heights to compute a weighted average of T\textsubscript{wall}, T\textsubscript{road}, and T\textsubscript{roof} for each sensor position at 30 minute intervals over the IOP.
+Temperatures are normalized by T\textsubscript{comp} and averaged at each time step over the IOP, the results of which are shown in Figures \ref{vf_all} and \ref{vf_all45}.
+As is the case with view factor proportions, T\textsubscript{hem, r} is highly dependent on sensor placement.
 For the lower sensor height, as predicted by view factor proportions, positions at the center of the canyon (positions 4, 5, and 6) and edge of the building (positions 1, 4, and 7) follow T\textsubscript{comp} most closely.
@@
 For these positions, although roof view factor proportion is well represented, road view factor is significantly overestimated, oversampling warmer nighttime road temperatures.
-For all other positions daytime overestimation and nighttime underestimation of T\textsubscript{comp} is reduced.
-As is the case with view factor proportions, T\textsubscript{hem, r} is more spatially stable with respect to sensor placement.
+For all other positions, daytime overestimation and nighttime underestimation of T\textsubscript{comp} is reduced.
+Thus, as is the case with view factor proportions, at the higher test height, T\textsubscript{hem, r} is more spatially stable.
 
@@
 For each time step, measured humidity is used to infer a hemispherical transmittance ($\overline{\tau_{\Phi}}$) for the approximated surface-sensor geometry.
-A parameterized T\textsubscript{hem, r} (T\textsubscript{hem, para}) can then be calculated by using measured (in this case, profile averaged) T\textsubscript{air} to decompose measured L\textsubscript{up} into irradiance received by the sensor from the surface $L_{surf}^{at-sensor}$ and from the atmosphere $L_{atm}^{at-sensor}$ via,
+A parameterized T\textsubscript{hem, r} can then be calculated by using measured (in this case, profile averaged) T\textsubscript{air} to decompose measured upwelling longwave into irradiance received by the sensor from the surface $L_{surf}^{at-sensor}$ and from the atmosphere $L_{atm}^{at-sensor}$ via,
 
@@
 	\label{para}
-	T_{hem, ~para} =\left(\displaystyle\frac{\biggl(\cfrac{\displaystyle L_{surf}^{at-sensor}}{ \displaystyle\tau}\biggr)}{ \displaystyle\sigma}\right)^{0.25}
+	T_{hem, ~} =\left(\displaystyle\frac{\biggl(\cfrac{\displaystyle L_{surf}^{at-sensor}}{ \displaystyle\tau}\biggr)}{ \displaystyle\sigma}\right)^{0.25}
 \end{equation}
@@
 
-Using this parameterization, a climatology of T\textsubscript{hem, para} for a given surface geometry can be retrieved from approximately 100 simulations as opposed to the many thousands needed for full hemispherical radiative transfer simulation.
+Using this parameterization, a climatology of T\textsubscript{hem, r} for a given surface geometry can be retrieved from approximately 100 simulations as opposed to the many thousands needed for full hemispherical radiative transfer simulation.
 A variety of online and offline, open and closed source resources are available for radiative transfer simulation \citep{Gastellu-Etchegorry1996,Berk1987,Buehler2005}, some of which are available for free or at a low cost (Spectral-Calc), all of which are likely to provide similar transmittances to those included in Appendix \ref{appendixa}.
-Figure \ref{para_eval} and Table \ref{parastats} summarize parameterization performance relative to corrected T\textsubscript{hem, r}.
-Errors are largest during the daytime hours (shaded in grey), where the profile averaged T\textsubscript{air} overestimates above canyon T\textsubscript{air} and underestimates near-surface T\textsubscript{air}.
+Figure \ref{para_eval} and Table \ref{parastats} summarize parameterization performance relative to T\textsubscript{hem, r} retrieved via the correction method.
+Errors are largest during the daytime hours, where the profile averaged T\textsubscript{air} overestimates above canyon T\textsubscript{air} and underestimates near-surface T\textsubscript{air}.
 Neutral stability in the nighttime and early morning hours reduces these errors significantly as the canyon T\textsubscript{air} profile is approximately isothermal.
@@
 Thus, a combination of satellite, aerial, and in-situ observational methods has been integral in isolating and elucidating the urban effect on surface temperature and its spatial, temporal, and geometric characteristics.
-However, neither technological advancement nor critical analysis of these methods has provided a satisfactory answer to the questions posed in Roth et al.
-1989 viz urban thermal remote sensing.
+However, neither technological advancement nor critical analysis of these methods has provided a satisfactory answer to the questions posed in \citet{Roth1989} viz urban thermal remote sensing.
 As such, the true spatial, temporal, and geometric nature of the urban effect on surface temperature remains unknown.
@@
 Irradiance - T\textsubscript{hem, r} pairings are aggregated into a LUT and the measured irradiance is matched with the closest modeled irradiance to retrieve T\textsubscript{hem, r} for a given time step.
-Repeated at 30-minute intervals, this method is used to derive an 8-month climatology of urban T\textsubscript{hem, r} from urban irradiances measured via a pyrgeometer at 2.17 times mean building height in the Sperrstrasse canyon instrument as a part of the BUBBLE campaign in Basel, Switzerland.
-T\textsubscript{hem, r} at each time step is a view factor weighted temperature representative of the surfaces contained within the sensor FOV.
+Repeated at 30-minute intervals, the method is used to derive an 8-month climatology of urban T\textsubscript{hem, r} from urban irradiances measured as a part of the BUBBLE campaign in Basel, Switzerland.
 
 Atmospheric corrections are large and vary significantly based on sensor-surface-sun geometry and synoptic conditions.
-Thus, for accurate intersite comparison, T\textsubscript{hem} derived from irradiances measured at different heights or above different surface geometries may have different atmospheric or dome effects that can confound results.
-This fact makes clear the need for robust atmospheric correction of T\textsubscript{hem} for sUHI analysis.
-Correction magnitudes are largest on hot, clear sky summertime days --- with large solar input and high T\textsubscript{hem, r} and T\textsubscript{air} --- and smallest by night and under overcast conditions, during which solar input is suppressed and T\textsubscript{hem, r} and T\textsubscript{air} are low.
-Correction magnitudes are most strongly correlated with the T\textsubscript{surf} - T\textsubscript{air} differential and incoming solar radiation and weakly correlated with T\textsubscript{hem, r} and T\textsubscript{air}.
+Thus, T\textsubscript{hem} derived from irradiances measured at different heights or above different surface geometries may have different atmospheric or dome effects that can confound results.
+This fact makes clear the need for robust atmospheric correction of T\textsubscript{hem} for sUHI analysis or to retrieve radiometric T\textsubscript{hem}.
+Correction magnitudes are largest on hot, clear sky summertime days --- with large solar input and high T\textsubscript{surf} and T\textsubscript{air} --- and smallest by night and under overcast conditions, during which solar input is suppressed and T\textsubscript{surf} and T\textsubscript{air} are low.
+Correction magnitudes are most strongly correlated with the T\textsubscript{surf} to T\textsubscript{air} differential and incoming solar radiation and weakly correlated with T\textsubscript{hem, r} and T\textsubscript{air}.
 Water vapor content did not appear to exert strong control over correction magnitudes.
@@
 Comparison of T\textsubscript{surf} from nadir, hemispherical, and complete representations of the Basel Sperrstrasse street canyon show that a hemispherical view is more geometrically representative of the complete surface temperature than a sensor viewing in the nadir.
-However, sensor placement sensitivity tests show that T\textsubscript{hem, r} is highly dependent on sensor placement and height.
-Overestimations by T\textsubscript{plan} and T\textsubscript{hem, r} are greatest under clear sky conditions, but is particularly large for T\textsubscript{plan}.
+However, sensor placement sensitivity tests show that T\textsubscript{hem, r} varies based on sensor placement and height.
+Overestimations of T\textsubscript{comp} by T\textsubscript{plan} and T\textsubscript{hem, r} are greatest under clear sky conditions and are much larger for T\textsubscript{plan}.
 This is particularly important when one considers that satellite observations of urban T\textsubscript{surf} are only possible under clear sky conditions (when T\textsubscript{plan} is least representative of T\textsubscript{comp}).

### 2017-05-11 paper2.tex
Add files via upload

@@
 Urban modification of surface geometry and thermal, radiative, moisture, and aerodynamic properties results in differential surface heating and cooling patterns and strong microscale spatiotemporal variations in urban surface temperature (T\textsubscript{surf}).
-Integrated up to larger scales, urban areas tend to store more heat relative to non-built surroundings; traffic, HVAC systems, and industrial processes contribute to an additional anthropogenic heat input.
-In addition, a reduced pervious (and vegetative) cover in urban areas shifts convective fluxes towards sensible and away from latent.
+Integrated up to larger scales, urban areas tend to store more heat relative to non-built surroundings; and traffic, HVAC systems, and industrial processes contribute to an additional anthropogenic heat input.
+In addition, a reduced pervious (and vegetative) cover in urban areas shifts convective fluxes towards sensible heat and away from latent heat.
 These manifest in elevated T\textsubscript{surf} and T\textsubscript{air} in cities, a phenomenon termed the urban heat island effect (UHI).
 To foster a more complete understanding of the effect of urban areas on climates across a range of scales, accurate, spatiotemporally continuous and geometrically representative characterization of T\textsubscript{surf} in cities has long been a goal in urban climatology.
-The proliferation of satellite and aerial thermal infrared (TIR) remote sensing has enabled spatially-extensive characterizations of surface climates at ever improving spatial and spectral resolutions.
-Such campaigns have elucidated urban T\textsubscript{surf} and surface urban heat island (sUHI) patterns globally at large spatial scales \cite{Peng2012,Zhao2014}.
+The proliferation of satellite and aerial thermal remote sensing has enabled spatially-extensive characterizations of surface climates at ever improving spatial and spectral resolutions.
+Such campaigns have elucidated urban T\textsubscript{surf} and surface urban heat island patterns globally at large spatial scales \citep{Peng2012,Zhao2014}.
 However, technological improvements in TIR remote sensing have yet to address three potential sources of error that arise when applied in urban areas:
@@
 	\item Satellite overpass cycles and sensor sampling regimes that result in temporal discontinuity of the surface temperature time series
-	\item Clear-sky bias that contribute to an overestimation of "all-sky" surface temperatures
+	\item Clear-sky bias that contributes to an overestimation of "all-sky" surface temperatures
 \end{enumerate}
@@
 atmospheric and emissivity effects).
-Meta-analysis of the air temperature (T\textsubscript{air}) UHI literature shows that these goals are rarely satisfied \citep{Stewart2011}.
+Meta-analysis of the aUHI literature shows that these goals are rarely satisfied \citep{Stewart2011}.
 Given the relative difficulty in retrieving accurate, representative urban T\textsubscript{surf}, similar conclusions are likely for sUHI analysis.
 In spite of this fact, and the short period over which large-scale, generalizable methods for measuring urban T\textsubscript{surf} have been available, study of sUHI via TIR remote sensing has expanded significantly in the last twenty years \citep{Peng2012,Voogt2003}.
-Using the method and parameterization described in Chapter \ref{paper1} twl long-term climatologies of hemispherical radiometric urban T\textsubscript{surf} (T\textsubscript{hem, r}) were derived from time continuous near-ground upwelling TIR as measured from an inverted pyrgeometer.
+
+%Using the method and parameterization described in Chapter \ref{paper1} two long-term climatologies of hemispherical radiometric urban T\textsubscript{surf} were derived from time continuous near-ground upwelling TIR measured from an inverted pyrgeometer.
+These climatologies represent the first time-continuous seasonal analysis of the urban effect on T\textsubscript{surf}.
 The method, and an associated parameterization scheme, was developed to address and overcome biases inherent in traditional methods for urban T\textsubscript{surf} retrieval by providing temporally continuous, geometrically representative\footnote{A hemispherical view is not perfectly geometrically representative of urban surface geometry.
@@
 Compared to flat terrain, complex urban surface geometry modifies receipt of incoming solar radiation and traps a portion of reflected solar and outgoing longwave radiation.
-The resulting microscale spatiotemporal contrasts in urban T\textsubscript{surf} create a directional dependence in observed urban T\textsubscript{surf} when measured from conventional narrow-field-of-view (FOV) TIR remote sensing platforms.
+The resulting microscale spatiotemporal contrasts in urban T\textsubscript{surf} create a directional dependence in observed urban T\textsubscript{surf} when measured from conventional narrow-FOV TIR remote sensing platforms.
 Thus, remote sensed urban T\textsubscript{surf} varies based on sensor FOV, viewing angle and direction, and sun-surface geometry.
@@
 How does it vary temporally and over a full range of weather conditions?
-To address these questions, this study presents the first time-continuous, climatological analysis of sUHI retrieved from radiometric hemispherical urban T\textsubscript{surf} retrieved via the correction method described in Chapter \ref{paper1}.
-These measures are used to assess the magnitude of and overcome geometric and temporal biases inherent in urban TIR remote sensing.
+To address these questions, this study presents the first time-continuous, climatological analysis of sUHI derived from geometrically representative\footnote{A hemispherical view is not perfectly geometrically representative of urban surface geometry.
+This is best illustrated by visualizing an urban area from the perspective of a downward facing ‘fish-eye’ camera.
+Geometric sampling biases are a result of lens distortions and the sensor cosine response.
+However, by sampling the surface 3-dimensionally, T\textsubscript{hem, r} is more representative of urban geometry than conventional 2-dimensional views of the surface.
+The geometric representivity of urban T\textsubscript{hem, r} is discussed in section \ref{Methodological limitations and considerations}.} radiometric hemispherical urban T\textsubscript{surf} retrieved via the correction method described in Chapter \ref{paper1}.
+sUHI calculated for the eight month BUBBLE campaign in Basel, Switzerland and a year long subset of the EPiCC campaign in Vancouver, Canada are used to analyze the diurnal and seasonal character of sUHI and to assess the magnitude of and overcome geometric and temporal biases inherent in urban TIR remote sensing.
 
@@
 
-This section describes the study area and provides a brief overview of the multiple line-of-sight atmospheric correction method used to derive a time-continuous eight month climatology of urban T\textsubscript{hem, r} for sUHI analysis.
+This section describes the study area and provides a brief overview of the multiple line-of-sight atmospheric correction method used to derive time-continuous urban T\textsubscript{hem, r} for sUHI analysis.
 A thorough discussion of the method is included in Chapter \ref{paper1}.
@@
 A pyrgeometer's broad waveband and wide FOV increase the potential for significant atmospheric influence on an observed TIR signal.
-These effects can result in differences of up to 8 \si{\kelvin} between the 'true' radiometric T\textsubscript{surf} and brightness T\textsubscript{surf} derived from remote sensed observations of upwelling TIR.
-Atmospheric effects are particularly important when comparing irradiances or remote sensed T\textsubscript{surf} across different study sites and times, as differences in surface geometry, instrument height, and ambient conditions can introduce large spatiotemporal contrasts in the magnitude of atmospheric influence.
-This is particularly salient for near-ground, broadband radiometers, as band by band spectral atmospheric transmittance can change significantly with path length, shown in Figure \ref{spectransheight}.
+These effects can result in differences of up to 10 \si{\kelvin} between the 'true' radiometric T\textsubscript{surf} and brightness T\textsubscript{surf} derived from remote sensed observations of upwelling TIR.
+Atmospheric effects are important when comparing irradiances or remote sensed T\textsubscript{surf} across different study sites and times, as differences in surface geometry, instrument height, and ambient conditions can introduce large spatiotemporal contrasts in the magnitude of atmospheric influence and degrade comparability.
+This is particularly salient for near-ground, broadband radiometers, as band-by-band spectral atmospheric transmittance can change significantly with path length, shown in Figure \ref{spectransheight}.
 As inter-site comparison is paramount in understanding the urban effect on T\textsubscript{surf} and is the basis of sUHI analysis, accurate atmospheric correction of remote sensed T\textsubscript{surf} is of prime importance.
-Thus, urban T\textsubscript{hem, r} in this study are derived using a multiple line-of-sight correction routine which accounts for the 3-dimensionality of the urban surface, changing atmospheric conditions, and spectrally non-uniform sensor response.
+Thus, urban T\textsubscript{hem, r} in this study are derived using a multiple LOS correction routine which accounts for the 3-dimensionality of the urban surface, changing atmospheric conditions, and spectrally non-uniform sensor response.
 For each time step, the method generates a LUT of modeled at-sensor irradiances for a range of potential T\textsubscript{hem, r} using profiles of measured T\textsubscript{air} and humidity.
@@
 
-First, surface-to-sensor path length and view factor geometries are calculated with the surface-sensor-sun urban model (SUM) \citep{Soux2004}.
+First, surface-to-sensor path length and view factor geometries are calculated with the SUM sensor view model \citep{Soux2004}.
 Using a digital building model (DBM), SUM calculates path lengths from the surface to the sensor and angular view factors by projecting the pyrgeometer FOV onto a simplified representation of the surface, determining which points in the DBM are 'seen' by the sensor, and calculating the distance from each 'seen' point to the sensor.
@@
 Runs are initialized using profiles of T\textsubscript{air} and water vapor content collected concurrently at each time step.
-Spectral radiances are modeled over a bandpass of 1 to 2500 \si{cm^{-1}} at an emissivity of 0.95 and convolved by an extended dome transmittance function to ensure that modeled radiances replicate the actual remote sensed signal.
-Consultation with radiometer manufacturers made clear the need to extend the modeled bandpass beyond a typical longwave bandpass (approximately 250 - 2500 \si{cm^{-1}}), as silicone domed pyrgeometers transmit radiation at much smaller wavenumbers (longer wavelengths) than 250 \si{cm^{-1}} (40 \si{\micro\meter}).
+Spectral radiances are modeled over a bandpass of 1 to 2300 \si{cm^{-1}} at an emissivity of 0.95 and convolved by an extended dome transmittance function to ensure that modeled radiances replicate the actual remote sensed signal.
+Consultation with radiometer manufacturers made clear the need to extend the modeled bandpass beyond a typical longwave bandpass (approximately 250 - 2300 \si{cm^{-1}}), as silicone domed pyrgeometers transmit radiation at much smaller wavenumbers (longer wavelengths) than 250 \si{cm^{-1}} (40 \si{\micro\meter}).
 
@@
 
-Although MODTRAN has been show to accurately model radiative transfer of broadband longwave radiation at urban canyon scale path lengths \citep{Hoch2005, Hoch2007}, the added complexity inherent in simulating fluxes of radiation hemispherically over complex urban terrain prompted an independent evaluation of the correction method using observations of upwelling longwave radiation (measured from 2 \si{\meter}, 10 \si{\meter}, and 30 \si{\meter} above ground level) over a simple, flat surface for a continuous 16-day period.
-In the evaluation, a modified version of the method uses T\textsubscript{hem, b} inferred from the lowest L\textsubscript{up} measurement along with profiles of T\textsubscript{air} and humidity to model irradiances at 10 \si{\meter} and 30 \si{\meter} at each time step.
-Modeled and measured irradiances match well, over wide range of mid-latitude atmospheric conditions, with results summarized in Table \ref{evalstats}.
+Although MODTRAN has been show to accurately model radiative transfer of broadband longwave radiation at urban canyon scale path lengths \citep{Hoch2005, Hoch2007}, the added complexity inherent in simulating fluxes of radiation hemispherically over complex urban terrain prompted a robust evaluation of the correction method using observations of upwelling longwave radiation (measured from 2 \si{\meter}, 10 \si{\meter}, and 30 \si{\meter} above ground level) over a simple, flat surface for a continuous 14-day period.
+In the evaluation, a modified version of the correction method uses T\textsubscript{hem, b} inferred from the lowest upwelling longwave measurement along with profiles of T\textsubscript{air} and humidity to model irradiances at 10 \si{\meter} and 30 \si{\meter} at each time step.
+Modeled and measured irradiances match well over wide range of day and night mid-latitude atmospheric conditions, with results summarized in Table \ref{evalstats}.
 
-Sensor placement sensitivity testing in section \ref{sensitivity} shows that T\textsubscript{hem, r} is spatially variant when measured over urban areas at typical pyrgeometer heights.
+Sensor placement sensitivity testing in Section \ref{sensitivity} shows that T\textsubscript{hem, r} is spatially variant when measured over urban areas at typical pyrgeometer heights.
 This introduces a potential bias in T\textsubscript{hem, r} as changing sensor position relative to canyon features alters view factor proportions of rooftops, roads, and walls.
-The sensitivity of T\textsubscript{hem, r} to sensor placement decreases with height and T\textsubscript{hem, r} becomes spatially invariant at approximately 3.5 times mean building height \citep{Roberts2010,Adderley2015}.
-As such, these biases can be significantly reduced by mounting downward facing pyrgeometers slightly higher than normal.
-Although pyrgeometers at both the Sperrstrasse and are mounted below the height at which T\textsubscript{hem,r} is spatially invariant, sensitivity tests in section \ref{sensitivity} show that the magnitude of this bias is relatively small for sensor placement at the Sperrstrasse canyon.
-Similar or lesser biases as the sensor is mounted offset from surrounding rooftops (which contribute to large diurnal amplitude in T\textsubscript{hem, r} and larger deviation from T\textsubscript{comp}).
+The sensitivity of T\textsubscript{hem, r} to sensor placement decreases as sensor height is increased, and T\textsubscript{hem, r} becomes spatially invariant at approximately 3.5 times mean building height \citep{Roberts2010,Adderley2015}.
+As such, these biases can be significantly reduced by mounting downward facing pyrgeometers slightly higher than typical.
+Although pyrgeometers in both studies utilized in this campaign are mounted below the height at which T\textsubscript{hem, r} is spatially invariant, sensitivity tests in Section \ref{sensitivity} show that the magnitude of this bias is relatively small for sensor placement in the BUBBLE campaign.
+Similar biases are expected in analysis of Vancouver sUHI as the sensor is mounted offset from surrounding rooftops (which contribute to large diurnal amplitude in T\textsubscript{hem, r} and larger deviation from T\textsubscript{comp}).
 In addition, T\textsubscript{hem, r} shows good performance relative to T\textsubscript{plan} when compared to T\textsubscript{comp}, particularly under clear sky conditions when microscale spatiotemporal variations in urban T\textsubscript{surf} and sUHI magnitudes are greatest.
-As such, the influence of biases in T\textsubscript{hem} in sUHI analysis for both BUBBLE and EPiCC campaigns is likely small.
+As such, the influence of biases in T\textsubscript{hem} in sUHI analysis for both BUBBLE and EPiCC campaigns is likely small and relatively consistent.
 
@@
 
-By virtue of requiring nearly 500 radiative transfer simulations per time step and significant post-processing, the method described in \ref{method2} is computationally expensive.
+By virtue of requiring nearly 500 radiative transfer simulations per time step and significant post-processing, the method described in Section \ref{method2} is computationally expensive.
 This prompted the development of a parameterization scheme to facilitate more widespread usage of urban T\textsubscript{hem, r} for climatological and comparative sUHI analysis.
@@
 Using the view factor weighted path length, a suite of radiative transfer simulations is run prior to correction to retrieve "hemispherical" broadband atmospheric transmittances for a range of potential humidities.
-Humidity-transmittance pairings are aggregated into a look-up table (LUT), where the measured humidity at each time step can be matched with the associated hemispherical transmittance.
-Using the hemispherical transmittance as the fraction of surface emitted L\textsubscript{up} reaching the sensor, measured L\textsubscript{up} can then be decomposed into its surface and atmospheric emitted components, from which a parameterized T\textsubscript{hem, r} is inferred.
+Humidity-transmittance pairings are aggregated into a LUT, where the measured humidity at each time step can be matched with the associated hemispherical transmittance.
+Using the hemispherical transmittance to calculate the fraction of surface emitted L\textsubscript{up} reaching the sensor, measured L\textsubscript{up} can then be decomposed into its surface and atmospheric emitted components, from which a parameterized T\textsubscript{hem, r} is inferred.
 Comparison of T\textsubscript{hem, r} derived from the correction method and the parameterization show good agreement, the results of which are included in Figure \ref{para_eval} and Table \ref{parastats}.
@@
 
-This study uses data from two long-term climatological campaigns: the Basel Urban Boundary Layer Experiment (BUBBLE) conducted over December 2001 through July 2002 in Basel, Switzerland \citep{Rotach2005}, and the Environmental Prediction in Canadian Cities (EPiCC) \citep{Voogt2007} campaign in Vancouver, Canada.
-Using the method described in section \ref{method2}, an eight month climatology of T\textsubscript{hem, r} are retrieved from radiation and meteorological data collected as a part of the BUBBLE campaign.
-Urban data were observed from a tower located near city center at the "Basel Sperrstrasse" site.
-Rural reference data were observed at the "Lange Erlen" site approximately 6 \si{\kilo\meter} to the north east in the outskirts of the city of Basel.
+This study uses data from two long-term climatological campaigns: the Basel Urban Boundary Layer Experiment (BUBBLE) conducted over December 2001 through July 2002 in Basel, Switzerland \citep{Rotach2005}, and a year long subset of the the Environmental Prediction in Canadian Cities (EPiCC) \citep{Voogt2007} campaign in Vancouver, Canada.
+Using the method described in Section \ref{method2}, an eight month climatology of T\textsubscript{hem, r} are retrieved from radiation and meteorological data collected as a part of the BUBBLE campaign.
+Urban data were observed from a tower located in a mid-rise, largely residential neighborhood near city center at the "Basel Sperrstrasse" site.
+Rural reference data were observed from the "Lange Erlen" site approximately 6 \si{\kilo\meter} to the north east in the outskirts of the city of Basel.
 
@@
 To supplement data from the BUBBLE campaign, radiation and meteorological variables observed as a part of the EPiCC project are used to derive a climatology of T\textsubscript{hem, r} for the year long period spanning October 2008 through September 2009 using the parameterization scheme.
-EPiCC urban and rural data were collected from an urban/suburban site in central Vancouver ("Sunset Tower") and from a rural site approximately 25 \si{\kilo\meter} to the SSW ("Westham Island").
-These data provide a second, long term, time continuous climatology of sUHI for comparison of seasonal and diurnal patterns, to understand the influence of urban and rural site selection and characteristics on sUHI, and to aid in developing a generalized model for sUHI development over multiple time scales.
+EPiCC urban and rural data were collected from an urban residential neighborhood site in central Vancouver (”Sunset Tower”) and from a rural site over a flat, non-irrigated, unmanaged field approximately 25 km to the SSW (”Westham Island”) near to the coast.
+These data provide a second, long term, time-continuous climatology of sUHI for comparison of seasonal and diurnal patterns, to understand the influence of urban and rural site selection and characteristics on sUHI, and to aid in developing a generalized model for sUHI development over multiple time scales.
 
@@
 
-sUHI and clhUHI magnitudes were calculated from T\textsubscript{hem, r} and T\textsubscript{air} (measured from 2 \si{\meter} above ground level) as $\Delta$T\textsubscript{(urban-rural)} at 30 \si{\min} intervals.
+sUHI and clUHI magnitudes were calculated from T\textsubscript{hem, r} and T\textsubscript{air} (measured from 2 \si{\meter} above ground level) as $\Delta$T\textsubscript{(urban-rural)} at 30 \si{\min} intervals.
 Rural T\textsubscript{hem, b} was calculated an expanded version of equation \ref{stefb1} to remove emissivity effects,
@@
 Observations at Village Neuf were intermittent and often discontinuous As such, a subset of nine partially-complete days was selected to represent T\textsubscript{hem, b} and sUHI over a range of cloud coverages, to investigate the influence of rural site characteristics and T\textsubscript{surf} regimes on sUHI.
-As sUHI develop, preference was given to days with continuous observations from early afternoon through the night.
-For each of the nine days sUHI\textsubscript{max} was used to infer a normalized sUHI magnitude for both Lange Erlen and Village Neuf sites at 30- \si{\minute} intervals to compare patterns of sUHI development for the two sites.
+As sUHI development is most variable in the late afternoon and evening hours, preference was given to days with continuous observations from early afternoon through the night.
+For each of the nine days sUHI\textsubscript{max} was used to infer a normalized sUHI magnitude for both Lange Erlen and Village Neuf sites at 30 \si{\minute} intervals to compare patterns of sUHI development for the two sites.
 Mean normalized sUHI was then calculated at each time step over the nine day period.
@@
 
-Figure \ref{heat_uhi_diff} shows seasonal variation in the difference between mean monthly hemispherical sUHI\textsubscript{norm} and clUHi\textsubscript{norm} for the BUBBLE campaign.
+Figure \ref{heat_uhi_diff} shows seasonal variation in the difference between mean monthly hemispherical sUHI\textsubscript{norm} and clUHI\textsubscript{norm} for the BUBBLE campaign.
 By restricting UHI values to a range between one and zero, diurnal patterns of sUHI and clUHI can be compared directly.
@@
 Figure \ref{vanc_suhi} shows seasonal variations in diurnal patterns of mean monthly hemispherical sUHI magnitudes at 30 minute intervals for the EPiCC campaign.
-The sUHI in Vancouver shows similar seasonal and diurnal patterns to the sUHI in Basel, but does not persist as far into the late evening hours.
+The sUHI in Vancouver shows similar seasonal and diurnal patterns to the sUHI in Basel, but degrades more quickly in the late evening hours.
 
@@
 Figure \ref{bxsUHI_season_bubble} shows hemispherical sUHI magnitudes for the eight month BUBBLE campaign and binned for winter (December through February) and summer (April through July) seasons.
-Diurnal variability in sUHI magnitudes is concentrated in summer months, as well as seasonal sUHI\textsubscript{max}.
+Diurnal variability in sUHI magnitudes and seasonal sUHI\textsubscript{max} is concentrated in summer months.
 
@@
 	\includegraphics[width=15cm,height=7in,keepaspectratio]{bxsUHI_iop_bubble}
-	\caption{A comparison of sUHI magnitudes from complete (sUHI\textsubscript{comp}), hemispherical (sUHI\textsubscript{hem}), and nadir (sUHI\textsubscript{nadir}) remote sensed representations of the Sperrstrasse canyon over the IOP.
+	\caption{A comparison of sUHI magnitudes from complete (sUHI\textsubscript{comp}), hemispherical (sUHI\textsubscript{hem}), and nadir (sUHI\textsubscript{nadir}) remote sensed representations of the Sperrstrasse canyon over the BUBBLE IOP.
 Each plot includes mean sUHI, as well as case days representing sUHI under clear sky and clouded sky conditions.}
@@
 
-Figure \ref{meteo_sol} shows mean and maximum sUHI magnitudes (sUHI\textsubscript{mean} and sUHI\textsubscript{max}) as a function of solar radiant exposure for each day of the BUBBLE campaign.
+\subsubsection{BUBBLE}
+
+Figure \ref{meteo_sol} shows mean and maximum hemispherical sUHI magnitudes (sUHI\textsubscript{mean} and sUHI\textsubscript{max}) as a function of solar radiant exposure for each day of the BUBBLE campaign.
 Point shading indicates the season, where black is winter (Nov - Feb), white is spring (Mar - Apr), and red is summer (May - Aug).
@@
 No discernible relationship is observed between atmospheric water vapor content and sUHI\textsubscript{mean} or sUHI\textsubscript{max}.
-In both, colors indicate solar radiant exposure accumulated over each day (yellow is large accumulated insolation, brown is low).
+In both figures, colors indicate solar radiant exposure accumulated over each day (yellow is large accumulated insolation, brown is small).
 
-Figured \ref{meteo_atuhi} shows hemispherical sUHI magnitudes as a function of water vapor content and clUHI measured over the summer months of the BUBBLE campaign (May through July 2002).
-Winter hours are omitted as sUHI is small and displays minimal variation in winter.
+Figured \ref{meteo_atuhi} shows hemispherical sUHI magnitudes as a function clUHI measured over the summer months of the BUBBLE campaign (May through July 2002).
+Winter hours are omitted as both sUHI and clUHI are small and display minimal variation in winter.
 The relationship between sUHI and clUHI is complex and highly dependent on time of day - owing largely to similar, but slightly offset patterns of sUHI and clUHI development.
@@
 	\includegraphics[width=15cm,height=7in,keepaspectratio]{meteo_kdown}
-	\caption{Maximum and mean 24 \si{\hour} hemispherical sUHI magnitude for each day of April through July versus integrated solar radiant exposure measured over the same 24 \si{\hour} period.
+	\caption{Maximum and mean 24 \si{\hour} hemispherical sUHI magnitude for each day of April through July during the BUBBLE campaign versus integrated solar radiant exposure measured over the same 24 \si{\hour} period.
 Color indicates season: black is winter (Nov - Feb), white is spring (Mar - Apr), and red is summer (May - Aug).
@@
 	\includegraphics[width=15cm,height=7in,keepaspectratio]{meteo_wv}
-	\caption{Maximum and mean 24 \si{\hour} hemispherical sUHI magnitude for each day versus mean wind velocity measured at approximately 2 \si{\meter} above ground at the Sperrstrasse site.
+	\caption{Maximum and mean 24 \si{\hour} hemispherical sUHI magnitude for each day of the BUBBLE campaign versus mean wind velocity measured at approximately 2 \si{\meter} above ground at the Sperrstrasse site.
 Coloring indicates integrated solar radiant exposure over each day.}
@@
 	\includegraphics[width=15cm,height=7in,keepaspectratio]{meteo_hum}
-	\caption{Maximum and mean 24 \si{\hour} sUHI magnitude versus atmospheric water vapor content measured at 2 \si{\meter} at the Sperrstrasse site.
+	\caption{Maximum and mean 24 \si{\hour} sUHI magnitude for each day of the BUBBLE campaign versus atmospheric water vapor content measured at 2 \si{\meter} at the Sperrstrasse site.
 Similar patterns are observed when rural water vapor content is substituted.
@@
 
+\subsubsection{EPiCC}
 %Vancouver
 
-Figure \ref{vanc_sol} shows mean and maximum sUHI magnitudes (sUHI\textsubscript{mean} and sUHI\textsubscript{max}) as a function of solar radiant exposure over each day of the year long EPiCC Vancouver campaign.
+Figure \ref{vanc_sol} shows hemispherical sUHI\textsubscript{mean} and sUHI\textsubscript{max} as a function of solar radiant exposure over each day of the year long EPiCC Vancouver campaign.
 Point shading indicates the season, where black is winter (Nov - Feb), white is spring (Mar - Apr), red is summer (May - Aug), and blue is fall (Sept - Oct).
@@
 In contrast to findings in the BUBBLE campaign, a weak relationship between sUHI and humidity is observed.
-However, the apparent relationship may be a product of Vancouver's proximity to the coast, as there is significant seasonal variation in both  and may be no actual causal relationship.
+However, the apparent relationship may be a product of Vancouver’s proximity to the coast and distinct wet and dry seasons, as there is significant seasonal variation in both sUHI and atmospheric water vapor content.
 In both, colors indicate solar radiant exposure accumulated over each day (yellow is large accumulated insolation, brown is low).
 
-Figured \ref{vanc_swc} shows hemispherical sUHI magnitudes as a function of water vapor content and clUHI measured for April and May 2009 of the EPiCC Vancouver canpaign.
-These months were selected to show a wide range of soil water contents and sUHI magnitudes.
-A negative relationship is observed between sUHI and soil water vapor content, particularly in May, where larger sUHI magnitudes often coincide with lower soil water vapor contents.
-However, some of this relationship may owe to diurnal patterns of soil saturation, which, contrary to sUHI magnitudes, tend to be highest in early morning and lower throughout the day.
-Color in both plots indicates time of day: shades of blue indicate early morning through midday hours, white is solar noon through early evening hours, and red is late evening through the nighttime.
+Figured \ref{vanc_swc} shows hemispherical sUHI magnitudes as a function of soil water vapor content measured over May 2009 of the EPiCC Vancouver campaign.
+May was selected to show a wide range of soil water contents and sUHI magnitudes.
+A negative relationship is observed between sUHI and soil water vapor content by day where larger sUHI magnitudes often occur on days with lower soil water vapor content.
+No discernible relationship between sUHI and soil water vapor content is observed at night.
 
@@
 	\centering
-	\includegraphics[width=15cm,height=7in,keepaspectratio]{vanc_swc}
-	\caption{sUHI\textsubscript{hemi} magnitude versus soil water content measured from 0.05 \si{\meter} below the surface at the Westham island site in Vancouver, Canada for April and May 2009.
-Color indicates time of day: blue is early morning through midday, white is solar noon through early evening, and red is late evening through the night.}
+	\includegraphics[width=15cm,height=7in,keepaspectratio]{vanc_swc_dn}
+	\caption{sUHI\textsubscript{hemi} magnitude versus volumetric soil water content measured from 0.05 \si{\meter} below the surface at the Westham island site in Vancouver, Canada for May 2009 binned for day (11:00 - 17:00 LST) and night (22:00 - 5:00 LST) hours.}
 	\label{vanc_swc}
@@
 The "extra", largely non-radiative cooling at the Lange Erlen site results in a strong second peak in sUHI that is not present in sUHI calculated using the Village Neuf site.
-In the context of the Lowry method discussed in Chapter \ref{intro}, $V_{M, \ R}$ measured from the Lange Erlen site and $V_{M, \ U}$ are not subject to the same $V_{L}$ and equation does not yield a perfectly isolated $V_{H}$ in the hours just after sunset.
+In the context of the Lowry method discussed in Chapter \ref{intro}, $V_{M, \ R}$ measured from the Lange Erlen site and $V_{M, \ U}$ are not subject to the same $V_{L}$.
+Thus, equation \ref{lowry} does not yield a perfectly isolated $V_{H}$ in the hours just after sunset.
 
@@
 Cooling rates at the Vancouver Westham island rural site are not affected by cold air drainage and, as a result, sUHI does not display two diurnal peaks in summer nor a coherent peak after sunset in winter months.
-Thus, observations of seasonal variations in hemispherical sUHI in Vancouver are likely highly generalizable to other mid-latitude cities with distinct winter and summer seasons.
+Thus, observations of seasonal variations in hemispherical sUHI in Vancouver are highly generalizable to other mid-latitude cities.
 
 The sUHI and the clUHI have similar micro-meteorological genesis, climatic controls, and display similar seasonal variability, albeit concentrated at different times of the day.
-This is evident when comparing normalized mean monthly hemispherical sUHI and clUHI - where the range of both types of sUHI is restricted to values between zero and one.
-In winter, low solar input and frequent cloud cover inhibit variability in both sUHI and clUHI and both are of similar normalized magnitudes.
+This is evident when comparing normalized mean monthly hemispherical sUHI and clUHI - where the range of both types of UHI are restricted to values between zero and one.
+In winter, low solar input and frequent cloud cover inhibit variability in sUHI and clUHI and both are of similar normalized magnitudes.
 In summer, larger solar input fosters contrasts in urban-rural heating and cooling rates and both clUHI and sUHI are most variable, with peak sUHI in the early afternoon when clUHI is smallest, and peak clUHI in the hours after sunset when sUHI is relatively low.
 These results suggest the following generalization: when conditions that foster the formation of strong urban-rural contrasts in T\textsubscript{air} and T\textsubscript{surf} are present (most notably a large solar input), sUHI peaks approximately two hours after solar noon and clUHI peaks approximately three to four hours after sunset.
-Thus, a surface urban heat island is not simply a stronger, more pronounced, canopy layer urban heat island.
+\textit{Thus, a surface urban heat island is not simply a stronger, more pronounced, canopy layer urban heat island.
+}
 
@@
 
-Results of sensor placement sensitivity testing in Section \ref{sensitivity} show that T\textsubscript{hem, r} is dependent on sensor position relative to canyon features.
+Results of sensor placement sensitivity testing in Section \ref{sensitivity} show that T\textsubscript{hem, r} is dependent on sensor position and height relative to canyon features for pyrgeometers mounted below 3 to 5 times mean building height.
+As pyrgeometers in both EPiCC and BUBBLE campaigns were mounted below these heights, derived T\textsubscript{hem, r} include some bias towards facets that make up a majority of their FOV.
 These effects are translated into sUHI\textsubscript{hem}.
-In general, when increasing sUHI magnitudes.
+This bias results in a small overestimation of daytime T\textsubscript{comp}, and an underestimation of nighttime T\textsubscript{comp}, manifesting as a small increase (decrease) in daytime (nighttime) sUHI\textsubscript{hem}.
+However, as these sensors are positioned offset from rooftops (the main source of positional bias), these biases are likely relatively small and consistent and do not exert a large influence on diurnal or seasonal patterns of sUHI\textsubscript{hem}.
 
@@
 As such, sUHI development is strongly tied to integrated solar radiant exposure accumulated over a day and wind velocity, which display positive and negative relationships with sUHI magnitudes respectively.
-Other variables also show weakly positive (mean atmospheric water vapor content) and negative (soil water content) relationships with sUHI, however, conclusions about causality are more difficult to make as they too display some dependence on the same variables.
+Other variables also show weakly positive (mean atmospheric water vapor content) and negative (soil water content) relationships with sUHI, however, for these variables, conclusions about causality are more difficult to make.
 Meteorological controls on sUHI are summarized in Table \ref{meteo_cont}.
@@
 This suppresses microscale spatiotemporal variations in T\textsubscript{surf} as hot sunlit surfaces experience greater rates of convective cooling under high wind conditions than cooler shaded surfaces.
-This reduces overall urban and rural T\textsubscript{surf} and inhibits sUHI development.
+Thus, high wind velocities reduce overall urban and rural T\textsubscript{surf} and inhibit sUHI development.
 Nighttime sUHI is also affected by wind velocity as the formation of a near surface thermal inversion, which increases radiative cooling in non-urban areas, requires calm winds.
@@
 
+Results in \citet{Runnalls2000} indicate that rural thermal admittance (which is largely a product of rural soil moisture) exerts strong negative control over clUHI.
+A similar relationship is observed for sUHI in Vancouver when analyzed over long time scales.
+However, examining the diurnal character reveals a more complex relationship between the two.
+A higher rural soil moisture content increases rural thermal admittance, suppressing rural heating and cooling rates and reducing the overall T\textsubscript{surf} amplitude at the rural site.
+Much of the urban surface is impervious and does not respond strongly to changes in soil moisture.
+Thus, \citet{Oke2017} suggests that with a higher rural soil moisture content, reduced nocturnal rural cooling rates could result in decreased nighttime sUHI.
+Conversely, by day, reduced rural heating rates may result in higher sUHI magnitudes.
+Results in Vancouver do not support these suggestions.
+Over the daytime hours, in Vancouver, a negative relationship is observed between sUHI and rural soil moisture where sUHI is largest with low rural soil moisture.
+At night, no clear relationship between the two emerges.
+As Vancouver has distinct wet and dry seasons and little variability in soil moisture during the summer months (when sUHI is highly variable), study of the effect in cities with wet summers may yield a more complete picture of the relationship between sUHI and rural soil moisture.
+
 \begin{table}[H]
 	\centering
-	\caption{A summary of the relationship between sUHI, meteorological variables, and clUHI.}
+	\caption{A summary of the relationship between sUHI, meteorological variables, and clUHI.
+These relationships are generalizable and represent the character of mean diurnal sUHI in response to changing conditions.}
 	\label{meteo_cont}
@@
 
-\section{Conclusion}
+\section{Conclusions}
 
@@
 Hemispherical sUHI magnitudes display significant seasonal and diurnal variation and are largest in summer months during the late afternoon during which solar heating of the surface is most intense, forcing the development of strong urban-rural contrasts in T\textsubscript{surf}.
-Seasonality in late afternoon sUHI is largely dependent on accumulated radiative exposure over the day which displays a strong positive relationship with sUHI magnitudes.
+Seasonality in late afternoon sUHI is largely dependent on accumulated radiative exposure over the day, which displays a strong positive relationship with sUHI magnitudes.
 In the winter months, a reduced overall solar input and increased cloud cover frequency suppresses daytime sUHI development resulting in less variability in sUHI, and lower overall sUHI magnitudes.
+Other meteorological controls have complex relationships with sUHI that often vary with time of day.
+Thus, sUHI development on any given day is the result of complex interplay between myriad meteorological controls.
 
 Remote sensed sUHI is highly dependent on sensor-surface-sun geometry.
-This results in significant variation in the character of urban T\textsubscript{surf} and sUHI when observed from sensors with different spatial, geometric, and temporal sampling regimes.
+This results in significant variation in the character of sUHI observed from sensors with different spatial, geometric, and temporal sampling regimes.
 Thus, traditional methods for thermal remote sensing of urban areas do not capture the true temporal and geometric nature of the urban effect on land T\textsubscript{surf}.
@@
 These geometric biases manifest not only in deviation between observed sUHI magnitudes and the 'true' complete sUHI, but also impact characterizations of diurnal and seasonal patterns of sUHI and urban T\textsubscript{surf}.
-Geometric biases in thermal remote sensing of urban area are compounded by myriad temporal biases across a range of temporal scales, as traditional methods for urban thermal remote sensing are not time-continuous and only operable under clear sky conditions.
+Geometric biases in thermal remote sensing of urban areas are compounded by myriad temporal biases across a range of temporal scales, as traditional methods for urban thermal remote sensing are not time-continuous and only operable under clear sky conditions.
 Both geometric and temporal biases affect the vast majority of sUHI study, most of which is performed using only a small number of satellite remote sensors, to which these biases are inherent.
 
-This study provides the first, long-term, temporally continuous, geometrically representative analysis of urban T\textsubscript{surf} and sUHI.
+This study provides the first, long-term, temporally continuous, geometrically representative analysis of sUHI.
 By sampling the urban surface in 3-dimensions and time-continuously, the method utilized in this chapter overcomes and quantifies spatial and temporal biases in the satellite thermal remote sensing record to better understand the geometric and temporal effects of a city on the surface climate.

### 2017-05-11 uw-ethesis-frontpgs.tex
Add files via upload

@@
 		T\textsubscript{air} & \si{\kelvin} & air temperature \\
+		T\textsubscript{comp} & \si{\kelvin} & complete urban surface temperature\\
+		T\textsubscript{hem} & \si{\kelvin} & urban surface temperature viewed from a hemispherical remote sensor\\
+		T\textsubscript{plan} & \si{\kelvin} & urban surface temperature viewed from a sensor in the nadir\\
+		T\textsubscript{road} & \si{\kelvin} &road surface temperature\\
+		T\textsubscript{roof} & \si{\kelvin} & roof surface temperature \\
 		T\textsubscript{surf} & \si{\kelvin} & surface temperature \\
-		T\textsubscript{comp} & \si{\kelvin} & complete urban surface temperature\\
-		T\textsubscript{plan} & \si{\kelvin} & urban surface temperature viewed from a sensor in the nadir\\
-		T\textsubscript{hem} & \si{\kelvin} & urban surface temperature viewed from a hemispherical remote sensor\\
+		T\textsubscript{wall} & \si{\kelvin} & wall surface temperature\\
+
 		aUHI & \si{\kelvin} & air temperature urban heat island effect \\

### 2017-05-11 conclusion.tex
Add files via upload

@@
 However, as sUHI is the result of differences in heating and cooling rates between the city and its non urban surroundings, its relationship to meteorological controls is complex.
-Analysis of sUHI development under a range of mid-latitude synoptic conditions lead to the development of the first observationally-supported, generalizable model for sUHI development presented in Section \ref{gen}.
+Analysis of sUHI development under a range of mid-latitude synoptic conditions lead to the development of the first observationally-supported, generalizable model for sUHI development presented in Section \ref{gen_mod}.
 
@@
 The complex nature of the urban surface introduces myriad shortcomings and biases into conventional remote sensing of urban thermal climates - this method is no exception.
-Sensor placement sensitivity tests in Section \ref{placement} corroborate findings in \citet{Roberts2010,Adderley2015} and indicate that longwave radiation upwelling from the urban surface as measured from a downward-facing radiometer is spatially variant when measured from heights below approximately three to five times mean building height.
+Sensor placement sensitivity tests in Section \ref{placement} corroborate findings in \citet{Roberts2010} and \citet{Adderley2015}.
+They indicate that longwave radiation upwelling from the urban surface as measured from a downward-facing radiometer is spatially variant when measured from heights below approximately three to five times mean building height.
 Most pyrgeometers (including those used in this study) are mounted below this height and are subject to a bias towards facets that make up a disproportionately large fraction of the sensor FOV.
@@
 In addition, many other datasets of urban energy balances from inactive and decommissioned towers and from campaigns not indexed by Urbanflux are available for analysis via this method.
-These datasets span several continents and climatic zones, and characterize energy balances for cities over a range of developing and developed nations, with many operational over very long (5+ years) time scales.
+These datasets span several continents and climatic zones, and characterize energy balances for cities over a range of developing and developed nations, with many operational over very long (5+ year) time scales.
 Thus, the application of this method over existing untapped datasets, and as a component of future energy balance assessments, can not only could provide a wealth of novel urban T\textsubscript{surf} and sUHI analyses for urban climatology and could also can provide for analysis outside the traditional purview of urban climatology.

### 2017-05-11 introduction.tex
Add files via upload

@@
 
-The temperature of a surface (urban or otherwise) at its interface with the atmosphere is defined as the net result of the surface energy balance \citep{Oke2017}, wherein change in temperature at time $t$ for a surface with thickness $z$ and heat capacity $C$ can be solved as,
+The temperature of a surface (urban or otherwise) at its interface with the atmosphere is defined as the net result of the surface energy balance \citep{Oke2017}, wherein change in temperature at time $t$ for a surface with thickness $z_t$ and heat capacity $C$ can be solved as,
 
 \begin{equation}
-	C~\frac{\partial \text{T\textsubscript{surf}}}{\partial t}~z = Q^*+ Q_F - Q_H - Q_E - \Delta Q_A - \Delta Q_G
+	C~\frac{\partial \text{T\textsubscript{surf}}}{\partial t}~z_t = Q^*+ Q_F - Q_H - Q_E - \Delta Q_A - \Delta Q_G
 \end{equation}
@@
 	keepaspectratio]{model}
-	\caption{Projected FOVs for a narrow-FOV sensor at nadir and oblique angles, and a wide-FOV hemispherical sensor viewing an idealized urban array with building width \textit{H}, height 2\textit{H}, and inter element spacing \textit{H}.
+	\caption{Projected FOVs for a narrow-FOV sensor at nadir and oblique angles, and a wide-FOV hemispherical sensor viewing an idealized urban array with building width \textit{W}, height 2\textit{W}, and inter element spacing \textit{W}.
 Figure from Adderley et al.

### 2017-05-11 paper2.tex
Add files via upload

@@
 The temperature of the surface is integral in understanding, predicting, and modeling boundary-layer air temperature patterns, surface energy balances, and, in urban areas, has important implications for human thermal comfort and building energy usage.
-Urban modification of surface geometry and thermal, radiative, moisture, and aerodynamic properties results in differential surface heating and cooling patterns and strong microscale spatiotemporal variations in urban surface temperature (T\textsubscript{surf}).
+Urban modification of surface geometry and thermal, radiative, moisture, and aerodynamic properties results in differential surface heating and cooling patterns and strong microscale spatiotemporal variations in urban surface temperature.
 Integrated up to larger scales, urban areas tend to store more heat relative to non-built surroundings; and traffic, HVAC systems, and industrial processes contribute to an additional anthropogenic heat input.
 In addition, a reduced pervious (and vegetative) cover in urban areas shifts convective fluxes towards sensible heat and away from latent heat.
-These manifest in elevated T\textsubscript{surf} and T\textsubscript{air} in cities, a phenomenon termed the urban heat island effect (UHI).
+These manifest in elevated T\textsubscript{surf} and T\textsubscript{air} in cities, a phenomenon termed the urban heat island effect.
 To foster a more complete understanding of the effect of urban areas on climates across a range of scales, accurate, spatiotemporally continuous and geometrically representative characterization of T\textsubscript{surf} in cities has long been a goal in urban climatology.
@@
 		Basel Sperrstrasse \newline \textit{Urban} \textit{street canyon} \newline LCZ: 2 & 47.57\si{\degree} N \newline 7.60\si{\degree} E \newline 255 \si{\meter} a.s.l.
-& $z_H$ = 14.6  \si{\meter} \newline $\sigma_H $ = 6.9 \si{\meter} \newline H/W = 0.54 \newline $\lambda_C $ = 0.37\newline $\lambda_S$ = 1.0 \newline $\alpha$ = 11.0\% & T\textsubscript{air} [7] \newline H [7] \newline WV [12] \newline WD [1] \newline P [1] & L\textsubscript{up} [3] \newline L\textsubscript{down} [5] \newline S\textsubscript{up} [2] \newline S\textsubscript{down} [3] \\
+& $z_H$ = 14.6  \si{\meter} \newline $\sigma_H $ = 6.9 \si{\meter} \newline H/W = 1.0 \newline $\lambda_C $ = 1.92\newline $\lambda_P$ = 0.54 \newline $\alpha$ = 11.0\% & T\textsubscript{air} [7] \newline H [7] \newline WV [12] \newline WD [1] \newline P [1] & L\textsubscript{up} [3] \newline L\textsubscript{down} [5] \newline K\textsubscript{up} [2] \newline K\textsubscript{down} [3] \\
 		& & & & \\
 		Lange Erlen \newline \textit{Rural parkland} \newline LCZ: B/D & 47.59\si{\degree} N \newline 7.65\si{\degree} E \newline 275 \si{\meter} a.s.l.
-& $\alpha$ = 21.4\%  & T\textsubscript{air} [4] \newline H [4] \newline WV [3] \newline WD [1] &  L\textsubscript{up} [1] \newline L\textsubscript{down} [1] \newline S\textsubscript{up} [1] \newline S\textsubscript{down} [1]  \\
+& $\alpha$ = 21.4\%  & T\textsubscript{air} [4] \newline H [4] \newline WV [3] \newline WD [1] &  L\textsubscript{up} [1] \newline L\textsubscript{down} [1] \newline K\textsubscript{up} [1] \newline K\textsubscript{down} [1]  \\
 		&&&&\\
@@
 		& T\textsubscript{air} [3] \newline H [3] \newline WV [2] \newline WD [1] &  L\textsubscript{up} [1] \newline L\textsubscript{down} [1] \newline
-		S\textsubscript{up} [2] \newline S\textsubscript{down} [2]  \\ 
+		K\textsubscript{up} [2] \newline K\textsubscript{down} [2]  \\ 
 		\bottomrule
@@
 		\textsuperscript{3} H: humidity, WV: wind velocity, WD: wind direction, P: pressure. \\
-		\textsuperscript{4} L\textsubscript{up}: upwelling longwave radiation, L\textsubscript{down}: downwelling longwave radiation, S\textsubscript{up}: upwelling shortwave radiation, S\textsubscript{down}: downwelling shortwave radiation.
+		\textsuperscript{4} L\textsubscript{up}: upwelling longwave radiation, L\textsubscript{down}: downwelling longwave radiation, K\textsubscript{up}: upwelling shortwave radiation, K\textsubscript{down}: downwelling shortwave radiation.
 \end{table}
@@
 		Sunset Tower \newline \textit{Urban/Suburban neighborhood} \newline LCZ: 6 & 49.23\si{\degree} N \newline 123.07\si{\degree} W \newline 2.0 \si{\meter} a.s.l.
-& $z_H$ = 5.3 \si{\meter} \newline $\lambda_B $ = 0.29 \newline $\lambda_T $ = 0.15\newline $\lambda_V$ = 0.22 \newline $\lambda_I$ = 0.37 \newline $\alpha$ = 14.7\% & T\textsubscript{air} [1] \newline H [1] \newline WV [1] \newline WD [1] \newline P [1] & L\textsubscript{up} [1] \newline L\textsubscript{down} [1] \newline S\textsubscript{up} [1] \newline S\textsubscript{down} [1] \\
+& $z_H$ = 5.3 \si{\meter} \newline $\lambda_B $ = 0.29 \newline $\lambda_T $ = 0.15\newline $\lambda_V$ = 0.22 \newline $\lambda_I$ = 0.37 \newline $\alpha$ = 14.7\% & T\textsubscript{air} [1] \newline H [1] \newline WV [1] \newline WD [1] \newline P [1] & L\textsubscript{up} [1] \newline L\textsubscript{down} [1] \newline K\textsubscript{up} [1] \newline K\textsubscript{down} [1] \\
 		& & & & \\
 		Westham Island \newline \textit{Rural, unmanaged} \newline LCZ: D & 49.09\si{\degree} N \newline 123.18\si{\degree} W \newline 91.4 \si{\meter} a.s.l.
-& $\alpha$ = 20.6\%  & T\textsubscript{air} [2] \newline H [2] \newline WV [2] \newline WD [2] &  L\textsubscript{up} [1] \newline L\textsubscript{down} [1] \newline S\textsubscript{up} [1] \newline S\textsubscript{down} [1]  \\
+& $\alpha$ = 20.6\%  & T\textsubscript{air} [2] \newline H [2] \newline WV [2] \newline WD [2] &  L\textsubscript{up} [1] \newline L\textsubscript{down} [1] \newline K\textsubscript{up} [1] \newline K\textsubscript{down} [1]  \\
 		\bottomrule
@@
 	\centering
-	\includegraphics[width=15cm,height=7in,keepaspectratio]{heatsuhi}
+	\includegraphics[width=15cm,height=7in,keepaspectratio]{heatsuhi5_11}
 	\caption{A heatmap of mean half hourly hemispherical sUHI for each month calculated at 30 \si{\min} intervals for the BUBBLE campaign. Results are interpolated between months.}
@@
 	\centering
-	\includegraphics[width=15cm,height=7in,keepaspectratio]{heatatuhi}
+	\includegraphics[width=15cm,height=7in,keepaspectratio]{heatatuhi5_11}
 	\caption{A heatmap of mean half hourly canopy layer UHI for each month calculated at 30 \si{\min} intervals for the BUBBLE campaign. Results are interpolated between months.}
@@
 	\centering
-	\includegraphics[width=15cm,height=7in,keepaspectratio]{heat_uhi_diff}
+	\includegraphics[width=15cm,height=7in,keepaspectratio]{heat_uhi_diff5_11}
 	\caption{A heatmap of the difference between mean half hourly hemispherical normalized sUHI and canopy layer UHI for each month calculated at 30 \si{\min} intervals over the BUBBLE campaign.
@@
 	\includegraphics[width=15cm,height=7in,keepaspectratio]{meteo_kdown}
-	\caption{Maximum and mean 24 \si{\hour} hemispherical sUHI magnitude for each day of April through July during the BUBBLE campaign versus integrated solar radiant exposure measured over the same 24 \si{\hour} period.
+	\caption{Maximum and mean 24 \si{\hour} hemispherical sUHI magnitude for each day of April through July during the BUBBLE campaign versus integrated solar radiant exposure measured at the Sperrstrasse site over the same 24 \si{\hour} period.
 Color indicates season: black is winter (Nov - Feb), white is spring (Mar - Apr), and red is summer (May - Aug).
@@
 	\includegraphics[width=15cm,height=7in,keepaspectratio]{meteo_atuhi}
-	\caption{sUHI magnitude versus clUHI magnitude binned for day and night calculated over the summer months of the BUBBLE campaign.
+	\caption{sUHI magnitude versus clUHI magnitude binned for day and night calculated over the summer months of the BUBBLE campaign calculated for the Sperrstrasse and Lange Erlen sites.
 Coloring in the daytime plot indicates time of day: Red is morning (5:00 - 9:00 LST), gray is midday hours (9:30 - 17:00 LST), and yellow is evening hours (17:30 - 20:00 LST).}
@@
 	\includegraphics[width=15cm,height=7in,keepaspectratio]{vanc_kdown}
-	\caption{Maximum and mean 24 \si{\hour} hemispherical sUHI magnitude for each day of the EPiCC climatology versus integrated solar radiant exposure measured over the same 24 \si{\hour} period.
+	\caption{Maximum and mean 24 \si{\hour} hemispherical sUHI magnitude for each day of the EPiCC climatology versus integrated solar radiant exposure measured at the Sunset tower site over the same 24 \si{\hour} period.
 Color indicates season: black is winter (Nov - Feb), white is spring (Mar - Apr), red is summer (May - Aug), and blue is fall (Sept - Oct).}
@@
 
-\subsection{The effect of rural/non-urban characteristics on sUHI and a generalizable case for sUHI}
+\subsection{The effect of rural/non-urban characteristics on sUHI}
 
@@
 Days were chosen to represent a wide range of summertime conditions and cloud coverages, however, observations at the Village Neuf site were intermittent and n = 9 only applies to hours after 12:00 LST when data could be continuously sampled.
-
-Figure \ref{gen_case} presents a schematic for summertime sUHI development constructed from mean urban and rural T\textsubscript{hem} over the summer months of the EPiCC Vancouver campaign.
-sUHI shows clear diurnal dependence on differential heating/cooling rates at the urban and rural sites, with greater urban heating rates observed from sunrise to approximately solar noon forcing the development of sUHI and greater urban cooling rates in the afternoon and evening hours forcing the destruction of daytime sUHI.
-These results are highly generalizable to other mid-latitude cities.
 
@@
 	\label{rur_uhi_means}
-\end{figure}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=15cm,height=7in,keepaspectratio]{gen_case}
-	\caption{Top: Mean T\textsubscript{hem, r} for the Sunset tower (urban) and Westham island (rural) sites in Vancouver, Canada over May - Aug 2009.
-Middle: Mean heading/cooling rates for urban and rural sites.
-Bottom: Mean sUHI\textsubscript{hemi} and hourly rate of change in sUHI\textsubscript{hemi}.
-n = 5905}
-	\label{gen_case}
 \end{figure}
@@
 For the Sperrstrasse canyon, the combination of a narrow but relatively low street canyon results in a slight reduction in Basel's sUHI as strong radiative cooling of rooftops is slightly offset by canyon radiation trapping.
-In Vancouver, urban cooling rates quickly overtake rural - owing to strong radiative cooling - and sUHI declines precipitously after sUHI\textsubscript{max}.
+In Vancouver, urban cooling rates quickly overtake rural rates - owing to strong radiative cooling - and sUHI declines precipitously after sUHI\textsubscript{max}.
 The dependence of sUHI development on canyon geometry is illustrated well by comparing nadir and complete representations of sUHI observed in Basel.
@@
 Both of these cases demonstrate the dependence of observed sUHI on urban and rural site characteristics and sensor-surface geometry illustrating the need for proper critical analysis of the ensemble of climactic forcings present at a given urban-rural pairing and the potential for influence from sensor or instrument bias.
+
+\subsubsection{The general temporal evolution of sUHI}
+\label{gen_mod}
+
+\citet{Oke1982} presents a generalizable model for diurnal clUHI development.
+A similar schematic is presented in \citet{Oke2017} for sUHI based on simulations in \citet{Stewart2014}, this study provides the first observational support of modeled assessments of typical diurnal sUHI development.
+To provide for sUHI what \citet{Oke1982} did for clUHI, Figure \ref{gen_case} presents the diurnal evolution of urban and rural T\textsubscript{hem, r}, heating and cooling rates, and the resulting sUHI development constructed from mean urban and rural T\textsubscript{hem} observed over the summer months of the EPiCC Vancouver.
+Mean aUHI calculated for the same time period is also included for reference.
+However, it should be noted that rural and urban T\textsubscript{air} were measured from 2 \si{\meter} and 26 \si{\meter} respectively, so do not constitute a true clUHI.
+
+Development regimes are similar to those found for simulations of diurnal sUHI in \citet{Oke2017}, both in terms of the diurnal evolution of sUHI and its dependence on urban to rural differentials in heating/cooling rates.
+sUHI shows clear diurnal dependence on differential heating/cooling rates at the urban and rural sites, with greater urban heating rates observed from sunrise to approximately solar noon forcing the development of sUHI and greater urban cooling rates in the afternoon and evening hours forcing the destruction of daytime sUHI.
+These results are generalizable to other mid-latitude cities for sUHI measured over low- to mid-rise neighborhoods under summertime conditions.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=15cm,height=7in,keepaspectratio]{gen_case1}
+	\caption{Top: Mean T\textsubscript{hem, r} for the Sunset tower (urban) and Westham island (rural) sites in Vancouver, Canada over May - Aug 2009.
+Middle: Mean heading/cooling rates for urban and rural sites.
+Bottom: Mean sUHI\textsubscript{hemi} and aUHI and the hourly rate of change in sUHI\textsubscript{hemi} and aUHI.
+n = 5905}
+	\label{gen_case}
+\end{figure}
 
@@
 Results of sensor placement sensitivity testing in Section \ref{sensitivity} show that T\textsubscript{hem, r} is dependent on sensor position and height relative to canyon features for pyrgeometers mounted below 3 to 5 times mean building height.
-As pyrgeometers in both EPiCC and BUBBLE campaigns were mounted below these heights, derived T\textsubscript{hem, r} include some bias towards facets that make up a majority of their FOV.
-These effects are translated into sUHI\textsubscript{hem}.
+As pyrgeometers in both EPiCC and BUBBLE campaigns were mounted below these heights, derived T\textsubscript{hem, r} include some bias towards facets that make up a majority of their FOV, these effects are translated into sUHI\textsubscript{hem}.
 This bias results in a small overestimation of daytime T\textsubscript{comp}, and an underestimation of nighttime T\textsubscript{comp}, manifesting as a small increase (decrease) in daytime (nighttime) sUHI\textsubscript{hem}.
@@
 As such, sUHI development is strongly tied to integrated solar radiant exposure accumulated over a day and wind velocity, which display positive and negative relationships with sUHI magnitudes respectively.
-Other variables also show weakly positive (mean atmospheric water vapor content) and negative (soil water content) relationships with sUHI, however, for these variables, conclusions about causality are more difficult to make.
+Other variables also show weak (mean atmospheric water vapor content) and weakly negative (soil water content) relationships with sUHI, however, for these variables, conclusions about causality are more difficult to make.
 Meteorological controls on sUHI are summarized in Table \ref{meteo_cont}.
@@
 Therefore, typical variations in urban and rural humidities have little effect on the fraction of outgoing longwave radiation available for heating the canopy layer air volume through absorption and for heating of the surface by atmospheric emission.
-As a result, a similar relationship is expected to that found with correction magnitudes in Chapter \ref{paper1} is expected between humidity and sUHI.
+As a result, the relationship between humidity and sUHI is expected to be similar that found between humidity and correction magnitudes in Chapter \ref{paper1}.
 A weak positive correlation between the two is therefore likely a result of coincident seasonality in both variables as both mean daytime sUHI and mean water vapor mass densities are large in summer and small in winter.
@@
 Conversely, by day, reduced rural heating rates may result in higher sUHI magnitudes.
-Results in Vancouver do not support these suggestions.
+Results in Vancouver do not support these suggestions when viewed in aggregate.
 Over the daytime hours, in Vancouver, a negative relationship is observed between sUHI and rural soil moisture where sUHI is largest with low rural soil moisture.
 At night, no clear relationship between the two emerges.
-As Vancouver has distinct wet and dry seasons and little variability in soil moisture during the summer months (when sUHI is highly variable), study of the effect in cities with wet summers may yield a more complete picture of the relationship between sUHI and rural soil moisture.
+However, this relationship should be viewed with caution as both variables respond strongly to incoming solar radiation.
+Soil moisture content in Vancouver is greatest just before morning (when sUHI is small) and lowest in late afternoon (when sUHI is large).
+This results in a negative correlation when the two are compared over a time series, but may not indicate causality.
+In addition, Vancouver has distinct wet and dry seasons and little variability in soil moisture during the summer months (when sUHI is highly variable), study of the effect in cities with wet summers may yield a more complete picture of the relationship between sUHI and rural soil moisture.
 

### 2017-05-11 uw-ethesis-frontpgs.tex
Add files via upload

--- 2017-05-10 uw-ethesis-frontpgs.tex
@@
 	\centering
-	\begin{tabular}{p{1.5cm}p{2cm}p{10cm}}
+	\begin{tabular}{p{1.5cm}p{3cm}p{10cm}}
 		Roman &&\\
 		\toprule
-
 		Symbol & Unit & Property \\
@@
 		FOV & \si{\deg} & sensor field of view \\
-		\textit{H} & \si{\meter} & building width \\
+		\textit{H} & \si{\meter} & building height \\
+		H & \si{\gram\per\meter\cubed} & water vapor mass density \\
+		K\textsubscript{up} & \si{\watt\per\square\meter} & upwelling shortwave radiation \\
+		K\textsubscript{down} & \si{\watt\per\square\meter} & downwelling shortwave radiation \\
+		$L_0$ & \si{\watt\per\square\meter} & irradiance at the surface\\
+		$L_z$ & \si{\watt\per\square\meter} & irradiance at height $z$\\
+		$L^{\downarrow}_{sky}$ & \si{\watt\per\square\meter} & upwelling longwave flux density emitted from the atmosphere\\
+		$L^{\uparrow}_z$ & \si{\watt\per\square\meter} & upwelling longwave flux density at height $z$\\
+		$L_{atm}^{at-sensor}$ & \si{\watt\per\square\meter} & the component of at-sensor irradiance from the atmosphere\\
+		$L_{surf}^{at-sensor}$ & \si{\watt\per\square\meter} & the component of at-sensor irradiance from the surface\\
+		$L_{total}^{at-sensor}$ & \si{\watt\per\square\meter} & total at-sensor irradiance\\
+		L\textsubscript{down} & \si{\watt\per\square\meter} & downwelling longwave flux density \\
+		L\textsubscript{up} & \si{\watt\per\square\meter} & upwelling longwave flux density \\
+		$L'_z$ & \si{\watt\per\square\meter\per\steradian} & directional radiance through path length $z$\\
+		P & \si{\hecto\pascal} & atmospheric pressure \\
+		\bottomrule
+	\end{tabular} 
+\end{table}
+
+\begin{table}[H]
+	\centering
+	\begin{tabular}{p{1.5cm}p{3cm}p{10cm}}
+		\toprule
+		Symbol & Unit & Property \\
+		\midrule
 		$Q*$ & \si{\watt\per\square\meter} & net radiation flux density \\
@@
 		$Q_H$ & \si{\watt\per\square\meter} & sensible heat flux density \\
+		$ r $ & - & spectral sensor response \\
+		$\overline{r}$ & - & Planck weighted mean broadband sensor response \\
+		$R$ & \si{\watt\per\square\meter\per\steradian\per\micro\meter} & spectral directional radiance \\
+		$R^{\uparrow}_0$ & \si{\watt\per\square\meter\per\steradian\per\micro\meter} & upwelling spectral directional radiance at height zero\\
+		$R^{\uparrow}_{atm}$ & \si{\watt\per\square\meter\per\steradian\per\micro\meter} & upwelling spectral directional radiance emitted from the atmosphere\\
+		$R^{\downarrow}_{sky}$ & \si{\watt\per\square\meter\per\steradian\per\micro\meter} & downwelling spectral directional radiance emitted from the sky/atmosphere\\
+		$R^{\uparrow}_z$ & \si{\watt\per\square\meter\per\steradian\per\micro\meter} & upwelling spectral directional radiance at height $z$\\
 		$t$ & \si{\minute} & time \\
+		T & \si{\kelvin} & temperature \\
+		T\textsubscript{adj} & \si{\kelvin} & dome adjusted T\textsubscript{hem, r} \\
 		T\textsubscript{air} & \si{\kelvin} & air temperature \\
+		T\textsubscript{bias} & \si{\kelvin} & bias in T\textsubscript{hem, r} in terms of temperature \\
+		T\textsubscript{bright} & \si{\kelvin} & surface brightness temperature\\
+		T\textsubscript{calib} & \si{\kelvin} & instrument calibration temperature\\
+		T\textsubscript{comp} & \si{\kelvin} & complete urban surface temperature\\
+		T\textsubscript{emiss} & \si{\kelvin} & emission temperature\\
+		T\textsubscript{hem} & \si{\kelvin} & urban surface temperature viewed from a hemispherical remote sensor\\
+		T\textsubscript{hem, b} & \si{\kelvin} & urban surface brightness temperature viewed from a hemispherical remote sensor\\
+		T\textsubscript{hem, r} & \si{\kelvin} & urban surface radiometric  temperature viewed from a hemispherical remote sensor\\
+		\bottomrule
+	\end{tabular} 
+\end{table}
+\begin{table}[H]
+	\centering
+	\begin{tabular}{p{1.5cm}p{3cm}p{10cm}}
+		\toprule
+		Symbol & Unit & Property \\
+		\midrule
+		T\textsubscript{plan} & \si{\kelvin} & urban surface temperature viewed from a sensor in the nadir\\
+		T\textsubscript{rad} & \si{\kelvin} & radiometric surface temperature\\
+		T\textsubscript{road} & \si{\kelvin} & road surface temperature\\
+		T\textsubscript{roof} & \si{\kelvin} & roof surface temperature \\
 		T\textsubscript{surf} & \si{\kelvin} & surface temperature \\
-		T\textsubscript{comp} & \si{\kelvin} & complete urban surface temperature\\
-		T\textsubscript{plan} & \si{\kelvin} & urban surface temperature viewed from a sensor in the nadir\\
-		T\textsubscript{hem} & \si{\kelvin} & urban surface temperature viewed from a hemispherical remote sensor\\
-		aUHI & \si{\kelvin} & air temperature urban heat island effect \\
-		clUHI & \si{\kelvin} & canopy layer air temperature urban heat island effect \\
-		sUHI & \si{\kelvin} & surface temperature urban heat island effect \\
+		T\textsubscript{wall} & \si{\kelvin} & wall surface temperature\\
+		aUHI & \si{\kelvin} & air temperature urban heat island magnitude \\
+		clUHI & \si{\kelvin} & canopy layer air temperature urban heat island magnitude \\
+		sUHI & \si{\kelvin} & surface temperature urban heat island magnitude \\
+		sUHI\textsubscript{comp}  & \si{\kelvin} & surface urban heat island magnitude calculated from T\textsubscript{comp} \\
+		sUHI\textsubscript{hem}  & \si{\kelvin} & surface urban heat island magnitude calculated from T\textsubscript{hem, r} \\
+		sUHI\textsubscript{plan}  & \si{\kelvin} & surface urban heat island magnitude calculated from T\textsubscript{plan} \\
 		$V_B$ & - & effects from background climate on observation of a given meteorological variable \\
@@
 		$V_{M  R}$ & - & observed meteorological variable (rural) \\
-		$z$ & \si{\meter} & layer thickness (air, surface, or subsurface)\\
+		\textit{W} & \si{\meter} & building width \\
+		WD & \si{\deg} & wind direction \\
+		WV & \si{\meter\per\second} & wind velocity \\
+		$z$ & \si{\meter} & height/path length \\
+		$z_H$ & \si{\meter} & mean building height \\
+		$z_t$ & \si{\meter} & layer thickness\\
 		\bottomrule
@@
 	\centering
-	\begin{tabular}{p{1.5cm}p{2cm}p{10cm}}
-		Roman &&\\
+	\begin{tabular}{p{1.5cm}p{3cm}p{10cm}}
+		Greek &&\\
 		\toprule
-
 		Symbol & Unit & Property \\
 		\midrule
-		$C$ & \si{\joule} & heat capacity \\
-		FOV & \si{\deg} & sensor field of view \\
-		\textit{H} & \si{\meter} & building width \\
-		$Q*$ & \si{\watt\per\square\meter} & net radiation flux density \\
-		$\Delta Q_A$ & \si{\watt\per\square\meter} & net advective heat flux density \\
-		$\Delta Q_G$ & \si{\watt\per\square\meter} & net heat storage flux density \\
-		$Q_E$ & \si{\watt\per\square\meter} & latent heat flux density \\
-		$Q_F$ & \si{\watt\per\square\meter} & anthropogenic heat flux density \\
-		$Q_H$ & \si{\watt\per\square\meter} & sensible heat flux density \\
-		$t$ & \si{\minute} & time \\
-		T\textsubscript{air} & \si{\kelvin} & air temperature \\
-		T\textsubscript{surf} & \si{\kelvin} & surface temperature \\
-		T\textsubscript{comp} & \si{\kelvin} & complete urban surface temperature\\
-		T\textsubscript{plan} & \si{\kelvin} & urban surface temperature viewed from a sensor in the nadir\\
-		T\textsubscript{hem} & \si{\kelvin} & urban surface temperature viewed from a hemispherical remote sensor\\
-		aUHI & \si{\kelvin} & air temperature urban heat island effect \\
-		clUHI & \si{\kelvin} & canopy layer air temperature urban heat island effect \\
-		sUHI & \si{\kelvin} & surface temperature urban heat island effect \\
-		$V_B$ & - & effects from background climate on observation of a given meteorological variable \\
-		$V_H$ & - & effects from the urban environment on observation of a given meteorological variable \\
-		$V_L$ & - & effects from local topography on observation of a given meteorological variable \\
-		$V_M$ & - & observed meteorological variable \\
-		$V_{M, U} $ & - & observed meteorological variable (urban) \\
-		$V_{M  R}$ & - & observed meteorological variable (rural) \\
-		$z$ & \si{\meter} & layer thickness (air, surface, or subsurface)\\
+		$\alpha $ & - & albedo \\
+		$\epsilon$ & - & emissivity \\
+		$\epsilon_\lambda$ & - & spectral emissivity \\
+		$\theta$ & \si{\deg} & zenith angle \\
+		$\lambda$ & \si{\micro\meter} & wavelength \\
+		$\lambda_B $ & - & plan area of buildings \\
+		$\lambda_C $ &- & complete aspect ratio \\
+		$\lambda_I $ & -& plan area of impervious ground \\
+		$ \lambda_P$ & -& plan aspect ratio\\
+		$ \lambda_T $ & -& plan area of trees \\
+		$ \lambda_V$ & -& plan area of ground vegetation \\
+		$\nu$ & \si{\per\centi\meter} & wavenumber \\
+		$\sigma$ & \si{\watt\meter^{-2}\kelvin^{-4}} & Stefan Boltzmann constant \\
+		$\sigma_H $ & \si{\meter} & standard deviation of building height \\
+		$\tau $ & - & atmospheric transmittance \\ 
+		$\tau_d $ & - & atmospheric transmittance \\ 
+		$\tau_\lambda $ & - & spectral atmospheric transmittance \\ 
+		$\tau_{\Phi} $ & - & hemispherical  atmospheric transmittance \\ 
+		$\phi$ & \si{\deg} & azimuth angle \\
+		$\Phi $ & - & view factor \\
 		\bottomrule
